### 2022-06-28 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@thomasckng}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit. 
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis. 
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna. 
    Donec vehicula augue eu neque, morbi tristique senectus et netus et. 
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum. 
    Phasellus eu tellus sit amet tortor gravida placerat. 
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. 
    Praesent eget sem vel leo ultrices bibendum. 
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac. 
    Curabitur auctor semper nulla donec varius orci eget risus. 
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam. 
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. 
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis. 
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna. 
Donec vehicula augue eu neque, morbi tristique senectus et netus et. 
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum. 
Phasellus eu tellus sit amet tortor gravida placerat. 
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. 
Praesent eget sem vel leo ultrices bibendum. 
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac. 
Curabitur auctor semper nulla donec varius orci eget risus. 
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam. 
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi. 
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae. 
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna. 
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante. 
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient. 
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis. 
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2022-06-28 src/tex/ms.tex
Update basic content
Replace default content

@@
-% Define document class
 \documentclass[twocolumn]{aastex631}
@@
 
-% Begin!
 \begin{document}
 
-% Title
-\title{An open source scientific article}
+\title{Constraining gravitational wave amplitude birefringence and Chern-Simons gravity with GWTC-3}
 
-% Author list
-\author{@thomasckng}
+\author{Thomas C.K. Ng}
 
-% Abstract with filler text
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit. 
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis. 
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna. 
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et. 
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum. 
-    Phasellus eu tellus sit amet tortor gravida placerat. 
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. 
-    Praesent eget sem vel leo ultrices bibendum. 
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac. 
-    Curabitur auctor semper nulla donec varius orci eget risus. 
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam. 
-    Duis eget orci sit amet orci dignissim rutrum.
+
 \end{abstract}
 
-% Main body with filler text
 \section{Introduction}
-\label{sec:intro}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. 
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis. 
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna. 
-Donec vehicula augue eu neque, morbi tristique senectus et netus et. 
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum. 
-Phasellus eu tellus sit amet tortor gravida placerat. 
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. 
-Praesent eget sem vel leo ultrices bibendum. 
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac. 
-Curabitur auctor semper nulla donec varius orci eget risus. 
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam. 
-Duis eget orci sit amet orci dignissim rutrum.
+\citep{Maria_2021}.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi. 
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae. 
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna. 
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante. 
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient. 
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis. 
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+\section{Method}
+
+\section{Results}
+
+\section{Discussion}
+
+\section{Acknowledgements}
 

### 2022-06-28 src/tex/ms.tex
Update ms.tex
Insert points for some sections and content for method

@@
 \usepackage{showyourwork}
+\usepackage{amsfonts,amssymb,amsmath}
 
@@
 
-\citep{Maria_2021}.
+Beyond-GR theory\\
+Birefringence\\
+\citep{Maria_2021}
+
+Bilby\\
+Bayesian
 
@@
 
+Mentioned above, the amount of amplitude birefringence is represented by the opacity parameter $\kappa$.
+We performed estimation on $\kappa$ by Bilby with data from the GWTC catalog to constrain the plausible range of $\kappa$.
+
+To implement the amplitude birefringence on Bilby, waveforms generated by different approximants are modified according to the following equation:
+\begin{equation}
+    h_{L,R}^{\textrm{new}}\left(f\right)=
+    h_{L,R}^{\textrm{old}}\left(f\right)\times
+    \exp\left(\pm\kappa\frac{d_C}{1\textrm{ Gpc}}\frac{f}{0.1\textrm{ kHz}}\right)
+\end{equation}, where $h_{L,R}$ is the amplitude of the left or right polarization of the generated waveform
+, $\kappa$ is the dimensionless opacity parameter, $d_C$ is the comoving distance, and $f$ is the frequency of the generated waveform.
+
+Bilby will then perform estimation on the original gravitational wave parameters as well as $\kappa$.
+
 \section{Results}
 
+GWTC-3
+
 \section{Discussion}
+
+Limitation\\
+orientation
+sensitivity
+
+Future Study\\
+Data from more detectors
 

### 2022-07-08 src/tex/ms.tex
New content

--- 2022-06-28 src/tex/ms.tex
@@
 
-\title{Constraining gravitational wave amplitude birefringence and Chern-Simons gravity with GWTC-3}
+\title{Constraining gravitational wave amplitude birefringence with GWTC-3}
 
@@
 
+\section{Everything}
+Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
+After a century, we know that GR do not agree with quantum theories at some length scale.
+To unify both theories, we need to study the possibility of different beyond-GR theories.
+Some beyond-GR theories suggested that gravitational waves (GW) amplitude birefringence is a property of space, which GR does not included.
+
+GW consist of two polarizations (i.e. $+$ and $\times$), which could be transformed into two circular polarizations (i.e. left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
+In GR, the amplitude ratio of the left and right polarisations only depends on $\iota$, the angle between our line of sight and the orbital angular momentum of the binary black holes (BH).
+\begin{equation}
+    \left(\frac{h_\mathrm{L}}{h_\mathrm{R}}\right)_\mathrm{GR}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
+\end{equation}where $h_L$ and $h_R$ are the amplitude of left and right polarizations of the GW respectively.
+For the birefringence property mentioned in \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$, but also the distance to the merger and the strength of the birefringence.
+\begin{equation}
+    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
+\end{equation}where $d_C$ is the comoving distance to the merger, $\widetilde{\kappa}$ is the opacity parameter that represent the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed amplitude of left and right polarizations of the GW respectively.
+In this paper, we considered the frequency dependence of the birefringence as well. Thus, the amplitude ratio of the left and right polarisations will be as follow.
+\begin{equation}
+    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{new}=\frac{\exp\left({\kappa\frac{d_C}{1\mathrm{ Gpc}}\frac{f}{100\mathrm{ Hz}}}\right)\left(1-\cos\iota\right)^2}{\exp\left({-\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}}\right)\left(1+\cos\iota\right)^2}\,,
+\end{equation}where $f$ is the frequency of the GW and $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence.
+This frequency dependence provides stronger enhancement or suppression from the birefrigence to high frequency components of the GW, and visa versa.
+
+We performed estimations on $\kappa$ using Bilby with data from the GWTC catalog to constrain the plausible range of $\kappa$.
+Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from interferometer and priors of the parameters provided. \citep{Ashton_2019}
+It starts by generating a waveform using the chosen approximant such as IMRPhenomPv2 with a set of parameters.
+Bilby will then compute the likelihood of getting the provided data from the interferometer if the waveform just generated is the true waveform.
+Using the Bayes' theorem, the posterior probability of the parameters is calculated.
+To obtain the whole posterior probability, a chosen sampler is used to run over the parameter space instead of calculating the posterior for the whole parameter space.
+
+To implement the amplitude birefringence on Bilby, waveforms generated by different approximants are modified before calculating the likelihood according to the following equation:
+\begin{equation}
+    h_\mathrm{L,R}^{\mathrm{new}}=
+    h_\mathrm{L,R}^{\mathrm{old}}\times
+    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{ Gpc}}\frac{f}{100\mathrm{ Hz}}\right)\,.
+\end{equation}
+This modification allows Bilby to perform estimations on $\kappa$, as the waveforms generated will depends on $\kappa$ as well.
+
+GWTC-3 Results
+
+Limitation\\
+orientation
+sensitivity
+
+Future Study\\
+Data from more detectors
+
 \section{Introduction}
-
-\citep{Maria_2021}.
 

### 2022-07-15 src/tex/ms.tex
Update ms.tex
update content

@@
 To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories suggested that gravitational waves (GW) amplitude birefringence is a property of space, which GR does not included.
+Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GW) amplitude birefringence is a property of space, which GR does not included.
 
-GW consist of two polarizations (i.e. $+$ and $\times$), which could be transformed into two circular polarizations (i.e. left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
-In GR, the amplitude ratio of the left and right polarisations only depends on $\iota$, the angle between our line of sight and the orbital angular momentum of the binary black holes (BH).
+GW consist of two linear polarizations (i.e.
+$+$ and $\times$) similar to electromagnetic waves, which could be transformed into two circular polarizations (i.e.
+left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
+In GR, the amplitude ratio of the left and right polarisations only depends on $\iota$, the angle between our line of sight and the orbital angular momentum of the binary black holes (BBH).
 \begin{equation}
@@
 \end{equation}where $h_L$ and $h_R$ are the amplitude of left and right polarizations of the GW respectively.
-For the birefringence property mentioned in \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$, but also the distance to the merger and the strength of the birefringence.
+For the birefringence property mentioned in \citet{Maria_2021} and other papers, the amplitude ratio depends on not only $\iota$, but also the distance to the merger and the strength of the birefringence.
 \begin{equation}
@@
 \end{equation}where $d_C$ is the comoving distance to the merger, $\widetilde{\kappa}$ is the opacity parameter that represent the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed amplitude of left and right polarizations of the GW respectively.
+This is a preliminary version of the birefringence model in CS, which assume there is no frequency dependence.
+This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.
+
 In this paper, we considered the frequency dependence of the birefringence as well. Thus, the amplitude ratio of the left and right polarisations will be as follow.
@@
 This frequency dependence provides stronger enhancement or suppression from the birefrigence to high frequency components of the GW, and visa versa.
+Also, the frequency dependence can break the degeneracy between kappa and iota, as this would make the effect of the birefringence affect the amplitude ratio differently compared to iota.
 
@@
 Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from interferometer and priors of the parameters provided. \citep{Ashton_2019}
-It starts by generating a waveform using the chosen approximant such as IMRPhenomPv2 with a set of parameters.
-Bilby will then compute the likelihood of getting the provided data from the interferometer if the waveform just generated is the true waveform.
-Using the Bayes' theorem, the posterior probability of the parameters is calculated.
-To obtain the whole posterior probability, a chosen sampler is used to run over the parameter space instead of calculating the posterior for the whole parameter space.
 
-To implement the amplitude birefringence on Bilby, waveforms generated by different approximants are modified before calculating the likelihood according to the following equation:
+We assume the GW are produced at the BBH following GR. As the GW propagate through space, the effect of the birefringence will be built up as the distance increases.
+To implement the amplitude birefringence on Bilby, waveforms are modified according to the following equation:
 \begin{equation}
-    h_\mathrm{L,R}^{\mathrm{new}}=
-    h_\mathrm{L,R}^{\mathrm{old}}\times
+    h_\mathrm{L,R}^{\mathrm{Biref}}=
+    h_\mathrm{L,R}^{\mathrm{GR}}\times
     \exp\left(\pm\kappa\frac{d_C}{1\mathrm{ Gpc}}\frac{f}{100\mathrm{ Hz}}\right)\,.
 \end{equation}
-This modification allows Bilby to perform estimations on $\kappa$, as the waveforms generated will depends on $\kappa$ as well.
+This modification allows Bilby to perform estimations on $\kappa$, as the waveforms will depends on $\kappa$ as well.
 

### 2022-07-19 src/tex/ms.tex
Update ms.tex
Update temporary text

@@
 To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GW) amplitude birefringence is a property of space, which GR does not included.
+Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GWs) amplitude birefringence is a property of space, which GR does not included.
 
@@
     \left(\frac{h_\mathrm{L}}{h_\mathrm{R}}\right)_\mathrm{GR}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
-\end{equation}where $h_L$ and $h_R$ are the amplitude of left and right polarizations of the GW respectively.
-For the birefringence property mentioned in \citet{Maria_2021} and other papers, the amplitude ratio depends on not only $\iota$, but also the distance to the merger and the strength of the birefringence.
+\end{equation}where $h_L$ and $h_R$ are the amplitude of left and right polarizations of the GWs respectively.
+
+For other studies on the birefringence property such as\citet{Maria_2021}, the amplitude ratio depends on not only $\iota$, but also the distance to the merger and the strength of the birefringence.
 \begin{equation}
@@
     \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{new}=\frac{\exp\left({\kappa\frac{d_C}{1\mathrm{ Gpc}}\frac{f}{100\mathrm{ Hz}}}\right)\left(1-\cos\iota\right)^2}{\exp\left({-\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}}\right)\left(1+\cos\iota\right)^2}\,,
-\end{equation}where $f$ is the frequency of the GW and $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence.
+\end{equation}where $f$ is the frequency of the GWs and $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence.
 This frequency dependence provides stronger enhancement or suppression from the birefrigence to high frequency components of the GW, and visa versa.
@@
 
-We performed estimations on $\kappa$ using Bilby with data from the GWTC catalog to constrain the plausible range of $\kappa$.
+Estimations on $\kappa$ using Bilby with data from the the third LIGO-Virgo catalog, GWTC-3, was performed to constrain the plausible range of $\kappa$.
 Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from interferometer and priors of the parameters provided. \citep{Ashton_2019}
 
-We assume the GW are produced at the BBH following GR. As the GW propagate through space, the effect of the birefringence will be built up as the distance increases.
+We assume GWs are produced at the BBH following GR. As the GWs propagate through space, the effect of the birefringence will be built up as the distance increases.
 To implement the amplitude birefringence on Bilby, waveforms are modified according to the following equation:
@@
 
-GWTC-3 Results
+Consider only GW150914, the first detected GW by LIGO. In
 
 Limitation\\
-orientation
+orientation\\
 sensitivity

### 2022-07-21 src/tex/ms.tex
Add GW150914 figure
Update Content
Create point form section

@@
 \author{Thomas C.K. Ng}
+\email{thomas.ng@link.cuhk.edu.hk}
+
+\date{\today}
 
@@
 
+\section{Points}
+
+1 Introduction\\
+1.1 Motivation\\
+1.1.1 GR broke down at some scale, find a new theory
+
+2 Method
+
+3 Results
+
+4 Discussion\\
+4.1 Limitation\\
+4.1.1 orientation of detector (polarisations)\\
+4.1.2 sensitivity of detector
+
+4.2 Future Study\\
+4.2.1 Data from better detector
+
 \section{Everything}
@@
 To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GWs) amplitude birefringence is a property of space, which GR does not included.
+Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GWs) amplitude birefringence is a property of space,
+which GR does not included.
 
-GW consist of two linear polarizations (i.e.
-$+$ and $\times$) similar to electromagnetic waves, which could be transformed into two circular polarizations (i.e.
-left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
-In GR, the amplitude ratio of the left and right polarisations only depends on $\iota$, the angle between our line of sight and the orbital angular momentum of the binary black holes (BBH).
+GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves,
+which could be transformed into two circular polarizations (i.e. left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
+In GR, the amplitude ratio of the left and right polarisations only depends on $\iota$,
+the angle between our line of sight and the orbital angular momentum of the binary black holes (BBH).
 \begin{equation}
@@
 
-For other studies on the birefringence property such as\citet{Maria_2021}, the amplitude ratio depends on not only $\iota$, but also the distance to the merger and the strength of the birefringence.
+For other studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
+but also the distance to the merger and the strength of the birefringence.
 \begin{equation}
     \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
-\end{equation}where $d_C$ is the comoving distance to the merger, $\widetilde{\kappa}$ is the opacity parameter that represent the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed amplitude of left and right polarizations of the GW respectively.
+\end{equation}where $d_C$ is the comoving distance to the merger, $\widetilde{\kappa}$ is the opacity parameter that represent
+the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are
+the observed amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is 0.
 This is a preliminary version of the birefringence model in CS, which assume there is no frequency dependence.
@@
 
-In this paper, we considered the frequency dependence of the birefringence as well. Thus, the amplitude ratio of the left and right polarisations will be as follow.
+In this paper, we considered the frequency dependence of the birefringence as well.
+Thus, the amplitude ratio of the left and right polarisations will be as follow.
 \begin{equation}
@@
 \end{equation}where $f$ is the frequency of the GWs and $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence.
-This frequency dependence provides stronger enhancement or suppression from the birefrigence to high frequency components of the GW, and visa versa.
-Also, the frequency dependence can break the degeneracy between kappa and iota, as this would make the effect of the birefringence affect the amplitude ratio differently compared to iota.
+This frequency dependence provides stronger enhancement or suppression from the birefringence to high frequency components of the GW, and visa versa.
+Also, the frequency dependence can break the degeneracy between kappa and iota,
+as this would make the effect of the birefringence affect the amplitude ratio differently compared to iota.
 
 Estimations on $\kappa$ using Bilby with data from the the third LIGO-Virgo catalog, GWTC-3, was performed to constrain the plausible range of $\kappa$.
-Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from interferometer and priors of the parameters provided. \citep{Ashton_2019}
+Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from
+interferometer and priors of the parameters provided. \citep{Ashton_2019}
 
-We assume GWs are produced at the BBH following GR. As the GWs propagate through space, the effect of the birefringence will be built up as the distance increases.
+We assume GWs are produced at the BBH following GR. As the GWs propagate through space,
+the effect of the birefringence will be built up as the distance increases.
 To implement the amplitude birefringence on Bilby, waveforms are modified according to the following equation:
@@
 
-Consider only GW150914, the first detected GW by LIGO. In
+Consider GW150914, the first detected GW by LIGO. In figure \ref{fig:GW150914_corner}, the posteriors of the parameters of the GW150914 is shown.
+With the frequency independent birefringence model, the posteriors look different from the posteriors with GR.
+This is because there is a degeneracy between $\kappa$ and $\iota$ without the frequency dependence,
+so $\iota$ was less constrained compared to the PE with GR.
+On the other hand, with the frequency dependent birefringence model, the posteriors look similar to the posteriors with GR.
+This is because the degeneracy was broken by the frequency dependence. In this case, the posteriors can recover the PE with GR,
+and the most probable value of $\kappa$ is close to 0, which means the birefringence is weak or absent, and GR can be recovered.
 
-Limitation\\
-orientation\\
-sensitivity
-
-Future Study\\
-Data from more detectors
+\begin{figure*}[h]
+    \script{GW150914_corner.py}
+    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
+    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
+    The three sets of plots are the parameter estimations (PE) done by LIGO with GR \citep{LIGO_2021},
+    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
+    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
+    which does not suggest GWs amplitude birefringence.}
+    \label{fig:GW150914_corner}
+\end{figure*}
 

### 2022-08-01 src/tex/ms.tex
Update ms.tex
Update list

@@
 
-1 Introduction\\
-1.1 Motivation\\
-1.1.1 GR broke down at some scale, find a new theory
-
-2 Method
-
-3 Results
-
-4 Discussion\\
-4.1 Limitation\\
-4.1.1 orientation of detector (polarisations)\\
-4.1.2 sensitivity of detector
-
-4.2 Future Study\\
-4.2.1 Data from better detector
+\begin{enumerate}
+    \item Introduction
+    \begin{enumerate}
+        \item Motivation
+        \begin{enumerate}
+            \item GR broke down at some scale, find a new theory
+            \item CS: beyond-GR theory
+            \item test CS property (GW amplitude birefringence)
+            \item GW have 2 polarisations
+            \item GR: polarisation ratio depends on iota
+        \end{enumerate}
+    \end{enumerate}
+    \item Method
+    \begin{enumerate}
+        \item
+    \end{enumerate}
+    \item Results
+    \begin{enumerate}
+        \item GW150914
+        \begin{enumerate}
+            \item
+        \end{enumerate}
+        \item GWTC-3
+        \begin{enumerate}
+            \item
+        \end{enumerate}
+    \end{enumerate}
+    \item Discussion
+    \begin{enumerate}
+        \item Limitation
+        \begin{enumerate}
+            \item orientation of detector (polarisations)
+            \item sensitivity of detector
+        \end{enumerate}
+        \item Future Study
+        \begin{enumerate}
+            \item Data from better detector
+        \end{enumerate}
+    \end{enumerate}
+\end{enumerate}
 

### 2022-08-01 src/tex/ms.tex
Update ms.tex

@@
 
-Estimations on $\kappa$ using Bilby with data from the the third LIGO-Virgo catalog, GWTC-3, was performed to constrain the plausible range of $\kappa$.
+Estimations on $\kappa$ using Bilby with data from the third LIGO-Virgo catalog, GWTC-3, was performed to constrain the plausible range of $\kappa$.
 Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from

### 2022-08-08 src/tex/ms.tex
Update ms.tex

@@
             \item test CS property (GW amplitude birefringence)
-            \item GW have 2 polarisations
-            \item GR: polarisation ratio depends on iota
+        \end{enumerate}
+        \item Amplitude Birefringence
+        \begin{enumerate}
+            \item GW amplitude birefringence is a property of space from beyond-GR theory
+            \item birefringence: enhance one and suppress another
+            \item GR predicts there is no birefringence
+            \item GR: amplitude ratio depends on $\iota$ only
+            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
         \end{enumerate}
@@
     \begin{enumerate}
+        \item Assume GW are generated as GR described
+        \item birefringence is considered during the propagation
         \item 
@@
         \begin{enumerate}
-            \item Data from better detector
+            \item Better data from detector
         \end{enumerate}
@@
 To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GWs) amplitude birefringence is a property of space,
-which GR does not included.
+Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GWs) amplitude birefringence is a property of space;
+in contrast, GR predicts there is no birefringence.
 
@@
 which could be transformed into two circular polarizations (i.e. left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
-In GR, the amplitude ratio of the left and right polarisations only depends on $\iota$,
-the angle between our line of sight and the orbital angular momentum of the binary black holes (BBH).
+In GR, the amplitude ratio of the left and right polarisations only depends on the inclination $\iota$ for a nonprecessing binary,
+the angle between our line of sight and the orbital angular momentum of the binary.
 \begin{equation}
     \left(\frac{h_\mathrm{L}}{h_\mathrm{R}}\right)_\mathrm{GR}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
-\end{equation}where $h_L$ and $h_R$ are the amplitude of left and right polarizations of the GWs respectively.
+\end{equation}where $h_L$ and $h_R$ are the Fourier amplitude of left and right polarizations of the GWs respectively.
 
@@
     \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
-\end{equation}where $d_C$ is the comoving distance to the merger, $\widetilde{\kappa}$ is the opacity parameter that represent
+\end{equation}where $d_C$ is the comoving distance to the binary, $\widetilde{\kappa}$ is the opacity parameter that represent
 the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are
-the observed amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is 0.
-This is a preliminary version of the birefringence model in CS, which assume there is no frequency dependence.
+the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is 0.
+This is a zeroth-order approximation of the birefringence model in CS, which assume there is no frequency dependence.
 This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.
@@
 
-We assume GWs are produced at the BBH following GR. As the GWs propagate through space,
+We assume GWs are generated at the binary as GR described. As the GWs propagate through space,
 the effect of the birefringence will be built up as the distance increases.
@@
 Consider GW150914, the first detected GW by LIGO. In figure \ref{fig:GW150914_corner}, the posteriors of the parameters of the GW150914 is shown.
-With the frequency independent birefringence model, the posteriors look different from the posteriors with GR.
-This is because there is a degeneracy between $\kappa$ and $\iota$ without the frequency dependence,
-so $\iota$ was less constrained compared to the PE with GR.
-On the other hand, with the frequency dependent birefringence model, the posteriors look similar to the posteriors with GR.
-This is because the degeneracy was broken by the frequency dependence. In this case, the posteriors can recover the PE with GR,
-and the most probable value of $\kappa$ is close to 0, which means the birefringence is weak or absent, and GR can be recovered.
+With the frequency independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
+This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
+To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$,
+or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
+As a result,many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
+
+On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
+This is because the degeneracy was broken by the frequency dependence, as the effect of the birefringence will be different from the one of changing $$\iota$$.
+In this case, the posteriors can recover the GR distributions for the binary parameters,
+and the most probable value of $$\kappa$$ is close to $$0$$, which means the birefringence is weak or absent, and GR can be recovered.
 

### 2022-08-09 src/tex/ms.tex
Update ms.tex

@@
         \begin{enumerate}
-            \item Better data from detector
+            \item More events from detector
+            \item Higher order terms
+            \item Longer sampling
         \end{enumerate}
@@
 To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GWs) amplitude birefringence is a property of space;
-in contrast, GR predicts there is no birefringence.
+Some beyond-GR theories such as Chern-Simons theory (CS) suggested that there is gravitational wave (GW) amplitude birefringence,
+while GR predicts there is no birefringence.
 
+GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
+To illustrate the property, \textbf{[use snapshot from the animation to explain both frequency-dependent and frequency-independent birefringence]}
+
+Before we discuss more about GW amplitude birefringence, we need to understand the case of GR.
 GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves,
@@
 For other studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
-but also the distance to the merger and the strength of the birefringence.
+but also the distance to the binary and the strength of the birefringence.
 \begin{equation}
@@
 
-Estimations on $\kappa$ using Bilby with data from the third LIGO-Virgo catalog, GWTC-3, was performed to constrain the plausible range of $\kappa$.
+Estimations on $\kappa$ using Bilby with data from the third LIGO-Virgo catalog \textbf{[cite LIGO]}, GWTC-3,
+was performed to constrain the plausible range of $\kappa$.
 Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from
@@
     \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-    The three sets of plots are the parameter estimations (PE) done by LIGO with GR \citep{LIGO_2021},
+    The three sets of plots are the parameter estimations (PEs) done by LIGO with GR \citep{LIGO_2021},
     the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
     Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
-    which does not suggest GWs amplitude birefringence.}
+    which does not suggest GW amplitude birefringence.}
     \label{fig:GW150914_corner}

### 2022-08-10 src/tex/ms.tex
Update content

@@
 
-\section{Points}
+\section{Introduction}
 
-\begin{enumerate}
-    \item Introduction
-    \begin{enumerate}
-        \item Motivation
-        \begin{enumerate}
-            \item GR broke down at some scale, find a new theory
-            \item CS: beyond-GR theory
-            \item test CS property (GW amplitude birefringence)
-        \end{enumerate}
-        \item Amplitude Birefringence
-        \begin{enumerate}
-            \item GW amplitude birefringence is a property of space from beyond-GR theory
-            \item birefringence: enhance one and suppress another
-            \item GR predicts there is no birefringence
-            \item GR: amplitude ratio depends on $\iota$ only
-            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
-        \end{enumerate}
-    \end{enumerate}
-    \item Method
-    \begin{enumerate}
-        \item Assume GW are generated as GR described
-        \item birefringence is considered during the propagation
-        \item 
-    \end{enumerate}
-    \item Results
-    \begin{enumerate}
-        \item GW150914
-        \begin{enumerate}
-            \item 
-        \end{enumerate}
-        \item GWTC-3
-        \begin{enumerate}
-            \item 
-        \end{enumerate}
-    \end{enumerate}
-    \item Discussion
-    \begin{enumerate}
-        \item Limitation
-        \begin{enumerate}
-            \item orientation of detector (polarisations)
-            \item sensitivity of detector
-        \end{enumerate}
-        \item Future Study
-        \begin{enumerate}
-            \item More events from detector
-            \item Higher order terms
-            \item Longer sampling
-        \end{enumerate}
-    \end{enumerate}
-\end{enumerate}
-
-\section{Everything}
 Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
@@
 
-For other studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
+For previous studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
 but also the distance to the binary and the strength of the birefringence.
@@
 the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are
-the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is 0.
+the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is $0$.
 This is a zeroth-order approximation of the birefringence model in CS, which assume there is no frequency dependence.
@@
 
-Estimations on $\kappa$ using Bilby with data from the third LIGO-Virgo catalog \textbf{[cite LIGO]}, GWTC-3,
-was performed to constrain the plausible range of $\kappa$.
-Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from
-interferometer and priors of the parameters provided. \citep{Ashton_2019}
+\section{Method}
+
+To constrain $\kappa$, we perform parameter estimation (PE) with data from the third LIGO-Virgo catalog \citep{LIGO_2021}, GWTC-3, using Bilby,
+a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data
+and priors for the parameters in question. \citep{Ashton_2019}
 
@@
 
+\section{Results}
+
+\subsection{Result for GW150914}
+
+\begin{figure*}[h]
+    \script{GW150914_corner.py}
+    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
+    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
+    The three sets of plots are the PEs done by LIGO with GR \citep{LIGO_2021},
+    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
+    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
+    which does not suggest GW amplitude birefringence.}
+    \label{fig:GW150914_corner}
+\end{figure*}
+
 Consider GW150914, the first detected GW by LIGO. In figure \ref{fig:GW150914_corner}, the posteriors of the parameters of the GW150914 is shown.
@@
 On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
-This is because the degeneracy was broken by the frequency dependence, as the effect of the birefringence will be different from the one of changing $$\iota$$.
+This is because the degeneracy was broken by the frequency dependence, as the effect of the birefringence will be different from the one of changing $\iota$.
 In this case, the posteriors can recover the GR distributions for the binary parameters,
-and the most probable value of $$\kappa$$ is close to $$0$$, which means the birefringence is weak or absent, and GR can be recovered.
+and the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent, and GR can be recovered.
 
-\begin{figure*}[h]
-    \script{GW150914_corner.py}
-    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
-    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-    The three sets of plots are the parameter estimations (PEs) done by LIGO with GR \citep{LIGO_2021},
-    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
-    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
-    which does not suggest GW amplitude birefringence.}
-    \label{fig:GW150914_corner}
-\end{figure*}
-
-\section{Introduction}
-
-\section{Method}
-
-\section{Results}
+\subsection{Result on GWTC-3}
 
 \section{Discussion}
+
+\subsection{Limitation}
+
+\subsection{Future Study}
 
@@
 
+\section{Points}
+
+\begin{enumerate}
+    \item Introduction
+    \begin{enumerate}
+        \item Motivation
+        \begin{enumerate}
+            \item GR broke down at some scale, find a new theory
+            \item CS: beyond-GR theory
+            \item test CS property (GW amplitude birefringence)
+        \end{enumerate}
+        \item Amplitude Birefringence
+        \begin{enumerate}
+            \item GW amplitude birefringence is a property of space
+            \item birefringence: enhance one and suppress another
+        \end{enumerate}
+        \item Amplitude Ratio
+        \begin{enumerate}
+            \item GR: amplitude ratio depends on $\iota$ only
+            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
+        \end{enumerate}
+    \end{enumerate}
+    \item Method
+    \begin{enumerate}
+        \item PE on GWTC-3 to constrain $\kappa$
+        \item Assume GW are generated as GR described
+        \item birefringence is considered during the propagation
+        \item waveform modification
+    \end{enumerate}
+    \item Results
+    \begin{enumerate}
+        \item GW150914
+        \begin{enumerate}
+            \item figure with frequency-dependent and frequency-independent
+            \item posterior from $\cos\iota$ looks different
+            \item frequency-dependent: degeneracy between $\iota$ and $\kappa$
+            \item frequency-independent: degeneracy broken and $\kappa$ near $0$
+        \end{enumerate}
+        \item GWTC-3
+        \begin{enumerate}
+            \item figure of $\kappa$ of all events
+            \item constraint on $\kappa$
+            \item improvement of constraint on $\kappa$ compare to previous studies
+        \end{enumerate}
+    \end{enumerate}
+    \item Discussion
+    \begin{enumerate}
+        \item Limitation
+        \begin{enumerate}
+            \item orientation of detector (polarisations)
+            \item sensitivity of detector
+        \end{enumerate}
+        \item Future Study
+        \begin{enumerate}
+            \item More events from detector
+            \item Higher order terms
+        \end{enumerate}
+    \end{enumerate}
+\end{enumerate}
+
 \end{document}

### 2022-08-10 src/scripts/GW150914_corner.py

@@ -37,6 +37,6 @@ g.axes[1,0].set_ylabel("$d_L$")
 g.axes[2,1].set_xlabel("$d_L$")
 g.axes[2,0].set_ylabel("$\cos\iota$")
 g.axes[2,2].set_xlabel("$\cos\iota$")
-g._legend.set_bbox_to_anchor((0.65, 0.85))
+g.fig.legends[0].set_bbox_to_anchor((0.65,0.8))
 
 g.savefig(fname=paths.figures/"GW150914_corner.png", bbox_inches="tight", dpi=300)

### 2022-08-10 src/tex/ms.tex
Update ms.tex

@@
 \email{thomas.ng@link.cuhk.edu.hk}
+\affiliation{Department of Physics and Institute of Theoretical Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
+
+\author{Maximiliano Isi}
+\email{misi@flatironinstitute.org}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, United States}
+
+\author{Kaze W. K. Wong}
+\email{kwong@flatironinstitute.org}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, United States}
+
+\author{Will M. Farr}
+\email{wfarr@flatironinstitute.org}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, United States}
+\affiliation{Department of Physics and Astronomy, Stony Brook University, Stony Brook NY 11794, United States}
 
@@
 while GR predicts there is no birefringence.
+GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
 
-GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
-To illustrate the property, \textbf{[use snapshot from the animation to explain both frequency-dependent and frequency-independent birefringence]}
+% To illustrate the effect of amplitude birefringence,
+% the amplitudes of two polarisations of a artificial GW signal are shown in figure \textbf{[X]}.
 
@@
             \item birefringence: enhance one and suppress another
+            \item figure to illustrate the effect of birefringence
         \end{enumerate}

### 2022-08-10 src/tex/ms.tex
Update ms.tex

@@
 while GR predicts there is no birefringence.
+
 GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
-
-% To illustrate the effect of amplitude birefringence,
-% the amplitudes of two polarisations of a artificial GW signal are shown in figure \textbf{[X]}.
+The longer the GW travels, the larger the effect of birefringence.
+To illustrate the effect of amplitude birefringence,
+the observed amplitudes of two polarisations of an artificial GW signal example are shown in figure \textbf{[X]}.
+\textbf{[add figure similar to the animation in the website]}
+With birefringence, the amplitude is modified during the propagation,
+and the observed amplitudes will be different compare to the amplitudes when they are generated at the binary.
 
@@
 
-\begin{figure*}[h]
+\begin{figure}[h!]
     \script{GW150914_corner.py}
-    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
+    \includegraphics[width=\columnwidth]{figures/GW150914_corner.png}
     \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
@@
     \label{fig:GW150914_corner}
-\end{figure*}
+\end{figure}
 
@@
 \subsection{Result on GWTC-3}
+
+We then perform PE on $71$ binary black hole merger events in GWTC-3 to further constrain $\kappa$.
+The posteriors for $\kappa$ of the events are shown in figure \textbf{[X]}, which shows $\kappa$ tend to be $0$.
+The combined posterior for $\kappa$ was calculated and shown in figure \textbf{[X]}, which allows us to constrain $\kappa$ within \textbf{[the number we get]}.
 
@@
     \begin{enumerate}
+        \item Parameter Estimation
+        \begin{enumerate}
             \item PE on GWTC-3 to constrain $\kappa$
@@
             \item waveform modification
+        \end{enumerate}
     \end{enumerate}

### 2022-08-11 src/tex/ms.tex
New Content & Graph

@@
 
+\begin{figure}[h!]
+    \script{birefringence.py}
+    \includegraphics[width=\columnwidth]{figures/birefringence.png}
+    \caption{
+        The observed amplitudes of two polarisations of an artificial GW signal under different model.
+        The top panel shows the case of GR, the amplitudes are unmodified.
+        For the middle and bottom panel, they shows the case of birefringence.
+        Left-handed polarisation is favored by the birefringence in the middle panel,
+        and right-handed polarisation is favored by the birefringence in the bottom panel.
+    }
+    \label{fig:birefringence}
+\end{figure}
+
 GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
@@
 To illustrate the effect of amplitude birefringence,
-the observed amplitudes of two polarisations of an artificial GW signal example are shown in figure \textbf{[X]}.
-\textbf{[add figure similar to the animation in the website]}
+the observed amplitudes of two polarisations of an artificial GW signal example are shown in figure \ref{fig:birefringence}.
 With birefringence, the amplitude is modified during the propagation,
@@
 
-To constrain $\kappa$, we perform parameter estimation (PE) with data from the third LIGO-Virgo catalog \citep{LIGO_2021}, GWTC-3, using Bilby,
+To constrain $\kappa$, we perform parameter estimation (PE) with data from the third LIGO-Virgo catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby,
 a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data
-and priors for the parameters in question. \citep{Ashton_2019}
+and priors for the parameters in question. \citep{Bilby}
 
@@
 
+After modifying Bilby, we set the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as using GR,
+to test if the PEs could recover the PE done by LIGO in \citet{GWTC-2.1, GWTC-3}. These control tests result can be found in the data release.
+We then set the prior on $\kappa$ to a uniform function between $-1$ and $1$, and perform PEs on the modified waveforms.
+In both case, we use the same configuration as in \citep{GWTC-2.1, GWTC-3} for each individual events.
+
 \section{Results}
@@
     \includegraphics[width=\columnwidth]{figures/GW150914_corner.png}
-    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-    The three sets of plots are the PEs done by LIGO with GR \citep{LIGO_2021},
+    \caption{
+        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
+    The three sets of plots are the PEs done by LIGO with GR \citep{GWTC-2.1, GWTC-3},
     the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
     Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
-    which does not suggest GW amplitude birefringence.}
+    which does not suggest GW amplitude birefringence.
+    }
     \label{fig:GW150914_corner}
@@
             \item waveform modification
+            \item GR control test
+            \item same configuration as LIGO
         \end{enumerate}

### 2022-08-11 src/scripts/birefringence.py

@@ -1,0 +1,54 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import bilby
+from gwpy.timeseries import TimeSeries
+import paths
+
+def smooth(y):
+    box_pts = 30
+    box = np.ones(box_pts)/box_pts
+    y_smooth = np.convolve(y, box, mode='same')
+    return y_smooth
+
+time_of_event = 1264316116.4
+detectors={'L1', 'V1'}
+ifos = bilby.gw.detector.InterferometerList(detectors)
+
+duration = 8.0
+post_trigger_duration = 2.
+analysis_start = time_of_event + post_trigger_duration - duration
+
+for detector in ifos:
+    analysis_data = TimeSeries.fetch_open_data(detector.name, analysis_start, analysis_start + duration, sample_rate=4096., cache=True)
+    detector.set_strain_data_from_gwpy_timeseries(analysis_data)
+
+plt.figure(figsize=(5,11))
+plt.subplot(311)
+plt.plot(ifos[0].frequency_array, smooth(np.sqrt(np.square(ifos[0].frequency_domain_strain))))
+plt.plot(ifos[1].frequency_array, smooth(np.sqrt(np.square(ifos[1].frequency_domain_strain))))
+plt.yscale('log')
+plt.xscale('log')
+plt.xlim(20,896)
+plt.ylim(10e-26,10e-20)
+plt.title("GR")
+plt.legend(["left-handed", "right-handed"])
+plt.subplot(312)
+plt.plot(ifos[0].frequency_array, smooth(np.sqrt(np.square(ifos[0].frequency_domain_strain)))*np.exp(0.3*0.9*ifos[0].frequency_array/100))
+plt.plot(ifos[1].frequency_array, smooth(np.sqrt(np.square(ifos[1].frequency_domain_strain)))*np.exp(-0.3*0.9*ifos[1].frequency_array/100))
+plt.yscale('log')
+plt.xscale('log')
+plt.xlim(20,896)
+plt.ylim(10e-26,10e-20)
+plt.ylabel("$h(f)$")
+plt.title("Birefringence (left-handed)")
+plt.subplot(313)
+plt.plot(ifos[0].frequency_array, smooth(np.sqrt(np.square(ifos[0].frequency_domain_strain)))*np.exp(-0.3*0.9*ifos[0].frequency_array/100))
[…]

### 2022-08-12 src/tex/ms.tex
Update Graphs

@@
 To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories such as Chern-Simons theory (CS) suggested that there is gravitational wave (GW) amplitude birefringence,
+Some beyond-GR theories such as Chern-Simons gravity suggested that there is gravitational wave (GW) amplitude birefringence,
 while GR predicts there is no birefringence.
@@
 the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is $0$.
-This is a zeroth-order approximation of the birefringence model in CS, which assume there is no frequency dependence.
+This is a zeroth-order approximation of the birefringence model in Chern-Simons gravity, which assume there is no frequency dependence.
 This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.

### 2022-08-12 src/scripts/GW150914_corner.py

@@ -5,6 +5,8 @@ import seaborn as sns
 from bilby.gw.result import CBCResult
 import paths
 
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
 nsamples = 5000
 
 with h5py.File(paths.data/"IGWN-GWTC2p1-v2-GW150914_095045_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
@@ -30,7 +32,7 @@ result = pd.concat([result_bilby,result_extra,result_ligo], ignore_index=True)
 g = sns.pairplot(result,
                  vars=['kappa', 'luminosity_distance', 'cos_iota'],
                  corner=True, kind='kde', hue='with',
-                 diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=5))
+                 diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[(1.-0.954499736),(1.-0.682689492)]))
 
 g.axes[2,0].set_xlabel("$\kappa$")
 g.axes[1,0].set_ylabel("$d_L$")

### 2022-08-12 src/scripts/birefringence.py

@@ -1,54 +1,79 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import bilby
-from gwpy.timeseries import TimeSeries
+import lalsimulation as lalsim
+import lal
+import seaborn as sns
 import paths
 
-def smooth(y):
-    box_pts = 30
-    box = np.ones(box_pts)/box_pts
-    y_smooth = np.convolve(y, box, mode='same')
-    return y_smooth
+sns.set_theme(palette='colorblind', font_scale=1.2)
 
-time_of_event = 1264316116.4
-detectors={'L1', 'V1'}
-ifos = bilby.gw.detector.InterferometerList(detectors)
+approximant = lalsim.SimInspiralGetApproximantFromString("IMRPhenomXPHM")
 
-duration = 8.0
-post_trigger_duration = 2.
-analysis_start = time_of_event + post_trigger_duration - duration
+# frequency array parameters
+df = 0.25
+f_min = 20
+f_max = 3000
+f_ref = 100
 
-for detector in ifos:
-    analysis_data = TimeSeries.fetch_open_data(detector.name, analysis_start, analysis_start + duration, sample_rate=4096., cache=True)
-    detector.set_strain_data_from_gwpy_timeseries(analysis_data)
+# source parameters
+m1_msun = 10
+m2_msun = 10
+chi1 = [0, 0, 0.5]
+chi2 = [0, 0, 0.5]
+dist_mpc = 1000
+inclination = np.pi * 0.9
+phi_ref = 0
+
+m1_kg = m1_msun*lal.MSUN_SI
+m2_kg = m2_msun*lal.MSUN_SI
+distance = dist_mpc*1e6*lal.PC_SI
+
+hp, hc = lalsim.SimInspiralChooseFDWaveform(m1_kg, m2_kg,
+                                            chi1[0], chi1[1], chi1[2],
+                                            chi2[0], chi2[1], chi2[2],
+                                            distance, inclination,
+                                            phi_ref, 0, 0., 0.,
+                                            df, f_min, f_max, f_ref,
+                                            None, approximant)
+
[…]

### 2022-08-12 src/tex/ms.tex
Update ms.tex

@@
         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-    The three sets of plots are the PEs done by LIGO with GR \citep{GWTC-2.1, GWTC-3},
+    The three sets of plots are the PEs with GR done by LIGO without cosmological reweighing \citep{GWTC-2.1, GWTC-3},
     the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.

### 2022-08-26 src/tex/ms.tex
Update Content

--- 2022-08-11 src/tex/ms.tex
@@
 \documentclass[twocolumn]{aastex631}
+
 \usepackage{showyourwork}
@@
 \section{Introduction}
+\label{sec:Introduction}
 
+% Motivation
 Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
-After a century, we know that GR do not agree with quantum theories at some length scale.
+After a century, we know that GR does not agree with quantum theories at some length scale.
 To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories such as Chern-Simons theory (CS) suggested that there is gravitational wave (GW) amplitude birefringence,
+Some beyond-GR theories such as Chern-Simons gravity suggests that there is gravitational wave (GW) amplitude birefringence,
 while GR predicts there is no birefringence.
+GW amplitude birefringence is a property of space-time which consists of the enhancement of one GW polarization over the other during the propagation of the wave.
+the local modification of the amplitude is small, but it grows with the number of cycles the GW is propagated.
 
-\begin{figure}[h!]
+% Previous studies
+Previous studies have constrained amplitude birefringence by performing Bayesian inference.
+In \citet{Maria_2021}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second LIGO-Virgo catalog,
+and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
+In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption and
+performed Bayesian inference on the events in GWTC-1, the first LIGO-Virgo catalog.
+
+% What's new?
+In this paper, we used a birefringence model with higher order terms which include the frequency dependence of amplitude birefringence
+to perform a more accurate parameter estimation (PE).
+We also performed Bayesian inference on events in GWTC-3, the third LIGO-Virgo catalog.
+We included all binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$, which are listed in \citet{GWTC_3_population}.
+
+% Section guide
+In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, and mention the configuration we used in the PE.
+In Sec.~\ref{sec:Results}, we show the combined result of the PE, and present the constrain on the GW amplitude birefringence we obtained.
+In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study, and provide suggestions for future studies.
+
+\section{Method}
+\label{sec:Method}
+
+% Waveform modification
+GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves,
+which could be transformed into two circular polarizations (i.e. left-handed and right-handed) by
+\begin{equation}
+    h_{\mathrm{L/R}} = \frac{h_+ \pm i h_\times}{\sqrt{2}}\,,
+\end{equation}
+where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and right-handed polarization of the waveform,
+$h_+$ and $h_\times$ are the Fourier amplitude of the plus and cross polarization of the waveform.
+According to \textbf{[modification reference]}, we modified the waveform model by
+\begin{equation}
+    h_\mathrm{L/R}^{\mathrm{br}}=
+    h_\mathrm{L/R}^{\mathrm{GR}}\times
+    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
+\end{equation}
+where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified waveform with amplitude birefringence,
+$h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform at the source,
+$\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence,
+$d_C$ is the comoving distance to the source, and $f$ is the Fourier frequency of the waveform.
+We assume GWs are generated by the binary as GR predicts, as the modification at the source is small before the effect accumulate during the propagation.
+As the GWs propagate, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the GWs.
+
+% Effect on inclination PE
+This modification changes the ratio of the left-handed and right-handed polarization of the waveform,
+which affects the PE on the inclination as well.
+In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
+The relationship between the amplitude ratio and the inclination is
+\begin{equation}
+    \frac{h_\mathrm{L}}{h_\mathrm{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
+\end{equation}
+where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed and right-handed polarizations of the GWs
+and $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
+PE on $\iota$ takes the amplitude ratio into account. Therefore, the PE on $\iota$ depends on the amplitude ratio.
+
+With the modification we implemented, the amplitude ratio of the left-handed and right-handed polarizations becomes 
+\begin{equation}
+    \frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}=\frac{\left(1-\cos\iota\right)^2}{\left(1+\cos\iota\right)^2}\exp\left({2\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}}\right)\,,
+\end{equation}
+where $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed Fourier amplitude of left-handed and right-handed polarizations of the GWs.
+As the amplitude ratio does not only depends on $iota$ with this modification, but also $\kappa$ and $d_C$,
+the PE on $\iota$ and $d_C$ are affected. Note that GR is recovered if $\kappa$ is $0$.
+
+For some previous studies on the birefringence property such as \citet{Maria_2021}, they assume the effect of birefringence is independent of the frequency,
+which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.
+To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$,
+or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
+As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
+
+In contrast, We included the frequency dependence in the modification, which is a first-order approximation of the birefringence model.
+This frequency dependence provides stronger enhancement or suppression from the birefringence to high frequency components of the GW, and visa versa.
+Also, the frequency dependence can break the degeneracy between kappa and iota,
+as this would make the effect of the birefringence affect the amplitude ratio differently compared to iota.
+To illustrate the frequency dependence of amplitude birefringence,
+the observed amplitudes of two polarizations of a simulated GW signal are shown in figure \ref{fig:birefringence}.
+
+\begin{figure}[h]
     \script{birefringence.py}
-    \includegraphics[width=\columnwidth]{figures/birefringence.png}
+    \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
     \caption{
-        The observed amplitudes of two polarisations of an artificial GW signal under different model.
-        The top panel shows the case of GR, the amplitudes are unmodified.
-        For the middle and bottom panel, they shows the case of birefringence.
-        Left-handed polarisation is favored by the birefringence in the middle panel,
-        and right-handed polarisation is favored by the birefringence in the bottom panel.
+        The observed Fourier amplitude of two polarizations of an simulated GW signal with and without birefringence.
+        The amplitudes as GR predicts are shown in dotted lines with blue and orange colors representing the left-handed and right-handed polarizations.
+        The amplitudes with birefringence are shown in solid lines with blue and orange colors representing the left-handed and right-handed polarizations.
+        This shows the effect of amplitude birefringence with the frequency dependence.
     }
@@
 
-GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
-The longer the GW travels, the larger the effect of birefringence.
-To illustrate the effect of amplitude birefringence,
-the observed amplitudes of two polarisations of an artificial GW signal example are shown in figure \ref{fig:birefringence}.
-With birefringence, the amplitude is modified during the propagation,
-and the observed amplitudes will be different compare to the amplitudes when they are generated at the binary.
-
-Before we discuss more about GW amplitude birefringence, we need to understand the case of GR.
-GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves,
-which could be transformed into two circular polarizations (i.e. left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
-In GR, the amplitude ratio of the left and right polarisations only depends on the inclination $\iota$ for a nonprecessing binary,
-the angle between our line of sight and the orbital angular momentum of the binary.
-\begin{equation}
-    \left(\frac{h_\mathrm{L}}{h_\mathrm{R}}\right)_\mathrm{GR}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
-\end{equation}where $h_L$ and $h_R$ are the Fourier amplitude of left and right polarizations of the GWs respectively.
-
-For previous studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
-but also the distance to the binary and the strength of the birefringence.
-\begin{equation}
-    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
-\end{equation}where $d_C$ is the comoving distance to the binary, $\widetilde{\kappa}$ is the opacity parameter that represent
-the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are
-the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is $0$.
-This is a zeroth-order approximation of the birefringence model in CS, which assume there is no frequency dependence.
-This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.
-
-In this paper, we considered the frequency dependence of the birefringence as well.
-Thus, the amplitude ratio of the left and right polarisations will be as follow.
-\begin{equation}
-    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{new}=\frac{\exp\left({\kappa\frac{d_C}{1\mathrm{ Gpc}}\frac{f}{100\mathrm{ Hz}}}\right)\left(1-\cos\iota\right)^2}{\exp\left({-\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}}\right)\left(1+\cos\iota\right)^2}\,,
-\end{equation}where $f$ is the frequency of the GWs and $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence.
-This frequency dependence provides stronger enhancement or suppression from the birefringence to high frequency components of the GW, and visa versa.
-Also, the frequency dependence can break the degeneracy between kappa and iota,
-as this would make the effect of the birefringence affect the amplitude ratio differently compared to iota.
-
-\section{Method}
-
-To constrain $\kappa$, we perform parameter estimation (PE) with data from the third LIGO-Virgo catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby,
+% Parameter estimation
+To constrain $\kappa$, we perform PE with data from the third LIGO-Virgo catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby,
 a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data
-and priors for the parameters in question. \citep{Bilby}
-
-We assume GWs are generated at the binary as GR described. As the GWs propagate through space,
-the effect of the birefringence will be built up as the distance increases.
-To implement the amplitude birefringence on Bilby, waveforms are modified according to the following equation:
-\begin{equation}
-    h_\mathrm{L,R}^{\mathrm{Biref}}=
-    h_\mathrm{L,R}^{\mathrm{GR}}\times
-    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{ Gpc}}\frac{f}{100\mathrm{ Hz}}\right)\,.
-\end{equation}
-This modification allows Bilby to perform estimations on $\kappa$, as the waveforms will depends on $\kappa$ as well.
-
-After modifying Bilby, we set the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as using GR,
-to test if the PEs could recover the PE done by LIGO in \citet{GWTC-2.1, GWTC-3}. These control tests result can be found in the data release.
-We then set the prior on $\kappa$ to a uniform function between $-1$ and $1$, and perform PEs on the modified waveforms.
-In both case, we use the same configuration as in \citep{GWTC-2.1, GWTC-3} for each individual events.
+and priors for the parameters in question \citep{Bilby}. 
+We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$
+\footnote{For \textbf{[some events]}, a uniform distribution between $-0.5$ and $0.5$ was used instead. This does not affect the posterior distribution.},
+and perform PE using the modified waveforms.
+In both case, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each individual events, excluding the calibration of the detectors.
+We also perform test to check if the PE could recover the PE done by LIGO in \citet{GWTC-2.1, GWTC-3}.
+We set the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as using GR.
+These control tests result can be found in the data release.
 
 \section{Results}
+\label{sec:Results}
 
-\subsection{Result for GW150914}
+\subsection{Result on GWTC-3}
+% Combined plot
 
-\begin{figure}[h!]
+% How to calculate the combined posterior
+
+% Constraint on $\kappa$
+
+% Violin plot
+
+% Trend (mass & distance)
+
+% Constraining power
+
+% Case study
+
+% Case 1: GW150914
+\subsection{Result on GW150914}
+
+\begin{figure}[h]
     \script{GW150914_corner.py}
-    \includegraphics[width=\columnwidth]{figures/GW150914_corner.png}
+    \includegraphics[width=\columnwidth]{figures/GW150914_corner.pdf}
     \caption{
         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-    The three sets of plots are the PEs done by LIGO with GR \citep{GWTC-2.1, GWTC-3},
+    The three colors in the plot are the PE with GR done by LIGO without cosmological reweighing \citep{GWTC-2.1, GWTC-3},
     the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
     Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
-    which does not suggest GW amplitude birefringence.
+    which does not suggest GW amplitude birefringence. The plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$,
+    while the frequency dependent birefringence can break it.
     }
@@
 
-Consider GW150914, the first detected GW by LIGO. In figure \ref{fig:GW150914_corner}, the posteriors of the parameters of the GW150914 is shown.
+Consider GW150914, the first GW detected by LIGO. In figure \ref{fig:GW150914_corner}, we show the posteriors of the parameters of GW150914.
 With the frequency independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
@@
 
-\subsection{Result on GWTC-3}
+\section{Discussion}
+\label{sec:Discussion}
 
-We then perform PE on $71$ binary black hole merger events in GWTC-3 to further constrain $\kappa$.
-The posteriors for $\kappa$ of the events are shown in figure \textbf{[X]}, which shows $\kappa$ tend to be $0$.
-The combined posterior for $\kappa$ was calculated and shown in figure \textbf{[X]}, which allows us to constrain $\kappa$ within \textbf{[the number we get]}.
+% Reasons for special events
 
-\section{Discussion}
+% BNS
 
-\subsection{Limitation}
+% Higher order terms
 
-\subsection{Future Study}
+% More observations with higher SNR
 
 \section{Acknowledgements}
+\label{sec:Acknowledgements}
 
@@
 
-\section{Points}
-
-\begin{enumerate}
-    \item Introduction
-    \begin{enumerate}
-        \item Motivation
-        \begin{enumerate}
-            \item GR broke down at some scale, find a new theory
-            \item CS: beyond-GR theory
-            \item test CS property (GW amplitude birefringence)
-        \end{enumerate}
-        \item Amplitude Birefringence
-        \begin{enumerate}
-            \item GW amplitude birefringence is a property of space
-            \item birefringence: enhance one and suppress another
-            \item figure to illustrate the effect of birefringence
-        \end{enumerate}
-        \item Amplitude Ratio
-        \begin{enumerate}
-            \item GR: amplitude ratio depends on $\iota$ only
-            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
-        \end{enumerate}
-    \end{enumerate}
-    \item Method
-    \begin{enumerate}
-        \item Parameter Estimation
-        \begin{enumerate}
-            \item PE on GWTC-3 to constrain $\kappa$
-            \item Assume GW are generated as GR described
-            \item birefringence is considered during the propagation
-            \item waveform modification
-            \item GR control test
-            \item same configuration as LIGO
-        \end{enumerate}
-    \end{enumerate}
-    \item Results
-    \begin{enumerate}
-        \item GW150914
-        \begin{enumerate}
-            \item figure with frequency-dependent and frequency-independent
-            \item posterior from $\cos\iota$ looks different
-            \item frequency-dependent: degeneracy between $\iota$ and $\kappa$
-            \item frequency-independent: degeneracy broken and $\kappa$ near $0$
-        \end{enumerate}
-        \item GWTC-3
-        \begin{enumerate}
-            \item figure of $\kappa$ of all events
-            \item constraint on $\kappa$
-            \item improvement of constraint on $\kappa$ compare to previous studies
-        \end{enumerate}
-    \end{enumerate}
-    \item Discussion
-    \begin{enumerate}
-        \item Limitation
-        \begin{enumerate}
-            \item orientation of detector (polarisations)
-            \item sensitivity of detector
-        \end{enumerate}
-        \item Future Study
-        \begin{enumerate}
-            \item More events from detector
-            \item Higher order terms
-        \end{enumerate}
-    \end{enumerate}
-\end{enumerate}
-
 \end{document}

### 2022-08-26 src/scripts/GW150914_corner.py

@@ -4,8 +4,12 @@ import numpy as np
 import seaborn as sns
 from bilby.gw.result import CBCResult
 import paths
+from matplotlib import rc
 
 sns.set_theme(palette='colorblind', font_scale=1.2)
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('text', usetex=True)
 
 nsamples = 5000
 
@@ -14,17 +18,17 @@ with h5py.File(paths.data/"IGWN-GWTC2p1-v2-GW150914_095045_PEDataRelease_mixed_n
     result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
 result_ligo = result_ligo.sample(n=nsamples)
 result_ligo['kappa'] = np.zeros(len(result_ligo))
-result_ligo['with'] = np.full(len(result_ligo), "GR (LIGO)")
+result_ligo['with'] = np.full(len(result_ligo), r"GR (LIGO)")
 
 result_bilby = CBCResult.from_json(filename=paths.data/"GW150914_birefringence.json").posterior
 result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), "Non-GR (frequency dependent)")
+result_bilby['with'] = np.full(len(result_bilby), r"Non-GR (frequency dependent)")
 result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
 
 result_extra = CBCResult.from_json(filename=paths.data/"GW150914_birefringence(frequency_independent).json").posterior
 result_extra = result_extra.sample(n=nsamples)
 result_extra['kappa'] = result_extra['kappa'] * 1000 # different scale was used in the frequency independent data
-result_extra['with'] = np.full(len(result_extra), "Non-GR (frequency independent)")
+result_extra['with'] = np.full(len(result_extra), r"Non-GR (frequency independent)")
 result_extra['cos_iota'] = np.cos(result_extra['iota'])
 
 result = pd.concat([result_bilby,result_extra,result_ligo], ignore_index=True)
@@ -32,13 +36,13 @@ result = pd.concat([result_bilby,result_extra,result_ligo], ignore_index=True)
 g = sns.pairplot(result,
                  vars=['kappa', 'luminosity_distance', 'cos_iota'],
[…]

### 2022-08-26 src/scripts/birefringence.py

@@ -4,22 +4,26 @@ import lalsimulation as lalsim
 import lal
 import seaborn as sns
 import paths
+from matplotlib import rc
 
 sns.set_theme(palette='colorblind', font_scale=1.2)
 
-approximant = lalsim.SimInspiralGetApproximantFromString("IMRPhenomXPHM")
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('text', usetex=True)
+
+approximant = lalsim.SimInspiralGetApproximantFromString("IMRPhenomD")
 
 # frequency array parameters
 df = 0.25
 f_min = 20
-f_max = 3000
+f_max = 2030
 f_ref = 100
 
 # source parameters
 m1_msun = 10
 m2_msun = 10
-chi1 = [0, 0, 0.5]
-chi2 = [0, 0, 0.5]
+chi1 = [0, 0, 0]
+chi2 = [0, 0, 0]
 dist_mpc = 1000
 inclination = np.pi * 0.9
 phi_ref = 0
@@ -36,44 +40,24 @@ hp, hc = lalsim.SimInspiralChooseFDWaveform(m1_kg, m2_kg,
                                             df, f_min, f_max, f_ref,
                                             None, approximant)
 
-hl = (hp.data.data - (hc.data.data * 1j)) /np.sqrt(2)
-hr = (hp.data.data + (hc.data.data * 1j)) /np.sqrt(2)
+hl = (hp.data.data + (hc.data.data * 1j)) /np.sqrt(2)
+hr = (hp.data.data - (hc.data.data * 1j)) /np.sqrt(2)
 
 freq = np.arange(len(hl))*df
 
-kappa = 0.5
+kappa = -0.5
 
 hl_bi_L = hl * np.exp(kappa*(dist_mpc/1000)*(freq/100))
 hr_bi_L = hr * np.exp(-kappa*(dist_mpc/1000)*(freq/100))
 
-hl_bi_R = hl * np.exp(-kappa*(dist_mpc/1000)*(freq/100))
-hr_bi_R = hr * np.exp(kappa*(dist_mpc/1000)*(freq/100))
-
-plt.figure(figsize=(5,11))
-
-plt.subplot(311)
-plt.loglog(freq, abs(hl), label="L")
-plt.loglog(freq, abs(hr), ls='--', label="R")
-plt.xlim(f_min,f_max)
-plt.ylim(1e-40,1e-20)
[…]

### 2022-08-26 src/scripts/GW150914_corner.py
Testing matplotlib

@@ -5,6 +5,9 @@ import seaborn as sns
 from bilby.gw.result import CBCResult
 import paths
 from matplotlib import rc
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
 
 sns.set_theme(palette='colorblind', font_scale=1.2)
 

### 2022-08-26 src/scripts/birefringence.py

@@ -5,6 +5,9 @@ import lal
 import seaborn as sns
 import paths
 from matplotlib import rc
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
 
 sns.set_theme(palette='colorblind', font_scale=1.2)
 

### 2022-08-26 src/scripts/paths.py

@@ -3,8 +3,6 @@ Exposes common paths useful for manipulating datasets and generating figures.
 
 """
 from pathlib import Path
-import os
-os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
 
 # Absolute path to the top level of the repository
 root = Path(__file__).resolve().parents[2].absolute()

### 2022-09-01 src/tex/ms.tex
Testing matplotlib
fix plot font
Update ms.tex

@@
-\documentclass[twocolumn]{aastex631}
+% \documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,floatfix,nofootinbib]{revtex4-2}
+\documentclass[reprint,amsmath,amssymb,aps,twocolumn]{aastex631}
 
@@
 where $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed Fourier amplitude of left-handed and right-handed polarizations of the GWs.
-As the amplitude ratio does not only depends on $iota$ with this modification, but also $\kappa$ and $d_C$,
+As the amplitude ratio does not only depends on $\iota$ with this modification, but also $\kappa$ and $d_C$,
 the PE on $\iota$ and $d_C$ are affected. Note that GR is recovered if $\kappa$ is $0$.
@@
 which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
-This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.
+This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.
 To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$,
@@
 This frequency dependence provides stronger enhancement or suppression from the birefringence to high frequency components of the GW, and visa versa.
-Also, the frequency dependence can break the degeneracy between kappa and iota,
-as this would make the effect of the birefringence affect the amplitude ratio differently compared to iota.
+Also, the frequency dependence can break the degeneracy between $\kappa$ and $\iota$,
+as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
 To illustrate the frequency dependence of amplitude birefringence,

### 2022-11-19 src/tex/ms.tex
Update ms.tex
New content

@@
 
-\author{Thomas C.K. Ng}
+\author{Thomas C. K. Ng}
 \email{thomas.ng@link.cuhk.edu.hk}
-\affiliation{Department of Physics and Institute of Theoretical Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
+\affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
 
@@
 To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories such as Chern-Simons gravity suggests that there is gravitational wave (GW) amplitude birefringence,
-while GR predicts there is no birefringence.
+Some beyond-GR theories such as Chern-Simons gravity suggests that there is gravitational wave (GW) amplitude birefringence, while GR predicts there is no birefringence.
 GW amplitude birefringence is a property of space-time which consists of the enhancement of one GW polarization over the other during the propagation of the wave.
@@
 Previous studies have constrained amplitude birefringence by performing Bayesian inference.
-In \citet{Maria_2021}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second LIGO-Virgo catalog,
-and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
-In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption and
-performed Bayesian inference on the events in GWTC-1, the first LIGO-Virgo catalog.
+In \citet{Maria_2021}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
+In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption and performed Bayesian inference on the events in GWTC-1, the first GW transient catalog.
 
 % What's new?
-In this paper, we used a birefringence model with higher order terms which include the frequency dependence of amplitude birefringence
-to perform a more accurate parameter estimation (PE).
-We also performed Bayesian inference on events in GWTC-3, the third LIGO-Virgo catalog.
-We included all binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$, which are listed in \citet{GWTC_3_population}.
+In this paper, we used a birefringence model with higher order terms which include the frequency dependence of amplitude birefringence to perform a more accurate parameter estimation (PE).
+We also performed Bayesian inference on events in GWTC-3, the third GW transient catalog.
+We included 70 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$, which are listed in \citet{GWTC_3_population}.\footnote{GW190720 is included in list, however, we could not successfully perform PE with it.
+Details on GW190720 will be mention in Discussion.}
 
@@
 % Waveform modification
-GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves,
-which could be transformed into two circular polarizations (i.e. left-handed and right-handed) by
+GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves, which could be transformed into two circular polarizations (i.e. left-handed and right-handed) by
 \begin{equation}
@@
 \end{equation}
-where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and right-handed polarization of the waveform,
-$h_+$ and $h_\times$ are the Fourier amplitude of the plus and cross polarization of the waveform.
+where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of the plus and cross polarization of the waveform.
 According to \textbf{[modification reference]}, we modified the waveform model by
@@
 \end{equation}
-where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified waveform with amplitude birefringence,
-$h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform at the source,
-$\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence,
-$d_C$ is the comoving distance to the source, and $f$ is the Fourier frequency of the waveform.
+where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform at the source, $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the Fourier frequency of the waveform.
 We assume GWs are generated by the binary as GR predicts, as the modification at the source is small before the effect accumulate during the propagation.
@@
 % Effect on inclination PE
-This modification changes the ratio of the left-handed and right-handed polarization of the waveform,
-which affects the PE on the inclination as well.
+This modification changes the ratio of the left-handed and right-handed polarization of the waveform, which affects the PE on the inclination as well.
 In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
@@
 \end{equation}
-where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed and right-handed polarizations of the GWs
-and $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
+where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed and right-handed polarizations of the GWs and $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
 PE on $\iota$ takes the amplitude ratio into account. Therefore, the PE on $\iota$ depends on the amplitude ratio.
@@
 where $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed Fourier amplitude of left-handed and right-handed polarizations of the GWs.
-As the amplitude ratio does not only depends on $\iota$ with this modification, but also $\kappa$ and $d_C$,
-the PE on $\iota$ and $d_C$ are affected. Note that GR is recovered if $\kappa$ is $0$.
+As the amplitude ratio does not only depends on $\iota$ with this modification, but also $\kappa$ and $d_C$, the PE on $\iota$ and $d_C$ are affected. Note that GR is recovered if $\kappa$ is $0$.
 
-For some previous studies on the birefringence property such as \citet{Maria_2021}, they assume the effect of birefringence is independent of the frequency,
-which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+For some previous studies on the birefringence property such as \citet{Maria_2021}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
 This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.
-To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$,
-or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
+To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
 As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
@@
 This frequency dependence provides stronger enhancement or suppression from the birefringence to high frequency components of the GW, and visa versa.
-Also, the frequency dependence can break the degeneracy between $\kappa$ and $\iota$,
-as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
-To illustrate the frequency dependence of amplitude birefringence,
-the observed amplitudes of two polarizations of a simulated GW signal are shown in figure \ref{fig:birefringence}.
+Also, the frequency dependence can break the degeneracy between $\kappa$ and $\iota$, as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
+To illustrate the frequency dependence of amplitude birefringence, the observed amplitudes of two polarizations of a simulated GW signal are shown in figure \ref{fig:birefringence}.
 
@@
 % Parameter estimation
-To constrain $\kappa$, we perform PE with data from the third LIGO-Virgo catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby,
-a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data
-and priors for the parameters in question \citep{Bilby}. 
-We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$
-\footnote{For \textbf{[some events]}, a uniform distribution between $-0.5$ and $0.5$ was used instead. This does not affect the posterior distribution.},
-and perform PE using the modified waveforms.
-In both case, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each individual events, excluding the calibration of the detectors.
-We also perform test to check if the PE could recover the PE done by LIGO in \citet{GWTC-2.1, GWTC-3}.
+The major objective of this study is to constrain GW amplitude birefringence, which is represented by the opacity parameter $\kappa$.
+To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question \citep{Bilby}.
+We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using the modified waveforms.
+In both case, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each individual events.
+We also perform test to check if the PE could recover the PE done by LIGO-Virgo-KAGRA Collaboration (LVK) in \citet{GWTC-2.1, GWTC-3}.
 We set the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as using GR.
@@
 \subsection{Result on GWTC-3}
-% Combined plot
+% Stacked kappa posterior plot
+In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$ of each events.
+
+\begin{figure}[h]
+    \script{kappa_stacked.py}
+    % \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
+    \caption{
+        The posterior of $kappa$ for all 70 events included in this study.
+        Each color represent a different event.
+        These posterior are used to calculate the constraint on $\kappa$ in the following.
+    }
+    \label{fig:kappa_stacked}
+\end{figure}
 
 % How to calculate the combined posterior
+
+We assume $\kappa$ is different for each event, and the population distribution of $\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
+We then performed hierarchical modeling to construct the likelihood function of $\mu$ and $\sigma$ given interferometer data.
+To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}, with prior of $\mu$ and $\sigma$ being uniform distributions.
+The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
+
+% Corner plot of mean and standard deviation of kappa
+
+\begin{figure}[h]
+    \script{corner_Gaussian.py}
+    % \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
+    \caption{
+        The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
+    }
+    \label{fig:corner_Gaussian}
+\end{figure}
 
@@
 % Violin plot
+
+\begin{figure}[h]
+    \script{violin_kappa.py}
+    % \includegraphics[width=\columnwidth]{figures/violin_kappa.pdf}
+    \caption{
+        The violin plot shows the posterior of $kappa$ for all 70 events included in this study.
+        Each violin represents a different event.
+        The violins are sorted by the quotient of median and standard deviation of the posterior.
+    }
+    \label{fig:violin_kappa}
+\end{figure}
 
@@
 \begin{figure}[h]
-    \script{GW150914_corner.py}
-    \includegraphics[width=\columnwidth]{figures/GW150914_corner.pdf}
+    \script{corner_GW150914.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
     \caption{
         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-    The three colors in the plot are the PE with GR done by LIGO without cosmological reweighing \citep{GWTC-2.1, GWTC-3},
-    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
-    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
-    which does not suggest GW amplitude birefringence. The plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$,
-    while the frequency dependent birefringence can break it.
+        The three colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
+        Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
+        The plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
     }
-    \label{fig:GW150914_corner}
+    \label{fig:corner_GW150914}
 \end{figure}
 
-Consider GW150914, the first GW detected by LIGO. In figure \ref{fig:GW150914_corner}, we show the posteriors of the parameters of GW150914.
+Consider GW150914, the first GW detected by LIGO.
+In figure \ref{fig:GW150914_corner}, we show the posteriors of the parameters of GW150914.
 With the frequency independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
 This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
-To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$,
-or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
-As a result,many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
+To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
+As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
 
 On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
-This is because the degeneracy was broken by the frequency dependence, as the effect of the birefringence will be different from the one of changing $\iota$.
+This is because the degeneracy was broken by the frequency dependence, as the effect of birefringence will be different from the one of changing $\iota$.
 In this case, the posteriors can recover the GR distributions for the binary parameters,
 and the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent, and GR can be recovered.
+
+% Case 2: GW190521 (massive BBH)
+
+\begin{figure}[h]
+    \script{corner_GW190521.py}
+    % \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
+    \caption{
+        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
+        The colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
+        The plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
+    }
+    \label{fig:corner_GW190521}
+\end{figure}
+
+Consider GW190521, the most massive binary black hole merger in the events we included.
+The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency dependent birefringence model, as shown in figure \ref{fig:corner_GW190521}.
+This is because the frequency range of the signal is much smaller due to the large mass of the binary.
+The effect of birefringence at different frequency within the range is similar, thus the degeneracy cannot be broken.
 
@@
 
+% Special event 1: GW190720
+
+% Special event 2: GW200129
+
+% Future work
+
 % BNS
-
-% Higher order terms
+A future work is to perform PE on binary neutron star mergers, such as GW170817.
+The frequency range of the signal is much larger compared to the binary black hole mergers, thus the difference of the effect of birefringence at different frequency within the range can be more significant.
+This can allow us to further constrain the birefringence effect, and the beyond-GR theories that predict it.
+However, performing PE on binary neutron star mergers requires much more computational resources.
+Therefore, we may need to wait for future PE methods and tools to further our work.
 
 % More observations with higher SNR
+Another future work is to apply the same method to more events with higher signal-to-noise ratio (SNR).
+LVK will release more events with higher SNR in the future.
 

### 2022-11-19 src/scripts/corner_GW150914.py

@@ -48,4 +48,4 @@ g.axes[2,0].set_ylabel(r"$\cos\iota$")
 g.axes[2,2].set_xlabel(r"$\cos\iota$")
 g.fig.legends[0].set_bbox_to_anchor((0.65,0.8))
 
-g.savefig(fname=paths.figures/"GW150914_corner.pdf", bbox_inches="tight", dpi=300)
+g.savefig(fname=paths.figures/"corner_GW150914.pdf", bbox_inches="tight", dpi=300)

### 2022-11-19 src/tex/ms.tex
Update figures

@@
 We also performed Bayesian inference on events in GWTC-3, the third GW transient catalog.
-We included 70 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$, which are listed in \citet{GWTC_3_population}.\footnote{GW190720 is included in list, however, we could not successfully perform PE with it.
+We included 70 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$, which are listed in \citet{GWTC_3_population}.\footnote{GW190720 is included in the list in \citet{GWTC_3_population}, however, we could not successfully perform PE with it.
 Details on GW190720 will be mention in Discussion.}
@@
     \script{kappa_stacked.py}
-    % \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
+    \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
     \caption{
-        The posterior of $kappa$ for all 70 events included in this study.
+        The posterior of $\kappa$ for all 70 events included in this study.
         Each color represent a different event.
@@
     \script{corner_Gaussian.py}
-    % \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
+    \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
     \caption{
@@
 
-\begin{figure}[h]
+\begin{figure*}[ht]
     \script{violin_kappa.py}
-    % \includegraphics[width=\columnwidth]{figures/violin_kappa.pdf}
+    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
     \caption{
-        The violin plot shows the posterior of $kappa$ for all 70 events included in this study.
+        The violin plot shows the posterior of $\kappa$ for all 70 events included in this study.
         Each violin represents a different event.
@@
     \label{fig:violin_kappa}
-\end{figure}
+\end{figure*}
 
@@
     \script{corner_GW190521.py}
-    % \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
+    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
     \caption{

### 2022-11-19 src/scripts/corner_GW190521.py

@@ -1,0 +1,45 @@
+import h5py
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"],
+})
+
+nsamples = 5000
+
+with h5py.File(paths.data/"IGWN-GWTC2p1-v2-GW190521_030229_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
+    f = f['C01:IMRPhenomXPHM']
+    result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
+result_ligo = result_ligo.sample(n=nsamples)
+result_ligo['kappa'] = np.zeros(len(result_ligo))
+result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
+
+result_bilby = pd.read_feather(paths.data/"samples_posterior_nonGR.feather")
+result_bilby = result_bilby[result_bilby.event == "GW190521"]
+result_bilby = result_bilby.sample(n=nsamples)
+result_bilby['with'] = np.full(len(result_bilby), r"Non-GR (frequency dependent)")
+result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+
+result = pd.concat([result_bilby,result_ligo], ignore_index=True)
+
+g = sns.pairplot(result,
+                 vars=['kappa', 'luminosity_distance', 'cos_iota'],
+                 corner=True, kind='kde', hue='with',
+                 diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[(1.-0.90),(1.-0.3935)]))
+
+g.axes[2,0].set_xlabel(r"$\kappa$")
+g.axes[1,0].set_ylabel(r"$d_L$")
+g.axes[2,1].set_xlabel(r"$d_L$")
+g.axes[2,0].set_ylabel(r"$\cos\iota$")
+g.axes[2,2].set_xlabel(r"$\cos\iota$")
+g.fig.legends[0].set_bbox_to_anchor((0.65,0.8))
+
+g.savefig(fname=paths.figures/"corner_GW190521.pdf", bbox_inches="tight", dpi=300)

### 2022-11-19 src/scripts/violin_kappa.py

@@ -1,0 +1,43 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"]
+})
+
+result_dict = {}
+median_std_dict = {}
+result_DataFrame = pd.read_feather(paths.data/"samples_posterior_nonGR.feather")
+for event in result_DataFrame['event'].unique():
+    result_dict[event] = result_DataFrame[result_DataFrame.event == event]
+    median_std_dict[event] = abs(result_dict[event]['kappa'].median()/result_dict[event]['kappa'].std())
+median_std_dict = {k: v for k, v in sorted(median_std_dict.items(), key=lambda item: item[1])}
+
+samples_Gaussian = np.load(paths.data/"samples_Gaussian.npz")["chains"]
+samples_Gaussian = samples_Gaussian.reshape(-1, 2)
+mean_median = np.median(samples_Gaussian[:,0])
+sd_median = np.median(samples_Gaussian[:,1])
+
+g = sns.violinplot(data=result_DataFrame, x="event", y="kappa", scale="width", inner=None, 
+                    order=list(median_std_dict.keys())
+                   )
+
+g.set_ylabel("$\kappa$", fontsize=20)
+g.set_ylim(-0.2,0.2)
+g.set_xlabel("")
+g.axhline(0,color=sns.color_palette()[3])
+x_low, x_high = g.get_xlim()
+g.fill_between([x_low-1,x_high+1], mean_median-sd_median, mean_median+sd_median, alpha=0.2)
+g.set_xlim(x_low-0.5,x_high+0.5)
+g.axhline(mean_median,color=sns.color_palette()[0])
+g.figure.autofmt_xdate(rotation=45)
+g.figure.set_size_inches(30,10)
+
+g.figure.savefig(fname=paths.figures/"violin_kappa.pdf", bbox_inches="tight", dpi=300)

### 2022-11-20 src/tex/ms.tex
Add content

@@
 \begin{abstract}
-
+    One of the major problems that physicists are facing nowadays is Einstein's theory of general relativity does not agree with quantum theories at some length scales.
+    In recent decades, theorist has been working on many beyond-GR theories to unify both of them.
+    Some of these theories such as Chern-Simons gravity suggest that there is gravitational wave amplitude birefringence.
+    In our study, we perform parameter estimation to constrain the strength of GW amplitude birefringence.
+    Compare to similar previous studies on the same topic, we perform PE on more events, including events in the third LIGO-Virgo catalog, and we used a more accurate model of birefringence to describe the phenomenon.
+    % results
 \end{abstract}
@@
 \label{sec:Introduction}
-
 % Motivation
@@
 \label{sec:Method}
-
 % Waveform modification
@@
 \label{sec:Results}
-
 \subsection{Result on GWTC-3}
 % Stacked kappa posterior plot
-In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$ of each events.
-
 \begin{figure}[h]
@@
         Each color represent a different event.
-        These posterior are used to calculate the constraint on $\kappa$ in the following.
+        These posterior are used to calculate the population constraint on $\kappa$ in the following.
     }
@@
 % How to calculate the combined posterior
-
+In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$ of each events obtained from PE.
 We assume $\kappa$ is different for each event, and the population distribution of $\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
@@
 % Corner plot of mean and standard deviation of kappa
-
 \begin{figure}[h]
@@
         The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
+        2D contour plot shows 1 $\sigma$ and 90\% confidence level.
+        Orange line represent the median value of the posterior.
+        The median of $\mu$ and $\sigma$ are \variable{output/mu_median.txt} and \variable{output/sigma_median.txt}, respectively.
     }
@@
 % Violin plot
-
+In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each events in the form of violin plot.
+There are some events with $\kappa$ distribution deviate from $\kappa=0$.
+Also, there are some events with bimodal $\kappa$ distribution.
+We are going to discuss some of these events in the following.
 \begin{figure*}[ht]
@@
         The violins are sorted by the quotient of median and standard deviation of the posterior.
+        Blue horizontal line represent the median value of $\mu$, and blue region represent 1 $\sigma$ confidence level.
+        Orange horizontal line represent $\kappa=0$.
+        This plot shows that there are some events with $\kappa$ distribution significantly deviate from $0$.
     }
@@
 % Case study
+% Case: GW170818
+\subsection{Result on GW170818}
 
-% Case 1: GW150914
+% Case: GW150914 (example of broken degeneracy)
 \subsection{Result on GW150914}
@@
         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-        The three colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
+        Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
+        2D contour plot shows 1 $\sigma$ and 90\% confidence level.
         Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
-        The plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
+        This plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
     }
@@
 
+Some events have bimodal $\kappa$ distribution, but before we show a bimodal event, we should understand the mechanism of a unimodal $\kappa$ distribution.
 Consider GW150914, the first GW detected by LIGO.
-In figure \ref{fig:GW150914_corner}, we show the posteriors of the parameters of GW150914.
+In figure \ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
 With the frequency independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
@@
 
-% Case 2: GW190521 (massive BBH)
-
+% Case: GW190521 (massive BBH)
+\subsection{Result on GW190521}
 \begin{figure}[h]
@@
         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
-        The colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
-        The plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
+        Colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
+        2D contour plot shows 1 $\sigma$ and 90\% confidence level.
+        This plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
     }
@@
 
+Now, we show a bimodal $\kappa$ distribution.
 Consider GW190521, the most massive binary black hole merger in the events we included.
@@
 The effect of birefringence at different frequency within the range is similar, thus the degeneracy cannot be broken.
+This can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
 
@@
 % Reasons for special events
-
 % Special event 1: GW190720
@@
 % Future work
-
 % BNS
@@
 LVK will release more events with higher SNR in the future.
+Data with higher SNR allows us to obtain more precise PE results, which also allows us to constrain the birefringence effect more precisely.
+And more events will allow us to calculate a more constrained population posterior of $\kappa$.
 

### 2022-11-20 src/scripts/corner_Gaussian.py

@@ -14,6 +14,13 @@ plt.rcParams.update({
 
 samples_Gaussian = np.load(paths.data/"samples_Gaussian.npz")["chains"]
 samples_Gaussian = samples_Gaussian.reshape(-1, 2)
+mu_median = np.median(samples_Gaussian[:,0])
+sigma_median = np.median(samples_Gaussian[:,1])
+
+with open(paths.output/"mu_median.txt", "w") as f:
+    f.write(f"${mu_median:.3f}\\pm{np.std(samples_Gaussian[:,1]):.3f}$")
+with open(paths.output/"sigma_median.txt", "w") as f:
+    f.write(f"${sigma_median:.3f}\\pm{np.std(samples_Gaussian[:,1]):.3f}$")
 
 df = pd.DataFrame()
 df['mu'] = samples_Gaussian.reshape(-1,2)[:,0]
@@ -25,6 +32,10 @@ g = sns.pairplot(df,
                 corner=True, kind='kde',
                 diag_kws=dict(common_norm=False, cut=0), plot_kws=dict(common_norm=False, levels=[
                         (1.-0.90),(1.-0.3935)], cut=0))
+g.axes[0,0].axvline(mu_median, color=sns.color_palette()[3])
+g.axes[1,1].axvline(sigma_median, color=sns.color_palette()[3])
+g.axes[1,0].axvline(mu_median, color=sns.color_palette()[3])
+g.axes[1,0].axhline(sigma_median, color=sns.color_palette()[3])
 
 g.axes[1,1].set_xlabel("$\\sigma$")
 g.axes[1,0].set_xlabel("$\\mu$")

### 2022-11-20 src/tex/ms.tex
Update ms.tex

@@
 \begin{abstract}
-    One of the major problems that physicists are facing nowadays is Einstein's theory of general relativity does not agree with quantum theories at some length scales.
+    One of the major problems that physicists are facing nowadays is Einstein's theory of general relativity (GR) does not agree with quantum theories at some length scales.
     In recent decades, theorist has been working on many beyond-GR theories to unify both of them.
-    Some of these theories such as Chern-Simons gravity suggest that there is gravitational wave amplitude birefringence.
-    In our study, we perform parameter estimation to constrain the strength of GW amplitude birefringence.
-    Compare to similar previous studies on the same topic, we perform PE on more events, including events in the third LIGO-Virgo catalog, and we used a more accurate model of birefringence to describe the phenomenon.
+    Some of these theories such as Chern-Simons gravity suggest that there is gravitational wave (GW) amplitude birefringence.
+    In our study, we perform parameter estimation (PE) to constrain the strength of GW amplitude birefringence.
+    Compare to similar previous studies on the same topic, we perform PE on more events, including events in the third LIGO-Virgo catalog (GWTC-3), and we used a more accurate model of birefringence to describe the phenomenon.
     % results

### 2022-11-22 src/tex/ms.tex
Remove glitched event

@@
 We also performed Bayesian inference on events in GWTC-3, the third GW transient catalog.
-We included 70 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$, which are listed in \citet{GWTC_3_population}.\footnote{GW190720 is included in the list in \citet{GWTC_3_population}, however, we could not successfully perform PE with it.
-Details on GW190720 will be mention in Discussion.}
+We included 69 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$.
+These events are also listed in TABLE I in \citet{GWTC_3_population}.\footnote{GW190720 and GW200129 are included in the list in \citet{GWTC_3_population}, but we did not include them in this study.
+Details in Discussion.}
 
@@
     \caption{
-        The posterior of $\kappa$ for all 70 events included in this study.
+        The posterior of $\kappa$ for all 69 events included in this study.
         Each color represent a different event.
@@
         The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
-        2D contour plot shows 1 $\sigma$ and 90\% confidence level.
+        2D contour plot shows 39.35\% and 90\% confidence level.
         Orange line represent the median value of the posterior.
-        The median of $\mu$ and $\sigma$ are \variable{output/mu_median.txt} and \variable{output/sigma_median.txt}, respectively.
+        The mean of $\mu$ and $\sigma$ are \variable{output/mu_mean.txt} and \variable{output/sigma_mean.txt}, respectively.
     }
@@
 % Constraint on $\kappa$
+
+% Comparison with previous studies
 
@@
     \caption{
-        The violin plot shows the posterior of $\kappa$ for all 70 events included in this study.
+        The violin plot shows the posterior of $\kappa$ for all 69 events included in this study.
         Each violin represents a different event.
@@
 % Case: GW170818
-\subsection{Result on GW170818}
 
@@
         Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
-        2D contour plot shows 1 $\sigma$ and 90\% confidence level.
+        2D contour plot shows 39.35\% and 90\% confidence level.
         Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
@@
         Colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
-        2D contour plot shows 1 $\sigma$ and 90\% confidence level.
+        2D contour plot shows 39.35\% and 90\% confidence level.
         This plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
@@
 % Reasons for special events
+\subsection{Special Events}
 % Special event 1: GW190720
+For GW190720, we could not perform PE with Bilby successfully.
+In LVK PE papers \citep{GWTC-2.1, GWTC-3}, the PE of some events were performed with LALInference \citep{lalsuite} instead of Bilby, GW190720 was one of them.
+We successfully migrate LALInference configuration to Bilby for most of these events and recover similar PE results, except for GW190720.
+Bilby could not evaluate the likelihood function on Virgo data.
+This could bias the population posterior.
 
 % Special event 2: GW200129
+For GW200129, past research suggest there could be a potential glitch in Virgo data. \citep{GW200129}
+We perform 3 sets of PE to check if the glitch would affect the population posterior of $\kappa$.
+In each set, we pick only 2 out of 3 detectors data to perform PE.
+The 2 sets of PE results with Virgo data return a $\kappa$ posterior which strongly suggest a negative $\kappa$.
+We removed this event from our analysis.
+Further investigation is needed to understand the glitch.
 
 % Future work
+\subsection{Future Work}
 % BNS

### 2022-11-22 src/scripts/corner_Gaussian.py

@@ -12,15 +12,15 @@ plt.rcParams.update({
     "font.serif": ["Computer Modern Roman"],
 })
 
-samples_Gaussian = np.load(paths.data/"samples_Gaussian.npz")["chains"]
+samples_Gaussian = np.load(paths.data/"samples_Gaussian_without_GW200129.npz")["chains"]
 samples_Gaussian = samples_Gaussian.reshape(-1, 2)
 mu_median = np.median(samples_Gaussian[:,0])
 sigma_median = np.median(samples_Gaussian[:,1])
 
-with open(paths.output/"mu_median.txt", "w") as f:
-    f.write(f"${mu_median:.3f}\\pm{np.std(samples_Gaussian[:,1]):.3f}$")
-with open(paths.output/"sigma_median.txt", "w") as f:
-    f.write(f"${sigma_median:.3f}\\pm{np.std(samples_Gaussian[:,1]):.3f}$")
+with open(paths.output/"mu_mean.txt", "w") as f:
+    f.write(f"${np.mean(samples_Gaussian[:,0]):.3f}\\pm{np.std(samples_Gaussian[:,0]):.3f}$")
+with open(paths.output/"sigma_mean.txt", "w") as f:
+    f.write(f"${np.mean(samples_Gaussian[:,1]):.3f}\\pm{np.std(samples_Gaussian[:,1]):.3f}$")
 
 df = pd.DataFrame()
 df['mu'] = samples_Gaussian.reshape(-1,2)[:,0]

### 2022-11-22 src/scripts/violin_kappa.py

@@ -15,28 +15,29 @@ plt.rcParams.update({
 result_dict = {}
 median_std_dict = {}
 result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_DataFrame = result_DataFrame[result_DataFrame.event != "GW200129"] # remove GW200129
 for event in result_DataFrame['event'].unique():
     result_dict[event] = result_DataFrame[result_DataFrame.event == event]
     median_std_dict[event] = abs(result_dict[event]['kappa'].median()/result_dict[event]['kappa'].std())
 median_std_dict = {k: v for k, v in sorted(median_std_dict.items(), key=lambda item: item[1])}
 
-samples_Gaussian = np.load(paths.data/"samples_Gaussian.npz")["chains"]
+samples_Gaussian = np.load(paths.data/"samples_Gaussian_without_GW200129.npz")["chains"]
 samples_Gaussian = samples_Gaussian.reshape(-1, 2)
-mean_median = np.median(samples_Gaussian[:,0])
-sd_median = np.median(samples_Gaussian[:,1])
+mu_median = np.median(samples_Gaussian[:,0])
+sigma_median = np.median(samples_Gaussian[:,1])
 
 g = sns.violinplot(data=result_DataFrame, x="event", y="kappa", scale="width", inner=None, 
-                    order=list(median_std_dict.keys())
+                    order=list(median_std_dict.keys()), palette="magma"
                    )
 
 g.set_ylabel("$\kappa$", fontsize=20)
 g.set_ylim(-0.2,0.2)
 g.set_xlabel("")
-g.axhline(0,color=sns.color_palette()[3])
+g.axhline(0, color=sns.color_palette()[3])
 x_low, x_high = g.get_xlim()
-g.fill_between([x_low-1,x_high+1], mean_median-sd_median, mean_median+sd_median, alpha=0.2)
+g.fill_between([x_low-1,x_high+1], mu_median-sigma_median, mu_median+sigma_median, alpha=0.3, color=sns.color_palette()[9])
 g.set_xlim(x_low-0.5,x_high+0.5)
-g.axhline(mean_median,color=sns.color_palette()[0])
+g.axhline(mu_median,color=sns.color_palette()[9])
 g.figure.autofmt_xdate(rotation=45)
 g.figure.set_size_inches(30,10)
 

### 2022-11-22 src/tex/ms.tex
New content

@@
         The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
-        2D contour plot shows 39.35\% and 90\% confidence level.
-        Orange line represent the median value of the posterior.
+        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        Orange line represent the median of $\mu$ and $\sigma$, which are \variable{output/mu_median.txt} and \variable{output/sigma_median.txt} respectively,.
         The mean of $\mu$ and $\sigma$ are \variable{output/mu_mean.txt} and \variable{output/sigma_mean.txt}, respectively.
@@
 % Constraint on $\kappa$
+We then calculate the constraint on $\kappa$ from the posterior of $\mu$ and $\sigma$.
+% constraint
 
 % Comparison with previous studies
+In \citet{Maria_2021}, the authors gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
+We convert the constraint to the same format as ours to make comparison.
+They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
 
@@
         Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
-        2D contour plot shows 39.35\% and 90\% confidence level.
+        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
         Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
@@
         Colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
-        2D contour plot shows 39.35\% and 90\% confidence level.
+        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
         This plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.

### 2022-11-22 src/scripts/Gaussian_stat.py

@@ -1,0 +1,14 @@
+import numpy as np
+import paths
+
+samples_Gaussian = np.load(paths.data/"samples_Gaussian_without_GW200129.npz")["chains"]
+samples_Gaussian = samples_Gaussian.reshape(-1, 2)
+
+with open(paths.output/"mu_mean.txt", "w") as f:
+    f.write(f"${np.mean(samples_Gaussian[:,0]):.3f}\\pm{np.std(samples_Gaussian[:,0]):.3f}$")
+with open(paths.output/"sigma_mean.txt", "w") as f:
+    f.write(f"${np.mean(samples_Gaussian[:,1]):.3f}\\pm{np.std(samples_Gaussian[:,1]):.3f}$")
+with open(paths.output/"mu_median.txt", "w") as f:
+    f.write(f"${np.median(samples_Gaussian[:,0]):.3f}$")
+with open(paths.output/"sigma_median.txt", "w") as f:
+    f.write(f"${np.median(samples_Gaussian[:,1]):.3f}$")

### 2022-11-22 src/scripts/corner_Gaussian.py

@@ -17,11 +17,6 @@ samples_Gaussian = samples_Gaussian.reshape(-1, 2)
 mu_median = np.median(samples_Gaussian[:,0])
 sigma_median = np.median(samples_Gaussian[:,1])
 
-with open(paths.output/"mu_mean.txt", "w") as f:
-    f.write(f"${np.mean(samples_Gaussian[:,0]):.3f}\\pm{np.std(samples_Gaussian[:,0]):.3f}$")
-with open(paths.output/"sigma_mean.txt", "w") as f:
-    f.write(f"${np.mean(samples_Gaussian[:,1]):.3f}\\pm{np.std(samples_Gaussian[:,1]):.3f}$")
-
 df = pd.DataFrame()
 df['mu'] = samples_Gaussian.reshape(-1,2)[:,0]
 df['sigma'] = samples_Gaussian.reshape(-1,2)[:,1]

### 2022-11-22 src/tex/ms.tex
New content

@@
     Compare to similar previous studies on the same topic, we perform PE on more events, including events in the third LIGO-Virgo catalog (GWTC-3), and we used a more accurate model of birefringence to describe the phenomenon.
-    % results
+    We were able to obtain a constrain an order of magnitude tighter than previous studies which used GWTC-2.
 \end{abstract}
@@
 We then calculate the constraint on $\kappa$ from the posterior of $\mu$ and $\sigma$.
-% constraint
+We generated a Gaussian distribution for each sample of $\mu$ and $\sigma$.
+The Gaussian distributions were then added up to obtain the population distribution of $\kappa$.
+The mean of the population distribution of $\kappa$ is \variable{output/kappa_mean.txt}.
+Note that GR is recovered if $\kappa$ is $0$.
 
@@
 They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
+We were able to obtain a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
+This is a order of magnitude improvement in constraining $\kappa$.
+This is due to the fact that we have much more events from GWTC-3 compared to GWTC-2.
 
@@
 We are going to discuss some of these events in the following.
+
 \begin{figure*}[ht]
@@
 \subsection{Result on GW150914}
-
-\begin{figure}[h]
-    \script{corner_GW150914.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
-    \caption{
-        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-        Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
-        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
-        This plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
-    }
-    \label{fig:corner_GW150914}
-\end{figure}
-
 Some events have bimodal $\kappa$ distribution, but before we show a bimodal event, we should understand the mechanism of a unimodal $\kappa$ distribution.
@@
 
+\begin{figure}[h]
+    \script{corner_GW150914.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
+    \caption{
+        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
+        Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with both frequency independent and dependent birefringence respectively.
+        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
+        This plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
+    }
+    \label{fig:corner_GW150914}
+\end{figure}
+
 % Case: GW190521 (massive BBH)
 \subsection{Result on GW190521}
+
+Now, we show a bimodal $\kappa$ distribution.
+Consider GW190521, the most massive binary black hole merger in the events we included.
+The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency dependent birefringence model, as shown in figure \ref{fig:corner_GW190521}.
+This is because the frequency range of the signal is much smaller due to the large mass of the binary.
+The effect of birefringence at different frequency within the range is similar, thus the degeneracy cannot be broken.
+This can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
+
 \begin{figure}[h]
@@
 \end{figure}
-
-Now, we show a bimodal $\kappa$ distribution.
-Consider GW190521, the most massive binary black hole merger in the events we included.
-The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency dependent birefringence model, as shown in figure \ref{fig:corner_GW190521}.
-This is because the frequency range of the signal is much smaller due to the large mass of the binary.
-The effect of birefringence at different frequency within the range is similar, thus the degeneracy cannot be broken.
-This can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
 

### 2022-11-22 src/scripts/Gaussian_stat.py

@@ -1,14 +1,32 @@
 import numpy as np
+import pandas as pd
 import paths
 
+def normal_distribution(x, mu, sigma):
+    return (1/(sigma*np.sqrt(2*np.pi)))*np.exp(-0.5*((x-mu)/sigma)**2)
+
 samples_Gaussian = np.load(paths.data/"samples_Gaussian_without_GW200129.npz")["chains"]
-samples_Gaussian = samples_Gaussian.reshape(-1, 2)
+df = pd.DataFrame()
+df['mu'] = samples_Gaussian.reshape(-1,2)[:,0]
+df['sigma'] = samples_Gaussian.reshape(-1,2)[:,1]
 
 with open(paths.output/"mu_mean.txt", "w") as f:
-    f.write(f"${np.mean(samples_Gaussian[:,0]):.3f}\\pm{np.std(samples_Gaussian[:,0]):.3f}$")
+    f.write(f"${np.mean(df['mu']):.3f}\\pm{np.std(df['mu']):.3f}$")
 with open(paths.output/"sigma_mean.txt", "w") as f:
-    f.write(f"${np.mean(samples_Gaussian[:,1]):.3f}\\pm{np.std(samples_Gaussian[:,1]):.3f}$")
+    f.write(f"${np.mean(df['sigma']):.3f}\\pm{np.std(df['sigma']):.3f}$")
 with open(paths.output/"mu_median.txt", "w") as f:
-    f.write(f"${np.median(samples_Gaussian[:,0]):.3f}$")
+    f.write(f"${np.median(df['mu']):.3f}$")
 with open(paths.output/"sigma_median.txt", "w") as f:
-    f.write(f"${np.median(samples_Gaussian[:,1]):.3f}$")
+    f.write(f"${np.median(df['sigma']):.3f}$")
+
+kappa = np.linspace(-1, 1, 50000)
+df = df.sample(10000)
+
+population_probability = np.mean([normal_distribution(kappa[i], df['mu'], df['sigma']) for i in range(len(kappa))], axis=1)
+population_probability = population_probability / np.trapz(population_probability, kappa)
+
+kappa_mean = np.trapz(kappa * population_probability, kappa)
+kappa_sd = np.sqrt(np.trapz(np.square(kappa - kappa_mean) * population_probability, kappa))
+
+with open(paths.output/"kappa_mean.txt", "w") as f:
+    f.write(f"${kappa_mean:.3f}\\pm{kappa_sd:.3f}$")

### 2022-11-22 src/scripts/corner_GW150914.py

@@ -26,13 +26,13 @@ result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
 result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
 result_bilby = result_bilby[result_bilby.event == "GW150914"]
 result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), r"Non-GR (frequency dependent)")
+result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
 result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
 
 result_extra = CBCResult.from_json(filename=paths.data/"GW150914_birefringence(frequency_independent).json").posterior
 result_extra = result_extra.sample(n=nsamples)
 result_extra['kappa'] = result_extra['kappa'] * 1000 # different scale was used in the frequency independent data
-result_extra['with'] = np.full(len(result_extra), r"Non-GR (frequency independent)")
+result_extra['with'] = np.full(len(result_extra), r"birefringence (frequency independent)")
 result_extra['cos_iota'] = np.cos(result_extra['iota'])
 
 result = pd.concat([result_bilby,result_extra,result_ligo], ignore_index=True)

### 2022-11-22 src/scripts/violin_kappa.py

@@ -15,7 +15,7 @@ plt.rcParams.update({
 result_dict = {}
 median_std_dict = {}
 result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_DataFrame = result_DataFrame[result_DataFrame.event != "GW200129"] # remove GW200129
+result_DataFrame = result_DataFrame[result_DataFrame.event != "GW200129_065458"] # remove GW200129
 for event in result_DataFrame['event'].unique():
     result_dict[event] = result_DataFrame[result_DataFrame.event == event]
     median_std_dict[event] = abs(result_dict[event]['kappa'].median()/result_dict[event]['kappa'].std())

### 2022-11-22 src/tex/ms.tex
Update ms.tex

@@
 In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, and mention the configuration we used in the PE.
-In Sec.~\ref{sec:Results}, we show the combined result of the PE, and present the constrain on the GW amplitude birefringence we obtained.
+In Sec.~\ref{sec:Results}, we show the combined result of the PE, and present the constrain on GW amplitude birefringence we obtained.
 In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study, and provide suggestions for future studies.
@@
 \label{sec:Acknowledgements}
+M.~I., K.~W.~K.~W.~ and W.~M.~F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute, which is supported by the Simons Foundation.
+The computational resources used in this work were provided by the Flatiron Institute.
 

### 2022-11-22 src/tex/ms.tex
Update content

@@
 The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
+The mean of $\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.
 
@@
         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        Orange line represent the median of $\mu$ and $\sigma$, which are \variable{output/mu_median.txt} and \variable{output/sigma_median.txt} respectively,.
-        The mean of $\mu$ and $\sigma$ are \variable{output/mu_mean.txt} and \variable{output/sigma_mean.txt}, respectively.
+        Orange line represent the median of $\mu$ and $\sigma$.
     }

### 2022-11-23 src/tex/ms.tex
Change document class
Change document class to REVTeX 4.2

@@
-% \documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,floatfix,nofootinbib]{revtex4-2}
-\documentclass[reprint,amsmath,amssymb,aps,twocolumn]{aastex631}
+\documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,floatfix,nofootinbib]{revtex4-2}
 
@@
 
+\maketitle
+
 \section{Introduction}
@@
 Previous studies have constrained amplitude birefringence by performing Bayesian inference.
-In \citet{Maria_2021}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
+In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
 In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption and performed Bayesian inference on the events in GWTC-1, the first GW transient catalog.
@@
 
-For some previous studies on the birefringence property such as \citet{Maria_2021}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+For some previous studies on the birefringence property such as \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
 This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.
@@
 % Comparison with previous studies
-In \citet{Maria_2021}, the authors gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
+In \citet{Okounkova_2022}, the authors gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
 We convert the constraint to the same format as ours to make comparison.

### 2022-11-24 src/tex/ms.tex
Add comments

@@
 \usepackage{amsfonts,amssymb,amsmath}
+
+\definecolor{rb4}{HTML}{27408B}
+\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
 
@@
 % Motivation
-Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
-After a century, we know that GR does not agree with quantum theories at some length scale.
-To unify both theories, we need to study the possibility of different beyond-GR theories.
-Some beyond-GR theories such as Chern-Simons gravity suggests that there is gravitational wave (GW) amplitude birefringence, while GR predicts there is no birefringence.
-GW amplitude birefringence is a property of space-time which consists of the enhancement of one GW polarization over the other during the propagation of the wave.
-the local modification of the amplitude is small, but it grows with the number of cycles the GW is propagated.
+Since Einstein proposed his theory of general relativity (GR), it was tested in
+a wide range of length scales. After a century, we know that GR does not agree
+with quantum theories at some length scale. To unify both theories, we need to
+study the possibility of different beyond-GR theories. \kw{I think now it kind
+of goes into birefringence too fast. Maybe talk about what effect one may expect
+to see in GW in general a bit?} Some beyond-GR theories such as Chern-Simons
+gravity suggests that there is gravitational wave (GW) amplitude birefringence,
+while GR predicts there is no birefringence. GW amplitude birefringence is a
+property of space-time which consists of the enhancement of one GW polarization
+over the other during the propagation of the wave. the local modification of the
+amplitude is small, but it grows with the number of cycles the GW is propagated.
 
 % Previous studies
-Previous studies have constrained amplitude birefringence by performing Bayesian inference.
-In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
-In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption and performed Bayesian inference on the events in GWTC-1, the first GW transient catalog.
+Previous studies have constrained amplitude birefringence by performing Bayesian
+inference\kw{This sentence looks odd to me}. In \citet{Okounkova_2022}, they
+considered the distribution of observed inclination of the GW events in GWTC-2,
+the second GW transient catalog, and use the distribution to constrain the
+amplitude birefringence effect with the assumption of frequency independence in
+GW amplitude birefringence. In \citet{Yamada_2020} and \citet{Wang_2021}, they
+both modeled the waveform under birefringence effect with a simplify assumption
+\kw{What is the simplify assumption} and performed Bayesian inference on the
+events in GWTC-1, the first GW transient catalog.
 
 % What's new?
-In this paper, we used a birefringence model with higher order terms which include the frequency dependence of amplitude birefringence to perform a more accurate parameter estimation (PE).
-We also performed Bayesian inference on events in GWTC-3, the third GW transient catalog.
-We included 69 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$.
-These events are also listed in TABLE I in \citet{GWTC_3_population}.\footnote{GW190720 and GW200129 are included in the list in \citet{GWTC_3_population}, but we did not include them in this study.
+In this paper, we used a birefringence model with higher order terms which
+include the frequency dependence of amplitude birefringence to perform a more
+accurate parameter estimation (PE). We also performed Bayesian inference on
+events in GWTC-3, the third GW transient catalog. We included 69 binary black
+hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$. These
+events are also listed in TABLE I in
+\citet{GWTC_3_population}.\footnote{GW190720 and GW200129 are included in the
+list in \citet{GWTC_3_population}, but we did not include them in this study.
 Details in Discussion.}
@@
 % Section guide
-In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, and mention the configuration we used in the PE.
-In Sec.~\ref{sec:Results}, we show the combined result of the PE, and present the constrain on GW amplitude birefringence we obtained.
-In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study, and provide suggestions for future studies.
+In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform
+model, and mention the configuration we used in the PE. In
+Sec.~\ref{sec:Results}, we show the combined result of the PE, and present the
+constrain on GW amplitude birefringence we obtained. In
+Sec.~\ref{sec:Discussion}, we discuss the limitation of this study, and provide
+suggestions for future studies.
 
@@
 % Waveform modification
-GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves, which could be transformed into two circular polarizations (i.e. left-handed and right-handed) by
+GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to
+electromagnetic waves, which could be transformed into two circular
+polarizations (i.e. left-handed and right-handed) by
 \begin{equation}
@@
 \end{equation}
-where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of the plus and cross polarization of the waveform.
-According to \textbf{[modification reference]}, we modified the waveform model by
+where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and
+right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier
+amplitude of the plus and cross polarization of the waveform. According to
+\textbf{[modification reference]}, we modified the waveform model by
 \begin{equation}
@@
 \end{equation}
-where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform at the source, $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the Fourier frequency of the waveform.
-We assume GWs are generated by the binary as GR predicts, as the modification at the source is small before the effect accumulate during the propagation.
-As the GWs propagate, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the GWs.
+where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified
+waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the
+Fourier amplitude of the waveform at the source, $\kappa$ is the dimensionless
+opacity parameter that represent the strength of the birefringence, $d_C$ is the
+comoving distance to the source, and $f$ is the Fourier frequency of the
+waveform. We assume GWs are generated by the binary as GR predicts, as the
+modification at the source is small before the effect accumulate during the
+propagation. As the GWs propagate, the effect of birefringence will be built up
+with the number of cycles which depends on the distance traveled and the
+frequency of the GWs.
 
@@
 % Stacked kappa posterior plot
-\begin{figure}[h]
-    \script{kappa_stacked.py}
-    \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
-    \caption{
-        The posterior of $\kappa$ for all 69 events included in this study.
-        Each color represent a different event.
-        These posterior are used to calculate the population constraint on $\kappa$ in the following.
-    }
-    \label{fig:kappa_stacked}
-\end{figure}
+% \begin{figure}[h]
+%     \script{kappa_stacked.py}
+%     \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
+%     \caption{
+%         The posterior of $\kappa$ for all 69 events included in this study.
+%         Each color represent a different event.
+%         These posterior are used to calculate the population constraint on $\kappa$ in the following.
+%     }
+%     \label{fig:kappa_stacked}
+% \end{figure}
 
@@
 % Corner plot of mean and standard deviation of kappa
-\begin{figure}[h]
-    \script{corner_Gaussian.py}
-    \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
-    \caption{
-        The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
-        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        Orange line represent the median of $\mu$ and $\sigma$.
-    }
-    \label{fig:corner_Gaussian}
-\end{figure}
+% \begin{figure}[h]
+%     \script{corner_Gaussian.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
+%     \caption{
+%         The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
+%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+%         Orange line represent the median of $\mu$ and $\sigma$.
+%     }
+%     \label{fig:corner_Gaussian}
+% \end{figure}
 
@@
 
-\begin{figure*}[ht]
-    \script{violin_kappa.py}
-    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
-    \caption{
-        The violin plot shows the posterior of $\kappa$ for all 69 events included in this study.
-        Each violin represents a different event.
-        The violins are sorted by the quotient of median and standard deviation of the posterior.
-        Blue horizontal line represent the median value of $\mu$, and blue region represent 1 $\sigma$ confidence level.
-        Orange horizontal line represent $\kappa=0$.
-        This plot shows that there are some events with $\kappa$ distribution significantly deviate from $0$.
-    }
-    \label{fig:violin_kappa}
-\end{figure*}
+% \begin{figure*}[ht]
+%     \script{violin_kappa.py}
+%     \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
+%     \caption{
+%         The violin plot shows the posterior of $\kappa$ for all 69 events included in this study.
+%         Each violin represents a different event.
+%         The violins are sorted by the quotient of median and standard deviation of the posterior.
+%         Blue horizontal line represent the median value of $\mu$, and blue region represent 1 $\sigma$ confidence level.
+%         Orange horizontal line represent $\kappa=0$.
+%         This plot shows that there are some events with $\kappa$ distribution significantly deviate from $0$.
+%     }
+%     \label{fig:violin_kappa}
+% \end{figure*}
 
@@
 
-\begin{figure}[h]
-    \script{corner_GW150914.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
-    \caption{
-        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-        Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with both frequency independent and dependent birefringence respectively.
-        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
-        This plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
-    }
-    \label{fig:corner_GW150914}
-\end{figure}
+% \begin{figure}[h]
+%     \script{corner_GW150914.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
+%     \caption{
+%         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
+%         Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with both frequency independent and dependent birefringence respectively.
+%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+%         Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
+%         This plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
+%     }
+%     \label{fig:corner_GW150914}
+% \end{figure}
 
@@
 
-\begin{figure}[h]
-    \script{corner_GW190521.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
-    \caption{
-        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
-        Colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
-        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        This plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
-    }
-    \label{fig:corner_GW190521}
-\end{figure}
+% \begin{figure}[h]
+%     \script{corner_GW190521.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
+%     \caption{
+%         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
+%         Colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
+%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+%         This plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
+%     }
+%     \label{fig:corner_GW190521}
+% \end{figure}
 

### 2022-11-24 src/tex/ms.tex
update

@@
 of goes into birefringence too fast. Maybe talk about what effect one may expect
-to see in GW in general a bit?} Some beyond-GR theories such as Chern-Simons
-gravity suggests that there is gravitational wave (GW) amplitude birefringence,
-while GR predicts there is no birefringence. GW amplitude birefringence is a
-property of space-time which consists of the enhancement of one GW polarization
-over the other during the propagation of the wave. the local modification of the
+to see in GW in general a bit? Say there is polarization modification, phasing
+modification, etc.} Some beyond-GR theories such as Chern-Simons gravity
+suggests that there is gravitational wave (GW) amplitude birefringence, while GR
+predicts there is no birefringence. GW amplitude birefringence is a property of
+space-time which consists of the enhancement of one GW polarization over the
+other during the propagation of the wave. the local modification of the
 amplitude is small, but it grows with the number of cycles the GW is propagated.
@@
 include the frequency dependence of amplitude birefringence to perform a more
-accurate parameter estimation (PE). We also performed Bayesian inference on
-events in GWTC-3, the third GW transient catalog. We included 69 binary black
-hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$. These
-events are also listed in TABLE I in
+accurate parameter estimation (PE) \kw{I think there is qualitative difference
+between this work and former works, so it is not just more accurate}. We also
+performed Bayesian inference on events in GWTC-3, the third GW transient
+catalog. We included 69 binary black hole merger events with smallest false
+alarm rate $\leq1\mathrm{yr^{-1}}$. These events are also listed in TABLE I in
 \citet{GWTC_3_population}.\footnote{GW190720 and GW200129 are included in the

### 2022-11-25 src/tex/ms.tex
update

@@
 electromagnetic waves, which could be transformed into two circular
-polarizations (i.e. left-handed and right-handed) by
+polarizations (i.e. left-handed and right-handed) \kw{GW consist of two
+independent polarization modes, which we can choose to represent them in either
+the linear or circular basis. Better rewrite this sentence.}by
 \begin{equation}
@@
 \end{equation}
-where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and
-right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier
-amplitude of the plus and cross polarization of the waveform. According to
+where $h_{\mathrm{L/R}}$ are the Fourier amplitude \kw{The equation above is not
+specifically in the Fourier space, right?} of the left-handed and right-handed
+polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of
+the plus and cross polarization of the waveform. According to
 \textbf{[modification reference]}, we modified the waveform model by
@@
 waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the
-Fourier amplitude of the waveform at the source, $\kappa$ is the dimensionless
-opacity parameter that represent the strength of the birefringence, $d_C$ is the
-comoving distance to the source, and $f$ is the Fourier frequency of the
-waveform. We assume GWs are generated by the binary as GR predicts, as the
-modification at the source is small before the effect accumulate during the
-propagation. As the GWs propagate, the effect of birefringence will be built up
-with the number of cycles which depends on the distance traveled and the
-frequency of the GWs.
+Fourier amplitude of the waveform at the source \kw{Just say this is the
+amplitude in the GR case.}, $\kappa$ is the dimensionless opacity parameter that
+represent the strength of the birefringence, $d_C$ is the comoving distance to
+the source, and $f$ is the Fourier frequency of the waveform. We assume GWs are
+generated by the binary as GR predicts, as the modification at the source is
+small before the effect accumulate during the propagation. As the GWs propagate,
+the effect of birefringence will be built up with the number of cycles which
+depends on the distance traveled and the frequency of the GWs.
 
@@
 This modification changes the ratio of the left-handed and right-handed polarization of the waveform, which affects the PE on the inclination as well.
-In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
+In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary. \kw{Swap the order of the two sentences.}
 The relationship between the amplitude ratio and the inclination is
@@
 \end{equation}
-where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed and right-handed polarizations of the GWs and $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
-PE on $\iota$ takes the amplitude ratio into account. Therefore, the PE on $\iota$ depends on the amplitude ratio.
+where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed
+and right-handed polarizations of the GWs \kw{You don't have to redefine the
+symbols everytime. Just do it once when you introduce the symbol.} and $\iota$
+is the inclination, the angle between our line of sight and the orbital angular
+momentum of the binary. PE on $\iota$ takes the amplitude ratio into account. \kw{You don't need the sentence in front of this comment.}
+Therefore, the PE on $\iota$ depends on the amplitude ratio.
 
@@
 
-For some previous studies on the birefringence property such as \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+For some previous studies on the birefringence property such as \citet{Okounkova_2022} \kw{In \citet{Okounkova_2022},}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
 This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.

### 2022-11-26 src/tex/ms.tex
Update ms.tex

--- 2022-11-24 src/tex/ms.tex
@@
 % Previous studies
-Previous studies have constrained amplitude birefringence by performing Bayesian
-inference\kw{This sentence looks odd to me}. In \citet{Okounkova_2022}, they
-considered the distribution of observed inclination of the GW events in GWTC-2,
+Previous studies have constrained amplitude birefringence by performing different statistical analysis.
+In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2,
 the second GW transient catalog, and use the distribution to constrain the
@@
 both modeled the waveform under birefringence effect with a simplify assumption
-\kw{What is the simplify assumption} and performed Bayesian inference on the
+\kw{What is the simplify assumption} and performed parameter estimation (PE) on the
 events in GWTC-1, the first GW transient catalog.
@@
 include the frequency dependence of amplitude birefringence to perform a more
-accurate parameter estimation (PE) \kw{I think there is qualitative difference
+accurate PE \kw{I think there is qualitative difference
 between this work and former works, so it is not just more accurate}. We also
@@
 % Waveform modification
-GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to
-electromagnetic waves, which could be transformed into two circular
-polarizations (i.e. left-handed and right-handed) by
+GW consist of two independent polarization modes similar to electromagnetic waves,
+which we can choose to represent them in either the linear basis (i.e. $+$ and $\times$)
+or the circular basis (i.e. left-handed and right-handed).
+They are related by
 \begin{equation}
@@
 \end{equation}
-where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and
-right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier
-amplitude of the plus and cross polarization of the waveform. According to
+where $h_{\mathrm{L/R}}$ are the amplitude of the left-handed and right-handed
+polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of
+the plus and cross polarization of the waveform. According to
 \textbf{[modification reference]}, we modified the waveform model by
@@
 waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the
-Fourier amplitude of the waveform at the source, $\kappa$ is the dimensionless
-opacity parameter that represent the strength of the birefringence, $d_C$ is the
-comoving distance to the source, and $f$ is the Fourier frequency of the
-waveform. We assume GWs are generated by the binary as GR predicts, as the
-modification at the source is small before the effect accumulate during the
-propagation. As the GWs propagate, the effect of birefringence will be built up
-with the number of cycles which depends on the distance traveled and the
-frequency of the GWs.
+Fourier amplitude of the waveform as GR predicts, $\kappa$ is the dimensionless opacity parameter that
+represent the strength of the birefringence, $d_C$ is the comoving distance to
+the source, and $f$ is the Fourier frequency of the waveform. We assume GWs are
+generated by the binary as GR predicts, as the modification at the source is
+small before the effect accumulate during the propagation. As the GWs propagate,
+the effect of birefringence will be built up with the number of cycles which
+depends on the distance traveled and the frequency of the GWs.
 
 % Effect on inclination PE
+In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
 This modification changes the ratio of the left-handed and right-handed polarization of the waveform, which affects the PE on the inclination as well.
-In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
 The relationship between the amplitude ratio and the inclination is
@@
 \end{equation}
-where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed and right-handed polarizations of the GWs and $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
-PE on $\iota$ takes the amplitude ratio into account. Therefore, the PE on $\iota$ depends on the amplitude ratio.
+where $\iota$ is the inclination, the angle between our line of sight and the orbital angular
+momentum of the binary. Therefore, the PE on $\iota$ depends on the amplitude ratio.
 
@@
 
-For some previous studies on the birefringence property such as \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+In \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
 This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.

### 2022-11-26 src/tex/ms.tex
Add comments

--- 2022-11-25 src/tex/ms.tex
@@
 In contrast, We included the frequency dependence in the modification, which is a first-order approximation of the birefringence model.
-    This frequency dependence provides stronger enhancement or suppression from the birefringence to high frequency components of the GW, and visa versa.
-Also, the frequency dependence can break the degeneracy between $\kappa$ and $\iota$, as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
+This frequency dependence provides stronger enhancement or suppression from the birefringence to high frequency components of the GW, and visa versa. \kw{Redundant.}
+Also, \kw{You don't need an also here since you are still talking about the same thing} the frequency dependence can break the degeneracy between $\kappa$ and $\iota$, as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
 To illustrate the frequency dependence of amplitude birefringence, the observed amplitudes of two polarizations of a simulated GW signal are shown in figure \ref{fig:birefringence}.
@@
 % Parameter estimation
-The major objective of this study is to constrain GW amplitude birefringence, which is represented by the opacity parameter $\kappa$.
-To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question \citep{Bilby}.
-We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using the modified waveforms.
-In both case, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each individual events.
-We also perform test to check if the PE could recover the PE done by LIGO-Virgo-KAGRA Collaboration (LVK) in \citet{GWTC-2.1, GWTC-3}.
-We set the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as using GR.
-These control tests result can be found in the data release.
+The major objective of this study is to constrain GW amplitude birefringence,
+which is represented by the opacity parameter $\kappa$. To constrain $\kappa$,
+we perform PE with data from the third GW transient catalog \citep{GWTC-2.1,
+GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which is
+able to calculate posteriors of GW parameters based on interferometer data and
+priors for the parameters in question \citep{Bilby}. We set the prior on
+$\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using
+the modified waveforms. In both case, we use the same PE configuration as in
+\citet{GWTC-2.1, GWTC-3} for each individual events. We also perform test to
+check if the PE could recover the PE done by LIGO-Virgo-KAGRA Collaboration
+\kw{State that this is a sanity check} (LVK) in \citet{GWTC-2.1, GWTC-3}. We set
+the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as
+using GR \kw{This sentence kind of read odd. Add conjunction to link it to the
+GR test.}. These control tests result can be found in the data release.
+\kw{Link}
 
@@
 \label{sec:Results}
+
+\kw{Looking at the current draft, I do think maybe we should first talk about
+ individual constraints first, then population constraints. Tho we should give a
+ brief summary of the structure at the beginning of this section}
 \subsection{Result on GWTC-3}
@@
 % How to calculate the combined posterior
-In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$ of each events obtained from PE.
-We assume $\kappa$ is different for each event, and the population distribution of $\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
-We then performed hierarchical modeling to construct the likelihood function of $\mu$ and $\sigma$ given interferometer data.
-To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}, with prior of $\mu$ and $\sigma$ being uniform distributions.
-The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
-The mean of $\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.
+In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$
+of each events obtained from PE. \kw{Rewrite the following paragraph in the
+following way: 1. There are a set of measured $\kappa$ 2. State that there can
+either be a single true $\kappa$ or a distribution of $\kappa$ 3. Describe
+hierarchical Bayesian, which is essentially stacking the individual results
+together. 4. Your choice of model and reason behind such choice.} We assume
+$\kappa$ is different for each event, and the population distribution of
+$\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation
+$\sigma$. We then performed hierarchical modeling \kw{Need to describe what
+hierarchical modelling is} to construct the likelihood function of $\mu$ and
+$\sigma$ given interferometer data. To sample the posterior distribution of
+$\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC},
+with prior of $\mu$ and $\sigma$ being uniform distributions. The posterior of
+$\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}. The mean of
+$\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.
 
@@
 \subsection{Result on GW150914}
-Some events have bimodal $\kappa$ distribution, but before we show a bimodal event, we should understand the mechanism of a unimodal $\kappa$ distribution.
-Consider GW150914, the first GW detected by LIGO.
-In figure \ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
-With the frequency independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
-This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
-To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
-As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
+Some events have bimodal $\kappa$ distribution, but before we show a bimodal
+event, we should understand the mechanism of a unimodal $\kappa$ distribution.
+\kw{Just say what you are trying to say, instead of leaving a cliff hanger}
+Consider GW150914, the first GW detected by LIGO. In figure
+\ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
+With the frequency independent birefringence model, the posteriors for
+$\cos\iota$ look different from the posteriors assuming GR. This is because, for
+a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if
+the frequency dependence is not included. To reconstruct the amplitude ratio
+from the interferometer data, an $\iota$ representing a more face-off inspiral
+can pair with a positive $\kappa$, or an $\iota$ representing a more face-on
+inspiral with a negative $\kappa$. As a result, many pairs of $\iota$ and
+$\kappa$ can be plausible with the frequency independence. \kw{Many pairs
+doesn't mean bimodelity, you should say why the middle part is constrained to be small as well.}
 
@@
 
+\kw{Mention two things that drive the population constrain to be better: better event (which you already did), and more events}
+
 \section{Acknowledgements}

### 2022-11-26 src/tex/ms.tex
Update

@@
 \usepackage{amsfonts,amssymb,amsmath}
+\usepackage{hyperref}
 
@@
 In contrast, We included the frequency dependence in the modification, which is a first-order approximation of the birefringence model.
-This frequency dependence provides stronger enhancement or suppression from the birefringence to high frequency components of the GW, and visa versa. \kw{Redundant.}
-Also, \kw{You don't need an also here since you are still talking about the same thing} the frequency dependence can break the degeneracy between $\kappa$ and $\iota$, as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
+The frequency dependence can break the degeneracy between $\kappa$ and $\iota$, as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
 To illustrate the frequency dependence of amplitude birefringence, the observed amplitudes of two polarizations of a simulated GW signal are shown in figure \ref{fig:birefringence}.
@@
 using GR \kw{This sentence kind of read odd. Add conjunction to link it to the
-GR test.}. These control tests result can be found in the data release.
-\kw{Link}
+GR test.}. These control tests result can be found in the \href{https://zenodo.org/record/7338924}{data release}.
 

### 2022-11-28 src/tex/ms.tex
Update ms.tex

@@
 % Motivation
-Since Einstein proposed his theory of general relativity (GR), it was tested in
-a wide range of length scales. After a century, we know that GR does not agree
-with quantum theories at some length scale. To unify both theories, we need to
-study the possibility of different beyond-GR theories. \kw{I think now it kind
-of goes into birefringence too fast. Maybe talk about what effect one may expect
-to see in GW in general a bit? Say there is polarization modification, phasing
-modification, etc.} Some beyond-GR theories such as Chern-Simons gravity
-suggests that there is gravitational wave (GW) amplitude birefringence, while GR
-predicts there is no birefringence. GW amplitude birefringence is a property of
-space-time which consists of the enhancement of one GW polarization over the
-other during the propagation of the wave. the local modification of the
-amplitude is small, but it grows with the number of cycles the GW is propagated.
+Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
+After a century, we know that GR does not agree with quantum theories at some length scale.
+To unify both theories, we need to study the possibility of different beyond-GR theories.
+\kw{I think now it kind of goes into birefringence too fast.
+Maybe talk about what effect one may expect to see in GW in general a bit?
+Say there is polarization modification, phasing modification, etc.}
+Some beyond-GR theories such as Chern-Simons gravity suggests that there is gravitational wave (GW) amplitude birefringence, while GR predicts there is no birefringence.
+GW amplitude birefringence is a property of space-time which consists of the enhancement of one GW polarization over the other during the propagation of the wave.
+The local modification of the amplitude is small, but it grows with the number of cycles the GW is propagated.
 
@@
 Previous studies have constrained amplitude birefringence by performing different statistical analysis.
-In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2,
-the second GW transient catalog, and use the distribution to constrain the
-amplitude birefringence effect with the assumption of frequency independence in
-GW amplitude birefringence. In \citet{Yamada_2020} and \citet{Wang_2021}, they
-both modeled the waveform under birefringence effect with a simplify assumption
-\kw{What is the simplify assumption} and performed parameter estimation (PE) on the
-events in GWTC-1, the first GW transient catalog.
+In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
+In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption.
+They assumed the modification of amplitude birefringence is small and performed parameter estimation (PE) on the events in GWTC-1, the first GW transient catalog.
 
 % What's new?
-In this paper, we used a birefringence model with higher order terms which
-include the frequency dependence of amplitude birefringence to perform a more
-accurate PE \kw{I think there is qualitative difference
-between this work and former works, so it is not just more accurate}. We also
-performed Bayesian inference on events in GWTC-3, the third GW transient
-catalog. We included 69 binary black hole merger events with smallest false
-alarm rate $\leq1\mathrm{yr^{-1}}$. These events are also listed in TABLE I in
-\citet{GWTC_3_population}.\footnote{GW190720 and GW200129 are included in the
-list in \citet{GWTC_3_population}, but we did not include them in this study.
+In this paper, we used a birefringence model with higher order terms which include the frequency dependence of amplitude birefringence to perform a more accurate PE.
+\kw{I think there is qualitative difference between this work and former works, so it is not just more accurate}
+We also performed Bayesian inference on events in GWTC-3, the third GW transient catalog.
+We included 69 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$.
+These events are also listed in TABLE I in \citet{GWTC_3_population}.
+~\footnote{GW190720 and GW200129 are included in the list in \citet{GWTC_3_population}, but we did not include them in this study.
 Details in Discussion.}
@@
 % Section guide
-In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform
-model, and mention the configuration we used in the PE. In
-Sec.~\ref{sec:Results}, we show the combined result of the PE, and present the
-constrain on GW amplitude birefringence we obtained. In
-Sec.~\ref{sec:Discussion}, we discuss the limitation of this study, and provide
-suggestions for future studies.
+In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, and mention the configuration we used in the PE.
+In Sec.~\ref{sec:Results}, we show the combined result of the PE, and present the constrain on GW amplitude birefringence we obtained.
+In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study, and provide suggestions for future studies.
 
@@
 % Waveform modification
-GW consist of two independent polarization modes similar to electromagnetic waves,
-which we can choose to represent them in either the linear basis (i.e. $+$ and $\times$)
-or the circular basis (i.e. left-handed and right-handed).
+GW consist of two independent polarization modes similar to electromagnetic waves, which we can choose to represent them in either the linear basis (i.e.
+$+$ and $\times$) or the circular basis (i.e.
+left-handed and right-handed).
 They are related by
@@
 \end{equation}
-where $h_{\mathrm{L/R}}$ are the amplitude of the left-handed and right-handed
-polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of
-the plus and cross polarization of the waveform. According to
-\textbf{[modification reference]}, we modified the waveform model by
+where $h_{\mathrm{L/R}}$ are the amplitude of the left-handed and right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of the plus and cross polarization of the waveform.
+According to \textbf{[modification reference]}, we modified the waveform model by
 \begin{equation}
@@
 \end{equation}
-where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified
-waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the
-Fourier amplitude of the waveform as GR predicts, $\kappa$ is the dimensionless opacity parameter that
-represent the strength of the birefringence, $d_C$ is the comoving distance to
-the source, and $f$ is the Fourier frequency of the waveform. We assume GWs are
-generated by the binary as GR predicts, as the modification at the source is
-small before the effect accumulate during the propagation. As the GWs propagate,
-the effect of birefringence will be built up with the number of cycles which
-depends on the distance traveled and the frequency of the GWs.
+where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform as GR predicts, $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the Fourier frequency of the waveform.
+We assume GWs are generated by the binary as GR predicts, as the modification at the source is small before the effect accumulate during the propagation.
+As the GWs propagate, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the GWs.
 
@@
 \end{equation}
-where $\iota$ is the inclination, the angle between our line of sight and the orbital angular
-momentum of the binary. Therefore, the PE on $\iota$ depends on the amplitude ratio.
+where $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
+Therefore, the PE on $\iota$ depends on the amplitude ratio.
 
@@
 where $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed Fourier amplitude of left-handed and right-handed polarizations of the GWs.
-As the amplitude ratio does not only depends on $\iota$ with this modification, but also $\kappa$ and $d_C$, the PE on $\iota$ and $d_C$ are affected. Note that GR is recovered if $\kappa$ is $0$.
+As the amplitude ratio does not only depends on $\iota$ with this modification, but also $\kappa$ and $d_C$, the PE on $\iota$ and $d_C$ are affected.
+Note that GR is recovered if $\kappa$ is $0$.
 
@@
 % Parameter estimation
-The major objective of this study is to constrain GW amplitude birefringence,
-which is represented by the opacity parameter $\kappa$. To constrain $\kappa$,
-we perform PE with data from the third GW transient catalog \citep{GWTC-2.1,
-GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which is
-able to calculate posteriors of GW parameters based on interferometer data and
-priors for the parameters in question \citep{Bilby}. We set the prior on
-$\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using
-the modified waveforms. In both case, we use the same PE configuration as in
-\citet{GWTC-2.1, GWTC-3} for each individual events. We also perform test to
-check if the PE could recover the PE done by LIGO-Virgo-KAGRA Collaboration
-\kw{State that this is a sanity check} (LVK) in \citet{GWTC-2.1, GWTC-3}. We set
-the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as
-using GR \kw{This sentence kind of read odd. Add conjunction to link it to the
-GR test.}. These control tests result can be found in the \href{https://zenodo.org/record/7338924}{data release}.
+The major objective of this study is to constrain GW amplitude birefringence, which is represented by the opacity parameter $\kappa$.
+To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question \citep{Bilby}.
+\
+We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using the modified waveforms.
+We use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each individual events.
+We also perform sanity test to check if our PE could recover the PE done by LIGO-Virgo-KAGRA Collaboration (LVK) in \citet{GWTC-2.1, GWTC-3}.
+To implement the sanity check, we need to assume there is no birefringence by setting the prior on $\kappa$ to be a $\delta$ function at $0$, which means GR is recovered.
+These sanity tests result can be found in the \href{https://zenodo.org/record/7338924}{data release}.
 
@@
 
-\kw{Looking at the current draft, I do think maybe we should first talk about
- individual constraints first, then population constraints. Tho we should give a
- brief summary of the structure at the beginning of this section}
+\kw{Looking at the current draft, I do think maybe we should first talk about individual constraints first, then population constraints.
+Tho we should give a brief summary of the structure at the beginning of this section}
 \subsection{Result on GWTC-3}
@@
 In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$
-of each events obtained from PE. \kw{Rewrite the following paragraph in the
-following way: 1. There are a set of measured $\kappa$ 2. State that there can
-either be a single true $\kappa$ or a distribution of $\kappa$ 3. Describe
-hierarchical Bayesian, which is essentially stacking the individual results
-together. 4. Your choice of model and reason behind such choice.} We assume
-$\kappa$ is different for each event, and the population distribution of
-$\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation
-$\sigma$. We then performed hierarchical modeling \kw{Need to describe what
-hierarchical modelling is} to construct the likelihood function of $\mu$ and
-$\sigma$ given interferometer data. To sample the posterior distribution of
-$\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC},
-with prior of $\mu$ and $\sigma$ being uniform distributions. The posterior of
-$\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}. The mean of
-$\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.
+of each events obtained from PE.
+
+\kw{Rewrite the following paragraph in the following way:
+1. There are a set of measured $\kappa$
+2. State that there can either be a single true $\kappa$ or a distribution of $\kappa$
+3. Describe hierarchical Bayesian, which is essentially stacking the individual results together.
+4. Your choice of model and reason behind such choice.}
+
+We assume $\kappa$ is different for each event, and the population distribution of $\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
+We then performed hierarchical modeling
+\kw{Need to describe what hierarchical modelling is}
+to construct the likelihood function of $\mu$ and $\sigma$ given interferometer data.
+To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}, with prior of $\mu$ and $\sigma$ being uniform distributions.
+The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
+The mean of $\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.
 
@@
 \subsection{Result on GW150914}
-Some events have bimodal $\kappa$ distribution, but before we show a bimodal
-event, we should understand the mechanism of a unimodal $\kappa$ distribution.
+Some events have bimodal $\kappa$ distribution, but before we show a bimodalcevent, we should understand the mechanism of a unimodal $\kappa$ distribution.
 \kw{Just say what you are trying to say, instead of leaving a cliff hanger}
-Consider GW150914, the first GW detected by LIGO. In figure
-\ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
-With the frequency independent birefringence model, the posteriors for
-$\cos\iota$ look different from the posteriors assuming GR. This is because, for
-a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if
-the frequency dependence is not included. To reconstruct the amplitude ratio
-from the interferometer data, an $\iota$ representing a more face-off inspiral
-can pair with a positive $\kappa$, or an $\iota$ representing a more face-on
-inspiral with a negative $\kappa$. As a result, many pairs of $\iota$ and
-$\kappa$ can be plausible with the frequency independence. \kw{Many pairs
-doesn't mean bimodelity, you should say why the middle part is constrained to be small as well.}
+Consider GW150914, the first GW detected by LIGO.
+In figure \ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
+With the frequency independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
+This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
+To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
+As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
+\kw{Many pairs doesn't mean bimodelity, you should say why the middle part is constrained to be small as well.}
 
@@
 This is because the degeneracy was broken by the frequency dependence, as the effect of birefringence will be different from the one of changing $\iota$.
-In this case, the posteriors can recover the GR distributions for the binary parameters,
-and the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent, and GR can be recovered.
+In this case, the posteriors can recover the GR distributions for the binary parameters, and the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent, and GR can be recovered.
 

### 2022-11-29 src/tex/ms.tex
Update ms.tex

@@
 \begin{abstract}
-    One of the major problems that physicists are facing nowadays is Einstein's theory of general relativity (GR) does not agree with quantum theories at some length scales.
-    In recent decades, theorist has been working on many beyond-GR theories to unify both of them.
-    Some of these theories such as Chern-Simons gravity suggest that there is gravitational wave (GW) amplitude birefringence.
+    One of the major problems physicists are facing nowadays is that Einstein's theory of general relativity (GR) does not agree with quantum theories at some length scales.
+    In recent decades, theorists have been working on many beyond-GR theories to unify both.
+    Some of these theories, such as Chern-Simons gravity, suggest that there is gravitational wave (GW) amplitude birefringence.
     In our study, we perform parameter estimation (PE) to constrain the strength of GW amplitude birefringence.
-    Compare to similar previous studies on the same topic, we perform PE on more events, including events in the third LIGO-Virgo catalog (GWTC-3), and we used a more accurate model of birefringence to describe the phenomenon.
-    We were able to obtain a constrain an order of magnitude tighter than previous studies which used GWTC-2.
+    Compared to similar previous studies on the same topic, we performed PE on more events, including events in the third LIGO-Virgo catalog (GWTC-3), and used a more accurate birefringence model to describe the phenomenon.
+    We obtained a constraint an order of magnitude tighter than previous studies which used GWTC-2.
 \end{abstract}
@@
 
-\kw{Looking at the current draft, I do think maybe we should first talk about individual constraints first, then population constraints.
-Tho we should give a brief summary of the structure at the beginning of this section}
+In this section, we will present the results of individual PE first.
+Then, we will show the population constraint we obtained from the PE of all events in GWTC-3.
+At last, we will discuss some of the special events.
+\kw{Looking at the current draft, I do think maybe we should first talk about individual constraints first, then population constraints.}
+
 \subsection{Result on GWTC-3}
@@
 % How to calculate the combined posterior
-In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$
-of each events obtained from PE.
-
-\kw{Rewrite the following paragraph in the following way:
-1. There are a set of measured $\kappa$
-2. State that there can either be a single true $\kappa$ or a distribution of $\kappa$
-3. Describe hierarchical Bayesian, which is essentially stacking the individual results together.
-4. Your choice of model and reason behind such choice.}
-
-We assume $\kappa$ is different for each event, and the population distribution of $\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
-We then performed hierarchical modeling
-\kw{Need to describe what hierarchical modelling is}
-to construct the likelihood function of $\mu$ and $\sigma$ given interferometer data.
+In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$ of each events obtained from PE.
+These individual posteriors can then be treated as a set of measured $\kappa$ posterior.
+We can either assume that $\kappa$ is a single value constant, or that $\kappa$ is a random variable with a probability distribution.
+In this study, we assume $\kappa$ is a random variable, and the population distribution of $\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
+With Bayesian hierarchical modelling, which is a method to combine multiple measurements of a random variable by assuming the measurements are independent and identically distributed, we can calculate the posterior of $\mu$ and $\sigma$ from the set of measured $\kappa$ posterior.
+The posterior of $\mu$ and $\sigma$ can be calculated by
+\begin{equation}
+    P(\mu,\sigma|\{d_i\})\propto P(\mu,\sigma)\prod_{i}\int\frac{P(\kappa|d_i)}{P(\kappa)}P(\kappa|\mu,\sigma)\,d\kappa\,,
+\end{equation}
+where function $P$ is the probability density function, and $\{d_i\}$ is the set of GW event data.
 To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}, with prior of $\mu$ and $\sigma$ being uniform distributions.
@@
 \subsection{Result on GW150914}
-Some events have bimodal $\kappa$ distribution, but before we show a bimodalcevent, we should understand the mechanism of a unimodal $\kappa$ distribution.
-\kw{Just say what you are trying to say, instead of leaving a cliff hanger}
-Consider GW150914, the first GW detected by LIGO.
+In this study, we included frequency dependence of birefringence.
+This would affect the posterior of $\kappa$ obtained from PE.
+Consider GW150914, the first GW detected by LIGO, as an example.
 In figure \ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
@@
 To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
-As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
-\kw{Many pairs doesn't mean bimodelity, you should say why the middle part is constrained to be small as well.}
+As a result, degeneracy between $\iota$ and $\kappa$ is formed with the frequency independence.
+\kw{you should say why the middle part is constrained to be small.}
 
@@
 \subsection{Result on GW190521}
-
-Now, we show a bimodal $\kappa$ distribution.
+Even with the frequency dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
 Consider GW190521, the most massive binary black hole merger in the events we included.
@@
 % More observations with higher SNR
-Another future work is to apply the same method to more events with higher signal-to-noise ratio (SNR).
+Another future work is to apply the same method to more events, and events with higher signal-to-noise ratio (SNR).
 LVK will release more events with higher SNR in the future.
-Data with higher SNR allows us to obtain more precise PE results, which also allows us to constrain the birefringence effect more precisely.
-And more events will allow us to calculate a more constrained population posterior of $\kappa$.
-
-\kw{Mention two things that drive the population constrain to be better: better event (which you already did), and more events}
+Using data with higher SNR allows us to obtain more precise PE results, which also allows us to constrain the birefringence effect more precisely.
+And using data from more events will allow us to calculate a more constrained population posterior of $\kappa$.
 

### 2022-11-30 src/tex/ms.tex
New Content

@@
 \begin{abstract}
-    One of the major problems physicists are facing nowadays is that Einstein's theory of general relativity (GR) does not agree with quantum theories at some length scales.
-    In recent decades, theorists have been working on many beyond-GR theories to unify both.
+    One of the major problems physicists face nowadays is that Einstein's theory of general relativity (GR) does not agree with quantum theories at some length scales.
+    In recent decades, theorists have worked on many beyond-GR theories to unify both.
     Some of these theories, such as Chern-Simons gravity, suggest that there is gravitational wave (GW) amplitude birefringence.
     In our study, we perform parameter estimation (PE) to constrain the strength of GW amplitude birefringence.
-    Compared to similar previous studies on the same topic, we performed PE on more events, including events in the third LIGO-Virgo catalog (GWTC-3), and used a more accurate birefringence model to describe the phenomenon.
-    We obtained a constraint an order of magnitude tighter than previous studies which used GWTC-2.
+    Compared to previous studies, we performed PE on more events, including events in the third LIGO-Virgo catalog (GWTC-3), and used a more realistic birefringence model to describe the phenomenon.
+    We obtained a constraint tighter than previous studies for an order of magnitude.
 \end{abstract}
@@
 % Motivation
-Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
+Since Einstein proposed his theory of general relativity (GR), it has been tested in various length scales.
 After a century, we know that GR does not agree with quantum theories at some length scale.
@@
 \kw{I think now it kind of goes into birefringence too fast.
-Maybe talk about what effect one may expect to see in GW in general a bit?
+Maybe talk about what effect one may expect to see in GW, in general, a bit?
 Say there is polarization modification, phasing modification, etc.}
-Some beyond-GR theories such as Chern-Simons gravity suggests that there is gravitational wave (GW) amplitude birefringence, while GR predicts there is no birefringence.
-GW amplitude birefringence is a property of space-time which consists of the enhancement of one GW polarization over the other during the propagation of the wave.
-The local modification of the amplitude is small, but it grows with the number of cycles the GW is propagated.
+Some beyond-GR theories, such as Chern-Simons gravity, suggest that there is gravitational wave (GW) amplitude birefringence, while GR predicts no birefringence.
+GW amplitude birefringence is a property of space-time that consists of the enhancement of one GW polarization over the other during wave propagation.
 
 % Previous studies
-Previous studies have constrained amplitude birefringence by performing different statistical analysis.
-In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
-In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption.
-They assumed the modification of amplitude birefringence is small and performed parameter estimation (PE) on the events in GWTC-1, the first GW transient catalog.
+Previous studies have constrained amplitude birefringence by performing different statistical analyses.
+In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and used the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
+In \citet{Yamada_2020} and \citet{Wang_2021}, they both performed parameter estimation (PE) on the events in GWTC-1, the first GW transient catalog.
 
 % What's new?
-In this paper, we used a birefringence model with higher order terms which include the frequency dependence of amplitude birefringence to perform a more accurate PE.
-\kw{I think there is qualitative difference between this work and former works, so it is not just more accurate}
-We also performed Bayesian inference on events in GWTC-3, the third GW transient catalog.
-We included 69 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$.
-These events are also listed in TABLE I in \citet{GWTC_3_population}.
-~\footnote{GW190720 and GW200129 are included in the list in \citet{GWTC_3_population}, but we did not include them in this study.
+In this study, we used a birefringence model with higher-order terms, including amplitude birefringence's frequency dependence.
+This model is more realistic than the frequency-independent model used in \citet{Okounkova_2022}.
+Compared to previous studies, we also performed PE on more events, including events in GWTC-3, the third GW transient catalog.
+We included 69 binary black hole merger events with the lowest false alarm rate $\leq1\mathrm{yr^{-1}}$.
+These events are also listed in TABLE I in \citet{GWTC_3_population}.\footnote{
+GW190720 and GW200129 are included in the list in \citet{GWTC_3_population}, but we did not include them in this study.
 Details in Discussion.}
@@
 % Section guide
-In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, and mention the configuration we used in the PE.
-In Sec.~\ref{sec:Results}, we show the combined result of the PE, and present the constrain on GW amplitude birefringence we obtained.
-In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study, and provide suggestions for future studies.
+In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, mention the configuration we used in the PE and show the method we used to obtain the population constraint on GW amplitude birefringence.
+In Sec.~\ref{sec:Results}, we present the population constraint on GW amplitude birefringence we obtained and show the results of individual PE.
+In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study and provide suggestions for future studies.
 
@@
 \label{sec:Method}
-% Waveform modification
-GW consist of two independent polarization modes similar to electromagnetic waves, which we can choose to represent them in either the linear basis (i.e.
+% GW polarization
+GW consists of two independent polarization modes similar to electromagnetic waves, which we can represent in either the linear basis (i.e.
 $+$ and $\times$) or the circular basis (i.e.
@@
 \end{equation}
-where $h_{\mathrm{L/R}}$ are the amplitude of the left-handed and right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of the plus and cross polarization of the waveform.
+where $h_{\mathrm{L/R}}$ are the amplitude of the left-handed and right-handed polarization of the waveform, $h_+$ and $h_\times$ are the amplitude of the plus and cross polarization of the waveform.
+
+% Waveform modification
+GW amplitude birefringence would enhance one polarization mode over the other.
 According to \textbf{[modification reference]}, we modified the waveform model by
@@
     \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
+    \label{eq:waveform_modification}
 \end{equation}
-where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform as GR predicts, $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the Fourier frequency of the waveform.
-We assume GWs are generated by the binary as GR predicts, as the modification at the source is small before the effect accumulate during the propagation.
+where $h_\mathrm{L/R}^{\mathrm{br}}$ is the amplitude of the modified waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the amplitude of the waveform as GR predicts, $\kappa$ is the dimensionless opacity parameter that represents the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the frequency of the waveform.
+We assume the modification at the source is small, which means the exponent in equation \ref{eq:waveform_modification} has to be small.
+As a result, we treat the binary generates GWs as GR predicts.
+
+Although the modification at the source is small, the effect accumulates as the GW propagates.
 As the GWs propagate, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the GWs.
+According to equation \ref{eq:waveform_modification}, a positive $\kappa$ means the left-handed polarization is enhanced over the right-handed polarization, while a negative $\kappa$ means the opposite.
+Note that when $\kappa=0$, the waveform is the same as GR predicts.
 
-% Effect on inclination PE
+% Effect on inclination PE results
 In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
-This modification changes the ratio of the left-handed and right-handed polarization of the waveform, which affects the PE on the inclination as well.
 The relationship between the amplitude ratio and the inclination is
@@
 where $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
-Therefore, the PE on $\iota$ depends on the amplitude ratio.
+The waveform modification by the birefringence effect changes the ratio of the left-handed and right-handed polarization of the waveform.
+Thus, the birefringence effect will affect the PE result on $\iota$.
 
@@
     \frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}=\frac{\left(1-\cos\iota\right)^2}{\left(1+\cos\iota\right)^2}\exp\left({2\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}}\right)\,,
+    \label{eq:modified_amplitude_ratio}
 \end{equation}
-where $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed Fourier amplitude of left-handed and right-handed polarizations of the GWs.
-As the amplitude ratio does not only depends on $\iota$ with this modification, but also $\kappa$ and $d_C$, the PE on $\iota$ and $d_C$ are affected.
-Note that GR is recovered if $\kappa$ is $0$.
+where $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed amplitude of left-handed and right-handed polarizations of the GWs.
+As the amplitude ratio does not only depends on $\iota$ with this modification but also $\kappa$ and $d_C$, the PE results on $\iota$ and $d_C$ are affected.
 
 In \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
-This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.
+This assumption would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio similarly.
 To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
-As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
+% As a result, degeneracy between $\iota$ and $\kappa$ is formed with frequency independence.
 
-In contrast, We included the frequency dependence in the modification, which is a first-order approximation of the birefringence model.
+In contrast, we included the frequency dependence in the equation \ref{eq:modified_amplitude_ratio}, which is a first-order approximation of the birefringence model.
 The frequency dependence can break the degeneracy between $\kappa$ and $\iota$, as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
@@
     \caption{
-        The observed Fourier amplitude of two polarizations of an simulated GW signal with and without birefringence.
-        The amplitudes as GR predicts are shown in dotted lines with blue and orange colors representing the left-handed and right-handed polarizations.
+        The observed Fourier amplitude of two polarizations of a simulated GW signal with and without birefringence.
+        The amplitudes, as GR predicts, are shown in dotted lines with blue and orange colors representing the left-handed and right-handed polarizations.
         The amplitudes with birefringence are shown in solid lines with blue and orange colors representing the left-handed and right-handed polarizations.
-        This shows the effect of amplitude birefringence with the frequency dependence.
+        This plot shows the effect of amplitude birefringence with frequency dependence.
     }
@@
 % Parameter estimation
-The major objective of this study is to constrain GW amplitude birefringence, which is represented by the opacity parameter $\kappa$.
-To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question \citep{Bilby}.
-\
+To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which can calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question \citep{Bilby}.
 We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using the modified waveforms.
-We use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each individual events.
-We also perform sanity test to check if our PE could recover the PE done by LIGO-Virgo-KAGRA Collaboration (LVK) in \citet{GWTC-2.1, GWTC-3}.
-To implement the sanity check, we need to assume there is no birefringence by setting the prior on $\kappa$ to be a $\delta$ function at $0$, which means GR is recovered.
+We use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each event.
+We also perform a sanity test to check if our PE could recover the PE results done by LIGO-Virgo-KAGRA Collaboration (LVK) in \citet{GWTC-2.1, GWTC-3}.
+To implement the sanity check, we need to assume there is no birefringence by setting the prior on $\kappa$ to be a $\delta$ function at $0$, which force the PE results to be consistent with GR.
 These sanity tests result can be found in the \href{https://zenodo.org/record/7338924}{data release}.
+
+% Bayesian Hierarchical Modeling
+With the PE results of the individual events, we can perform Bayesian hierarchical modeling to constrain the birefringence effect.
+We treated $\kappa$ posteriors from all events as a set of measured $\kappa$ from an underlying distribution.
+The underlying distribution can be assumed to be a Dirac delta function, which means there is a single true value of $\kappa$, or other distributions, which means there is a range of plausible values of $\kappa$.
+In this study, we assume the underlying distribution is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
+\kw{explain why we chose Gaussian}
+With Bayesian hierarchical modeling, which is a method to combine multiple measurements of a random variable by assuming the measurements are independent and identically distributed, we can calculate the posterior of $\mu$ and $\sigma$ from the set of measured $\kappa$ posterior.
+The posterior of $\mu$ and $\sigma$ can be calculated by
+\begin{equation}
+    P(\mu,\sigma|\{d_i\})\propto P(\mu,\sigma)\prod_{i}\int\frac{P(\kappa|d_i)}{P(\kappa)}P(\kappa|\mu,\sigma)\,d\kappa\,,
+    \label{eq:posterior_of_mu_sigma}
+\end{equation}
+where function $P$ is the probability density function, and $\{d_i\}$ is the set of GW event data.
+Note that $P(\kappa)$ is the prior on $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
+We chose the prior of $\mu$ and $\sigma$ being uniform distributions.
+Equation \ref{eq:posterior_of_mu_sigma} can then be simplified to
+\begin{equation}
+    P(\mu,\sigma|\{d_i\})\propto\prod_{i}\int P(\kappa|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa\,.
+\end{equation}
+To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}.
+We then calculate the population constraint on $\kappa$ from the posterior of $\mu$ and $\sigma$.
+The constraint was calculated by
+\begin{equation}
+    P(\kappa|\{d_i\})=\int \mathcal{N}(\mu,\sigma)P(\mu,\sigma|\{d_i\})\,d\mu\,d\sigma\,.
+\end{equation}
 
@@
 \label{sec:Results}
-
-In this section, we will present the results of individual PE first.
-Then, we will show the population constraint we obtained from the PE of all events in GWTC-3.
-At last, we will discuss some of the special events.
-\kw{Looking at the current draft, I do think maybe we should first talk about individual constraints first, then population constraints.}
+In this section, we present the results of our study.
+We first show the PE results of all events and present the results of Bayesian hierarchical modeling.
+Then, we will discuss some special events individually.
 
@@
 % Stacked kappa posterior plot
-% \begin{figure}[h]
-%     \script{kappa_stacked.py}
-%     \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
-%     \caption{
-%         The posterior of $\kappa$ for all 69 events included in this study.
-%         Each color represent a different event.
-%         These posterior are used to calculate the population constraint on $\kappa$ in the following.
-%     }
-%     \label{fig:kappa_stacked}
-% \end{figure}
+In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$ of each event obtained from PE.
+The posteriors of $\kappa$ for most events are concentrated around $0$, which means the PE results are consistent with GR.
+With Bayesian hierarchical modeling, these posteriors are used to calculate the population constraint on $\kappa$.
 
-% How to calculate the combined posterior
-In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$ of each events obtained from PE.
-These individual posteriors can then be treated as a set of measured $\kappa$ posterior.
-We can either assume that $\kappa$ is a single value constant, or that $\kappa$ is a random variable with a probability distribution.
-In this study, we assume $\kappa$ is a random variable, and the population distribution of $\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
-With Bayesian hierarchical modelling, which is a method to combine multiple measurements of a random variable by assuming the measurements are independent and identically distributed, we can calculate the posterior of $\mu$ and $\sigma$ from the set of measured $\kappa$ posterior.
-The posterior of $\mu$ and $\sigma$ can be calculated by
-\begin{equation}
-    P(\mu,\sigma|\{d_i\})\propto P(\mu,\sigma)\prod_{i}\int\frac{P(\kappa|d_i)}{P(\kappa)}P(\kappa|\mu,\sigma)\,d\kappa\,,
-\end{equation}
-where function $P$ is the probability density function, and $\{d_i\}$ is the set of GW event data.
-To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}, with prior of $\mu$ and $\sigma$ being uniform distributions.
+\begin{figure}[h]
+    \script{kappa_stacked.py}
+    \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
+    \caption{
+        The posterior of $\kappa$ for all 69 events included in this study.
+        Each color represents a different event.
+    }
+    \label{fig:kappa_stacked}
+\end{figure}
+
+% Corner plot of the mean and standard deviation of kappa
 The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
 The mean of $\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.
+The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.
 
-% Corner plot of mean and standard deviation of kappa
-% \begin{figure}[h]
-%     \script{corner_Gaussian.py}
-%     \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
-%     \caption{
-%         The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
-%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-%         Orange line represent the median of $\mu$ and $\sigma$.
-%     }
-%     \label{fig:corner_Gaussian}
-% \end{figure}
+\begin{figure}[h]
+    \script{corner_Gaussian.py}
+    \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
+    \caption{
+        The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
+        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        The orange line represents the median of $\mu$ and $\sigma$.
+        The plot shows that the population constraint on $\kappa$ is consistent with GR.
+    }
+    \label{fig:corner_Gaussian}
+\end{figure}
 
 % Constraint on $\kappa$
-We then calculate the constraint on $\kappa$ from the posterior of $\mu$ and $\sigma$.
-We generated a Gaussian distribution for each sample of $\mu$ and $\sigma$.
-The Gaussian distributions were then added up to obtain the population distribution of $\kappa$.
+With the posteriors of $\mu$ and $\sigma$, the population constraint on $\kappa$ was calculated.
 The mean of the population distribution of $\kappa$ is \variable{output/kappa_mean.txt}.
-Note that GR is recovered if $\kappa$ is $0$.
+Note that the result is consistent with GR when $\kappa=0$.
 
@@
 In \citet{Okounkova_2022}, the authors gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
-We convert the constraint to the same format as ours to make comparison.
+We convert the constraint to the same format as ours to make a comparison.
 They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
-We were able to obtain a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
-This is a order of magnitude improvement in constraining $\kappa$.
-This is due to the fact that we have much more events from GWTC-3 compared to GWTC-2.
+We obtained a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
+Our result is an order of magnitude improvement in constraining $\kappa$.
+The main reason is that we have more events from GWTC-3 than GWTC-2.
 
 % Violin plot
-In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each events in the form of violin plot.
-There are some events with $\kappa$ distribution deviate from $\kappa=0$.
+In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each event in the form of a violin plot.
+Some events with $\kappa$ distribution deviate from $\kappa=0$.
 Also, there are some events with bimodal $\kappa$ distribution.
@@
 
-% \begin{figure*}[ht]
-%     \script{violin_kappa.py}
-%     \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
-%     \caption{
-%         The violin plot shows the posterior of $\kappa$ for all 69 events included in this study.
-%         Each violin represents a different event.
-%         The violins are sorted by the quotient of median and standard deviation of the posterior.
-%         Blue horizontal line represent the median value of $\mu$, and blue region represent 1 $\sigma$ confidence level.
-%         Orange horizontal line represent $\kappa=0$.
-%         This plot shows that there are some events with $\kappa$ distribution significantly deviate from $0$.
-%     }
-%     \label{fig:violin_kappa}
-% \end{figure*}
+\begin{figure*}[ht]
+    \script{violin_kappa.py}
+    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
+    \caption{
+        The violin plot shows the posterior of $\kappa$ for all 69 events in this study.
+        Each violin represents a different event.
+        The violins are sorted by the quotient of the median and standard deviation of the posterior.
+        The blue horizontal line represents the median value of $\mu$, and the blue region represents 1 $\sigma$ confidence level.
+        The Orange horizontal line represents $\kappa=0$.
+        This plot shows that some events with $\kappa$ distribution significantly deviate from $0$.
+    }
+    \label{fig:violin_kappa}
+\end{figure*}
 
@@
 \subsection{Result on GW150914}
-In this study, we included frequency dependence of birefringence.
-This would affect the posterior of $\kappa$ obtained from PE.
+In this study, we included frequency dependence of birefringence, which would affect the posterior of $\kappa$ obtained from PE.
 Consider GW150914, the first GW detected by LIGO, as an example.
 In figure \ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
-With the frequency independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
+With the frequency-independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
 This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
-To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
-As a result, degeneracy between $\iota$ and $\kappa$ is formed with the frequency independence.
 \kw{you should say why the middle part is constrained to be small.}
@@
 On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
-This is because the degeneracy was broken by the frequency dependence, as the effect of birefringence will be different from the one of changing $\iota$.
-In this case, the posteriors can recover the GR distributions for the binary parameters, and the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent, and GR can be recovered.
+This is because the frequency dependence broke the degeneracy, as the effect of birefringence will differ from the effect of changing $\iota$.
+In this case, the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent.
+Therefore, the PE result with frequency dependence is consistent with GR.
 
-% \begin{figure}[h]
-%     \script{corner_GW150914.py}
-%     \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
-%     \caption{
-%         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-%         Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with both frequency independent and dependent birefringence respectively.
-%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-%         Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
-%         This plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
-%     }
-%     \label{fig:corner_GW150914}
-% \end{figure}
+\begin{figure}[h]
+    \script{corner_GW150914.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
+    \caption{
+        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
+        Colors in the plot represent the PE result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE results done by us with both frequency independent and dependent birefringence respectively.
+        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        Note that there is no posterior of $\kappa$ for the PE result from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
+        This plot shows that frequency-independent birefringence creates a degeneracy between $\kappa$ and $\iota$, while the frequency-dependent birefringence can break it.
+    }
+    \label{fig:corner_GW150914}
+\end{figure}
 
@@
 \subsection{Result on GW190521}
-Even with the frequency dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
+Even with the frequency-dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
 Consider GW190521, the most massive binary black hole merger in the events we included.
-The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency dependent birefringence model, as shown in figure \ref{fig:corner_GW190521}.
-This is because the frequency range of the signal is much smaller due to the large mass of the binary.
-The effect of birefringence at different frequency within the range is similar, thus the degeneracy cannot be broken.
-This can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
+The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in figure \ref{fig:corner_GW190521}.
+The reason could be this GW event's frequency range is too narrow to break the degeneracy.
+The effect of birefringence at different frequencies within the range is similar, so the degeneracy cannot be broken.
+Therefore, the narrow frequency range can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
 
-% \begin{figure}[h]
-%     \script{corner_GW190521.py}
-%     \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
-%     \caption{
-%         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
-%         Colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
-%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-%         This plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
-%     }
-%     \label{fig:corner_GW190521}
-% \end{figure}
+\begin{figure}[h]
+    \script{corner_GW190521.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
+    \caption{
+        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
+        Colors in the plot are the PE result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE result done by us with the frequency-dependent birefringence.
+        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
+    }
+    \label{fig:corner_GW190521}
+\end{figure}
 
@@
 For GW190720, we could not perform PE with Bilby successfully.
-In LVK PE papers \citep{GWTC-2.1, GWTC-3}, the PE of some events were performed with LALInference \citep{lalsuite} instead of Bilby, GW190720 was one of them.
-We successfully migrate LALInference configuration to Bilby for most of these events and recover similar PE results, except for GW190720.
+In LVK PE papers \citep{GWTC-2.1, GWTC-3}, the PE of some events was performed with LALInference \citep{lalsuite} instead of Bilby.
+GW190720 was one of them.
+We successfully migrated the LALInference configuration to Bilby for most of these events and recovered similar PE results, except for GW190720.
 Bilby could not evaluate the likelihood function on Virgo data.
-This could bias the population posterior.
+As a result, we excluded this event from our analysis.
 
 % Special event 2: GW200129
-For GW200129, past research suggest there could be a potential glitch in Virgo data. \citep{GW200129}
-We perform 3 sets of PE to check if the glitch would affect the population posterior of $\kappa$.
-In each set, we pick only 2 out of 3 detectors data to perform PE.
-The 2 sets of PE results with Virgo data return a $\kappa$ posterior which strongly suggest a negative $\kappa$.
-We removed this event from our analysis.
+For GW200129, past research suggests a potential glitch in Virgo data. \citep{GW200129}
+We perform three sets of PE to check if the glitch would affect the population posterior of $\kappa$.
+We pick only two out of three detector data in each set to perform PE.
+The two sets of PE results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
 Further investigation is needed to understand the glitch.
+Thus, we did not include this event in our analysis.
 
@@
 % BNS
-A future work is to perform PE on binary neutron star mergers, such as GW170817.
-The frequency range of the signal is much larger compared to the binary black hole mergers, thus the difference of the effect of birefringence at different frequency within the range can be more significant.
-This can allow us to further constrain the birefringence effect, and the beyond-GR theories that predict it.
+Future work is to perform PE on binary neutron star mergers, such as GW170817.
+The frequency range of the signal is much wider compared to the binary black hole mergers.
+Thus, the difference in the effect of birefringence at different frequencies within the range can be more significant.
+The PE on binary neutron star mergers can allow us to further constrain the birefringence effect and the beyond-GR theories that predict it.
 However, performing PE on binary neutron star mergers requires much more computational resources.
@@
 % More observations with higher SNR
-Another future work is to apply the same method to more events, and events with higher signal-to-noise ratio (SNR).
+Another future work is to apply the same method to more GW events and events with a higher signal-to-noise ratio (SNR).
 LVK will release more events with higher SNR in the future.
-Using data with higher SNR allows us to obtain more precise PE results, which also allows us to constrain the birefringence effect more precisely.
+Using data with higher SNR allows us to obtain more precise PE results and constrain the birefringence effect more precisely.
 And using data from more events will allow us to calculate a more constrained population posterior of $\kappa$.
@@
 \label{sec:Acknowledgements}
-M.~I., K.~W.~K.~W.~ and W.~M.~F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute, which is supported by the Simons Foundation.
-The computational resources used in this work were provided by the Flatiron Institute.
+M.~I., K.~W.~K.~W.~, and W.~M.~F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute.
+The Flatiron Institute provided the computational resources used in this work.
 

### 2022-11-30 src/scripts/birefringence.py

@@ -49,13 +49,13 @@ freq = np.arange(len(hl))*df
 
 kappa = -0.5
 
-hl_bi_L = hl * np.exp(kappa*(dist_mpc/1000)*(freq/100))
-hr_bi_L = hr * np.exp(-kappa*(dist_mpc/1000)*(freq/100))
+hl_bi_with_freq = hl * np.exp(kappa*(dist_mpc/1000)*(freq/100))
+hr_bi_with_freq = hr * np.exp(-kappa*(dist_mpc/1000)*(freq/100))
 
-plt.loglog(freq, abs(hl), ls='--', color=sns.color_palette()[0], label=r"L (GR)")
-plt.loglog(freq, abs(hr), ls='--', color=sns.color_palette()[3], label=r"R (GR)")
-plt.loglog(freq, abs(hl_bi_L), color=sns.color_palette()[0], label=r"L (birefringence)")
-plt.loglog(freq, abs(hr_bi_L), color=sns.color_palette()[3], label=r"R (birefringence)")
+plt.loglog(freq, abs(hl), ls=':', color=sns.color_palette()[0], label=r"L (GR)")
+plt.loglog(freq, abs(hr), ls=':', color=sns.color_palette()[3], label=r"R (GR)")
+plt.loglog(freq, abs(hl_bi_with_freq), color=sns.color_palette()[0], label=r"L (birefringence)")
+plt.loglog(freq, abs(hr_bi_with_freq), color=sns.color_palette()[3], label=r"R (birefringence)")
 plt.xlim(f_min+1,f_max-10)
 plt.ylim(1e-38,1e-20)
 plt.legend()

### 2022-12-03 src/tex/ms.tex
Update content

@@
 \usepackage{amsfonts,amssymb,amsmath}
-\usepackage{hyperref}
-
-\definecolor{rb4}{HTML}{27408B}
-\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
 
@@
 % Motivation
-Since Einstein proposed his theory of general relativity (GR), it has been tested in various length scales.
-After a century, we know that GR does not agree with quantum theories at some length scale.
-To unify both theories, we need to study the possibility of different beyond-GR theories.
-\kw{I think now it kind of goes into birefringence too fast.
-Maybe talk about what effect one may expect to see in GW, in general, a bit?
-Say there is polarization modification, phasing modification, etc.}
-Some beyond-GR theories, such as Chern-Simons gravity, suggest that there is gravitational wave (GW) amplitude birefringence, while GR predicts no birefringence.
-GW amplitude birefringence is a property of space-time that consists of the enhancement of one GW polarization over the other during wave propagation.
+Gravitational waves (GWs) have been used to test Einstein's theory of general relativity (GR) since the first detection in 2015.
+To study the possibility of different beyond-GR theories, we can observe if GWs have any deviation from GR.
+Some beyond-GR theories, such as Chern-Simons gravity, suggest that there is gravitational wave amplitude birefringence, while GR predicts no birefringence.
+GW amplitude birefringence is a property of space-time that consists of the enhancement of one GW polarization over the other as the wave propagates.
 
@@
 Previous studies have constrained amplitude birefringence by performing different statistical analyses.
-In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and used the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
-In \citet{Yamada_2020} and \citet{Wang_2021}, they both performed parameter estimation (PE) on the events in GWTC-1, the first GW transient catalog.
+% cite ligo, virgo
+\citet{Yamada_2020} and \citet{Wang_2021} both performed parameter estimation (PE) on the events in GWTC-1, the first GW transient catalog, using birefringence templates.
+% cite GWTC-1
+\citet{Okounkova_2022} considered the distribution of observed inclinations of the GW events in GWTC-2, the second GW transient catalog, to look for signs of birefringence.
+% cite GWTC-2
 
 % What's new?
-In this study, we used a birefringence model with higher-order terms, including amplitude birefringence's frequency dependence.
-This model is more realistic than the frequency-independent model used in \citet{Okounkova_2022}.
+In this study, we used a frequency-dependent birefringence model to constrain the strength of GW amplitude birefringence.
+This model is a better approximation of the birefringence effect than the frequency-independent model used in \citet{Okounkova_2022}.
 Compared to previous studies, we also performed PE on more events, including events in GWTC-3, the third GW transient catalog.
-We included 69 binary black hole merger events with the lowest false alarm rate $\leq1\mathrm{yr^{-1}}$.
+We consider 70 binary black hole merger events with a false alarm rate $\leq1\mathrm{yr^{-1}}$.
 These events are also listed in TABLE I in \citet{GWTC_3_population}.\footnote{
 GW190720 and GW200129 are included in the list in \citet{GWTC_3_population}, but we did not include them in this study.
+% remove GW200129 from the footnote
 Details in Discussion.}
+% what we will do with the results
 
@@
 \label{sec:Method}
+
+\subsection{Waveform modification}
 % GW polarization
-GW consists of two independent polarization modes similar to electromagnetic waves, which we can represent in either the linear basis (i.e.
-$+$ and $\times$) or the circular basis (i.e.
-left-handed and right-handed).
+GWs comprise two independent polarization modes, just like electromagnetic waves, which are usually represented in the linear states (i.e. $+$ and $\times$).
+For a given Fourier mode, they can be combined into left and right circular states (i.e. left-handed and right-handed).
 They are related by
@@
 GW amplitude birefringence would enhance one polarization mode over the other.
-According to \textbf{[modification reference]}, we modified the waveform model by
+In theories like Chern- Simons gravity, the observed Fourier waveform takes the form of
 \begin{equation}
@@
 \end{equation}
-where $h_\mathrm{L/R}^{\mathrm{br}}$ is the amplitude of the modified waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the amplitude of the waveform as GR predicts, $\kappa$ is the dimensionless opacity parameter that represents the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the frequency of the waveform.
-We assume the modification at the source is small, which means the exponent in equation \ref{eq:waveform_modification} has to be small.
-As a result, we treat the binary generates GWs as GR predicts.
+where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the observed waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform emitted, $\kappa$ is the dimensionless opacity parameter that represents the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the frequency of the waveform.
+We assume the intrinsic modification is small, so we can use GR waveform templates to describe the emitted waveform.
 
-Although the modification at the source is small, the effect accumulates as the GW propagates.
+Although the intrinsic modification is small, the effect accumulates as the GW propagates.
 As the GWs propagate, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the GWs.
@@
 
+\subsection{Effect on inclination}
 % Effect on inclination PE results
 In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
+% For a nonprecessing binary, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination. dominated by quadrupolar mode radiation. 2-2 mode
 The relationship between the amplitude ratio and the inclination is
@@
 The waveform modification by the birefringence effect changes the ratio of the left-handed and right-handed polarization of the waveform.
+% birfringence can alter the ratio of the polarization of the waveform
 Thus, the birefringence effect will affect the PE result on $\iota$.
+% will -> could/ without freq the effect will affect
+%base on masha
 
-With the modification we implemented, the amplitude ratio of the left-handed and right-handed polarizations becomes 
+Based on the modification we implemented, the amplitude ratio of the left-handed and right-handed polarizations would be given by
 \begin{equation}
@@
 
-In \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+\citet{Okounkova_2022} assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
 This assumption would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio similarly.
 To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
-% As a result, degeneracy between $\iota$ and $\kappa$ is formed with frequency independence.
 
@@
 
+\subsection{Parameter estimation}
 % Parameter estimation
 To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which can calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question \citep{Bilby}.
-We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using the modified waveforms.
+We modified Bilby to include the birefringence effect in the likelihood function.
+We then perform PE on intrinsic and extrinsic parameters including $\kappa$.
+We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
 We use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each event.
+% cite GWOSC
 We also perform a sanity test to check if our PE could recover the PE results done by LIGO-Virgo-KAGRA Collaboration (LVK) in \citet{GWTC-2.1, GWTC-3}.
-To implement the sanity check, we need to assume there is no birefringence by setting the prior on $\kappa$ to be a $\delta$ function at $0$, which force the PE results to be consistent with GR.
+To implement the sanity check, we need to assume there is no birefringence by setting $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
 These sanity tests result can be found in the \href{https://zenodo.org/record/7338924}{data release}.
+% cite the zenodo data release in bib
 
+\subsection{Bayesian hierarchical modeling}
 % Bayesian Hierarchical Modeling
+
+%cite Max's paper
+
+% under XXX, kappa should share the same value
+% to understand more, the true value of kappa is drawn from a distribution
+% the distribution should be consistent with a delta function
+% if it is not a delta function, there should be more study (model not fully correct, new physics, or systematic error)
+
+% we expect the kappa to be a single value for all events at the beginning
 With the PE results of the individual events, we can perform Bayesian hierarchical modeling to constrain the birefringence effect.
 We treated $\kappa$ posteriors from all events as a set of measured $\kappa$ from an underlying distribution.
+% instead of assuming all event share one value, we assume the kappa is a true value from a distribution.
 The underlying distribution can be assumed to be a Dirac delta function, which means there is a single true value of $\kappa$, or other distributions, which means there is a range of plausible values of $\kappa$.
+% we expect the delta should be 
 In this study, we assume the underlying distribution is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
-\kw{explain why we chose Gaussian}
+% Gaussian max entropy, whatever the true distribution is, the mean and the standard deviation will be true.
 With Bayesian hierarchical modeling, which is a method to combine multiple measurements of a random variable by assuming the measurements are independent and identically distributed, we can calculate the posterior of $\mu$ and $\sigma$ from the set of measured $\kappa$ posterior.
@@
 \begin{equation}
-    P(\mu,\sigma|\{d_i\})\propto P(\mu,\sigma)\prod_{i}\int\frac{P(\kappa|d_i)}{P(\kappa)}P(\kappa|\mu,\sigma)\,d\kappa\,,
+    P(\mu,\sigma|\{d_i\})\propto P(\mu,\sigma)\prod_{i}\int\frac{P(\kappa_i|d_i)}{P(\kappa_i}P(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
     \label{eq:posterior_of_mu_sigma}
@@
 \begin{equation}
-    P(\mu,\sigma|\{d_i\})\propto\prod_{i}\int P(\kappa|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa\,.
+    P(\mu,\sigma|\{d_i\})\propto\prod_{i}\int P(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
 \end{equation}
@@
 \end{equation}
+% np.choose and weight
 
@@
 \subsection{Result on GWTC-3}
-% Stacked kappa posterior plot
-In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$ of each event obtained from PE.
+% Violin plot
+In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each event in the form of a violin plot.
 The posteriors of $\kappa$ for most events are concentrated around $0$, which means the PE results are consistent with GR.
-With Bayesian hierarchical modeling, these posteriors are used to calculate the population constraint on $\kappa$.
+Some events with $\kappa$ distribution deviate from $\kappa=0$.
+Also, there are some events with bimodal $\kappa$ distribution.
+We are going to discuss some of these events in the following.
+% mention about GW200129
 
-\begin{figure}[h]
-    \script{kappa_stacked.py}
-    \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
+\begin{figure*}[ht]
+    \script{violin_kappa.py}
+    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
     \caption{
-        The posterior of $\kappa$ for all 69 events included in this study.
-        Each color represents a different event.
+        The violin plot shows the posterior of $\kappa$ for all 69 events in this study.
+        Each violin represents a different event.
+        The violins are sorted by the quotient of the median and standard deviation of the posterior.
+        The blue horizontal line represents the median value of $\mu$, and the blue region represents 1 $\sigma$ confidence level.
+        The Orange horizontal line represents $\kappa=0$.
+        This plot shows that most events are consistent with GR.
     }
-    \label{fig:kappa_stacked}
-\end{figure}
+    \label{fig:violin_kappa}
+\end{figure*}
+% order time, color median/std
+% symmetrical credible interval 5 to 95 percentile (np.percentile)
 
+\subsection{Bayesian hierarchical modeling}
 % Corner plot of the mean and standard deviation of kappa
@@
 
+\subsection{Constaint on $\kappa$}
 % Constraint on $\kappa$
@@
 Note that the result is consistent with GR when $\kappa=0$.
-
-% Comparison with previous studies
-In \citet{Okounkova_2022}, the authors gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
-We convert the constraint to the same format as ours to make a comparison.
-They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
-We obtained a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
-Our result is an order of magnitude improvement in constraining $\kappa$.
-The main reason is that we have more events from GWTC-3 than GWTC-2.
-
-% Violin plot
-In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each event in the form of a violin plot.
-Some events with $\kappa$ distribution deviate from $\kappa=0$.
-Also, there are some events with bimodal $\kappa$ distribution.
-We are going to discuss some of these events in the following.
-
-\begin{figure*}[ht]
-    \script{violin_kappa.py}
-    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
-    \caption{
-        The violin plot shows the posterior of $\kappa$ for all 69 events in this study.
-        Each violin represents a different event.
-        The violins are sorted by the quotient of the median and standard deviation of the posterior.
-        The blue horizontal line represents the median value of $\mu$, and the blue region represents 1 $\sigma$ confidence level.
-        The Orange horizontal line represents $\kappa=0$.
-        This plot shows that some events with $\kappa$ distribution significantly deviate from $0$.
-    }
-    \label{fig:violin_kappa}
-\end{figure*}
-
-% Trend (mass & distance)
-
-% Constraining power
+% include both Hierarchical and stacking
 
@@
 This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
-\kw{you should say why the middle part is constrained to be small.}
+% explain the kappa gap
 
@@
 
+% Comparison with previous studies\
+\citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
+We convert the constraint to the same format as ours to make a comparison.
+They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
+We obtained a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
+Our result is an order of magnitude improvement in constraining $\kappa$.
+The main reason is that we have more events from GWTC-3 than GWTC-2.
+
 % Reasons for special events
 \subsection{Special Events}
+
 % Special event 1: GW190720
@@
 Thus, we did not include this event in our analysis.
+% show result included it first
 
@@
 
+\pagebreak
 \bibliography{bib}

### 2022-12-03 src/scripts/kappa_stacked.py

@@ -1,27 +1,0 @@
-import matplotlib.pyplot as plt
-import pandas as pd
-import seaborn as sns
-import paths
-
-sns.set_theme(palette='colorblind', font_scale=1.2)
-
-plt.rcParams.update({
-    "text.usetex": True,
-    "font.family": "serif",
-    "font.serif": ["Computer Modern Roman"],
-})
-
-result_dict = {}
-result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_DataFrame = result_DataFrame[result_DataFrame.event != "GW200129_065458"] # remove GW200129
-for event in result_DataFrame['event'].unique():
-    result_dict[event] = result_DataFrame[result_DataFrame.event == event]
-
-for event in result_DataFrame['event'].unique():
-    g = sns.kdeplot(result_dict[event]['kappa'], fill=True)
-
-g.set_xlabel("$\kappa$")
-g.set_ylabel("$p(\kappa|d)$")
-g.set_xlim(-1,1)
-
-g.figure.savefig(fname=paths.figures/"kappa_stacked.pdf", bbox_inches="tight", dpi=300)

### 2022-12-04 src/tex/ms.tex
Update content

@@
 % Motivation
-Gravitational waves (GWs) have been used to test Einstein's theory of general relativity (GR) since the first detection in 2015.
+Gravitational waves (GWs) have been used to test Einstein's theory of general relativity (GR) with detections by LIGO-Virgo-KAGRA Collaboration (LVK) \citep{LIGO, Virgo, KAGRA}.
 To study the possibility of different beyond-GR theories, we can observe if GWs have any deviation from GR.
@@
 Previous studies have constrained amplitude birefringence by performing different statistical analyses.
-% cite ligo, virgo
-\citet{Yamada_2020} and \citet{Wang_2021} both performed parameter estimation (PE) on the events in GWTC-1, the first GW transient catalog, using birefringence templates.
-% cite GWTC-1
-\citet{Okounkova_2022} considered the distribution of observed inclinations of the GW events in GWTC-2, the second GW transient catalog, to look for signs of birefringence.
-% cite GWTC-2
+\citet{Yamada_2020} and \citet{Wang_2021} both performed parameter estimation (PE) on the events in the first GW transient catalog \citep{GWTC-1}, GWTC-1, using birefringence templates.
+\citet{Okounkova_2022} considered the distribution of observed inclinations of the GW events in the second GW transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.
 
 % What's new?
-In this study, we used a frequency-dependent birefringence model to constrain the strength of GW amplitude birefringence.
+In this study, we used a frequency-dependent birefringence model to constrain the strength of GW amplitude birefringence by performing PE.
 This model is a better approximation of the birefringence effect than the frequency-independent model used in \citet{Okounkova_2022}.
-Compared to previous studies, we also performed PE on more events, including events in GWTC-3, the third GW transient catalog.
+Compared to previous studies, we also performed PE on more events, including events in the third GW transient catalog \citep{GWTC-3}, GWTC-3.
 We consider 70 binary black hole merger events with a false alarm rate $\leq1\mathrm{yr^{-1}}$.
-These events are also listed in TABLE I in \citet{GWTC_3_population}.\footnote{
-GW190720 and GW200129 are included in the list in \citet{GWTC_3_population}, but we did not include them in this study.
-% remove GW200129 from the footnote
+These events are also listed in TABLE I in \citet{GWTC-3_population}.\footnote{
+GW190720 is included in the list in \citet{GWTC-3_population}, but we did not include it in this study.
 Details in Discussion.}
-% what we will do with the results
+We then give the population constraint on the strength of GW amplitude birefringence by using the PE results.
 
@@
 % Effect on inclination PE results
-In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
-% For a nonprecessing binary, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination. dominated by quadrupolar mode radiation. 2-2 mode
+In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for the dominant $(2,\pm 2)$ angular mode of nonprecessing binaries.
 The relationship between the amplitude ratio and the inclination is
@@
 where $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
-The waveform modification by the birefringence effect changes the ratio of the left-handed and right-handed polarization of the waveform.
-% birfringence can alter the ratio of the polarization of the waveform
-Thus, the birefringence effect will affect the PE result on $\iota$.
-% will -> could/ without freq the effect will affect
-%base on masha
+Birefringence can alter the ratio between polarizations of the waveform.
+Thus, the birefringence effect could affect the PE result on $\iota$ based on \citet{Okounkova_2022}.
 
@@
 % Parameter estimation
-To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which can calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question \citep{Bilby}.
+To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby \citep{Bilby}, a Bayesian toolkit for GW data analysis which can calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question.
 We modified Bilby to include the birefringence effect in the likelihood function.
-We then perform PE on intrinsic and extrinsic parameters including $\kappa$.
-We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
-We use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each event.
-% cite GWOSC
-We also perform a sanity test to check if our PE could recover the PE results done by LIGO-Virgo-KAGRA Collaboration (LVK) in \citet{GWTC-2.1, GWTC-3}.
+We then perform PE on intrinsic and extrinsic parameters including $\kappa$ with open data from the Gravitational Wave Open Science Center \citep{GWOSC}.
+For PE configuration, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each event and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+We also perform a sanity test to check if our PE could recover the PE results done by LVK in \citet{GWTC-2.1, GWTC-3}.
 To implement the sanity check, we need to assume there is no birefringence by setting $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
-These sanity tests result can be found in the \href{https://zenodo.org/record/7338924}{data release}.
-% cite the zenodo data release in bib
+These sanity tests result can be found in \citet{dataset}.
 
@@
 % Special event 2: GW200129
-For GW200129, past research suggests a potential glitch in Virgo data. \citep{GW200129}
+For GW200129, past research suggests a potential glitch in Virgo data. \citep{GW200129_glitch}
 We perform three sets of PE to check if the glitch would affect the population posterior of $\kappa$.
@@
 
+This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gwosc.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
+LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
+Additional support for Advanced LIGO was provided by the Australian Research Council.
+Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
+KAGRA is supported by Ministry of Education, Culture, Sports, Science and Technology (MEXT), Japan Society for the Promotion of Science (JSPS) in Japan; National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea; Academia Sinica (AS) and National Science and Technology Council (NSTC) in Taiwan.
+
 \pagebreak

### 2022-12-08 src/tex/ms.tex
Updated content

@@
 For PE configuration, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each event and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
-We also perform a sanity test to check if our PE could recover the PE results done by LVK in \citet{GWTC-2.1, GWTC-3}.
-To implement the sanity check, we need to assume there is no birefringence by setting $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
-These sanity tests result can be found in \citet{dataset}.
+The PE results are published in \citet{dataset}.
+
+We assume $\kappa$ is the same for all events and calculate a posterior distribution of $\kappa$.
+The posterior distribution is given by
+\begin{equation}
+    p(\kappa|\{d\})\propto \prod{i}p(\kappa|d_i)\,,
+    \label{eq:restricted_posterior}
+\end{equation}
+where $p$ is the probability density function, $i$ is the index of the event, $\{d\}$ is the set of data including all events, and $d_i$ is the data of a single event.
+Equation \ref{eq:restricted_posterior} shows that the posterior distribution of $\kappa$ is the product of the posterior distribution of $\kappa$ for each event.
+This posterior distribution is a restricted posterior distribution because we assume $\kappa$ is the same for all events.
+
+Note that we also perform a sanity test to check if our PE could recover the PE results done by LVK in \citet{GWTC-2.1, GWTC-3}.
+To implement the sanity check, we set $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
+These sanity test results are also attached in \citet{dataset}.
 
@@
 
-%cite Max's paper
+We then perform Bayesian hierarchical modeling to further analyze the constraint on $\kappa$.
+Bayesian hierarchical modeling is a method to check if our waveform model can fully describe the amplitude birefringence effect.
+Here, instead of assuming all events share the same value of $\kappa$, we assume $\kappa$ from each event is a true value drawn from a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
+We chose a Gaussian distribution as the underlying distribution of $\kappa$ because it captures the correct mean and the standard deviation of the distribution no matter what the underlying distribution is.
 
-% under XXX, kappa should share the same value
-% to understand more, the true value of kappa is drawn from a distribution
-% the distribution should be consistent with a delta function
-% if it is not a delta function, there should be more study (model not fully correct, new physics, or systematic error)
-
-% we expect the kappa to be a single value for all events at the beginning
-With the PE results of the individual events, we can perform Bayesian hierarchical modeling to constrain the birefringence effect.
-We treated $\kappa$ posteriors from all events as a set of measured $\kappa$ from an underlying distribution.
-% instead of assuming all event share one value, we assume the kappa is a true value from a distribution.
-The underlying distribution can be assumed to be a Dirac delta function, which means there is a single true value of $\kappa$, or other distributions, which means there is a range of plausible values of $\kappa$.
-% we expect the delta should be 
-In this study, we assume the underlying distribution is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
-% Gaussian max entropy, whatever the true distribution is, the mean and the standard deviation will be true.
-With Bayesian hierarchical modeling, which is a method to combine multiple measurements of a random variable by assuming the measurements are independent and identically distributed, we can calculate the posterior of $\mu$ and $\sigma$ from the set of measured $\kappa$ posterior.
-The posterior of $\mu$ and $\sigma$ can be calculated by
+With the posterior of $\kappa$ from PE on each event, the posterior of $\mu$ and $\sigma$ can be calculated by
 \begin{equation}
-    P(\mu,\sigma|\{d_i\})\propto P(\mu,\sigma)\prod_{i}\int\frac{P(\kappa_i|d_i)}{P(\kappa_i}P(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
+    p(\mu,\sigma|\{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i|d_i)}{p(\kappa_i}p(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
     \label{eq:posterior_of_mu_sigma}
 \end{equation}
-where function $P$ is the probability density function, and $\{d_i\}$ is the set of GW event data.
-Note that $P(\kappa)$ is the prior on $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
+Note that $p(\kappa)$ is the prior of $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
 We chose the prior of $\mu$ and $\sigma$ being uniform distributions.
@@
 \begin{equation}
-    P(\mu,\sigma|\{d_i\})\propto\prod_{i}\int P(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
+    p(\mu,\sigma|\{d_i\})\propto\prod_{i}\int p(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
 \end{equation}
-To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}.
-We then calculate the population constraint on $\kappa$ from the posterior of $\mu$ and $\sigma$.
+To sample the posterior distribution of $\mu$ and $\sigma$, we used a sampling package, flowMC \citep{flowMC}.
+
+If our model can fully describe the birefringence effect, the true value of $\kappa$ should be a single value, and the Gaussian distribution should have a standard deviation of $0$.
+If the standard deviation obtained is not $0$, it means our model cannot fully describe the birefringence effect.
+It could be due to our model does not fully capture the physics of birefringence, or there is a systematic error in the data.
+
+We then calculate the generic posterior distribution of $\kappa$ from the posterior of $\mu$ and $\sigma$.
 The constraint was calculated by
 \begin{equation}
-    P(\kappa|\{d_i\})=\int \mathcal{N}(\mu,\sigma)P(\mu,\sigma|\{d_i\})\,d\mu\,d\sigma\,.
+    p(\kappa|\{d_i\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma|\{d_i\})\,d\mu\,d\sigma\,.
 \end{equation}
-% np.choose and weight
 
@@
 Then, we will discuss some special events individually.
+
+% np.choose and weight for drawing samples
 

### 2022-12-08 src/tex/ms.tex
Update content

@@
 GW190720 is included in the list in \citet{GWTC-3_population}, but we did not include it in this study.
-Details in Discussion.}
+Details in Sec.~\ref{sec:GW190720}.}
 We then give the population constraint on the strength of GW amplitude birefringence by using the PE results.
@@
 \begin{equation}
-    p(\kappa|\{d\})\propto \prod{i}p(\kappa|d_i)\,,
+    p(\kappa|\{d\})\propto \prod_{i}p(\kappa|d_i)\,,
     \label{eq:restricted_posterior}
@@
 \begin{equation}
-    p(\mu,\sigma|\{d_i\})\propto\prod_{i}\int p(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
+    p(\mu,\sigma|\{d\})\propto\prod_{i}\int p(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
 \end{equation}
@@
 \begin{equation}
-    p(\kappa|\{d_i\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma|\{d_i\})\,d\mu\,d\sigma\,.
+    p(\kappa|\{d\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma|\{d\})\,d\mu\,d\sigma\,.
+    \label{eq:generic_posterior}
 \end{equation}
@@
 
-% np.choose and weight for drawing samples
-
 \subsection{Result on GWTC-3}
@@
 In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each event in the form of a violin plot.
-The posteriors of $\kappa$ for most events are concentrated around $0$, which means the PE results are consistent with GR.
-Some events with $\kappa$ distribution deviate from $\kappa=0$.
-Also, there are some events with bimodal $\kappa$ distribution.
-We are going to discuss some of these events in the following.
-% mention about GW200129
+The median and confidence interval of $\kappa$ are calculated with the restricted posterior distribution of $\kappa$, which is given by equation \ref{eq:restricted_posterior}.
+We excluded GW200129 from the calculation of the median and confidence interval of $\kappa$.
+Details of the exclusion are discussed in Sec.~\ref{sec:GW200129}.
 
@@
     \caption{
-        The violin plot shows the posterior of $\kappa$ for all 69 events in this study.
+        The violin plot shows the posterior of $\kappa$ for 69 events in this study.
         Each violin represents a different event.
-        The violins are sorted by the quotient of the median and standard deviation of the posterior.
-        The blue horizontal line represents the median value of $\mu$, and the blue region represents 1 $\sigma$ confidence level.
-        The Orange horizontal line represents $\kappa=0$.
-        This plot shows that most events are consistent with GR.
+        The color of the violins represents the quotient of the median and standard deviation of the posterior.
+        The orange horizontal solid line represents the median of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
+        The yellow horizontal solid line represents $\kappa=0$.
     }
@@
 \end{figure*}
-% order time, color median/std
-% symmetrical credible interval 5 to 95 percentile (np.percentile)
 
@@
 The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
-The mean of $\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.
+The median of $\mu$ within $90\%$ confidence interval is \variable{output/mu_median.txt}.
+The median of $\sigma$ within $90\%$ confidence interval is \variable{output/sigma_median.txt}.
 The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.
@@
 % Constraint on $\kappa$
-With the posteriors of $\mu$ and $\sigma$, the population constraint on $\kappa$ was calculated.
-The mean of the population distribution of $\kappa$ is \variable{output/kappa_mean.txt}.
+In figure \ref{fig:posterior_kappa}, we show both the restricted posterior of $\kappa$ and the generic posterior of $\kappa$.
+The restricted posterior of $\kappa$ is calculated by equation \ref{eq:restricted_posterior} with the assumption that $\kappa$ is a single value constant.
+The generic posterior of $\kappa$ is calculated by equation \ref{eq:generic_posterior} with the assumption that $\kappa$ is a true value drawn from a Gaussian distribution.
+The median of the restricted posterior of $\kappa$ within $90\%$ confidence interval is \variable{output/restricted_kappa_median.txt}.
+The median of the generic posterior of $\kappa$ within $90\%$ confidence interval is \variable{output/generic_kappa_median.txt}.
 Note that the result is consistent with GR when $\kappa=0$.
-% include both Hierarchical and stacking
+
+\begin{figure}[h]
+    \script{posterior_kappa.py}
+    \includegraphics[width=\columnwidth]{figures/posterior_kappa.pdf}
+    \caption{
+        The generic and restricted posterior of $\kappa$.
+        The orange solid line shows the restricted posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
+        The blue solid line shows the generic posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
+        The pink vertical solid line represents $\kappa=0$.
+    }
+    \label{fig:posterior_kappa}
+\end{figure}
 
@@
         Note that there is no posterior of $\kappa$ for the PE result from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
-        This plot shows that frequency-independent birefringence creates a degeneracy between $\kappa$ and $\iota$, while the frequency-dependent birefringence can break it.
+        This plot shows that frequency-independent birefringence creates a degeneracy between $\kappa$ and $\iota$, while frequency-dependent birefringence can break it.
     }
@@
 
-\section{Discussion}
-\label{sec:Discussion}
+% Case: GW200129 (glitch in Virgo data)
+\subsection{Exclusion of GW200129}
+\label{sec:GW200129}
+The reason for the exclusion is that past research suggests a potential glitch in GW200129 Virgo data. \citep{GW200129_glitch}
+We perform three sets of PE to check if the glitch would affect the population posterior of $\kappa$.
+We pick only two out of three detector data in each set to perform PE.
+The two sets of PE results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
+Further investigation is needed to understand the glitch.
+Thus, we did not include this event in our analysis.
+% show result included and excluded
 
-% Comparison with previous studies\
-\citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
-We convert the constraint to the same format as ours to make a comparison.
-They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
-We obtained a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
-Our result is an order of magnitude improvement in constraining $\kappa$.
-The main reason is that we have more events from GWTC-3 than GWTC-2.
-
-% Reasons for special events
-\subsection{Special Events}
-
-% Special event 1: GW190720
+% Case: GW190720 (failed PE)
+\subsection{Exclusion of GW190720}
+\label{sec:GW190720}
 For GW190720, we could not perform PE with Bilby successfully.
@@
 
-% Special event 2: GW200129
-For GW200129, past research suggests a potential glitch in Virgo data. \citep{GW200129_glitch}
-We perform three sets of PE to check if the glitch would affect the population posterior of $\kappa$.
-We pick only two out of three detector data in each set to perform PE.
-The two sets of PE results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
-Further investigation is needed to understand the glitch.
-Thus, we did not include this event in our analysis.
-% show result included it first
+\section{Discussion}
+\label{sec:Discussion}
+
+% Comparison with previous studies
+\subsection{Comparison with Previous Studies}
+\citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
+We convert the constraint to the same format as ours to make a comparison.
+They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
+We obtained a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
+Our result is an order of magnitude improvement in constraining $\kappa$.
+The main reason is that we have more events from GWTC-3 than GWTC-2.
 

### 2022-12-08 src/scripts/kappa_constraint.py

@@ -1,0 +1,43 @@
+import numpy as np
+import pandas as pd
+import scipy
+import paths
+
+def normal_distribution(x, mu, sigma):
+    return (1/(sigma*np.sqrt(2*np.pi)))*np.exp(-0.5*((x-mu)/sigma)**2)
+
+result_dict = {}
+result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_DataFrame = result_DataFrame[result_DataFrame.event != "GW200129_065458"]
+events = result_DataFrame['event'].unique()
+for event in events:
+    result_dict[event] = result_DataFrame[result_DataFrame.event == event]
+
+samples_original = np.load(paths.data/"samples_Gaussian_without_GW200129.npz")["chains"]
+samples_all = samples_original.reshape(-1, 2)
+df_Gaussian_samples_all = pd.DataFrame()
+df_Gaussian_samples_all['mu'] = samples_all[:,0]
+df_Gaussian_samples_all['sigma'] = samples_all[:,1]
+
+samples_reweighted = []
+for _ in range(1000):
+    Gaussian = np.full(shape=(len(result_DataFrame), 2), fill_value=df_Gaussian_samples_all.sample(1))
+    weight = normal_distribution(result_DataFrame['kappa'], Gaussian[:,0], Gaussian[:,1])
+    weight /= weight.sum()
+    samples_reweighted.append(np.random.choice(a=np.array(result_DataFrame['kappa']), size=1, p=weight))
+
+kernels = [scipy.stats.gaussian_kde(result_dict[event]['kappa']) for event in events[events != "GW200129_065458"]]
+kappa = np.linspace(-.1, .1, 1000)
+ll = [np.sum([np.log(ker(k)) for ker in kernels]) for k in kappa]
+ll = ll - np.max(ll)
+likelihood = np.exp(ll)
+likelihood = likelihood / np.trapz(likelihood, x=kappa)
+
+restricted_kappa_median = np.interp(0.5,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)
+generic_kappa_median = np.median(np.array(samples_reweighted).reshape(-1))
+
+with open(paths.output/"restricted_kappa_median.txt", "w") as f:
[…]

### 2022-12-09 src/tex/ms.tex
Formatting

@@
         The color of the violins represents the quotient of the median and standard deviation of the posterior.
-        The orange horizontal solid line represents the median of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
-        The yellow horizontal solid line represents $\kappa=0$.
+        The blue horizontal solid line represents the median of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
+        The pink horizontal solid line represents $\kappa=0$.
     }
@@
 In figure \ref{fig:posterior_kappa}, we show both the restricted posterior of $\kappa$ and the generic posterior of $\kappa$.
-The restricted posterior of $\kappa$ is calculated by equation \ref{eq:restricted_posterior} with the assumption that $\kappa$ is a single value constant.
-The generic posterior of $\kappa$ is calculated by equation \ref{eq:generic_posterior} with the assumption that $\kappa$ is a true value drawn from a Gaussian distribution.
-The median of the restricted posterior of $\kappa$ within $90\%$ confidence interval is \variable{output/restricted_kappa_median.txt}.
-The median of the generic posterior of $\kappa$ within $90\%$ confidence interval is \variable{output/generic_kappa_median.txt}.
+The restricted posterior is calculated by equation \ref{eq:restricted_posterior} with the assumption that $\kappa$ is a single value constant.
+The generic posterior is calculated by equation \ref{eq:generic_posterior} with the assumption that $\kappa$ is a true value drawn from a Gaussian distribution.
+The median of the restricted posterior within $90\%$ confidence interval is \variable{output/restricted_kappa_median.txt}.
+The median of the generic posterior within $90\%$ confidence interval is \variable{output/generic_kappa_median.txt}.
 Note that the result is consistent with GR when $\kappa=0$.
@@
 
-\pagebreak
+% \pagebreak
 \bibliography{bib}

### 2022-12-09 src/scripts/Gaussian_result.py

@@ -10,7 +10,7 @@ mu_median = np.median(df['mu'])
 sigma_median = np.median(df['sigma'])
 
 with open(paths.output/"mu_median.txt", "w") as f:
-    f.write(rf"${mu_median:.3f}^{{{(np.percentile(a=df['mu'],q=95)-mu_median):.3f}}}_{{{(np.percentile(a=df['mu'],q=5)-mu_median):.3f}}}$")
+    f.write(rf"${mu_median:.3f}^{{+{(np.percentile(a=df['mu'],q=95)-mu_median):.3f}}}_{{{(np.percentile(a=df['mu'],q=5)-mu_median):.3f}}}$")
 
 with open(paths.output/"sigma_median.txt", "w") as f:
-    f.write(rf"${sigma_median:.3f}^{{{(np.percentile(a=df['sigma'],q=95)-sigma_median):.3f}}}_{{{(np.percentile(a=df['sigma'],q=5)-sigma_median):.3f}}}$")
+    f.write(rf"${sigma_median:.3f}^{{+{(np.percentile(a=df['sigma'],q=95)-sigma_median):.3f}}}_{{{(np.percentile(a=df['sigma'],q=5)-sigma_median):.3f}}}$")

### 2022-12-09 src/scripts/corner_Gaussian.py

@@ -36,4 +36,6 @@ g.axes[1,1].set_xlabel("$\\sigma$")
 g.axes[1,0].set_xlabel("$\\mu$")
 g.axes[1,0].set_ylabel("$\\sigma$")
 
+g.axes[1,1].set_xlim(0.0)
+
 g.savefig(fname=paths.figures/"corner_Gaussian.pdf", bbox_inches="tight", dpi=300)

### 2022-12-09 src/scripts/kappa_constraint.py

@@ -37,7 +37,7 @@ restricted_kappa_median = np.interp(0.5,[np.trapz(likelihood[0:i],kappa[0:i]) fo
 generic_kappa_median = np.median(np.array(samples_reweighted).reshape(-1))
 
 with open(paths.output/"restricted_kappa_median.txt", "w") as f:
-    f.write(f"${restricted_kappa_median:.3f}^{{{(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)-restricted_kappa_median):.3f}}}_ {{{(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)-restricted_kappa_median):.3f}}}$")
+    f.write(f"${restricted_kappa_median:.3f}^{{+{(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)-restricted_kappa_median):.3f}}}_ {{{(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)-restricted_kappa_median):.3f}}}$")
 
 with open(paths.output/"generic_kappa_median.txt", "w") as f:
-    f.write(f"${generic_kappa_median:.3f}^{{{(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95)-generic_kappa_median):.3f}}}_{{{(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5)-generic_kappa_median):.3f}}}$")
+    f.write(f"${generic_kappa_median:.3f}^{{+{(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95)-generic_kappa_median):.3f}}}_{{{(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5)-generic_kappa_median):.3f}}}$")

### 2022-12-09 src/scripts/posterior_kappa.py

@@ -44,14 +44,15 @@ ll = ll - np.max(ll)
 likelihood = np.exp(ll)
 likelihood = likelihood / np.trapz(likelihood, x=kappa)
 
+plt.axvline(0, color=sns.color_palette()[4])
 sns.kdeplot(np.array(samples_reweighted).reshape(-1), label="generic", color=sns.color_palette()[0])
 plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5), color=sns.color_palette()[0], linestyle=':')
 plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95), color=sns.color_palette()[0], linestyle=':')
 plt.plot(kappa,likelihood, label="restricted", color=sns.color_palette()[3])
 plt.axvline(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[3], linestyle=':')
 plt.axvline(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[3], linestyle=':')
-plt.axvline(0, color=sns.color_palette()[4])
-plt.xlim(-.2, .2)
+
+plt.xlim(-.15, .15)
 plt.ylabel("$p(\kappa|\{d\})$")
 plt.legend()
 

### 2022-12-10 src/tex/ms.tex
Formatting
Update content

@@
 \begin{equation}
-    p(\mu,\sigma|\{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i|d_i)}{p(\kappa_i}p(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
+    p(\mu,\sigma|\{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i|d_i)}{p(\kappa_i)}p(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
     \label{eq:posterior_of_mu_sigma}
@@
 % Corner plot of the mean and standard deviation of kappa
-The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
+We then show the result of Bayesian hierarchical modeling in figure \ref{fig:corner_Gaussian}.
 The median of $\mu$ within $90\%$ confidence interval is \variable{output/mu_median.txt}.
@@
         The generic and restricted posterior of $\kappa$.
-        The orange solid line shows the restricted posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
-        The blue solid line shows the generic posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
+        The blue solid line shows the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
+        The orange solid line shows the generic posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
         The pink vertical solid line represents $\kappa=0$.
@@
 % Comparison with previous studies
-\subsection{Comparison with Previous Studies}
+\subsection{Comparison with previous studies}
 \citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
@@
 % Future work
-\subsection{Future Work}
+\subsection{Future work}
 % BNS
@@
 
-\section{Acknowledgements}
-\label{sec:Acknowledgements}
+\begin{acknowledgments}
 M.~I., K.~W.~K.~W.~, and W.~M.~F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute.
@@
 KAGRA is supported by Ministry of Education, Culture, Sports, Science and Technology (MEXT), Japan Society for the Promotion of Science (JSPS) in Japan; National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea; Academia Sinica (AS) and National Science and Technology Council (NSTC) in Taiwan.
+\end{acknowledgments}
 
-% \pagebreak
 \bibliography{bib}

### 2022-12-10 src/scripts/posterior_kappa.py

@@ -45,15 +45,16 @@ likelihood = np.exp(ll)
 likelihood = likelihood / np.trapz(likelihood, x=kappa)
 
 plt.axvline(0, color=sns.color_palette()[4])
-sns.kdeplot(np.array(samples_reweighted).reshape(-1), label="generic", color=sns.color_palette()[0])
-plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5), color=sns.color_palette()[0], linestyle=':')
-plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95), color=sns.color_palette()[0], linestyle=':')
-plt.plot(kappa,likelihood, label="restricted", color=sns.color_palette()[3])
-plt.axvline(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[3], linestyle=':')
-plt.axvline(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[3], linestyle=':')
+sns.kdeplot(np.array(samples_reweighted).reshape(-1), label="generic", color=sns.color_palette()[3])
+plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5), color=sns.color_palette()[3], linestyle=':')
+plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95), color=sns.color_palette()[3], linestyle=':')
+plt.plot(kappa,likelihood, label="restricted", color=sns.color_palette()[0])
+plt.axvline(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[0], linestyle=':')
+plt.axvline(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[0], linestyle=':')
 
 plt.xlim(-.15, .15)
 plt.ylabel("$p(\kappa|\{d\})$")
+plt.xlabel("$\kappa$")
 plt.legend()
 
 plt.savefig(fname=paths.figures/"posterior_kappa.pdf", bbox_inches="tight", dpi=300)

### 2022-12-10 src/scripts/violin_kappa.py

@@ -51,7 +51,7 @@ g.axhline(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(10
 g.axhline(np.interp(0.5,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa),color=sns.color_palette()[0])
 g.axhline(0, color=sns.color_palette()[4])
 g.figure.autofmt_xdate(rotation=45)
-g.figure.set_size_inches(30,10)
+g.figure.set_size_inches(40,10)
 g.figure.colorbar(plt.cm.ScalarMappable(norm=mpl.colors.Normalize(color_DataFrame['median_over_std'].min(),
                                                                   color_DataFrame['median_over_std'].max()),
                                         cmap=sns.color_palette("rocket_r", as_cmap=True)),

### 2022-12-10 src/tex/ms.tex
Update ms.tex

@@
 The median and confidence interval of $\kappa$ are calculated with the restricted posterior distribution of $\kappa$, which is given by equation \ref{eq:restricted_posterior}.
-We excluded GW200129 from the calculation of the median and confidence interval of $\kappa$.
+We excluded GW200129 from the calculation of the median and confidence interval of $\kappa$ in figure \ref{fig:violin_kappa}.
 Details of the exclusion are discussed in Sec.~\ref{sec:GW200129}.
@@
     \caption{
-        The violin plot shows the posterior of $\kappa$ for 69 events in this study.
+        The violin plot shows the posterior of $\kappa$ for 70 events in this study.
         Each violin represents a different event.

### 2023-01-30 src/tex/ms.tex
Update ms.tex

--- 2022-12-10 src/tex/ms.tex
@@
 % Bayesian Hierarchical Modeling
-
 We then perform Bayesian hierarchical modeling to further analyze the constraint on $\kappa$.
@@
         The color of the violins represents the quotient of the median and standard deviation of the posterior.
-        The blue horizontal solid line represents the median of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
+        The blue horizontal solid line represents the median of the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval of the restricted posterior of $\kappa$.
         The pink horizontal solid line represents $\kappa=0$.
@@
 \citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
-We convert the constraint to the same format as ours to make a comparison.
+We convert the constraint they gave to the same units as ours to make a comparison.
 They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.

### 2023-01-30 src/tex/ms.tex
Add to-do items
abstract

--- 2023-01-30 src/tex/ms.tex
@@
 \begin{abstract}
-    One of the major problems physicists face nowadays is that Einstein's theory of general relativity (GR) does not agree with quantum theories at some length scales.
-    In recent decades, theorists have worked on many beyond-GR theories to unify both.
-    Some of these theories, such as Chern-Simons gravity, suggest that there is gravitational wave (GW) amplitude birefringence.
-    In our study, we perform parameter estimation (PE) to constrain the strength of GW amplitude birefringence.
-    Compared to previous studies, we performed PE on more events, including events in the third LIGO-Virgo catalog (GWTC-3), and used a more realistic birefringence model to describe the phenomenon.
-    We obtained a constraint tighter than previous studies for an order of magnitude.
+    The propagation of gravitational waves can reveal fundamental feature of the structure of spacetime.
+    For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun of parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
+    Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
+    From 70 binary-black-hole signals, we obtain the most precise constraints on amplitude birefringence yet, an order of magnitude more stringent than previous results.
 \end{abstract}

### 2023-01-30 src/tex/ms.tex
intro and acronym package

@@
 \usepackage{amsfonts,amssymb,amsmath}
+\usepackage[nolist,nohyperlinks]{acronym}
 
@@
 
+\begin{acronym}
+\acro{GW}{gravitational wave}
+\acro{GR}{general relativity}
+\acro{CBC}{compact-binary coalescence}
+\acro{BH}{black hole}
+\acro{BBH}{binary black hole}
+\acro{LVK}{LIGO-Virgo-KAGRA Collaboration}
+\acro{PE}{parameter estimation}
+\end{acronym}
+
 \section{Introduction}
@@
 % Motivation
-Gravitational waves (GWs) have been used to test Einstein's theory of general relativity (GR) with detections by LIGO-Virgo-KAGRA Collaboration (LVK) \citep{LIGO, Virgo, KAGRA}.
-To study the possibility of different beyond-GR theories, we can observe if GWs have any deviation from GR.
-Some beyond-GR theories, such as Chern-Simons gravity, suggest that there is gravitational wave amplitude birefringence, while GR predicts no birefringence.
-GW amplitude birefringence is a property of space-time that consists of the enhancement of one GW polarization over the other as the wave propagates.
+\Ac{GW} detections by the \ac{LVK} \citep{LIGO, Virgo, KAGRA} are now routinely used to test various aspects of Einstein's theory of \ac{GR} \citep{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}.
+Among those, measurements of the basic properties of \acp{GW}, like their speed and polarization, can directly probe the fundamental symmetries of the underlying theory of gravity \citep{Will:2018bme}.
+For instance, the unequal propagation of \ac{GW} polarization states would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm, Alexander:2009tp, Sopuerta:2009iy}.
+Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3.
+% GW amplitude birefringence is a property of space-time that consists of the enhancement of one GW polarization over the other as the wave propagates.
 
@@
 Previous studies have constrained amplitude birefringence by performing different statistical analyses.
-\citet{Yamada_2020} and \citet{Wang_2021} both performed parameter estimation (PE) on the events in the first GW transient catalog \citep{GWTC-1}, GWTC-1, using birefringence templates.
-\citet{Okounkova_2022} considered the distribution of observed inclinations of the GW events in the second GW transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.
+\citet{Yamada_2020} and \citet{Wang_2021} both performed \ac{PE} on the events in the first \ac{GW} transient catalog \citep{GWTC-1}, GWTC-1, using birefringent templates.
+\citet{Okounkova_2022} considered the distribution of observed inclinations of the \ac{GW} events in the second \ac{GW} transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.
 
 % What's new?
-In this study, we used a frequency-dependent birefringence model to constrain the strength of GW amplitude birefringence by performing PE.
+In this study, we used a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE}.
 This model is a better approximation of the birefringence effect than the frequency-independent model used in \citet{Okounkova_2022}.
-Compared to previous studies, we also performed PE on more events, including events in the third GW transient catalog \citep{GWTC-3}, GWTC-3.
+Compared to previous studies, we also performed \ac{PE} on more events, including events in the third \ac{GW} transient catalog \citep{GWTC-3}, GWTC-3.
 We consider 70 binary black hole merger events with a false alarm rate $\leq1\mathrm{yr^{-1}}$.
-These events are also listed in TABLE I in \citet{GWTC-3_population}.\footnote{
+These events are also listed in Table I in \citet{GWTC-3_population}.\footnote{
 GW190720 is included in the list in \citet{GWTC-3_population}, but we did not include it in this study.
 Details in Sec.~\ref{sec:GW190720}.}
-We then give the population constraint on the strength of GW amplitude birefringence by using the PE results.
+We then give the population constraint on the strength of \ac{GW} amplitude birefringence by using the \ac{PE} results.
 
 % Section guide
-In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, mention the configuration we used in the PE and show the method we used to obtain the population constraint on GW amplitude birefringence.
-In Sec.~\ref{sec:Results}, we present the population constraint on GW amplitude birefringence we obtained and show the results of individual PE.
+In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, mention the configuration we used in the \ac{PE} and show the method we used to obtain the population constraint on \ac{GW} amplitude birefringence.
+In Sec.~\ref{sec:Results}, we present the population constraint on \ac{GW} amplitude birefringence we obtained and show the results of individual \ac{PE}.
 In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study and provide suggestions for future studies.

### 2023-01-30 src/tex/ms.tex
background sect A

@@
 \usepackage[nolist,nohyperlinks]{acronym}
+
+\newcommand*{\mi}[1]{\textsf{\color{magenta} [\textbf{MAX:} #1]}}
 
@@
 Among those, measurements of the basic properties of \acp{GW}, like their speed and polarization, can directly probe the fundamental symmetries of the underlying theory of gravity \citep{Will:2018bme}.
-For instance, the unequal propagation of \ac{GW} polarization states would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm, Alexander:2009tp, Sopuerta:2009iy}.
+For instance, unequal propagation of \ac{GW} polarization states would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm, Alexander:2009tp, Sopuerta:2009iy}.
 Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3.
@@
 
-\section{Method}
-\label{sec:Method}
+\section{Background}
+\label{sec:Background}
 
-\subsection{Waveform modification}
+\subsection{Birefringence}
+\label{sec:waveform}
+
 % GW polarization
-GWs comprise two independent polarization modes, just like electromagnetic waves, which are usually represented in the linear states (i.e. $+$ and $\times$).
-For a given Fourier mode, they can be combined into left and right circular states (i.e. left-handed and right-handed).
-They are related by
+In \ac{GR}, \acp{GW} are comprised of two independent polarization modes, usually represented in the linear basis of plus ($+$) and cross ($\times$) states.
+In the Fourier domain, they can be combined into left-handed (L) and right-handed (R) circular states \cite{Isi:2022mbx},
 \begin{equation}
-    h_{\mathrm{L/R}} = \frac{h_+ \pm i h_\times}{\sqrt{2}}\,,
+    h_{\mathrm{L/R}} = \frac{1}{\sqrt{2}}\left(h_+ \pm i h_\times\right)\,,
 \end{equation}
-where $h_{\mathrm{L/R}}$ are the amplitude of the left-handed and right-handed polarization of the waveform, $h_+$ and $h_\times$ are the amplitude of the plus and cross polarization of the waveform.
+where $h$ is the frequency domain strain, with the plus and minus signs for L and R respectively.
+These circular modes represent eigenstates of the helicity operator (helicity $\pm2$) and posses a definite parity.
+Einstein's theory, which is parity even, predicts no difference in the dynamics of these two states.
 
 % Waveform modification
-GW amplitude birefringence would enhance one polarization mode over the other.
-In theories like Chern- Simons gravity, the observed Fourier waveform takes the form of
+Yet, parity odd extensions of \ac{GR} may make distinctions between the two circular polarizations, which can manifest in both the generation and propagation of \acp{GW}.
+The latter can manifest in changes to the relative amplitude and phase of the polarizations that accrue as the wave propagates, giving us hope to detect initially small effects that compound over long propagation distances.
+
+In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
+To first order in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source takes the form of
 \begin{equation}
-    h_\mathrm{L/R}^{\mathrm{br}}=
-    h_\mathrm{L/R}^{\mathrm{GR}}\times
+    h_\mathrm{L/R}^{\mathrm{br}}(f) =
+    h_\mathrm{L/R}^{\mathrm{GR}}(f) \times
     \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
@@
 \end{equation}
-where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the observed waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform emitted, $\kappa$ is the dimensionless opacity parameter that represents the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the frequency of the waveform.
-We assume the intrinsic modification is small, so we can use GR waveform templates to describe the emitted waveform.
+where the emitted waveform $h_\mathrm{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_\mathrm{L/R}^{\mathrm{br}}$.
+The overall magnitude of this effect for a given distance and frequency $f$ is set by a dimensionless opacity parameter, $\kappa$, which encodes the intrinsic strength of the birefringence.
+The emitted waveform for a given source (i.e., the waveform observed very close to the source) will generally differ from the analogous waveform predicted by \ac{GR}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small, it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation $h^{\rm GR}$ above).
 
-Although the intrinsic modification is small, the effect accumulates as the GW propagates.
-As the GWs propagate, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the GWs.
-According to equation \ref{eq:waveform_modification}, a positive $\kappa$ means the left-handed polarization is enhanced over the right-handed polarization, while a negative $\kappa$ means the opposite.
-Note that when $\kappa=0$, the waveform is the same as GR predicts.
+Although the intrinsic modification is small, the effect targeted by Eq.~\eqref{eq:waveform_modification} accumulates as the \ac{GW} propagates.
+During propagation, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the \acp{GW}.
+According to Eq.~\eqref{eq:waveform_modification}, a positive $\kappa$ means the left-handed polarization is enhanced over the right-handed polarization, while a negative $\kappa$ means the opposite;
+when $\kappa=0$, the observed waveform is the same as \ac{GR} predicts, meaning there is no birefringence.
+
+Equation \eqref{eq:waveform_modification} can be derived as the first order effect in an expansion for multiple extensions of \ac{GR}.
+In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity.
+Since it originates from an expansion, Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, $\left|\kappa\right| \left(d_C/1\mathrm{Gpc}\right) \left(f/100\mathrm{Hz}\right) < 1$.
+Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
+\mi{specify what the parameter for the expansion is, add citations, and check with a theorist}
 
@@
 \end{figure}
+
+\section{Method}
+\label{sec:Method}
 

### 2023-01-30 src/tex/ms.tex
sec IIB

@@
 Among those, measurements of the basic properties of \acp{GW}, like their speed and polarization, can directly probe the fundamental symmetries of the underlying theory of gravity \citep{Will:2018bme}.
-For instance, unequal propagation of \ac{GW} polarization states would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm, Alexander:2009tp, Sopuerta:2009iy}.
+For instance, unequal propagation of \ac{GW} polarization states would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm,Alexander:2009tp,Sopuerta:2009iy}.
 Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3.
@@
 \begin{equation}
-    h_{\mathrm{L/R}} = \frac{1}{\sqrt{2}}\left(h_+ \pm i h_\times\right)\,,
+    h_{{L/R}} = \frac{1}{\sqrt{2}}\left(h_+ \pm i h_\times\right)\,,
 \end{equation}
@@
 \begin{equation}
-    h_\mathrm{L/R}^{\mathrm{br}}(f) =
-    h_\mathrm{L/R}^{\mathrm{GR}}(f) \times
-    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
+    h_{L/R}^{\mathrm{br}}(f) =
+    h_{L/R}^{\mathrm{GR}}(f) \times
+    \exp\left(\pm\kappa\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\,\mathrm{Hz}}\right)\,,
     \label{eq:waveform_modification}
 \end{equation}
-where the emitted waveform $h_\mathrm{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_\mathrm{L/R}^{\mathrm{br}}$.
+where the emitted waveform $h_{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_{L/R}^{\mathrm{br}}$.
 The overall magnitude of this effect for a given distance and frequency $f$ is set by a dimensionless opacity parameter, $\kappa$, which encodes the intrinsic strength of the birefringence.
-The emitted waveform for a given source (i.e., the waveform observed very close to the source) will generally differ from the analogous waveform predicted by \ac{GR}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small, it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation $h^{\rm GR}$ above).
+The emitted waveform for a given source (i.e., the waveform observed very close to the source) will generally differ from the analogous waveform predicted by \ac{GR}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small, it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
 
@@
 In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity.
-Since it originates from an expansion, Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, $\left|\kappa\right| \left(d_C/1\mathrm{Gpc}\right) \left(f/100\mathrm{Hz}\right) < 1$.
+Since it originates from an expansion, Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
+\begin{equation}
+\left|\kappa\right| \left(d_C/1\,\mathrm{Gpc}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, .
+\end{equation}
 Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
@@
 
-\subsection{Effect on inclination}
+\subsection{Relation to inclination}
+
 % Effect on inclination PE results
-In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for the dominant $(2,\pm 2)$ angular mode of nonprecessing binaries.
-The relationship between the amplitude ratio and the inclination is
+Under certain conditions, the effect of birefringence can be degenerate with a change in the orientation of the source with respect to the line of sight.
+Concretely, for a nonprecessing compact binary inspiral in \ac{GR}, the observed amplitude ratio of the left-handed and right-handed polarizations is only a function of the inclination $\iota$, the angle between the orbital angular momentum of the source and the line of sight.
+For  the dominant $\ell = |m| = 2$ angular mode of the radiation, the relation between the amplitude ratio and the inclination is
 \begin{equation}
-    \frac{h_\mathrm{L}}{h_\mathrm{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
+    \frac{h_{L}^\mathrm{GR}}{h^\mathrm{GR}_{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,
 \end{equation}
-where $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
-Birefringence can alter the ratio between polarizations of the waveform.
-Thus, the birefringence effect could affect the PE result on $\iota$ based on \citet{Okounkova_2022}.
+for all frequencies (see, e.g., Sec.~IIIC in \cite{Isi:2022mbx}).
 
-Based on the modification we implemented, the amplitude ratio of the left-handed and right-handed polarizations would be given by
+Since birefringence impacts the observed amplitude ratio of left- and right-handed modes, it could also affect inferences about the source inclination \cite{Alexander:2009tp}.
+However, the two effects are degenerate only if the frequency dependence of Eq.~\eqref{eq:waveform_modification} is neglected.
+This is easy to see from Eq.~\eqref{eq:waveform_modification}, since the implied polarization ratio for the $\ell = |m| = 2$ mode of a nonprecessing source is
 \begin{equation}
-    \frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}=\frac{\left(1-\cos\iota\right)^2}{\left(1+\cos\iota\right)^2}\exp\left({2\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}}\right)\,,
+    \frac{h_{L}^\mathrm{br}}{h_{R}^\mathrm{br}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2
+    \exp\left({2\kappa\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\, \mathrm{Hz}}}\right)\, .
     \label{eq:modified_amplitude_ratio}
 \end{equation}
-where $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed amplitude of left-handed and right-handed polarizations of the GWs.
-As the amplitude ratio does not only depends on $\iota$ with this modification but also $\kappa$ and $d_C$, the PE results on $\iota$ and $d_C$ are affected.
+For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will generate dephasing in the time domain waveform (Fig~\ref{fig:birefringence}).
 
-\citet{Okounkova_2022} assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
-This assumption would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio similarly.
-To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
-
-In contrast, we included the frequency dependence in the equation \ref{eq:modified_amplitude_ratio}, which is a first-order approximation of the birefringence model.
-The frequency dependence can break the degeneracy between $\kappa$ and $\iota$, as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
-To illustrate the frequency dependence of amplitude birefringence, the observed amplitudes of two polarizations of a simulated GW signal are shown in figure \ref{fig:birefringence}.
-
-\begin{figure}[h]
+\begin{figure}
     \script{birefringence.py}
@@
 \end{figure}
+
+\citet{Okounkova_2022} took the effect of birefringence to be independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+This assumption results in a full degeneracy between $\kappa$ and $\iota$:
+to reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
+That fact can be used to constrain frequency-independent birefringence by searching for features in the distribution of inferred inclinations \cite{Okounkova_2022}.
+
+By implementing Eq.~\eqref{eq:modified_amplitude_ratio}, which is a first-order approximation of the birefringence model, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
+Nevertheless, there exist systems for which the degeneracy cannot be perfectly broken because not enough frequencies are available in the data.
+As we will see in Sec.~\ref{sec:Results}, this is the case for heavy \acp{BBH}, which are only in the sensitive band of the detectors for a few cycles.
 

### 2023-01-30 src/tex/ms.tex
Sec IIIA

@@
 \acro{PE}{parameter estimation}
+\acro{FAR}{false-alarm rate}
+\acro{GWOSC}{the Gravitational Wave Open Science Center}
 \end{acronym}
@@
 Compared to previous studies, we also performed \ac{PE} on more events, including events in the third \ac{GW} transient catalog \citep{GWTC-3}, GWTC-3.
-We consider 70 binary black hole merger events with a false alarm rate $\leq1\mathrm{yr^{-1}}$.
+We consider 70 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$.
 These events are also listed in Table I in \citet{GWTC-3_population}.\footnote{
@@
 
-By implementing Eq.~\eqref{eq:modified_amplitude_ratio}, which is a first-order approximation of the birefringence model, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
+By implementing Eq.~\eqref{eq:waveform_modification}, which is a first-order approximation of the birefringence model, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
 Nevertheless, there exist systems for which the degeneracy cannot be perfectly broken because not enough frequencies are available in the data.
@@
 
-\subsection{Parameter estimation}
-% Parameter estimation
-To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby \citep{Bilby}, a Bayesian toolkit for GW data analysis which can calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question.
-We modified Bilby to include the birefringence effect in the likelihood function.
-We then perform PE on intrinsic and extrinsic parameters including $\kappa$ with open data from the Gravitational Wave Open Science Center \citep{GWOSC}.
-For PE configuration, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each event and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+\subsection{Single-event parameter estimation}
+
+To constrain birefringence, we reanalyze events from GWTC-3 \citep{GWTC-2.1, GWTC-3} implementing Eq.~\eqref{eq:waveform_modification} to directly obtain a posterior on $\kappa$ from the strain of each event.
+We analyze the 70 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$; to avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
+We procure strain data from \ac{GWOSC} \citep{GWOSC}.
+
+We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
+We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3} \mi{cite actual zenodo link}, with \textsc{IMRPhenomXPHM} as the reference waveform.
+We apply a distance prior uniform in comoving volume \mi{is this right?}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+
+\mi{list here any special considerations for individual events}
+
 The PE results are published in \citet{dataset}.
+
+\subsection{Collective analysis}
 
@@
 
-\subsection{Bayesian hierarchical modeling}
 % Bayesian Hierarchical Modeling

### 2023-01-30 src/tex/ms.tex
collective results

@@
 
-The PE results are published in \citet{dataset}.
+In order to validate our \ac{PE} implementation, we reproduce the \ac{LVK} \ac{PE} results obtained assuming \ac{GR} by enforcing $\kappa = 0$; this also has the advantage of producing \ac{GR} runs that are directly comparable to our birefringent runs.
+All our \ac{PE} results, including the \ac{GR} valudation runs, are published in \citet{dataset}.
 
@@
 
-We assume $\kappa$ is the same for all events and calculate a posterior distribution of $\kappa$.
-The posterior distribution is given by
+\subsubsection{Shared birefringence parameter}
+
+Since birefringence is a property of spacetime and we are targeting a propagation effect not intrinsic to any source, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
+\mi{caveats?}
+Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply the product of the individual posteriors,%
+\footnote{Had the $\kappa$ prior been nonuniform, then the individual \emph{likelihoods} would need to be multiplied, as opposed to the posteriors.}
 \begin{equation}
-    p(\kappa|\{d\})\propto \prod_{i}p(\kappa|d_i)\,,
+    p(\kappa \mid \{d_i\})\propto \prod_{i}p(\kappa \mid d_i)\,,
     \label{eq:restricted_posterior}
 \end{equation}
-where $p$ is the probability density function, $i$ is the index of the event, $\{d\}$ is the set of data including all events, and $d_i$ is the data of a single event.
-Equation \ref{eq:restricted_posterior} shows that the posterior distribution of $\kappa$ is the product of the posterior distribution of $\kappa$ for each event.
-This posterior distribution is a restricted posterior distribution because we assume $\kappa$ is the same for all events.
+where $d_i$ is the strain data for the $i$\textsuperscript{th} event.
+We use this expression to obtain the primary constraint presented in this work.
 
-Note that we also perform a sanity test to check if our PE could recover the PE results done by LVK in \citet{GWTC-2.1, GWTC-3}.
-To implement the sanity check, we set $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
-These sanity test results are also attached in \citet{dataset}.
+\subsubsection{Nonshared birefringence parameters}
 
 % Bayesian Hierarchical Modeling
-We then perform Bayesian hierarchical modeling to further analyze the constraint on $\kappa$.
-Bayesian hierarchical modeling is a method to check if our waveform model can fully describe the amplitude birefringence effect.
+In order to further characterize our set of measurements, and identify potential outliers, we carry out an additional collective analysis that allows $\kappa$ to take different values for each events.
+To do this, we make use of hierarchical Bayesian inference to model the distribuion of $\kappa$'s consistent with the observed data.
 Here, instead of assuming all events share the same value of $\kappa$, we assume $\kappa$ from each event is a true value drawn from a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.

### 2023-01-31 src/tex/ms.tex
hierarchical

@@
         This plot shows the effect of amplitude birefringence with frequency dependence.
+        \mi{@Thomas can you add a panel with the time-domain waveforms (maybe just for L or R)?}
     }
@@
 We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
-We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3} \mi{cite actual zenodo link}, with \textsc{IMRPhenomXPHM} as the reference waveform.
-We apply a distance prior uniform in comoving volume \mi{is this right?}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3} \mi{@Thomas cite actual zenodo link}, with \textsc{IMRPhenomXPHM} as the reference waveform.
+We apply a distance prior uniform in comoving volume \mi{@Thomas is this right?}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
 
-\mi{list here any special considerations for individual events}
+\mi{@Thomas list here any special considerations for individual events}
 
@@
 Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply the product of the individual posteriors,%
-\footnote{Had the $\kappa$ prior been nonuniform, then the individual \emph{likelihoods} would need to be multiplied, as opposed to the posteriors.}
+\footnote{Had the $\kappa$ prior been nonuniform, then we would need to multiply the individual \emph{likelihoods}, not to the posteriors.}
 \begin{equation}
@@
 % Bayesian Hierarchical Modeling
-In order to further characterize our set of measurements, and identify potential outliers, we carry out an additional collective analysis that allows $\kappa$ to take different values for each events.
-To do this, we make use of hierarchical Bayesian inference to model the distribuion of $\kappa$'s consistent with the observed data.
-Here, instead of assuming all events share the same value of $\kappa$, we assume $\kappa$ from each event is a true value drawn from a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
-We chose a Gaussian distribution as the underlying distribution of $\kappa$ because it captures the correct mean and the standard deviation of the distribution no matter what the underlying distribution is.
+In order to further characterize our set of measurements, and identify potential outliers, we carry out an additional collective analysis that allows $\kappa$ to take different values for each event \cite{Zimmerman:2019wzo,Isi:2022cii}.
+To do this, we apply hierarchical Bayesian inference \cite{Loredo:2004nn} to model the distribution of $\kappa$'s consistent with the observed data:
+we posit that there is a specific value of the parameter, $\kappa_i$, associated to each event, drawn from some unknown distribution of underlying values; from the imperfect measurements of $\kappa_i$ for each event, we may reconstruct the underlying distribution.
+If we are interested in constraining the first two moments of the distribution, it is convenient to parametrize the $\kappa_i$'s as drawn from a Gaussian with unknown mean $\mu$ and variance $\sigma^2$, i.e., $\kappa_i \sim \mathcal{N}(\mu, \sigma^2)$ \cite{Isi:2019asy}, and measure those hyperparameters from the collection of observed data.
+
+If \ac{GR} is correct and there is no birefringence, we should find the observed $\kappa$ distribution to be consistent with a delta function at the origin ($\kappa_i = 0$ for all $i$, or $\mu=\sigma=0$); on the other hand, if spacetime is birefringent, we expect to find a delta function at some nonzero value ($\kappa_i = \kappa \neq 0$, or $\mu = \kappa$ and $\sigma=0$).
+But this analysis also has the power to reveal unexpected physics or systematics in our measurements: if $\sigma$ is confidently found to be nonzero, this would imply that our set of measurements are statistically unlikely to originate from a unique $\kappa$ value.
+This could signal further physics than is implied by Eq.~\eqref{eq:waveform_modification} \mi{which physics?
+@Will, any ideas?} or, more prosaically, that there are outliers in our measurements due to mismodeling, e.g., in the waveform approximant or the noise of the detector.
 
@@
 
-If our model can fully describe the birefringence effect, the true value of $\kappa$ should be a single value, and the Gaussian distribution should have a standard deviation of $0$.
-If the standard deviation obtained is not $0$, it means our model cannot fully describe the birefringence effect.
-It could be due to our model does not fully capture the physics of birefringence, or there is a systematic error in the data.
-
 We then calculate the generic posterior distribution of $\kappa$ from the posterior of $\mu$ and $\sigma$.
@@
 
-\subsection{Result on GWTC-3}
-% Violin plot
-In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each event in the form of a violin plot.
-The median and confidence interval of $\kappa$ are calculated with the restricted posterior distribution of $\kappa$, which is given by equation \ref{eq:restricted_posterior}.
-We excluded GW200129 from the calculation of the median and confidence interval of $\kappa$.
-Details of the exclusion are discussed in Sec.~\ref{sec:GW200129}.
-
-\begin{figure*}[ht]
+\begin{figure*}
     \script{violin_kappa.py}
@@
 
+\subsection{Result on GWTC-3}
+% Violin plot
+In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each event in the form of a violin plot.
+The median and confidence interval of $\kappa$ are calculated with the restricted posterior distribution of $\kappa$, which is given by equation \ref{eq:restricted_posterior}.
+We excluded GW200129 from the calculation of the median and confidence interval of $\kappa$.
+Details of the exclusion are discussed in Sec.~\ref{sec:GW200129}.
+
 \subsection{Bayesian hierarchical modeling}
@@
 
-\begin{figure}[h]
+\begin{figure}
     \script{corner_Gaussian.py}
@@
 
-\begin{figure}[h]
+\begin{figure}
     \script{posterior_kappa.py}
@@
 
-\begin{figure}[h]
+\begin{figure}
     \script{corner_GW150914.py}

### 2023-01-31 src/tex/ms.tex

@@
 \mi{caveats?}
-Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply the product of the individual posteriors,%
-\footnote{Had the $\kappa$ prior been nonuniform, then we would need to multiply the individual \emph{likelihoods}, not to the posteriors.}
+Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply obtained from the product of the individual likelihoods, $p(d_i \mid \kappa) \propto p(\kappa \mid d_i)/p(\kappa)$, such that
 \begin{equation}
-    p(\kappa \mid \{d_i\})\propto \prod_{i}p(\kappa \mid d_i)\,,
+    p(\kappa \mid \{d_i\})\propto p(\kappa) \prod_{i}\frac{p(\kappa \mid d_i)}{p(\kappa)}\,,
     \label{eq:restricted_posterior}
 \end{equation}
-where $d_i$ is the strain data for the $i$\textsuperscript{th} event.
-We use this expression to obtain the primary constraint presented in this work.
+where $d_i$ is the strain data for the $i$\textsuperscript{th} event, and $p(\kappa)$ is the prior on $\kappa$; since the prior is uniform, in our case Eq.~\eqref{eq:restricted_posterior} reduces to the product of the posteriors, namely $p(\kappa \mid \{d_i\}) \propto \prod_{i}p(\kappa \mid d_i)$.
+We use Eq.~\eqref{eq:restricted_posterior} to obtain the primary constraint presented in this work.
 
@@
 
-With the posterior of $\kappa$ from PE on each event, the posterior of $\mu$ and $\sigma$ can be calculated by
+Starting from the posterior on $\kappa$ from each event, the posterior on the hyperparameters $\mu$ and $\sigma$ can be calculated by
 \begin{equation}
-    p(\mu,\sigma|\{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i|d_i)}{p(\kappa_i)}p(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
+    p(\mu,\sigma \mid \{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i\mid d_i)}{p(\kappa_i)}p(\kappa_i\mid\mu,\sigma)\,d\kappa_i\,,
     \label{eq:posterior_of_mu_sigma}
 \end{equation}
-Note that $p(\kappa)$ is the prior of $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
-We chose the prior of $\mu$ and $\sigma$ being uniform distributions.
-Equation \ref{eq:posterior_of_mu_sigma} can then be simplified to
+where $p(\kappa)$ is the prior initially applied to $\kappa$ during \ac{PE}, which in our case is a uniform distribution $\mathcal{U}[-1,1]$.
+Furhter choosing the hyperpriors on $\mu$ and $\sigma$ to be uniform, Eq.~\eqref{eq:posterior_of_mu_sigma} simplifies to
 \begin{equation}
-    p(\mu,\sigma|\{d\})\propto\prod_{i}\int p(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
+    p(\mu,\sigma\mid\{d\})\propto\prod_{i}\int p(\kappa_i\mid d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
 \end{equation}
-To sample the posterior distribution of $\mu$ and $\sigma$, we used a sampling package, flowMC \citep{flowMC}.
-
-We then calculate the generic posterior distribution of $\kappa$ from the posterior of $\mu$ and $\sigma$.
-The constraint was calculated by
+%
+We also calculate the expected population distribution of $\kappa$ marginalized over $\mu$ and $\sigma$.
+This is given by
 \begin{equation}
-    p(\kappa|\{d\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma|\{d\})\,d\mu\,d\sigma\,.
+            p(\kappa\mid \{d\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma\mid \{d\})\,d\mu\,d\sigma\, , 
     \label{eq:generic_posterior}
 \end{equation}
+and represents our overall expectation for $\kappa$, given the observed set of individual measurements.
+To sample the posterior distribution of $\mu$ and $\sigma$, we use the sampling package \textsc{flowMC} \citep{flowMC}.
 

### 2023-01-31 src/tex/ms.tex
Add citation

--- 2023-01-30 src/tex/ms.tex
@@
 \usepackage{amsfonts,amssymb,amsmath}
-\usepackage[nolist,nohyperlinks]{acronym}
-
-\newcommand*{\mi}[1]{\textsf{\color{magenta} [\textbf{MAX:} #1]}}
 
@@
 \begin{abstract}
-    The propagation of gravitational waves can reveal fundamental feature of the structure of spacetime.
-    For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun of parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
-    Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
-    From 70 binary-black-hole signals, we obtain the most precise constraints on amplitude birefringence yet, an order of magnitude more stringent than previous results.
+    One of the major problems physicists face nowadays is that Einstein's theory of general relativity (GR) does not agree with quantum theories at some length scales.
+    In recent decades, theorists have worked on many beyond-GR theories to unify both.
+    Some of these theories, such as Chern-Simons gravity, suggest that there is gravitational wave (GW) amplitude birefringence.
+    In our study, we perform parameter estimation (PE) to constrain the strength of GW amplitude birefringence.
+    Compared to previous studies, we performed PE on more events, including events in the third LIGO-Virgo catalog (GWTC-3), and used a more realistic birefringence model to describe the phenomenon.
+    We obtained a constraint tighter than previous studies for an order of magnitude.
 \end{abstract}
@@
 
-\begin{acronym}
-\acro{GW}{gravitational wave}
-\acro{GR}{general relativity}
-\acro{CBC}{compact-binary coalescence}
-\acro{BH}{black hole}
-\acro{BBH}{binary black hole}
-\acro{LVK}{LIGO-Virgo-KAGRA Collaboration}
-\acro{PE}{parameter estimation}
-\acro{FAR}{false-alarm rate}
-\acro{GWOSC}{the Gravitational Wave Open Science Center}
-\end{acronym}
-
 \section{Introduction}
@@
 % Motivation
-\Ac{GW} detections by the \ac{LVK} \citep{LIGO, Virgo, KAGRA} are now routinely used to test various aspects of Einstein's theory of \ac{GR} \citep{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}.
-Among those, measurements of the basic properties of \acp{GW}, like their speed and polarization, can directly probe the fundamental symmetries of the underlying theory of gravity \citep{Will:2018bme}.
-For instance, unequal propagation of \ac{GW} polarization states would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm,Alexander:2009tp,Sopuerta:2009iy}.
-Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3.
-% GW amplitude birefringence is a property of space-time that consists of the enhancement of one GW polarization over the other as the wave propagates.
+Gravitational waves (GWs) have been used to test Einstein's theory of general relativity (GR) with detections by LIGO-Virgo-KAGRA Collaboration (LVK) \citep{LIGO, Virgo, KAGRA}.
+To study the possibility of different beyond-GR theories, we can observe if GWs have any deviation from GR.
+Some beyond-GR theories, such as Chern-Simons gravity, suggest that there is gravitational wave amplitude birefringence, while GR predicts no birefringence.
+GW amplitude birefringence is a property of space-time that consists of the enhancement of one GW polarization over the other as the wave propagates.
 
@@
 Previous studies have constrained amplitude birefringence by performing different statistical analyses.
-\citet{Yamada_2020} and \citet{Wang_2021} both performed \ac{PE} on the events in the first \ac{GW} transient catalog \citep{GWTC-1}, GWTC-1, using birefringent templates.
-\citet{Okounkova_2022} considered the distribution of observed inclinations of the \ac{GW} events in the second \ac{GW} transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.
+\citet{Yamada_2020} and \citet{Wang_2021} both performed parameter estimation (PE) on the events in the first GW transient catalog \citep{GWTC-1}, GWTC-1, using birefringence templates.
+\citet{Okounkova_2022} considered the distribution of observed inclinations of the GW events in the second GW transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.
 
 % What's new?
-In this study, we used a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE}.
+In this study, we used a frequency-dependent birefringence model to constrain the strength of GW amplitude birefringence by performing PE.
 This model is a better approximation of the birefringence effect than the frequency-independent model used in \citet{Okounkova_2022}.
-Compared to previous studies, we also performed \ac{PE} on more events, including events in the third \ac{GW} transient catalog \citep{GWTC-3}, GWTC-3.
-We consider 70 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$.
-These events are also listed in Table I in \citet{GWTC-3_population}.\footnote{
+Compared to previous studies, we also performed PE on more events, including events in the third GW transient catalog \citep{GWTC-3}, GWTC-3.
+We consider 70 binary black hole merger events with a false alarm rate $\leq1\mathrm{yr^{-1}}$.
+These events are also listed in TABLE I in \citet{GWTC-3_population}.\footnote{
 GW190720 is included in the list in \citet{GWTC-3_population}, but we did not include it in this study.
 Details in Sec.~\ref{sec:GW190720}.}
-We then give the population constraint on the strength of \ac{GW} amplitude birefringence by using the \ac{PE} results.
+We then give the population constraint on the strength of GW amplitude birefringence by using the PE results.
 
 % Section guide
-In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, mention the configuration we used in the \ac{PE} and show the method we used to obtain the population constraint on \ac{GW} amplitude birefringence.
-In Sec.~\ref{sec:Results}, we present the population constraint on \ac{GW} amplitude birefringence we obtained and show the results of individual \ac{PE}.
+In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, mention the configuration we used in the PE and show the method we used to obtain the population constraint on GW amplitude birefringence.
+In Sec.~\ref{sec:Results}, we present the population constraint on GW amplitude birefringence we obtained and show the results of individual PE.
 In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study and provide suggestions for future studies.
 
-\section{Background}
-\label{sec:Background}
+\section{Method}
+\label{sec:Method}
 
-\subsection{Birefringence}
-\label{sec:waveform}
-
+\subsection{Waveform modification}
 % GW polarization
-In \ac{GR}, \acp{GW} are comprised of two independent polarization modes, usually represented in the linear basis of plus ($+$) and cross ($\times$) states.
-In the Fourier domain, they can be combined into left-handed (L) and right-handed (R) circular states \cite{Isi:2022mbx},
+GWs comprise two independent polarization modes, just like electromagnetic waves, which are usually represented in the linear states (i.e. $+$ and $\times$).
+For a given Fourier mode, they can be combined into left and right circular states (i.e. left-handed and right-handed).
+They are related by
 \begin{equation}
-    h_{{L/R}} = \frac{1}{\sqrt{2}}\left(h_+ \pm i h_\times\right)\,,
+    h_{\mathrm{L/R}} = \frac{h_+ \pm i h_\times}{\sqrt{2}}\,,
 \end{equation}
-where $h$ is the frequency domain strain, with the plus and minus signs for L and R respectively.
-These circular modes represent eigenstates of the helicity operator (helicity $\pm2$) and posses a definite parity.
-Einstein's theory, which is parity even, predicts no difference in the dynamics of these two states.
+where $h_{\mathrm{L/R}}$ are the amplitude of the left-handed and right-handed polarization of the waveform, $h_+$ and $h_\times$ are the amplitude of the plus and cross polarization of the waveform.
 
 % Waveform modification
-Yet, parity odd extensions of \ac{GR} may make distinctions between the two circular polarizations, which can manifest in both the generation and propagation of \acp{GW}.
-The latter can manifest in changes to the relative amplitude and phase of the polarizations that accrue as the wave propagates, giving us hope to detect initially small effects that compound over long propagation distances.
-
-In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
-To first order in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source takes the form of
+GW amplitude birefringence would enhance one polarization mode over the other.
+In theories like Chern- Simons gravity, the observed Fourier waveform takes the form of
 \begin{equation}
-    h_{L/R}^{\mathrm{br}}(f) =
-    h_{L/R}^{\mathrm{GR}}(f) \times
-    \exp\left(\pm\kappa\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\,\mathrm{Hz}}\right)\,,
+    h_\mathrm{L/R}^{\mathrm{br}}=
+    h_\mathrm{L/R}^{\mathrm{GR}}\times
+    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
     \label{eq:waveform_modification}
 \end{equation}
-where the emitted waveform $h_{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_{L/R}^{\mathrm{br}}$.
-The overall magnitude of this effect for a given distance and frequency $f$ is set by a dimensionless opacity parameter, $\kappa$, which encodes the intrinsic strength of the birefringence.
-The emitted waveform for a given source (i.e., the waveform observed very close to the source) will generally differ from the analogous waveform predicted by \ac{GR}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small, it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
+where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the observed waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the Fourier amplitude of the waveform emitted, $\kappa$ is the dimensionless opacity parameter that represents the strength of the birefringence, $d_C$ is the comoving distance to the source, and $f$ is the frequency of the waveform.
+We assume the intrinsic modification is small, so we can use GR waveform templates to describe the emitted waveform.
 
-Although the intrinsic modification is small, the effect targeted by Eq.~\eqref{eq:waveform_modification} accumulates as the \ac{GW} propagates.
-During propagation, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the \acp{GW}.
-According to Eq.~\eqref{eq:waveform_modification}, a positive $\kappa$ means the left-handed polarization is enhanced over the right-handed polarization, while a negative $\kappa$ means the opposite;
-when $\kappa=0$, the observed waveform is the same as \ac{GR} predicts, meaning there is no birefringence.
+Although the intrinsic modification is small, the effect accumulates as the GW propagates.
+As the GWs propagate, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the GWs.
+According to equation \ref{eq:waveform_modification}, a positive $\kappa$ means the left-handed polarization is enhanced over the right-handed polarization, while a negative $\kappa$ means the opposite.
+Note that when $\kappa=0$, the waveform is the same as GR predicts.
 
-Equation \eqref{eq:waveform_modification} can be derived as the first order effect in an expansion for multiple extensions of \ac{GR}.
-In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity.
-Since it originates from an expansion, Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
+\subsection{Effect on inclination}
+% Effect on inclination PE results
+In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for the dominant $(2,\pm 2)$ angular mode of nonprecessing binaries.
+The relationship between the amplitude ratio and the inclination is
 \begin{equation}
-\left|\kappa\right| \left(d_C/1\,\mathrm{Gpc}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, .
+    \frac{h_\mathrm{L}}{h_\mathrm{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
 \end{equation}
-Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
-\mi{specify what the parameter for the expansion is, add citations, and check with a theorist}
+where $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
+Birefringence can alter the ratio between polarizations of the waveform.
+Thus, the birefringence effect could affect the PE result on $\iota$ based on \citet{Okounkova_2022}.
 
-\subsection{Relation to inclination}
-
-% Effect on inclination PE results
-Under certain conditions, the effect of birefringence can be degenerate with a change in the orientation of the source with respect to the line of sight.
-Concretely, for a nonprecessing compact binary inspiral in \ac{GR}, the observed amplitude ratio of the left-handed and right-handed polarizations is only a function of the inclination $\iota$, the angle between the orbital angular momentum of the source and the line of sight.
-For  the dominant $\ell = |m| = 2$ angular mode of the radiation, the relation between the amplitude ratio and the inclination is
+Based on the modification we implemented, the amplitude ratio of the left-handed and right-handed polarizations would be given by
 \begin{equation}
-    \frac{h_{L}^\mathrm{GR}}{h^\mathrm{GR}_{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,
-\end{equation}
-for all frequencies (see, e.g., Sec.~IIIC in \cite{Isi:2022mbx}).
-
-Since birefringence impacts the observed amplitude ratio of left- and right-handed modes, it could also affect inferences about the source inclination \cite{Alexander:2009tp}.
-However, the two effects are degenerate only if the frequency dependence of Eq.~\eqref{eq:waveform_modification} is neglected.
-This is easy to see from Eq.~\eqref{eq:waveform_modification}, since the implied polarization ratio for the $\ell = |m| = 2$ mode of a nonprecessing source is
-\begin{equation}
-    \frac{h_{L}^\mathrm{br}}{h_{R}^\mathrm{br}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2
-    \exp\left({2\kappa\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\, \mathrm{Hz}}}\right)\, .
+    \frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}=\frac{\left(1-\cos\iota\right)^2}{\left(1+\cos\iota\right)^2}\exp\left({2\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}}\right)\,,
     \label{eq:modified_amplitude_ratio}
 \end{equation}
-For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will generate dephasing in the time domain waveform (Fig~\ref{fig:birefringence}).
+where $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed amplitude of left-handed and right-handed polarizations of the GWs.
+As the amplitude ratio does not only depends on $\iota$ with this modification but also $\kappa$ and $d_C$, the PE results on $\iota$ and $d_C$ are affected.
 
-\begin{figure}
+\citet{Okounkova_2022} assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+This assumption would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio similarly.
+To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
+
+In contrast, we included the frequency dependence in the equation \ref{eq:modified_amplitude_ratio}, which is a first-order approximation of the birefringence model.
+The frequency dependence can break the degeneracy between $\kappa$ and $\iota$, as this would make the effect of the birefringence affect the amplitude ratio differently compared to $\iota$.
+To illustrate the frequency dependence of amplitude birefringence, the observed amplitudes of two polarizations of a simulated GW signal are shown in figure \ref{fig:birefringence}.
+
+\begin{figure}[h]
     \script{birefringence.py}
@@
 
-\citet{Okounkova_2022} took the effect of birefringence to be independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
-This assumption results in a full degeneracy between $\kappa$ and $\iota$:
-to reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
-That fact can be used to constrain frequency-independent birefringence by searching for features in the distribution of inferred inclinations \cite{Okounkova_2022}.
+\subsection{Parameter estimation}
+% Parameter estimation
+To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby \citep{Bilby}, a Bayesian toolkit for GW data analysis which can calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question.
+We modified Bilby to include the birefringence effect in the likelihood function.
+We then perform PE on intrinsic and extrinsic parameters including $\kappa$ with open data from the Gravitational Wave Open Science Center \citep{GWOSC}.
+For PE configuration, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} for each event and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+The PE results are published in \citet{dataset}.
 
-By implementing Eq.~\eqref{eq:waveform_modification}, which is a first-order approximation of the birefringence model, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
-Nevertheless, there exist systems for which the degeneracy cannot be perfectly broken because not enough frequencies are available in the data.
-As we will see in Sec.~\ref{sec:Results}, this is the case for heavy \acp{BBH}, which are only in the sensitive band of the detectors for a few cycles.
-
-\section{Method}
-\label{sec:Method}
-
-\subsection{Single-event parameter estimation}
-
-To constrain birefringence, we reanalyze events from GWTC-3 \citep{GWTC-2.1, GWTC-3} implementing Eq.~\eqref{eq:waveform_modification} to directly obtain a posterior on $\kappa$ from the strain of each event.
-We analyze the 70 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$; to avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
-We procure strain data from \ac{GWOSC} \citep{GWOSC}.
-
-We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
-We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3} \mi{cite actual zenodo link}, with \textsc{IMRPhenomXPHM} as the reference waveform.
-We apply a distance prior uniform in comoving volume \mi{is this right?}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
-
-\mi{list here any special considerations for individual events}
-
-In order to validate our \ac{PE} implementation, we reproduce the \ac{LVK} \ac{PE} results obtained assuming \ac{GR} by enforcing $\kappa = 0$; this also has the advantage of producing \ac{GR} runs that are directly comparable to our birefringent runs.
-All our \ac{PE} results, including the \ac{GR} valudation runs, are published in \citet{dataset}.
-
-\subsection{Collective analysis}
-
-\subsubsection{Shared birefringence parameter}
-
-Since birefringence is a property of spacetime and we are targeting a propagation effect not intrinsic to any source, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
-\mi{caveats?}
-Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply the product of the individual posteriors,%
-\footnote{Had the $\kappa$ prior been nonuniform, then the individual \emph{likelihoods} would need to be multiplied, as opposed to the posteriors.}
+We assume $\kappa$ is the same for all events and calculate a posterior distribution of $\kappa$.
+The posterior distribution is given by
 \begin{equation}
-    p(\kappa \mid \{d_i\})\propto \prod_{i}p(\kappa \mid d_i)\,,
+    p(\kappa|\{d\})\propto \prod_{i}p(\kappa|d_i)\,,
     \label{eq:restricted_posterior}
 \end{equation}
-where $d_i$ is the strain data for the $i$\textsuperscript{th} event.
-We use this expression to obtain the primary constraint presented in this work.
+where $p$ is the probability density function, $i$ is the index of the event, $\{d\}$ is the set of data including all events, and $d_i$ is the data of a single event.
+Equation \ref{eq:restricted_posterior} shows that the posterior distribution of $\kappa$ is the product of the posterior distribution of $\kappa$ for each event.
+This posterior distribution is a restricted posterior distribution because we assume $\kappa$ is the same for all events.
 
-\subsubsection{Nonshared birefringence parameters}
+Note that we also perform a sanity test to check if our PE could recover the PE results done by LVK in \citet{GWTC-2.1, GWTC-3}.
+To implement the sanity check, we set $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
+These sanity test results are also attached in \citet{dataset}.
 
+\subsection{Bayesian hierarchical modeling}
 % Bayesian Hierarchical Modeling
-In order to further characterize our set of measurements, and identify potential outliers, we carry out an additional collective analysis that allows $\kappa$ to take different values for each events.
-To do this, we make use of hierarchical Bayesian inference to model the distribuion of $\kappa$'s consistent with the observed data.
+We then perform Bayesian hierarchical modeling to further analyze the constraint on $\kappa$.
+Bayesian hierarchical modeling is a method to check if our waveform model can fully describe the amplitude birefringence effect.
 Here, instead of assuming all events share the same value of $\kappa$, we assume $\kappa$ from each event is a true value drawn from a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
@@
 Then, we will discuss some special events individually.
+% What special treatment do we give to events
 
@@
 The median and confidence interval of $\kappa$ are calculated with the restricted posterior distribution of $\kappa$, which is given by equation \ref{eq:restricted_posterior}.
-We excluded GW200129 from the calculation of the median and confidence interval of $\kappa$.
+We excluded GW200129 from the calculation of the median and confidence interval of $\kappa$ in figure \ref{fig:violin_kappa}.
 Details of the exclusion are discussed in Sec.~\ref{sec:GW200129}.
+% mention excluded from all combined results
 
@@
 \end{figure*}
+% show the excluded events visually different
 
@@
 The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.
+% talk about what events pull sigma in figure 2
+% create a session for all the outliers
 
@@
 % Case study
+
 % Case: GW170818
@@
 Therefore, the narrow frequency range can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
+% 0 is still in the 90%. there is much probability that kappa is not 0 due to degeneracy.
 
@@
 % show result included and excluded
+% exclusion from what?
 
@@
 The main reason is that we have more events from GWTC-3 than GWTC-2.
+
+% Comparison with the previous paper with the equation related to the parameter they used
 
@@
 
+\appendix
+\section{}
+
 \bibliography{bib}

### 2023-01-31 src/tex/ms.tex

--- 2023-01-31 src/tex/ms.tex
@@
 \begin{abstract}
-    The propagation of gravitational waves can reveal fundamental feature of the structure of spacetime.
-    For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun of parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
+    The propagation of gravitational waves can reveal the fundamental feature of the structure of spacetime.
+    For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun for parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
     Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
@@
 We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
-We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3} \mi{@Thomas cite actual zenodo link}, with \textsc{IMRPhenomXPHM} as the reference waveform.
+We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, with \textsc{IMRPhenomXPHM} as the reference waveform.
 We apply a distance prior uniform in comoving volume \mi{@Thomas is this right?}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.

### 2023-01-31 src/tex/ms.tex
Update content

@@
 \begin{abstract}
-    The propagation of gravitational waves can reveal the fundamental feature of the structure of spacetime.
+    The propagation of gravitational waves can reveal fundamental features of the structure of spacetime.
     For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun for parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
@@
 Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3.
-% GW amplitude birefringence is a property of space-time that consists of the enhancement of one GW polarization over the other as the wave propagates.
 
@@
 % Section guide
+In Sec.~\ref{sec:Background}, we briefly review the background of \ac{GW} amplitude birefringence.
 In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, mention the configuration we used in the \ac{PE} and show the method we used to obtain the population constraint on \ac{GW} amplitude birefringence.
@@
 where $h$ is the frequency domain strain, with the plus and minus signs for L and R respectively.
-These circular modes represent eigenstates of the helicity operator (helicity $\pm2$) and posses a definite parity.
+These circular modes represent eigenstates of the helicity operator (helicity $\pm2$) and possess a definite parity.
 Einstein's theory, which is parity even, predicts no difference in the dynamics of these two states.
@@
 We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, with \textsc{IMRPhenomXPHM} as the reference waveform.
-We apply a distance prior uniform in comoving volume \mi{@Thomas is this right?}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+We apply a distance prior uniform in comoving volume instead of a power law distribution used in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
 
-\mi{@Thomas list here any special considerations for individual events}
+For GW190521, we increase the maximum distance of the distance prior to $1.5$ times the original value, as the birefringence effect enables the probability of a larger distance.
 
 In order to validate our \ac{PE} implementation, we reproduce the \ac{LVK} \ac{PE} results obtained assuming \ac{GR} by enforcing $\kappa = 0$; this also has the advantage of producing \ac{GR} runs that are directly comparable to our birefringent runs.
-All our \ac{PE} results, including the \ac{GR} valudation runs, are published in \citet{dataset}.
+All our \ac{PE} results, including the \ac{GR} validation runs, are published in \citet{dataset}.
 
@@
 \end{equation}
-%
+
 We also calculate the expected population distribution of $\kappa$ marginalized over $\mu$ and $\sigma$.
@@
 Then, we will discuss some special events individually.
-% What special treatment do we give to events
 
@@
 The median and confidence interval of $\kappa$ are calculated with the restricted posterior distribution of $\kappa$, which is given by equation \ref{eq:restricted_posterior}.
-We excluded GW200129 from the calculation of the median and confidence interval of $\kappa$.
+We excluded GW200129 from the calculation of collective analysis.
 Details of the exclusion are discussed in Sec.~\ref{sec:GW200129}.
@@
 The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.
-% talk about what events pull sigma in figure 2
-% create a session for all the outliers
 
@@
 Note that the result is consistent with GR when $\kappa=0$.
+% talk about what events pull sigma in figure 2
 
@@
 
+% Create sessions for all the outliers
+
 % Case: GW170818
+\subsection{Result on GW170818}
 
@@
 Therefore, the narrow frequency range can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
-% 0 is still in the 90%. there is much probability that kappa is not 0 due to degeneracy.
+Note that $\kappa=0$ is still within the $90\%$ confidence interval, with a non-negligible probability.
+The bimodal $\kappa$ distribution only means that the degeneracy between $\kappa$ and $\iota$ cannot be broken in this case.
 
@@
 We pick only two out of three detector data in each set to perform PE.
+(i.e.
+LIGO-Hanford and LIGO-Livingston for the first set, LIGO-Hanford and Virgo for the second set, and LIGO-Livingston and Virgo for the third set.)
 The two sets of PE results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
 Further investigation is needed to understand the glitch.
-Thus, we did not include this event in our analysis.
-% show result included and excluded
-% exclusion from what?
+Thus, we did not include this event in our collective analysis.
+% show results included and excluded
 

### 2023-02-07 src/scripts/corner_GW170818.py
Add placeholder to Appendix

@@ -1,0 +1,71 @@
+import h5py
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"],
+})
+
+nsamples = 5000
+
+with h5py.File(paths.data/"IGWN-GWTC2p1-v2-GW170818_022509_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
+    f = f['C01:IMRPhenomXPHM']
+    result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
+result_ligo = result_ligo.sample(n=nsamples)
+result_ligo['kappa'] = np.zeros(len(result_ligo))
+result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
+
+result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_bilby = result_bilby[result_bilby.event == "GW170818"]
+result_bilby = result_bilby.sample(n=nsamples)
+result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
+result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+
+result = pd.concat([result_bilby,result_ligo], ignore_index=True)
+
+g = sns.pairplot(result,
+                    vars=['kappa', 'mass_ratio', 'chirp_mass', 'luminosity_distance',
+                            'dec', 'ra', 'theta_jn', 'psi', 'a_1','a_2','tilt_1','tilt_2','phi_12','phi_jl',
+                            'phase', 'geocent_time', 'log_likelihood',
+                            ],
+                    corner=True, kind='kde', hue='with',
+                    diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[
+                            (1.-0.90),(1.-0.3935)]))
+
[…]

### 2023-02-07 src/scripts/corner_GW200129.py

@@ -1,0 +1,71 @@
+import h5py
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"],
+})
+
+nsamples = 5000
+
+with h5py.File(paths.data/"IGWN-GWTC3p0-v1-GW200129_065458_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
+    f = f['C01:IMRPhenomXPHM']
+    result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
+result_ligo = result_ligo.sample(n=nsamples)
+result_ligo['kappa'] = np.zeros(len(result_ligo))
+result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
+
+result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_bilby = result_bilby[result_bilby.event == "GW200129_065458"]
+result_bilby = result_bilby.sample(n=nsamples)
+result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
+result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+
+result = pd.concat([result_bilby,result_ligo], ignore_index=True)
+
+g = sns.pairplot(result,
+                    vars=['kappa', 'mass_ratio', 'chirp_mass', 'luminosity_distance',
+                            'dec', 'ra', 'theta_jn', 'psi', 'a_1','a_2','tilt_1','tilt_2','phi_12','phi_jl',
+                            'phase', 'geocent_time', 'log_likelihood',
+                            ],
+                    corner=True, kind='kde', hue='with',
+                    diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[
+                            (1.-0.90),(1.-0.3935)]))
+
[…]

### 2023-02-07 src/scripts/corner_GW200224.py

@@ -1,0 +1,71 @@
+import h5py
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"],
+})
+
+nsamples = 5000
+
+with h5py.File(paths.data/"IGWN-GWTC3p0-v1-GW200224_222234_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
+    f = f['C01:IMRPhenomXPHM']
+    result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
+result_ligo = result_ligo.sample(n=nsamples)
+result_ligo['kappa'] = np.zeros(len(result_ligo))
+result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
+
+result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_bilby = result_bilby[result_bilby.event == "GW200224_222234"]
+result_bilby = result_bilby.sample(n=nsamples)
+result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
+result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+
+result = pd.concat([result_bilby,result_ligo], ignore_index=True)
+
+g = sns.pairplot(result,
+                    vars=['kappa', 'mass_ratio', 'chirp_mass', 'luminosity_distance',
+                            'dec', 'ra', 'theta_jn', 'psi', 'a_1','a_2','tilt_1','tilt_2','phi_12','phi_jl',
+                            'phase', 'geocent_time', 'log_likelihood',
+                            ],
+                    corner=True, kind='kde', hue='with',
+                    diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[
+                            (1.-0.90),(1.-0.3935)]))
+
[…]

### 2023-02-07 src/scripts/corner_GW170818.py
Update violin plot

@@ -13,7 +13,7 @@ plt.rcParams.update({
     "font.serif": ["Computer Modern Roman"],
 })
 
-nsamples = 5000
+nsamples = 3000
 
 with h5py.File(paths.data/"IGWN-GWTC2p1-v2-GW170818_022509_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
     f = f['C01:IMRPhenomXPHM']

### 2023-02-07 src/scripts/corner_GW200129.py

@@ -13,7 +13,7 @@ plt.rcParams.update({
     "font.serif": ["Computer Modern Roman"],
 })
 
-nsamples = 5000
+nsamples = 3000
 
 with h5py.File(paths.data/"IGWN-GWTC3p0-v1-GW200129_065458_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
     f = f['C01:IMRPhenomXPHM']

### 2023-02-07 src/scripts/violin_kappa.py

@@ -58,5 +58,7 @@ g.figure.colorbar(plt.cm.ScalarMappable(norm=mpl.colors.Normalize(color_DataFram
                   label="$|\mathrm{median}/\sigma|$",
                   ax=g.axes
                  )
+g.axes.collections[59].set_linewidth(3) # GW200129_065458
+g.axes.collections[59].set_edgecolor(sns.color_palette()[9])
 
 g.figure.savefig(fname=paths.figures/"violin_kappa.pdf", bbox_inches="tight", dpi=300)

### 2023-02-09 src/tex/ms.tex
Add placeholder to Appendix
Update violin plot
Edit figures

@@
 \begin{figure}
-    \script{birefringence.py}
-    \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
+    \script{birefringence_frequency_domain.py}
+    \includegraphics[width=\columnwidth]{figures/birefringence_frequency_domain.pdf}
     \caption{
@@
     }
-    \label{fig:birefringence}
+    \label{fig:birefringence_frequency_domain}
 \end{figure}
@@
 We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, with \textsc{IMRPhenomXPHM} as the reference waveform.
-We apply a distance prior uniform in comoving volume instead of a power law distribution used in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+We apply a distance prior which is uniform in comoving volume and source frame time instead of a power law distribution used in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
 
@@
         The pink horizontal solid line represents $\kappa=0$.
+        GW200129 is shown with light blue edges, as it is excluded from the calculation of collective analysis.
     }
@@
 \end{figure*}
-% show the excluded events visually different
 
@@
 \appendix
-\section{}
+\label{sec:corner_plots}
+\begin{figure}[h]
+    \script{corner_GW170818.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
+    \caption{Corner plot of GW170818.}
+    \label{fig:corner_GW170818}
+\end{figure}
+
+\begin{figure}[h]
+    \script{corner_GW200129.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
+    \caption{Corner plot of GW200129.}
+    \label{fig:corner_GW200129}
+\end{figure}
+
+\begin{figure}[h]
+    \script{corner_GW200224.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW200224.pdf}
+    \caption{Corner plot of GW200224.}
+    \label{fig:corner_GW200224}
+\end{figure}
+
+% \begin{figure}
+%     \script{birefringence_time_domain.py}
+%     \includegraphics[width=\columnwidth]{figures/birefringence_time_domain.pdf}
+%     \caption{
+%         % The observed Fourier amplitude of two polarizations of a simulated GW signal with and without birefringence.
+%         % The signal, as GR predicts, is shown in the upper panel, and the signal with birefringence is shown in the lower panel.
+%         % The signals in the upper panel are shown in dashed blue and dotted orange lines representing the left-handed and right-handed polarizations.
+%         % Note that the 
+%         % The amplitudes with birefringence are shown in solid lines with blue and orange colors representing the left-handed and right-handed polarizations.
+%         This plot shows the effect of amplitude birefringence with frequency dependence.
+%     }
+%     \label{fig:birefringence_time_domain}
+% \end{figure}
 

### 2023-02-09 src/scripts/birefringence_time_domain.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import lalsimulation as lalsim
+import lal
+import seaborn as sns
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"],
+})
+
+approximant = lalsim.SimInspiralGetApproximantFromString("IMRPhenomD")
+
+# frequency array parameters
+f_min = 20
+f_ref = 100
+
+# source parameters
+m1_msun = 10
+m2_msun = 10
+chi1 = [0, 0, 0]
+chi2 = [0, 0, 0]
+dist_mpc = 1000
+inclination = np.pi * 0.9
+phi_ref = 0
+
+m1_kg = m1_msun*lal.MSUN_SI
+m2_kg = m2_msun*lal.MSUN_SI
+distance = dist_mpc*1e6*lal.PC_SI
+
+# time array parameters
+dt = 1e-4
+
+hp_time, hc_time = lalsim.SimInspiralChooseTDWaveform(m1_kg, m2_kg,
+                                            chi1[0], chi1[1], chi1[2],
+                                            chi2[0], chi2[1], chi2[2],
+                                            distance, inclination,
+                                            phi_ref, 0, 0., 0.,
+                                            dt, f_min, f_ref,
+                                            None, approximant)
+
+hl_time = (hp_time.data.data + (hc_time.data.data * 1j)) /np.sqrt(2)
+hr_time = (hp_time.data.data - (hc_time.data.data * 1j)) /np.sqrt(2)
+
+time = np.arange(len(hl_time))*dt
+
+hp_freq_fft = np.fft.fft(hp_time.data.data, norm="ortho")
+hc_freq_fft = np.fft.fft(hc_time.data.data, norm="ortho")
+
+n = np.fft.fftfreq(len(hp_freq_fft))
+freq_fft = n[:len(n)//2]*len(n)//4
+
+hl_freq_fft = (hp_freq_fft + (hc_freq_fft * 1j)) /np.sqrt(2)
+hr_freq_fft = (hp_freq_fft - (hc_freq_fft * 1j)) /np.sqrt(2)
+
[…]

### 2023-02-09 src/scripts/corner_Gaussian.py

@@ -2,6 +2,7 @@ import matplotlib.pyplot as plt
 import pandas as pd
 import numpy as np
 import seaborn as sns
+import kde_contour as kde_contour
 import paths
 
 sns.set_theme(palette='colorblind', font_scale=1.2)
@@ -22,20 +23,48 @@ df['mu'] = samples_Gaussian.reshape(-1,2)[:,0]
 df['sigma'] = samples_Gaussian.reshape(-1,2)[:,1]
 df = df.sample(10000)
 
-g = sns.pairplot(df,
-                vars=['mu', 'sigma'],
-                corner=True, kind='kde',
-                diag_kws=dict(common_norm=False, cut=0), plot_kws=dict(common_norm=False, levels=[
-                        (1.-0.90),(1.-0.3935)], cut=0))
-g.axes[0,0].axvline(mu_median, color=sns.color_palette()[3])
-g.axes[1,1].axvline(sigma_median, color=sns.color_palette()[3])
-g.axes[1,0].axvline(mu_median, color=sns.color_palette()[3])
-g.axes[1,0].axhline(sigma_median, color=sns.color_palette()[3])
+kde_1d_sigma = kde_contour.Bounded_1d_kde(df['sigma'], xlow=0)
+df['sigma_kde'] = kde_1d_sigma(df['sigma'])
+df = df.sort_values(['sigma'])
 
-g.axes[1,1].set_xlabel("$\\sigma$")
-g.axes[1,0].set_xlabel("$\\mu$")
-g.axes[1,0].set_ylabel("$\\sigma$")
+fig = plt.figure(figsize=(5,5))
+ax1 = fig.add_subplot(221)
+ax3 = fig.add_subplot(223)
+ax4 = fig.add_subplot(224)
 
-g.axes[1,1].set_xlim(0.0)
+ax1.axvline(mu_median, color=sns.color_palette()[3])
+sns.kdeplot(df['mu'], fill=True, ax=ax1)
 
-g.savefig(fname=paths.figures/"corner_Gaussian.pdf", bbox_inches="tight", dpi=300)
+ax1.set_xlabel("")
+ax1.set_ylabel("")
+ax1.set_xlim(-.07,.07)
+ax1.set_ylim(0)
+ax1.set_xticks([])
+ax1.set_yticks([])
+ax1.set_box_aspect(1)
+
[…]

### 2023-02-09 src/scripts/kde_contour.py

@@ -1,0 +1,245 @@
+# The following script was copied from
+# https://github.com/maxisi/ringdown/blob/c5388c010049dea59b1b54d02989ea9c19f4b12d/ringdown/kde_contour.py
+
+__all__ = ['Bounded_2d_kde', 'Bounded_1d_kde', 'kdeplot_2d_clevels']
+
+from pylab import *
+import scipy.stats as ss
+import seaborn as sns
+
+# The following routine, Bounded_2d_kde, was copied from
+# https://git.ligo.org/publications/gw190412/gw190412-discovery/-/blob/851f91431b7c36e7ea66fa47e8516f2aef9d7daf/scripts/bounded_2d_kde.py
+class Bounded_2d_kde(ss.gaussian_kde):
+    r"""Represents a two-dimensional Gaussian kernel density estimator
+    for a probability distribution function that exists on a bounded
+    domain."""
+
+    def __init__(self, pts, xlow=None, xhigh=None, ylow=None, yhigh=None,
+                 *args, **kwargs):
+        """Initialize with the given bounds.  Either ``low`` or
+        ``high`` may be ``None`` if the bounds are one-sided.  Extra
+        parameters are passed to :class:`gaussian_kde`.
+
+        :param xlow: The lower x domain boundary.
+
+        :param xhigh: The upper x domain boundary.
+
+        :param ylow: The lower y domain boundary.
+
+        :param yhigh: The upper y domain boundary.
+        """
+        pts = np.atleast_2d(pts)
+
+        assert pts.ndim == 2, 'Bounded_kde can only be two-dimensional'
+
+        super(Bounded_2d_kde, self).__init__(pts.T, *args, **kwargs)
+
+        self._xlow = xlow
+        self._xhigh = xhigh
+        self._ylow = ylow
+        self._yhigh = yhigh
+
+    @property
+    def xlow(self):
+        """The lower bound of the x domain."""
+        return self._xlow
+
+    @property
+    def xhigh(self):
+        """The upper bound of the x domain."""
+        return self._xhigh
+
+    @property
+    def ylow(self):
+        """The lower bound of the y domain."""
+        return self._ylow
+
+    @property
+    def yhigh(self):
+        """The upper bound of the y domain."""
+        return self._yhigh
+
+    def evaluate(self, pts):
+        """Return an estimate of the density evaluated at the given
+        points."""
+        pts = np.atleast_2d(pts)
[…]

### 2023-02-10 src/tex/ms.tex
Add time domain figure
Added derivation of M_PV

@@
 \begin{equation}
-    h_{{L/R}} = \frac{1}{\sqrt{2}}\left(h_+ \pm i h_\times\right)\,,
+    h_{L/R} = \frac{1}{\sqrt{2}}\left(h_+ \pm i h_\times\right)\,,
 \end{equation}
@@
 \end{equation}
-For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will generate dephasing in the time domain waveform (Fig~\ref{fig:birefringence}).
+For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will generate dephasing in the time domain waveform (Fig~\ref{fig:birefringence_frequency_domain}).
 
@@
 
+% Case: GW190412
+\subsection{Result on GW190412}
+
+% Case: GW200224
+\subsection{Result on GW200224}
+
 % Case: GW150914 (example of broken degeneracy)
@@
 
-% Comparison with the previous paper with the equation related to the parameter they used
+\citet{Wang_2021} performed PE on GWTC-1 events with a frequency-dependent birefringence model.
+They formulated the birefringence effect as corrections to the GW waveform on a linear basis.
+We convert their constraint to the same units as ours with the derivation in Appendix~\ref{sec:M_PV_derivation}.
+% Discussion on the difference between our work and theirs
 
@@
 \appendix
+
+\section{Relation between $\kappa$ and $M_{PV}$}
+\label{sec:M_PV_derivation}
+
+\citet{Wang_2021} formulated the birefringence effect as
+\begin{equation}
+    h_{+/\times}^{PV}(f) = h_{+/\times}^{GR}(f)\mp h_{\times/+}^{GR}(f)(i\delta h-\delta\Psi)\,,
+\end{equation}
+where $h_{+/\times}^{PV}(f)$ is the modified \ac{GW} waveform in the linear basis, $h_{+/\times}^{GR}(f)$ is the \ac{GW} waveform in the linear basis as \ac{GR} predicts, $\delta h$ is the amplitude modification, and $\delta\Psi$ is the phase modification.
+We can rewrite the equation as
+\begin{equation}
+\begin{split}
+    h_{L/R}^{PV}(f)&=h_{L/R}^{GR}(f)(1\mp\delta h\mp i\delta\Psi)\\
+    &\approx h_{L/R}^{GR}(f)e^{(\mp\delta h\mp i\delta\Psi)}
+\end{split}
+\,,
+\end{equation}
+where $h_{L/R}^{PV}(f)$ is the modified \ac{GW} waveform in the circular basis, $h_{L/R}^{GR}(f)$ is the \ac{GW} waveform in the circular basis as \ac{GR} predicts.
+This equation applies to any birefringence model that allows the assumption that the modification to \ac{GR} is small when observed very near the source.
+To make a comparison with our work, we can choose $\delta\Psi=0$, which is a case that \citet{Wang_2021} considered.
+In their work, they parametrized $\delta h=-A_\nu\pi f$, while
+\begin{equation}
+    A_\nu=M_{PV}^{-1}(\alpha_\nu(z=0)-\alpha_\nu(z)(1+z))\,,
+\end{equation}
+where $A_\nu$ is the parity-violating parameter for amplitude birefringence, $M_{PV}$ is the characteristic energy scale, and $\alpha_\nu$ is a arbitrary functions determined by the birefringence model.
+They then chose $\alpha_\nu=1$.
+Therefore, we can rewrite the equation as
+\begin{equation}
+    h_{L/R}^{PV}(f)= h_{L/R}^{GR}(f)\exp\left(\mp\frac{\pi zf}{M_{PV}}\right)\,.
+\end{equation}
+This equation takes the same form as \eqref{eq:waveform_modification}.
+We can then find the relation between $M_{PV}$ and $\kappa$ by
+\begin{equation}
+    \left|\mp\frac{\pi zf}{M_{PV}}\right|=\left|\pm\kappa\frac{d_C}{1\,\mathrm{Gpc}}\frac{f}{100\,\mathrm{Hz}}\right|\,.
+\end{equation}
+Using $H_0d_C=cz$, where $H_0$ is the Hubble constant and $c$ is the speed of light, we can rewrite the equation as
+\begin{equation}
+    \kappa=\frac{\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{PV}^{-1}\,.
+\end{equation}
+
+\section{Illustration of birefringence}
+\label{sec:time_domain}
+
+\begin{figure}[h]
+    \script{birefringence_time_domain.py}
+    \includegraphics[width=\columnwidth]{figures/birefringence_time_domain.pdf}
+    \caption{The observed time domain signal with and without birefringence.}
+    \label{fig:birefringence_time_domain}
+\end{figure}
+
+\section{Corner plots of special events}
 \label{sec:corner_plots}
+
 \begin{figure}[h]
@@
 
-% \begin{figure}
-%     \script{birefringence_time_domain.py}
-%     \includegraphics[width=\columnwidth]{figures/birefringence_time_domain.pdf}
-%     \caption{
-%         % The observed Fourier amplitude of two polarizations of a simulated GW signal with and without birefringence.
-%         % The signal, as GR predicts, is shown in the upper panel, and the signal with birefringence is shown in the lower panel.
-%         % The signals in the upper panel are shown in dashed blue and dotted orange lines representing the left-handed and right-handed polarizations.
-%         % Note that the 
-%         % The amplitudes with birefringence are shown in solid lines with blue and orange colors representing the left-handed and right-handed polarizations.
-%         This plot shows the effect of amplitude birefringence with frequency dependence.
-%     }
-%     \label{fig:birefringence_time_domain}
-% \end{figure}
-
 \bibliography{bib}

### 2023-02-10 src/tex/ms.tex
Update ms.tex
WIP updated Fig 1

@@
 Among those, measurements of the basic properties of \acp{GW}, like their speed and polarization, can directly probe the fundamental symmetries of the underlying theory of gravity \citep{Will:2018bme}.
-For instance, unequal propagation of \ac{GW} polarization states would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm,Alexander:2009tp,Sopuerta:2009iy}.
+For instance, unequal propagation of \ac{GW} polarization eigenstates would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm,Alexander:2009tp,Sopuerta:2009iy}.
 Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3.
@@
 % What's new?
-In this study, we used a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE}.
+In this study, we used a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE} on \ac{LVK} binaries.
 This model is a better approximation of the birefringence effect than the frequency-independent model used in \citet{Okounkova_2022}.
 Compared to previous studies, we also performed \ac{PE} on more events, including events in the third \ac{GW} transient catalog \citep{GWTC-3}, GWTC-3.
-We consider 70 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$.
-These events are also listed in Table I in \citet{GWTC-3_population}.\footnote{
-GW190720 is included in the list in \citet{GWTC-3_population}, but we did not include it in this study.
-Details in Sec.~\ref{sec:GW190720}.}
-We then give the population constraint on the strength of \ac{GW} amplitude birefringence by using the \ac{PE} results.
+We consider 70 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$, as listed in Table I of \citet{GWTC-3_population}.
+We use the results from individual events to place a collective population constraint on the strength of \ac{GW} amplitude birefringence from GWTC-3.
 
@@
 The overall magnitude of this effect for a given distance and frequency $f$ is set by a dimensionless opacity parameter, $\kappa$, which encodes the intrinsic strength of the birefringence.
-The emitted waveform for a given source (i.e., the waveform observed very close to the source) will generally differ from the analogous waveform predicted by \ac{GR}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small, it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
+The emitted waveform for a given source (i.e., the waveform observed in the near zone, very close to the source) will generally differ from the analogous waveform predicted by \ac{GR}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small, it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
 
@@
 
-Equation \eqref{eq:waveform_modification} can be derived as the first order effect in an expansion for multiple extensions of \ac{GR}.
+Equation \eqref{eq:waveform_modification} can be derived as the first order effect in an expansion away from \ac{GR} under multiple frameworks.
 In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity.
-Since it originates from an expansion, Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
+Since it originates from a truncated series expansion, Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
 \begin{equation}
@@
     h_{L/R}^{PV}(f)&=h_{L/R}^{GR}(f)(1\mp\delta h\mp i\delta\Psi)\\
-    &\approx h_{L/R}^{GR}(f)e^{(\mp\delta h\mp i\delta\Psi)}
+    &\approx h_{L/R}^{GR}(f)\exp\left(\mp\delta h\mp i\delta\Psi\right)
 \end{split}

### 2023-02-10 src/scripts/birefringence_frequency_domain.py

@@ -16,9 +16,9 @@ plt.rcParams.update({
 approximant = lalsim.SimInspiralGetApproximantFromString("IMRPhenomD")
 
 # frequency array parameters
-df = 0.25
+df = 0.125
 f_min = 20
-f_max = 2030
+f_max = 2048
 f_ref = 100
 
 # source parameters
@@ -27,8 +27,11 @@ m2_msun = 10
 chi1 = [0, 0, 0]
 chi2 = [0, 0, 0]
 dist_mpc = 400
-inclination = np.pi / 4
+inclination = np.pi/2
 phi_ref = 0
+
+# birefringence
+kappa = 0.6
 
 m1_kg = m1_msun*lal.MSUN_SI
 m2_kg = m2_msun*lal.MSUN_SI
@@ -42,25 +45,63 @@ hp_freq, hc_freq = lalsim.SimInspiralChooseFDWaveform(m1_kg, m2_kg,
                                             df, f_min, f_max, f_ref,
                                             None, approximant)
 
-hl_freq = (hp_freq.data.data + (hc_freq.data.data * 1j)) /np.sqrt(2)
-hr_freq = (hp_freq.data.data - (hc_freq.data.data * 1j)) /np.sqrt(2)
+freq = np.arange(len(hp_freq.data.data))*df
 
-freq = np.arange(len(hl_freq))*df
+# let's shift the peak of the waveform `tshift` s before the end of the segment
+tshift = 1
+hp_fd = hp_freq.data.data * np.exp(2j*np.pi*freq*tshift)
+hc_fd = hc_freq.data.data * np.exp(2j*np.pi*freq*tshift)
 
-kappa = -0.5
+hl_fd = (hp_fd + (hc_fd * 1j)) /np.sqrt(2)
+hr_fd = (hp_fd - (hc_fd * 1j)) /np.sqrt(2)
 
-hl_bi_freq = hl_freq * np.exp(kappa*(dist_mpc/1000)*(freq/100))
-hr_bi_freq = hr_freq * np.exp(-kappa*(dist_mpc/1000)*(freq/100))
+# apply birefringent transformation and get corresponding plus and cross
+hl_bi_fd = hl_fd * np.exp(kappa*(dist_mpc/1000)*(freq/100))
[…]

### 2023-02-11 src/tex/ms.tex
Better naming

@@
 \end{equation}
-For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will generate dephasing in the time domain waveform (Fig~\ref{fig:birefringence_frequency_domain}).
+For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will generate dephasing in the time domain waveform (Fig~\ref{fig:birefringence}).
 
 \begin{figure}
-    \script{birefringence_frequency_domain.py}
-    \includegraphics[width=\columnwidth]{figures/birefringence_frequency_domain.pdf}
+    \script{birefringence.py}
+    \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
     \caption{
@@
     }
-    \label{fig:birefringence_frequency_domain}
+    \label{fig:birefringence}
 \end{figure}
@@
 \section{Illustration of birefringence}
-\label{sec:time_domain}
+\label{sec:birefringence_appendix}
 
 \begin{figure}[h]
-    \script{birefringence_time_domain.py}
-    \includegraphics[width=\columnwidth]{figures/birefringence_time_domain.pdf}
+    \script{birefringence_appendix.py}
+    \includegraphics[width=\columnwidth]{figures/birefringence_appendix.pdf}
     \caption{The observed time domain signal with and without birefringence.}
-    \label{fig:birefringence_time_domain}
+    \label{fig:birefringence_appendix}
 \end{figure}

### 2023-02-16 src/tex/ms.tex
Add GW190720

@@
     Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
-    From 70 binary-black-hole signals, we obtain the most precise constraints on amplitude birefringence yet, an order of magnitude more stringent than previous results.
+    From 71 binary-black-hole signals, we obtain the most precise constraints on amplitude birefringence yet, an order of magnitude more stringent than previous results.
 \end{abstract}
@@
 Compared to previous studies, we also performed \ac{PE} on more events, including events in the third \ac{GW} transient catalog \citep{GWTC-3}, GWTC-3.
-We consider 70 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$, as listed in Table I of \citet{GWTC-3_population}.
+We consider 71 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$, as listed in Table I of \citet{GWTC-3_population}.
 We use the results from individual events to place a collective population constraint on the strength of \ac{GW} amplitude birefringence from GWTC-3.
@@
         This plot shows the effect of amplitude birefringence with frequency dependence.
-        \mi{@Thomas can you add a panel with the time-domain waveforms (maybe just for L or R)?}
     }
@@
 To constrain birefringence, we reanalyze events from GWTC-3 \citep{GWTC-2.1, GWTC-3} implementing Eq.~\eqref{eq:waveform_modification} to directly obtain a posterior on $\kappa$ from the strain of each event.
-We analyze the 70 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$; to avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
+We analyze the 71 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$; to avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
 We procure strain data from \ac{GWOSC} \citep{GWOSC}.
@@
     \caption{
-        The violin plot shows the posterior of $\kappa$ for 70 events in this study.
+        The violin plot shows the posterior of $\kappa$ for 71 events in this study.
         Each violin represents a different event.
@@
 
-% Case: GW190720 (failed PE)
-\subsection{Exclusion of GW190720}
-\label{sec:GW190720}
-For GW190720, we could not perform PE with Bilby successfully.
-In LVK PE papers \citep{GWTC-2.1, GWTC-3}, the PE of some events was performed with LALInference \citep{lalsuite} instead of Bilby.
-GW190720 was one of them.
-We successfully migrated the LALInference configuration to Bilby for most of these events and recovered similar PE results, except for GW190720.
-Bilby could not evaluate the likelihood function on Virgo data.
-As a result, we excluded this event from our analysis.
-
 \section{Discussion}
@@
 
-\begin{figure}[h]
+\begin{figure}
     \script{corner_GW170818.py}
@@
 
-\begin{figure}[h]
+\begin{figure}
+    \script{corner_GW190720.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW190720.pdf}
+    \caption{Corner plot of GW190720.}
+    \label{fig:corner_GW190720}
+\end{figure}
+
+\begin{figure}
     \script{corner_GW200129.py}
@@
 
-\begin{figure}[h]
+\begin{figure}
     \script{corner_GW200224.py}

### 2023-02-16 src/tex/ms.tex
Remove corner plot of GW190720
fig 1 caption

@@
 For instance, unequal propagation of \ac{GW} polarization eigenstates would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm,Alexander:2009tp,Sopuerta:2009iy}.
-Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3.
+Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3 \citep{GWTC-3}.
 
@@
 % What's new?
-In this study, we used a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE} on \ac{LVK} binaries.
-This model is a better approximation of the birefringence effect than the frequency-independent model used in \citet{Okounkova_2022}.
-Compared to previous studies, we also performed \ac{PE} on more events, including events in the third \ac{GW} transient catalog \citep{GWTC-3}, GWTC-3.
+In this study, we use a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE} on \ac{LVK} binaries.
+This model is a better approximation of the birefringence effect expected from theory than the frequency-independent model used in \citet{Okounkova_2022}.
+Compared to other studies, we perform \ac{PE} on more events, including events new to the third \ac{GW} transient catalog \citep{GWTC-3}, GWTC-3.
 We consider 71 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$, as listed in Table I of \citet{GWTC-3_population}.
@@
 In \ac{GR}, \acp{GW} are comprised of two independent polarization modes, usually represented in the linear basis of plus ($+$) and cross ($\times$) states.
-In the Fourier domain, they can be combined into left-handed (L) and right-handed (R) circular states \cite{Isi:2022mbx},
+In the Fourier domain, these can be combined into left-handed (L) and right-handed (R) circular states \cite{Isi:2022mbx},
 \begin{equation}
@@
 % Waveform modification
-Yet, parity odd extensions of \ac{GR} may make distinctions between the two circular polarizations, which can manifest in both the generation and propagation of \acp{GW}.
+Yet, parity odd extensions of \ac{GR} may make distinctions between the two circular polarizations, which appear in both the generation and propagation of \acp{GW}.
 The latter can manifest in changes to the relative amplitude and phase of the polarizations that accrue as the wave propagates, giving us hope to detect initially small effects that compound over long propagation distances.
@@
 In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
-To first order in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source takes the form of
+To first order in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
 \begin{equation}
@@
 \end{equation}
-For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will generate dephasing in the time domain waveform (Fig~\ref{fig:birefringence}).
+For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will affect the time domain waveform nontrivially (Fig~\ref{fig:birefringence}).
 
@@
     \caption{
-        The observed Fourier amplitude of two polarizations of a simulated GW signal with and without birefringence.
-        The amplitudes, as GR predicts, are shown in dotted lines with blue and orange colors representing the left-handed and right-handed polarizations.
-        The amplitudes with birefringence are shown in solid lines with blue and orange colors representing the left-handed and right-handed polarizations.
-        This plot shows the effect of amplitude birefringence with frequency dependence.
+        \emph{Illustration of amplitude birefringence.} The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is lineraly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
+        However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
+        In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
+        For this example, the black holes do not spin and have equal masses $m_1 = m_2 = 10\, M_\odot$, and we have chosen a luminosity distance $d_L = 400\, {\rm Mpc}$ and $\kappa = 0.6$.
         }
@@
 \begin{figure}
-    \script{corner_GW190720.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW190720.pdf}
-    \caption{Corner plot of GW190720.}
-    \label{fig:corner_GW190720}
-\end{figure}
-
-\begin{figure}
     \script{corner_GW200129.py}

### 2023-02-16 src/scripts/birefringence.py

@@ -5,7 +5,7 @@ import lal
 import seaborn as sns
 import paths
 
-sns.set_theme(palette='colorblind', font_scale=1.2)
+sns.set_theme(palette='colorblind', font_scale=1.5)
 
 plt.rcParams.update({
     "text.usetex": True,
@@ -86,7 +86,7 @@ axs[0].set_ylim(1e-30,1e-20)
 axs[0].legend()
 axs[0].set_ylabel(r"$|\tilde{h}|$")
 axs[0].set_xlabel(r"$f\mathrm{(Hz)}$")
-axs[0].set_yticks([])
+#axs[0].set_yticks([])
 
 axs[1].plot(time, hp_td, ls=':', c=c1, label="GR")
 axs[1].plot(time, hp_bi_td, ls='-', c=c2, label="BR")
@@ -100,7 +100,7 @@ for ax in axs[1:]:
     ax.legend()
     ax.set_xlabel(r"$t\mathrm{(s)}$")
     ax.set_xlim(-0.05, 0.01)
-    ax.set_yticks([])
+ #   ax.set_yticks([])
 
 fig.tight_layout(pad=0.3)
 fig.savefig(paths.figures/"birefringence.pdf", bbox_inches="tight")

### 2023-02-16 src/tex/ms.tex
statement about GW190720 seglen

@@
 We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
-We depart from the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, with \textsc{IMRPhenomXPHM} as the reference waveform.
-We apply a distance prior which is uniform in comoving volume and source frame time instead of a power law distribution used in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+We take the \ac{PE} configuration as in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
+We apply a distance prior which is uniform in comoving volume and source frame time instead of a power law distribution used in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
 
-For GW190521, we increase the maximum distance of the distance prior to $1.5$ times the original value, as the birefringence effect enables the probability of a larger distance.
+For GW190521, we increase the maximum distance of the distance prior to $1.5\times$ the original value, as the birefringence effect enables the probability of a larger distance.
+For GW190720, we decrease the analysis segment to 8 s from 16 s, in order to accomodate missing data near the edges of the 16 s segment in Virgo.
 

### 2023-02-16 src/tex/ms.tex
vertical violin plot

@@
 
-\begin{figure*}
+\begin{figure}
     \script{violin_kappa.py}
-    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
+    \includegraphics[width=\columnwidth]{figures/violin_kappa.pdf}
     \caption{
-        The violin plot shows the posterior of $\kappa$ for 71 events in this study.
-        Each violin represents a different event.
-        The color of the violins represents the quotient of the median and standard deviation of the posterior.
-        The blue horizontal solid line represents the median of the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval of the restricted posterior of $\kappa$.
-        The pink horizontal solid line represents $\kappa=0$.
-        GW200129 is shown with light blue edges, as it is excluded from the calculation of collective analysis.
+        Individual-event $\kappa$ posteriors (distributions), and joint measurement (blue band, 90\% CI; blue line, median).
+%        Each violin represents a different event.
+%        The color of the violins represents the quotient of the median and standard deviation of the posterior.
+%        The blue horizontal solid line represents the median of the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval of the restricted posterior of $\kappa$.
+%        The pink horizontal solid line represents $\kappa=0$.
+%        GW200129 is shown with light blue edges, as it is excluded from the calculation of collective analysis.
     }
     \label{fig:violin_kappa}
-\end{figure*}
+\end{figure}
 

### 2023-02-16 src/scripts/violin_kappa_horizontal.py

@@ -1,0 +1,64 @@
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import scipy
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"]
+})
+
+result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+events = result_DataFrame['event'].unique()
+
+result_dict = {}
+color_DataFrame = pd.DataFrame(columns=['event', 'median_over_std'])
+i = 0
+for event in events:
+    df = pd.DataFrame()
+    result_dict[event] = result_DataFrame[result_DataFrame.event == event]
+    color_DataFrame.loc[i] = [event, abs(result_dict[event]['kappa'].median()/result_dict[event]['kappa'].std())]
+    i += 1
+color_DataFrame['color_index'] = [round(np.interp(color_DataFrame.loc[i]['median_over_std'], np.array([color_DataFrame['median_over_std'].min(), color_DataFrame['median_over_std'].max()]), np.array([0,255]))) for i in range(len(color_DataFrame))]
+color_dict={}
+for i in range(len(color_DataFrame)):
+    color_dict[color_DataFrame.loc[i]['event']] = sns.color_palette("rocket_r", as_cmap=True).colors[color_DataFrame.loc[i]['color_index']]
+
+kernels = [scipy.stats.gaussian_kde(result_dict[event]['kappa']) for event in events[events != "GW200129_065458"]]
+kappa = np.linspace(-.1, .1, 1000)
+ll = [np.sum([np.log(ker(k)) for ker in kernels]) for k in kappa]
+ll = ll - np.max(ll)
+likelihood = np.exp(ll)
+likelihood = likelihood / np.trapz(likelihood, x=kappa)
+
+g = sns.violinplot(data=result_DataFrame, x="event", y="kappa",
+                    scale="width", inner=None,
+                    palette=color_dict
+                   )
+g.set_ylabel("$\kappa$", fontsize=20)
+g.set_ylim(-0.2,0.2)
+g.set_xlabel("")
+x_low, x_high = g.get_xlim()
+g.set_xlim(x_low-0.5,x_high+0.5)
[…]

### 2023-02-16 src/tex/ms.tex
text on individual events

@@
 \label{sec:Results}
+
 In this section, we present the results of our study.
@@
 \subsection{Result on GWTC-3}
+
 % Violin plot
-In figure \ref{fig:violin_kappa}, we show the posterior of $\kappa$ for each event in the form of a violin plot.
-The median and confidence interval of $\kappa$ are calculated with the restricted posterior distribution of $\kappa$, which is given by equation \ref{eq:restricted_posterior}.
+We present the result of our catalog analysis in Fig.~\ref{fig:violin_kappa}.
+This shows the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate), colored by the distance of the respective posterior mean from the origin in units of standard deviation, i.e., $|\mu/\sigma|$.
+The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line);
+this joint measurement is fully consistent with $\kappa = 0$ (dashed line), meaning we find no evidence for birefringence.
+
+Figure \ref{fig:violin_kappa} makes it clear that not all \acp{BBH} in GWTC-3 are equally informative about birefringence.
+When considered individually, the events that best constrain $\kappa$ are \mi{THOMAS: list 3-5 events with the narrowest posterior---maybe also add a little table showing their median $\pm$ their 90\% CL, like $\tilde{\kappa}^{+x}_{-y}$ where $\tilde{\kappa}$ is the posterior median value, $x$ is the distance from the median to the 95\% percentile and $y$ is the distance to the 5\% percentile}
+
+The event that shows the largest tension with $\kappa = 0$ is GW170818, for which $|\mu / \sigma| = \dots$ \mi{THOMAS: fill in these values}
+
 We excluded GW200129 from the calculation of collective analysis.

### 2023-02-17 src/tex/ms.tex
removed redundant fig from appendix
Add reweighed kappa distributions
results text

@@
 \subsection{Relation to inclination}
+\label{sec:inclination}
 
@@
 The event that shows the largest tension with $\kappa = 0$ is GW170818, for which $|\mu / \sigma| = \dots$ \mi{THOMAS: fill in these values}
+Additionally, the GW200129 result shows features that could be indicative of contamination from non-Gaussian instrumental noise, so we exclude it from the collective analysis.
+We discuss both these events in detail below.
 
-We excluded GW200129 from the calculation of collective analysis.
-Details of the exclusion are discussed in Sec.~\ref{sec:GW200129}.
+Another set of events stand out in Fig.~\ref{fig:violin_kappa} due evident bimodality in the $\kappa$ posterior.
+That is the case for GW170104, GW190413\_134308, GW190521, GW190805\_211137, and GW191105\_143521.
+\mi{THOMAS: can you list the total masses of these events?}
+We understand this effect to be a consequence of the narrow sets of frequencies that are visible for these heavy events, as we mentioned in Sec.~\ref{sec:inclination} and further demonstrate below.
 
 \subsection{Bayesian hierarchical modeling}
+
+Figure \ref{fig:violin_kappa} shows some variance in the distribution of $\kappa$ posteriors, including some apparent outliers like GW170818.
+This is not unexpected: assuming independent Gaussian noise, we might expect up to ${\sim}3$ events in a sample of 71 events to deviate from the true value of $\kappa$ by ${\sim}2\sigma$.
+\mi{WIP}
+
 % Corner plot of the mean and standard deviation of kappa
@@
 
-\begin{figure}[h]
-    \script{birefringence_appendix.py}
-    \includegraphics[width=\columnwidth]{figures/birefringence_appendix.pdf}
-    \caption{The observed time domain signal with and without birefringence.}
-    \label{fig:birefringence_appendix}
-\end{figure}
-
 \section{Corner plots of special events}
@@
 
+\section{Reweighed kappa distributions}
+\label{sec:reweighed_kappa}
+
+\begin{figure}
+    \script{reweighed_kappa.py}
+    \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
+    \caption{Reweighed $\kappa$ distributions.}
+    \label{fig:reweighed_kappa}
+\end{figure}
+
 \bibliography{bib}

### 2023-02-22 src/tex/ms.tex
Add tables

@@
     \caption{
-        \emph{Illustration of amplitude birefringence.} The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is lineraly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
+        \emph{Illustration of amplitude birefringence.} The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
         However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
@@
 For GW190521, we increase the maximum distance of the distance prior to $1.5\times$ the original value, as the birefringence effect enables the probability of a larger distance.
-For GW190720, we decrease the analysis segment to 8 s from 16 s, in order to accomodate missing data near the edges of the 16 s segment in Virgo.
+For GW190720, we decrease the analysis segment to 8 s from 16 s, in order to accommodate missing data near the edges of the 16 s segment in Virgo.
 
@@
 If \ac{GR} is correct and there is no birefringence, we should find the observed $\kappa$ distribution to be consistent with a delta function at the origin ($\kappa_i = 0$ for all $i$, or $\mu=\sigma=0$); on the other hand, if spacetime is birefringent, we expect to find a delta function at some nonzero value ($\kappa_i = \kappa \neq 0$, or $\mu = \kappa$ and $\sigma=0$).
-But this analysis also has the power to reveal unexpected physics or systematics in our measurements: if $\sigma$ is confidently found to be nonzero, this would imply that our set of measurements are statistically unlikely to originate from a unique $\kappa$ value.
+But this analysis also has the power to reveal unexpected physics or systematics in our measurements: if $\sigma$ is confidently found to be nonzero, this would imply that our set of measurements is statistically unlikely to originate from a unique $\kappa$ value.
 This could signal further physics than is implied by Eq.~\eqref{eq:waveform_modification} \mi{which physics?
@@
 Figure \ref{fig:violin_kappa} makes it clear that not all \acp{BBH} in GWTC-3 are equally informative about birefringence.
-When considered individually, the events that best constrain $\kappa$ are \mi{THOMAS: list 3-5 events with the narrowest posterior---maybe also add a little table showing their median $\pm$ their 90\% CL, like $\tilde{\kappa}^{+x}_{-y}$ where $\tilde{\kappa}$ is the posterior median value, $x$ is the distance from the median to the 95\% percentile and $y$ is the distance to the 5\% percentile}
+When considered individually, the events that best constrain $\kappa$ are listed in Table~\ref{tab:best_events}.
+\begin{table}
+    \begin{ruledtabular}
+        \begin{tabular}{ll}
+            Event & $\tilde{\kappa}$\\
+            \hline
+            \variable{output/best_events.txt}
+        \end{tabular}
+    \end{ruledtabular}
+    \caption{
+        Events that best constrain $\kappa$.
+        $\tilde{\kappa}$ is the posterior median of $\kappa$.
+    }
+    \label{tab:best_events}
+\end{table}
 
-The event that shows the largest tension with $\kappa = 0$ is GW170818, for which $|\mu / \sigma| = \dots$ \mi{THOMAS: fill in these values}
+The event that shows the largest tension with $\kappa = 0$ is GW170818, for which $|\mu / \sigma| =$ \variable{output/GW170818_constraint.txt}.
 Additionally, the GW200129 result shows features that could be indicative of contamination from non-Gaussian instrumental noise, so we exclude it from the collective analysis.
-We discuss both these events in detail below.
+We discuss both of these events in detail below.
 
-Another set of events stand out in Fig.~\ref{fig:violin_kappa} due evident bimodality in the $\kappa$ posterior.
+Another set of events stands out in Fig.~\ref{fig:violin_kappa} due to evident bimodality in the $\kappa$ posterior.
 That is the case for GW170104, GW190413\_134308, GW190521, GW190805\_211137, and GW191105\_143521.
-\mi{THOMAS: can you list the total masses of these events?}
+\begin{table}\
+    \begin{ruledtabular}
+        \begin{tabular}{ll}
+            Event & Total Mass ($(M_{\odot})$)\\
+            \hline
+            \variable{output/bimodal_events_mass.txt}
+        \end{tabular}
+    \end{ruledtabular}
+    \caption{
+        Events that show bimodality in the $\kappa$ posterior.
+        The total mass of the binary is given in solar masses.
+    }
+    \label{tab:bimodal_events_mass}
+\end{table}
 We understand this effect to be a consequence of the narrow sets of frequencies that are visible for these heavy events, as we mentioned in Sec.~\ref{sec:inclination} and further demonstrate below.
@@
         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-        Colors in the plot represent the PE result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE results done by us with both frequency independent and dependent birefringence respectively.
+        Colors in the plot represent the PE result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE results done by us with both frequency-independent and dependent birefringence respectively.
         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
@@
 
-\section{Illustration of birefringence}
-\label{sec:birefringence_appendix}
-
 \section{Corner plots of special events}

### 2023-02-22 src/scripts/bimodal_events_mass.py

@@ -1,0 +1,9 @@
+import pandas as pd
+import paths
+
+result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+
+with open(paths.output/"bimodal_events_mass.txt", "w") as f:
+    for event in ["GW170104", "GW190413_134308", "GW190521", "GW190805_211137", "GW191105_143521"]:
+        name = event.replace("_", r"\_")
+        f.write(rf"{name} & ${(result_DataFrame[result_DataFrame.event == event]['mass_1'].median() + result_DataFrame[result_DataFrame.event == event]['mass_2'].median()):.1f}$\\")

### 2023-02-22 src/tex/ms.tex
Formatting

@@
 
-Equation \eqref{eq:waveform_modification} can be derived as the first order effect in an expansion away from \ac{GR} under multiple frameworks.
+Eq.~\eqref{eq:waveform_modification} can be derived as the first order effect in an expansion away from \ac{GR} under multiple frameworks.
 In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity.
@@
 \end{equation}
-For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will affect the time domain waveform nontrivially (Fig~\ref{fig:birefringence}).
+For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will affect the time domain waveform nontrivially (Fig.~\ref{fig:birefringence}).
 
@@
 
-Figure \ref{fig:violin_kappa} makes it clear that not all \acp{BBH} in GWTC-3 are equally informative about birefringence.
-When considered individually, the events that best constrain $\kappa$ are listed in Table~\ref{tab:best_events}.
+Fig.~\ref{fig:violin_kappa} makes it clear that not all \acp{BBH} in GWTC-3 are equally informative about birefringence.
+When considered individually, the events that best constrain $\kappa$ are listed in Table~\ref{tab:best_events_kappa}.
 \begin{table}
+    \caption{Events that best constrain $\kappa$, where $\tilde{\kappa}$ is the posterior median of $\kappa$.} 
     \begin{ruledtabular}
-        \begin{tabular}{ll}
-            Event & $\tilde{\kappa}$\\
-            \hline
-            \variable{output/best_events.txt}
-        \end{tabular}
+        \variable{output/best_events_kappa.txt}
     \end{ruledtabular}
-    \caption{
-        Events that best constrain $\kappa$.
-        $\tilde{\kappa}$ is the posterior median of $\kappa$.
-    }
-    \label{tab:best_events}
+    \label{tab:best_events_kappa}
 \end{table}
@@
 Another set of events stands out in Fig.~\ref{fig:violin_kappa} due to evident bimodality in the $\kappa$ posterior.
-That is the case for GW170104, GW190413\_134308, GW190521, GW190805\_211137, and GW191105\_143521.
-\begin{table}\
+That is the case for the events listed in Table~\ref{tab:bimodal_events_mass}.
+\begin{table}
+    \caption{Events that show bimodality in the $\kappa$ posterior, where the total mass of the binary is given in solar masses.}
     \begin{ruledtabular}
-        \begin{tabular}{ll}
-            Event & Total Mass ($(M_{\odot})$)\\
-            \hline
         \variable{output/bimodal_events_mass.txt}
-        \end{tabular}
     \end{ruledtabular}
-    \caption{
-        Events that show bimodality in the $\kappa$ posterior.
-        The total mass of the binary is given in solar masses.
-    }
     \label{tab:bimodal_events_mass}
@@
 
-Figure \ref{fig:violin_kappa} shows some variance in the distribution of $\kappa$ posteriors, including some apparent outliers like GW170818.
+Fig.~\ref{fig:violin_kappa} shows some variance in the distribution of $\kappa$ posteriors, including some apparent outliers like GW170818.
 This is not unexpected: assuming independent Gaussian noise, we might expect up to ${\sim}3$ events in a sample of 71 events to deviate from the true value of $\kappa$ by ${\sim}2\sigma$.
@@
 % Corner plot of the mean and standard deviation of kappa
-We then show the result of Bayesian hierarchical modeling in figure \ref{fig:corner_Gaussian}.
+We then show the result of Bayesian hierarchical modeling in Fig.~\ref{fig:corner_Gaussian}.
 The median of $\mu$ within $90\%$ confidence interval is \variable{output/mu_median.txt}.
@@
 % Constraint on $\kappa$
-In figure \ref{fig:posterior_kappa}, we show both the restricted posterior of $\kappa$ and the generic posterior of $\kappa$.
-The restricted posterior is calculated by equation \ref{eq:restricted_posterior} with the assumption that $\kappa$ is a single value constant.
-The generic posterior is calculated by equation \ref{eq:generic_posterior} with the assumption that $\kappa$ is a true value drawn from a Gaussian distribution.
+In Fig.~\ref{fig:posterior_kappa}, we show both the restricted posterior of $\kappa$ and the generic posterior of $\kappa$.
+The restricted posterior is calculated by Eq.~\eqref{eq:restricted_posterior} with the assumption that $\kappa$ is a single value constant.
+The generic posterior is calculated by Eq.~\eqref{eq:generic_posterior} with the assumption that $\kappa$ is a true value drawn from a Gaussian distribution.
 The median of the restricted posterior within $90\%$ confidence interval is \variable{output/restricted_kappa_median.txt}.
@@
 % Case: GW170818
-\subsection{Result on GW170818}
-
-% Case: GW190412
-\subsection{Result on GW190412}
-
-% Case: GW200224
-\subsection{Result on GW200224}
+% \subsection{Result on GW170818}
 
@@
 Consider GW150914, the first GW detected by LIGO, as an example.
-In figure \ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
+In Fig.~\ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
 With the frequency-independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
@@
 Consider GW190521, the most massive binary black hole merger in the events we included.
-The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in figure \ref{fig:corner_GW190521}.
+The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in Fig.~\ref{fig:corner_GW190521}.
 The reason could be this GW event's frequency range is too narrow to break the degeneracy.
@@
 \end{equation}
-This equation takes the same form as \eqref{eq:waveform_modification}.
+This equation takes the same form as Eq.~\eqref{eq:waveform_modification}.
 We can then find the relation between $M_{PV}$ and $\kappa$ by

### 2023-02-22 src/tex/ms.tex
improved reweighted posteriors

@@
 
-Eq.~\eqref{eq:waveform_modification} can be derived as the first order effect in an expansion away from \ac{GR} under multiple frameworks.
+Equation \eqref{eq:waveform_modification} can be derived as the first order effect in an expansion away from \ac{GR} under multiple frameworks.
 In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity.
@@
 \subsubsection{Nonshared birefringence parameters}
+\label{sec:method:hier}
 
@@
 
-Fig.~\ref{fig:violin_kappa} makes it clear that not all \acp{BBH} in GWTC-3 are equally informative about birefringence.
+Figure \ref{fig:violin_kappa} makes it clear that not all \acp{BBH} in GWTC-3 are equally informative about birefringence.
 When considered individually, the events that best constrain $\kappa$ are listed in Table~\ref{tab:best_events_kappa}.
@@
 That is the case for the events listed in Table~\ref{tab:bimodal_events_mass}.
+
 \begin{table}
@@
 
-Fig.~\ref{fig:violin_kappa} shows some variance in the distribution of $\kappa$ posteriors, including some apparent outliers like GW170818.
-This is not unexpected: assuming independent Gaussian noise, we might expect up to ${\sim}3$ events in a sample of 71 events to deviate from the true value of $\kappa$ by ${\sim}2\sigma$.
-\mi{WIP}
+Figure \ref{fig:violin_kappa} shows a degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
+This is not unexpected: assuming independent Gaussian noise for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$.
 
-% Corner plot of the mean and standard deviation of kappa
-We then show the result of Bayesian hierarchical modeling in Fig.~\ref{fig:corner_Gaussian}.
-The median of $\mu$ within $90\%$ confidence interval is \variable{output/mu_median.txt}.
-The median of $\sigma$ within $90\%$ confidence interval is \variable{output/sigma_median.txt}.
-The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.
+To further assess the statistical properties of the set of posteriors in Fig.~\ref{fig:violin_kappa}, we apply the hierarchical analysis described in Sec.~\ref{sec:method:hier}.
+By characterizing the population mean and standard deviation over $\kappa$, this also allows us to obtain a collective measurement that does not assume all events share the same value of $\kappa$.
 
@@
     \label{fig:corner_Gaussian}
+\end{figure}
+
+We summarize the result of this exercise in Fig.~\ref{fig:corner_Gaussian}, which shows the posterior on the population mean $\mu$ and standard deviation $\sigma$ inferred from the collection of observations in Fig.~\ref{fig:violin_kappa}.
+This posterior supports lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals, we constrain $\mu =$ \variable{output/mu_median.txt} and $\sigma =$ \variable{output/sigma_median.txt}, for 90\%-credible symmetric intervals around the median \mi{THOMAS: for $\sigma$ it would make more sense to report the one-sided 90\%-credible upper limit here, i.e.i, the 90th percentile, as $\sigma < XX$}.
+However, even though $\sigma = 0$ is well supported, the $\sigma$ posterior peaks away from the origin, indicating some preference for a nonzero variance.
+This could be a sign of the presence of outliers in our sample.
+
+As a check for outliers, we reconsider the set of measurements in Fig.~\ref{fig:violin_kappa} in light of the hierarchical result for $\mu$ and $\sigma$ in Fig.~\ref{fig:corner_Gaussian}.
+This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Callister:T2100301}.
+
+\begin{figure}
+    \script{reweighed_kappa.py}
+    \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
+    \caption{Reweighted $\kappa$ distributions.}
+    \label{fig:reweighted_kappa}
 \end{figure}
@@
 
-\section{Reweighed kappa distributions}
-\label{sec:reweighed_kappa}
-
-\begin{figure}
-    \script{reweighed_kappa.py}
-    \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
-    \caption{Reweighed $\kappa$ distributions.}
-    \label{fig:reweighed_kappa}
-\end{figure}
-
 \bibliography{bib}

### 2023-02-22 src/tex/ms.tex
Change to one-sided limit for Gaussian sigma

@@
 We summarize the result of this exercise in Fig.~\ref{fig:corner_Gaussian}, which shows the posterior on the population mean $\mu$ and standard deviation $\sigma$ inferred from the collection of observations in Fig.~\ref{fig:violin_kappa}.
-This posterior supports lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals, we constrain $\mu =$ \variable{output/mu_median.txt} and $\sigma =$ \variable{output/sigma_median.txt}, for 90\%-credible symmetric intervals around the median \mi{THOMAS: for $\sigma$ it would make more sense to report the one-sided 90\%-credible upper limit here, i.e.i, the 90th percentile, as $\sigma < XX$}.
+This posterior supports the lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals, we constrain $\mu =$ \variable{output/mu_median.txt} for 90\%-credible symmetric intervals around the median, and $\sigma <$ \variable{output/sigma_median.txt} for the 90\%-credible one-sided upper limit.
 However, even though $\sigma = 0$ is well supported, the $\sigma$ posterior peaks away from the origin, indicating some preference for a nonzero variance.

### 2023-02-22 src/scripts/Gaussian_result.py

@@ -13,4 +13,4 @@ with open(paths.output/"mu_median.txt", "w") as f:
     f.write(rf"${mu_median:.3f}^{{+{(np.percentile(a=df['mu'],q=95)-mu_median):.3f}}}_{{{(np.percentile(a=df['mu'],q=5)-mu_median):.3f}}}$")
 
 with open(paths.output/"sigma_median.txt", "w") as f:
-    f.write(rf"${sigma_median:.3f}^{{+{(np.percentile(a=df['sigma'],q=95)-sigma_median):.3f}}}_{{{(np.percentile(a=df['sigma'],q=5)-sigma_median):.3f}}}$")
+    f.write(rf"${np.percentile(a=df['sigma'],q=90):.3f}$")

### 2023-02-22 src/scripts/corner_Gaussian_appendix_1.py
Added alternative hierarchical modeling results

@@ -1,0 +1,64 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import kde_contour as kde_contour
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"],
+})
+
+samples_Gaussian = np.load(paths.data/"samples_Gaussian.npz")["chains"]
+samples_Gaussian = samples_Gaussian.reshape(-1, 2)
+mu_median = np.median(samples_Gaussian[:,0])
+sigma_median = np.median(samples_Gaussian[:,1])
+
+df = pd.DataFrame()
+df['mu'] = samples_Gaussian.reshape(-1,2)[:,0]
+df['sigma'] = samples_Gaussian.reshape(-1,2)[:,1]
+df = df.sample(10000)
+
+kde_1d_sigma = kde_contour.Bounded_1d_kde(df['sigma'], xlow=0)
+df['sigma_kde'] = kde_1d_sigma(df['sigma'])
+df = df.sort_values(['sigma'])
+
+fig, axs = plt.subplots(2,2,sharex='col')
+fig.set_size_inches(5,5)
+
+axs[0,0].axvline(mu_median, color=sns.color_palette()[3])
+sns.kdeplot(df['mu'], fill=True, ax=axs[0,0])
+
+axs[0,0].set_ylabel("")
+axs[0,0].set_xlim(-.07,.07)
+axs[0,0].set_ylim(0)
+axs[0,0].set_yticks([])
+
+axs[1,1].axvline(sigma_median, color=sns.color_palette()[3])
+axs[1,1].fill_between(df['sigma'], df['sigma_kde'], np.zeros(len(df['sigma'])), alpha=0.2)
+axs[1,1].plot(df['sigma'], df['sigma_kde'])
+
+axs[1,1].set_xlabel("$\\sigma$")
+axs[1,1].set_xlim(0)
+axs[1,1].set_ylim(0)
+axs[1,1].set_yticks([])
+axs[1,1].get_lines()[1].set_linewidth(1)
+
+axs[1,0].axvline(mu_median, color=sns.color_palette()[3])
+axs[1,0].axhline(sigma_median, color=sns.color_palette()[3])
[…]

### 2023-02-22 src/tex/ms.tex
Added alternative hierarchical modeling results
edits to table caption

@@
 \begin{table}
-    \caption{Events that best constrain $\kappa$, where $\tilde{\kappa}$ is the posterior median of $\kappa$.} 
+    \caption{Events that best constrain $\kappa$. \mi{THOMAS: please just use $\kappa$, no need for the $\tilde{\kappa}$ notation}}
     \begin{ruledtabular}
@@
 \begin{table}
-    \caption{Events that show bimodality in the $\kappa$ posterior, where the total mass of the binary is given in solar masses.}
+    \caption{Events that show bimodality in the $\kappa$ posterior tend to have high total masses.}
     \begin{ruledtabular}
@@
 
+\section{Alternative results for Bayesian hierarchical modeling}
+\label{sec:alternative_results}
+
+\begin{figure}
+    \script{corner_Gaussian_appendix_1.py}
+    \includegraphics[width=\columnwidth]{figures/corner_Gaussian_appendix_1.pdf}
+    \caption{
+        Kept all events.
+        % The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
+        % 2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        % The orange line represents the median of $\mu$ and $\sigma$.
+        % The plot shows that the population constraint on $\kappa$ is consistent with GR.
+    }
+    \label{fig:corner_Gaussian_appendix_1}
+\end{figure}
+
+\begin{figure}
+    \script{corner_Gaussian_appendix_2.py}
+    \includegraphics[width=\columnwidth]{figures/corner_Gaussian_appendix_2.pdf}
+    \caption{
+        Excluded GW170818 and GW200129.
+        % The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
+        % 2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        % The orange line represents the median of $\mu$ and $\sigma$.
+        % The plot shows that the population constraint on $\kappa$ is consistent with GR.
+    }
+    \label{fig:corner_Gaussian_appendix_2}
+\end{figure}
+
 \bibliography{bib}

### 2023-02-23 src/scripts/best_events_kappa.py
remove tilde

@@ -16,7 +16,7 @@ for event in events:
 result_dict = {k: v for k, v in sorted(result_dict.items(), key=lambda item: item[1]['kappa_std'])}
 
 with open(paths.output/"best_events_kappa.txt", "w") as f:
-    f.write(r"\begin{tabular}{lr}Event & $\tilde{\kappa}$ (dimensionless)\\ \hline ")
+    f.write(r"\begin{tabular}{lr}Event & $\kappa$ (dimensionless)\\ \hline ")
     for i, event in enumerate(result_dict):
         f.write(rf"{result_dict[event]['event']} & ${result_dict[event]['kappa_median']:.3f}^{{+{result_dict[event]['kappe_95']:.3f}}}_{{-{result_dict[event]['kappe_5']:.3f}}}$")
         if i != 4:

### 2023-02-23 src/tex/ms.tex
remove tilde
Added section on GW170818

@@
 We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
-We take the \ac{PE} configuration as in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
-We apply a distance prior which is uniform in comoving volume and source frame time instead of a power law distribution used in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+We take the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
+We apply a distance prior which is uniform in comoving volume and source frame time instead of a power law distribution used in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
 
@@
 \begin{table}
-    \caption{Events that best constrain $\kappa$. \mi{THOMAS: please just use $\kappa$, no need for the $\tilde{\kappa}$ notation}}
+    \caption{Events that best constrain $\kappa$.}
     \begin{ruledtabular}
@@
 
-% Create sessions for all the outliers
+% Case: GW170818
+\subsection{Result on GW170818}
+GW170818 gave the largest tension with $\kappa=0$.
+Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$.
+This region corresponds to a larger distance and a smaller inclination angle.
+This is a sensible result because a smaller inclination angle will make the amplitude ratio between the two circular polarizations smaller.
+Then the birefringence effect will increase the amplitude ratio during the propagation.
+The larger distance is for compensating the combined effect of a nonzero $\kappa$ and a smaller inclination angle.
+We can see this effect in other events as well, but not as strongly as in GW170818.
 
-% Case: GW170818
-% \subsection{Result on GW170818}
+\begin{figure}
+    \script{corner_GW170818.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
+    \caption{Corner plot of GW170818.}
+    \label{fig:corner_GW170818}
+\end{figure}
 
@@
 
-\section{Corner plots of special events}
+\section{Special events}
 \label{sec:corner_plots}
@@
 \begin{figure}
-    \script{corner_GW170818.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
+    \script{corner_GW170818_appendix.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW170818_appendix.pdf}
     \caption{Corner plot of GW170818.}
-    \label{fig:corner_GW170818}
+    \label{fig:corner_GW170818_appendix}
 \end{figure}

### 2023-02-23 src/scripts/corner_GW170818_appendix.py

@@ -1,0 +1,71 @@
+import h5py
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.2)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"],
+})
+
+nsamples = 3000
+
+with h5py.File(paths.data/"IGWN-GWTC2p1-v2-GW170818_022509_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
+    f = f['C01:IMRPhenomXPHM']
+    result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
+result_ligo = result_ligo.sample(n=nsamples)
+result_ligo['kappa'] = np.zeros(len(result_ligo))
+result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
+
+result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_bilby = result_bilby[result_bilby.event == "GW170818"]
+result_bilby = result_bilby.sample(n=nsamples)
+result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
+result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+
+result = pd.concat([result_bilby,result_ligo], ignore_index=True)
+
+g = sns.pairplot(result,
+                    vars=['kappa', 'mass_ratio', 'chirp_mass', 'luminosity_distance',
+                            'dec', 'ra', 'theta_jn', 'psi', 'a_1','a_2','tilt_1','tilt_2','phi_12','phi_jl',
+                            'phase', 'geocent_time', 'log_likelihood',
+                            ],
+                    corner=True, kind='kde', hue='with',
+                    diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[
+                            (1.-0.90),(1.-0.3935)]))
+
[…]

### 2023-02-23 src/tex/ms.tex
Include GW200129 in the default result

@@
 To do this, we apply hierarchical Bayesian inference \cite{Loredo:2004nn} to model the distribution of $\kappa$'s consistent with the observed data:
-we posit that there is a specific value of the parameter, $\kappa_i$, associated to each event, drawn from some unknown distribution of underlying values; from the imperfect measurements of $\kappa_i$ for each event, we may reconstruct the underlying distribution.
+we posit that there is a specific value of the parameter, $\kappa_i$, associated with each event, drawn from some unknown distribution of underlying values; from the imperfect measurements of $\kappa_i$ for each event, we may reconstruct the underlying distribution.
 If we are interested in constraining the first two moments of the distribution, it is convenient to parametrize the $\kappa_i$'s as drawn from a Gaussian with unknown mean $\mu$ and variance $\sigma^2$, i.e., $\kappa_i \sim \mathcal{N}(\mu, \sigma^2)$ \cite{Isi:2019asy}, and measure those hyperparameters from the collection of observed data.
@@
 In this section, we present the results of our study.
-We first show the PE results of all events and present the results of Bayesian hierarchical modeling.
+We first show the \ac{PE} results of all events and present the results of Bayesian hierarchical modeling.
 Then, we will discuss some special events individually.
@@
 The event that shows the largest tension with $\kappa = 0$ is GW170818, for which $|\mu / \sigma| =$ \variable{output/GW170818_constraint.txt}.
-Additionally, the GW200129 result shows features that could be indicative of contamination from non-Gaussian instrumental noise, so we exclude it from the collective analysis.
+Additionally, the GW200129 result shows features that could be indicative of contamination from non-Gaussian instrumental noise.
 We discuss both of these events in detail below.
@@
 \subsection{Result on GW150914}
-In this study, we included frequency dependence of birefringence, which would affect the posterior of $\kappa$ obtained from PE.
+In this study, we included frequency dependence of birefringence, which would affect the posterior of $\kappa$ obtained from \ac{PE}.
 Consider GW150914, the first GW detected by LIGO, as an example.
@@
 In this case, the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent.
-Therefore, the PE result with frequency dependence is consistent with GR.
+Therefore, the \ac{PE} result with frequency dependence is consistent with GR.
 
@@
         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-        Colors in the plot represent the PE result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE results done by us with both frequency-independent and dependent birefringence respectively.
+        Colors in the plot represent the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, the \ac{PE} results done by us with both frequency-independent and dependent birefringence respectively.
         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        Note that there is no posterior of $\kappa$ for the PE result from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
+        Note that there is no posterior of $\kappa$ for the \ac{PE} result from LVK, as the LVK \ac{PE} is based on GR, which does not suggest GW amplitude birefringence.
         This plot shows that frequency-independent birefringence creates a degeneracy between $\kappa$ and $\iota$, while frequency-dependent birefringence can break it.
@@
         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
-        Colors in the plot are the PE result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE result done by us with the frequency-dependent birefringence.
+        Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with the frequency-dependent birefringence.
         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
@@
 % Case: GW200129 (glitch in Virgo data)
-\subsection{Exclusion of GW200129}
+\subsection{Contamination of GW200129}
 \label{sec:GW200129}
-The reason for the exclusion is that past research suggests a potential glitch in GW200129 Virgo data. \citep{GW200129_glitch}
-We perform three sets of PE to check if the glitch would affect the population posterior of $\kappa$.
-We pick only two out of three detector data in each set to perform PE.
+Past research suggests a potential glitch in GW200129 Virgo data. \citep{GW200129_glitch}
+We perform three sets of \ac{PE} to check if the glitch would affect the population posterior of $\kappa$.
+We pick only two out of three detector data in each set to perform \ac{PE}.
 (i.e.
 LIGO-Hanford and LIGO-Livingston for the first set, LIGO-Hanford and Virgo for the second set, and LIGO-Livingston and Virgo for the third set.)
-The two sets of PE results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
+The two sets of \ac{PE} results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
 Further investigation is needed to understand the glitch.
-Thus, we did not include this event in our collective analysis.
-% show results included and excluded
+% Thus, we did not include this event in our collective analysis.
+
+Figure~\ref{fig:corner_GW200129} shows the posteriors of the parameters of GW200129 are strongly affected by the Virgo data.
+
+\begin{figure}
+    \script{corner_GW200129.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
+    \caption{
+        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW200129.
+        Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with and without Virgo data.
+        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        This plot shows that Virgo data may contain a glitch, which affects the \ac{PE} result.
+    }
+    \label{fig:corner_GW200129}
+\end{figure}
 
@@
 
-\citet{Wang_2021} performed PE on GWTC-1 events with a frequency-dependent birefringence model.
+\citet{Wang_2021} performed \ac{PE} on GWTC-1 events with a frequency-dependent birefringence model.
 They formulated the birefringence effect as corrections to the GW waveform on a linear basis.
@@
 % BNS
-Future work is to perform PE on binary neutron star mergers, such as GW170817.
+Future work is to perform \ac{PE} on binary neutron star mergers, such as GW170817.
 The frequency range of the signal is much wider compared to the binary black hole mergers.
 Thus, the difference in the effect of birefringence at different frequencies within the range can be more significant.
-The PE on binary neutron star mergers can allow us to further constrain the birefringence effect and the beyond-GR theories that predict it.
-However, performing PE on binary neutron star mergers requires much more computational resources.
-Therefore, we may need to wait for future PE methods and tools to further our work.
+The \ac{PE} on binary neutron star mergers can allow us to further constrain the birefringence effect and the beyond-GR theories that predict it.
+However, performing \ac{PE} on binary neutron star mergers requires much more computational resources.
+Therefore, we may need to wait for future \ac{PE} methods and tools to further our work.
 
@@
 LVK will release more events with higher SNR in the future.
-Using data with higher SNR allows us to obtain more precise PE results and constrain the birefringence effect more precisely.
+Using data with higher SNR allows us to obtain more precise \ac{PE} results and constrain the birefringence effect more precisely.
 And using data from more events will allow us to calculate a more constrained population posterior of $\kappa$.
@@
 
-\section{Special events}
-\label{sec:corner_plots}
-
-\begin{figure}
-    \script{corner_GW170818_appendix.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW170818_appendix.pdf}
-    \caption{Corner plot of GW170818.}
-    \label{fig:corner_GW170818_appendix}
-\end{figure}
-
-\begin{figure}
-    \script{corner_GW200129.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
-    \caption{Corner plot of GW200129.}
-    \label{fig:corner_GW200129}
-\end{figure}
-
-\begin{figure}
-    \script{corner_GW200224.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW200224.pdf}
-    \caption{Corner plot of GW200224.}
-    \label{fig:corner_GW200224}
-\end{figure}
-
-\section{Alternative results for Bayesian hierarchical modeling}
-\label{sec:alternative_results}
-
-\begin{figure}
-    \script{corner_Gaussian_appendix_1.py}
-    \includegraphics[width=\columnwidth]{figures/corner_Gaussian_appendix_1.pdf}
-    \caption{
-        Kept all events.
-        % The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
-        % 2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        % The orange line represents the median of $\mu$ and $\sigma$.
-        % The plot shows that the population constraint on $\kappa$ is consistent with GR.
-    }
-    \label{fig:corner_Gaussian_appendix_1}
-\end{figure}
-
-\begin{figure}
-    \script{corner_Gaussian_appendix_2.py}
-    \includegraphics[width=\columnwidth]{figures/corner_Gaussian_appendix_2.pdf}
-    \caption{
-        Excluded GW170818 and GW200129.
-        % The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
-        % 2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        % The orange line represents the median of $\mu$ and $\sigma$.
-        % The plot shows that the population constraint on $\kappa$ is consistent with GR.
-    }
-    \label{fig:corner_Gaussian_appendix_2}
-\end{figure}
-
 \bibliography{bib}

### 2023-02-23 src/scripts/corner_GW170818_appendix.py

@@ -1,71 +1,0 @@
-import h5py
-import matplotlib.pyplot as plt
-import pandas as pd
-import numpy as np
-import seaborn as sns
-import paths
-
-sns.set_theme(palette='colorblind', font_scale=1.2)
-
-plt.rcParams.update({
-    "text.usetex": True,
-    "font.family": "serif",
-    "font.serif": ["Computer Modern Roman"],
-})
-
-nsamples = 3000
-
-with h5py.File(paths.data/"IGWN-GWTC2p1-v2-GW170818_022509_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
-    f = f['C01:IMRPhenomXPHM']
-    result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
-result_ligo = result_ligo.sample(n=nsamples)
-result_ligo['kappa'] = np.zeros(len(result_ligo))
-result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
-
-result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_bilby = result_bilby[result_bilby.event == "GW170818"]
-result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
-result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
-
-result = pd.concat([result_bilby,result_ligo], ignore_index=True)
-
-g = sns.pairplot(result,
-                    vars=['kappa', 'mass_ratio', 'chirp_mass', 'luminosity_distance',
-                            'dec', 'ra', 'theta_jn', 'psi', 'a_1','a_2','tilt_1','tilt_2','phi_12','phi_jl',
-                            'phase', 'geocent_time', 'log_likelihood',
-                            ],
-                    corner=True, kind='kde', hue='with',
-                    diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[
-                            (1.-0.90),(1.-0.3935)]))
-
[…]

### 2023-02-23 src/scripts/corner_GW200129.py

@@ -1,8 +1,8 @@
-import h5py
 import matplotlib.pyplot as plt
 import pandas as pd
 import numpy as np
 import seaborn as sns
+from bilby.gw.result import CBCResult
 import paths
 
 sns.set_theme(palette='colorblind', font_scale=1.2)
@@ -13,59 +13,41 @@ plt.rcParams.update({
     "font.serif": ["Computer Modern Roman"],
 })
 
-nsamples = 3000
+nsamples = 5000
 
-with h5py.File(paths.data/"IGWN-GWTC3p0-v1-GW200129_065458_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
-    f = f['C01:IMRPhenomXPHM']
-    result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
-result_ligo = result_ligo.sample(n=nsamples)
-result_ligo['kappa'] = np.zeros(len(result_ligo))
-result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
+result = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result = result[result.event == "GW200129_065458"]
+result = result.sample(n=nsamples)
+result['with'] = np.full(len(result), r"birefringence (H1, L1, V1)")
+result['cos_iota'] = np.cos(result['iota'])
 
-result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_bilby = result_bilby[result_bilby.event == "GW200129_065458"]
-result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
-result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+result_without_H1 = CBCResult.from_json(filename=paths.data/"GW200129_065458_birefringence_without_H1.json.gz").posterior
+result_without_H1 = result_without_H1.sample(n=nsamples)
+result_without_H1['with'] = np.full(len(result_without_H1), r"birefringence (L1, V1)")
+result_without_H1['cos_iota'] = np.cos([float(result_without_H1['iota'][i]) for i in result_without_H1.index])
 
-result = pd.concat([result_bilby,result_ligo], ignore_index=True)
[…]

### 2023-02-23 src/scripts/corner_GW200224.py

@@ -1,71 +1,0 @@
-import h5py
-import matplotlib.pyplot as plt
-import pandas as pd
-import numpy as np
-import seaborn as sns
-import paths
-
-sns.set_theme(palette='colorblind', font_scale=1.2)
-
-plt.rcParams.update({
-    "text.usetex": True,
-    "font.family": "serif",
-    "font.serif": ["Computer Modern Roman"],
-})
-
-nsamples = 3000
-
-with h5py.File(paths.data/"IGWN-GWTC3p0-v1-GW200224_222234_PEDataRelease_mixed_nocosmo.h5", 'r') as f:
-    f = f['C01:IMRPhenomXPHM']
-    result_ligo = pd.DataFrame.from_records(f["posterior_samples"][()])
-result_ligo = result_ligo.sample(n=nsamples)
-result_ligo['kappa'] = np.zeros(len(result_ligo))
-result_ligo['with'] = np.full(len(result_ligo), r"GR (LVK)")
-
-result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_bilby = result_bilby[result_bilby.event == "GW200224_222234"]
-result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
-result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
-
-result = pd.concat([result_bilby,result_ligo], ignore_index=True)
-
-g = sns.pairplot(result,
-                    vars=['kappa', 'mass_ratio', 'chirp_mass', 'luminosity_distance',
-                            'dec', 'ra', 'theta_jn', 'psi', 'a_1','a_2','tilt_1','tilt_2','phi_12','phi_jl',
-                            'phase', 'geocent_time', 'log_likelihood',
-                            ],
-                    corner=True, kind='kde', hue='with',
-                    diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[
-                            (1.-0.90),(1.-0.3935)]))
-
[…]

### 2023-02-23 src/scripts/corner_Gaussian.py

@@ -13,7 +13,7 @@ plt.rcParams.update({
     "font.serif": ["Computer Modern Roman"],
 })
 
-samples_Gaussian = np.load(paths.data/"samples_Gaussian_without_GW200129.npz")["chains"]
+samples_Gaussian = np.load(paths.data/"samples_Gaussian.npz")["chains"]
 samples_Gaussian = samples_Gaussian.reshape(-1, 2)
 mu_median = np.median(samples_Gaussian[:,0])
 sigma_median = np.median(samples_Gaussian[:,1])
@@ -27,20 +27,39 @@ kde_1d_sigma = kde_contour.Bounded_1d_kde(df['sigma'], xlow=0)
 df['sigma_kde'] = kde_1d_sigma(df['sigma'])
 df = df.sort_values(['sigma'])
 
+samples_Gaussian_without_GW200129 = np.load(paths.data/"samples_Gaussian_without_GW200129.npz")["chains"]
+samples_Gaussian_without_GW200129 = samples_Gaussian_without_GW200129.reshape(-1, 2)
+mu_median_without_GW200129 = np.median(samples_Gaussian_without_GW200129[:,0])
+sigma_median_without_GW200129 = np.median(samples_Gaussian_without_GW200129[:,1])
+
+df_without_GW200129 = pd.DataFrame()
+df_without_GW200129['mu'] = samples_Gaussian_without_GW200129.reshape(-1,2)[:,0]
+df_without_GW200129['sigma'] = samples_Gaussian_without_GW200129.reshape(-1,2)[:,1]
+df_without_GW200129 = df_without_GW200129.sample(10000)
+
+kde_1d_sigma_without_GW200129 = kde_contour.Bounded_1d_kde(df['sigma'], xlow=0)
+df_without_GW200129['sigma_kde'] = kde_1d_sigma_without_GW200129(df['sigma'])
+df_without_GW200129 = df_without_GW200129.sort_values(['sigma'])
+
 fig, axs = plt.subplots(2,2,sharex='col')
 fig.set_size_inches(5,5)
 
-axs[0,0].axvline(mu_median, color=sns.color_palette()[3])
+axs[0,0].axvline(mu_median, color=sns.color_palette()[0])
 sns.kdeplot(df['mu'], fill=True, ax=axs[0,0])
[…]

### 2023-02-23 src/scripts/posterior_kappa.py

@@ -18,11 +18,10 @@ def normal_distribution(x, mu, sigma):
 
 result_dict = {}
 result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_DataFrame = result_DataFrame[result_DataFrame.event != "GW200129_065458"]
 for event in result_DataFrame['event'].unique():
     result_dict[event] = result_DataFrame[result_DataFrame.event == event]
 
-samples_original = np.load(paths.data/"samples_Gaussian_without_GW200129.npz")["chains"]
+samples_original = np.load(paths.data/"samples_Gaussian.npz")["chains"]
 samples_all = samples_original.reshape(-1, 2)
 
 df_Gaussian_samples_all = pd.DataFrame()
@@ -37,7 +36,7 @@ for _ in range(1000):
     samples_reweighted.append(np.random.choice(a=np.array(result_DataFrame['kappa']), size=1, p=weight))
 
 events = result_DataFrame['event'].unique()
-kernels = [scipy.stats.gaussian_kde(result_dict[event]['kappa']) for event in events[events != "GW200129_065458"]]
+kernels = [scipy.stats.gaussian_kde(result_dict[event]['kappa']) for event in events]
 kappa = np.linspace(-.1, .1, 1000)
 ll = [np.sum([np.log(ker(k)) for ker in kernels]) for k in kappa]
 ll = ll - np.max(ll)

### 2023-02-28 src/tex/ms.tex
Update ms.tex
GW200129 plot

@@
     \caption{
-        The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
-        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
-        The orange line represents the median of $\mu$ and $\sigma$.
-        The plot shows that the population constraint on $\kappa$ is consistent with GR.
+        The posterior of the $\kappa$ population hyperparameters $\mu$ and $\sigma$, including (blue) and excluding (orange) GW200129 from the collection of events.
+        The 2D contours correspond to the $39.35\%$ and $90\%$ credible levels.
+        The plot shows that the population constraint on $\kappa$ is consistent with no birefringence ($\mu=\sigma=0$) at the 90\% credible level.
     }
@@
 Note that the result is consistent with GR when $\kappa=0$.
-% talk about what events pull sigma in figure 2
 
@@
         Colors in the plot represent the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, the \ac{PE} results done by us with both frequency-independent and dependent birefringence respectively.
-        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
         Note that there is no posterior of $\kappa$ for the \ac{PE} result from LVK, as the LVK \ac{PE} is based on GR, which does not suggest GW amplitude birefringence.
@@
         Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with the frequency-dependent birefringence.
-        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
         This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
@@
         Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with and without Virgo data.
-        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
+        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
         This plot shows that Virgo data may contain a glitch, which affects the \ac{PE} result.

### 2023-02-28 src/tex/ms.tex
minor wording

@@
 % Violin plot
-We present the result of our catalog analysis in Fig.~\ref{fig:violin_kappa}.
-This shows the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate), colored by the distance of the respective posterior mean from the origin in units of standard deviation, i.e., $|\mu/\sigma|$.
+Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis.
+as represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate), colored by the distance of the respective posterior mean from the origin in units of standard deviation, i.e., $|\mu/\sigma|$.
 The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line);
@@
 When considered individually, the events that best constrain $\kappa$ are listed in Table~\ref{tab:best_events_kappa}.
+
 \begin{table}
@@
 The event that shows the largest tension with $\kappa = 0$ is GW170818, for which $|\mu / \sigma| =$ \variable{output/GW170818_constraint.txt}.
+\mi{THOMAS: we don't need that many significant figures, $2.0$ would do}
 Additionally, the GW200129 result shows features that could be indicative of contamination from non-Gaussian instrumental noise.
@@
 Figure \ref{fig:violin_kappa} shows a degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
-This is not unexpected: assuming independent Gaussian noise for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$.
+This is not unexpected: assuming independent Gaussian noise instantiations for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$.
 

### 2023-02-28 src/tex/ms.tex
minor text change

@@
 Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis.
-as represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate), colored by the distance of the respective posterior mean from the origin in units of standard deviation, i.e., $|\mu/\sigma|$.
+As represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate), colored by the distance of the respective posterior mean from the origin in units of standard deviation, i.e., $|\mu/\sigma|$.
 The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line);
@@
 The event that shows the largest tension with $\kappa = 0$ is GW170818, for which $|\mu / \sigma| =$ \variable{output/GW170818_constraint.txt}.
-\mi{THOMAS: we don't need that many significant figures, $2.0$ would do}
 Additionally, the GW200129 result shows features that could be indicative of contamination from non-Gaussian instrumental noise.

### 2023-02-28 src/scripts/GW170818_constraint.py

@@ -6,4 +6,4 @@ result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.f
 mean_over_std_GW170818 = result_DataFrame[result_DataFrame.event == "GW170818"]['kappa'].mean() / result_DataFrame[result_DataFrame.event == "GW170818"]['kappa'].std()
 
 with open(paths.output/"GW170818_constraint.txt", "w") as f:
-    f.write(rf"${mean_over_std_GW170818:.3f}$")
+    f.write(rf"${mean_over_std_GW170818:.1f}$")

### 2023-02-28 src/tex/ms.tex
notable events

@@
 % Violin plot
-Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis.
-As represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate), colored by the distance of the respective posterior mean from the origin in units of standard deviation, i.e., $|\mu/\sigma|$.
+Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis
+as represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate).
+Posteriors are colored by the distance of the respective posterior mean from the origin in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.%
+\footnote{We use the $i$ notation for individual events here to disambiguate from the population mean $\mu$ and standard deviation $\sigma$ of Eq.~\eqref{eq:posterior_of_mu_sigma}.}
 The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line);
@@
 
+The two events that show the largest tension with $\kappa = 0$ is GW170818, for which $|\mu_i / \sigma_i| = \variable{output/GW170818_constraint.txt}$, and GW200129\_065458 (henceforth GW200129), for which $|\mu_i / \sigma_i| = \dots$.
+\mi{THOMAS: please add the number for GW200129. Also, we don't need that many significant figures, one decimal will do.}
+However, as we discuss further below, we have reason to think that the GW200129 preference for $\kappa < 0$ might be driven by noise anomalies in the Virgo detector; for that reason we consider the effect of excluding it from the joint result below.
+
 \begin{table}
-    \caption{Events that best constrain $\kappa$.}
+    \caption{Events that best constrain $\kappa$, sorted by posterior standard deviation $\sigma_i$.}
     \begin{ruledtabular}
@@
 
-The event that shows the largest tension with $\kappa = 0$ is GW170818, for which $|\mu / \sigma| =$ \variable{output/GW170818_constraint.txt}.
-Additionally, the GW200129 result shows features that could be indicative of contamination from non-Gaussian instrumental noise.
-We discuss both of these events in detail below.
-
-Another set of events stands out in Fig.~\ref{fig:violin_kappa} due to evident bimodality in the $\kappa$ posterior.
+Finally, a set of events stands out in Fig.~\ref{fig:violin_kappa} due to evident bimodality in the $\kappa$ posterior.
 That is the case for the events listed in Table~\ref{tab:bimodal_events_mass}.
+We understand this effect to be a consequence of the narrow sets of frequencies that are visible for these heavy events, as we mentioned in Sec.~\ref{sec:inclination} and further demonstrate below.
 
@@
 \end{table}
-We understand this effect to be a consequence of the narrow sets of frequencies that are visible for these heavy events, as we mentioned in Sec.~\ref{sec:inclination} and further demonstrate below.
 
-\subsection{Bayesian hierarchical modeling}
+\subsection{Hierarchical modeling}
 
-Figure \ref{fig:violin_kappa} shows a degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
+Figure \ref{fig:violin_kappa} shows certain variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
 This is not unexpected: assuming independent Gaussian noise instantiations for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$.
@@
 To further assess the statistical properties of the set of posteriors in Fig.~\ref{fig:violin_kappa}, we apply the hierarchical analysis described in Sec.~\ref{sec:method:hier}.
-By characterizing the population mean and standard deviation over $\kappa$, this also allows us to obtain a collective measurement that does not assume all events share the same value of $\kappa$.
+By characterizing the population mean and standard deviation over events, this also allows us to obtain a collective measurement that does not assume all events share the same value of $\kappa$.
 
@@
 We summarize the result of this exercise in Fig.~\ref{fig:corner_Gaussian}, which shows the posterior on the population mean $\mu$ and standard deviation $\sigma$ inferred from the collection of observations in Fig.~\ref{fig:violin_kappa}.
-This posterior supports the lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals, we constrain $\mu =$ \variable{output/mu_median.txt} for 90\%-credible symmetric intervals around the median, and $\sigma <$ \variable{output/sigma_median.txt} for the 90\%-credible one-sided upper limit.
+The figure shows two distributions, which result from analyses with (blue) and without (orange) the potentially-contaminated event GW200129; the difference between the two is mainly restricted to a slight shift in $\mu$, indicating that GW200129 has a small effect on our overall population inference.
+
+Both versions of the posterior support the lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals of the result including all events, we constrain $\mu =$ \variable{output/mu_median.txt} for 90\%-credible symmetric intervals around the median, and $\sigma <$ \variable{output/sigma_median.txt} for the 90\%-credible one-sided upper limit.
 However, even though $\sigma = 0$ is well supported, the $\sigma$ posterior peaks away from the origin, indicating some preference for a nonzero variance.
@@
 
-As a check for outliers, we reconsider the set of measurements in Fig.~\ref{fig:violin_kappa} in light of the hierarchical result for $\mu$ and $\sigma$ in Fig.~\ref{fig:corner_Gaussian}.
+As a check for outliers, we reconsider the set of measurements in Fig.~\ref{fig:violin_kappa} in light of the hierarchical result for $\mu$ and $\sigma$ in Fig.~\ref{fig:corner_Gaussian}, including GW200129 (blue curve).
 This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Callister:T2100301}.
+Figure \ref{fig:corner_Gaussian} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its preference for $\kappa > 0$.
+This feature appears to offset a few other events which tend to favor $\kappa < 0$.
+The interaction between these distributions leads to a hyperposterior that is fully consistent with $\mu = 0$ while indicating some support for $\sigma > 0$ (Fig.~\ref{fig:corner_Gaussian}).
+Future observations will determine whether there is truly evidence for a nonvanishing variance in this population.
 
@@
     \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
-    \caption{Reweighted $\kappa$ distributions.}
+    \caption{Individual-event $\kappa_i$ distributions of Fig.~\ref{fig:violin_kappa} reweighted in light of the population-level inference of Fig.~\ref{fig:corner_Gaussian}, marginalizing over $\mu$ and $\sigma$.
+Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i$; we label the events most in tension with $\kappa_i = 0$ by that same measure.}
     \label{fig:reweighted_kappa}
@@
 
-\subsection{Constaint on $\kappa$}
 % Constraint on $\kappa$
-In Fig.~\ref{fig:posterior_kappa}, we show both the restricted posterior of $\kappa$ and the generic posterior of $\kappa$.
-The restricted posterior is calculated by Eq.~\eqref{eq:restricted_posterior} with the assumption that $\kappa$ is a single value constant.
-The generic posterior is calculated by Eq.~\eqref{eq:generic_posterior} with the assumption that $\kappa$ is a true value drawn from a Gaussian distribution.
-The median of the restricted posterior within $90\%$ confidence interval is \variable{output/restricted_kappa_median.txt}.
-The median of the generic posterior within $90\%$ confidence interval is \variable{output/generic_kappa_median.txt}.
-Note that the result is consistent with GR when $\kappa=0$.
+Finally, the hierarchical result in Fig.~\ref{fig:corner_Gaussian} can be translated into an overall expectation for $\kappa_i$ via Eq.~\eqref{eq:generic_posterior}.
+We show the result of doing this in Fig.~\ref{fig:posterior_kappa}, where we also compare to the posterior on $\kappa$ obtained by assuming a shared value across events (same result shown as a blue band in Fig.~\ref{fig:violin_kappa}).
+The hierarchical measurement leads to an expectation that $\kappa_i = $\variable{output/restricted_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies $\kappa = $\variable{output/generic_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
 
@@
         The orange solid line shows the generic posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
-        The pink vertical solid line represents $\kappa=0$.
+        The dashed line marks the absence of birefringence ($\kappa=0$).
+        \mi{THOMAS: please add a second set of curves \emph{without} GW200129; use the same colors you have now but plot them with a dashed trace}
     }
@@
 % Case: GW170818
-\subsection{Result on GW170818}
+\subsection{Notable events}
+
+Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on three events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, and GW190521.
+When considered in isolation, the first of these is the event with the most significant support for $\kappa \neq 0$; the second shows sign of potential noise anomalies; and the third is representative of a class of events with strongly bimodal $\kappa$ posteriors.
+
+\subsubsection{GW170818}
 GW170818 gave the largest tension with $\kappa=0$.
-Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$.
-This region corresponds to a larger distance and a smaller inclination angle.
-This is a sensible result because a smaller inclination angle will make the amplitude ratio between the two circular polarizations smaller.
+Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$ for
+ larger distance and a smaller inclination angles than would be allowed in the absence of birefringence.
+We can make sense of this result because a smaller inclination angle will make the amplitude ratio between the two circular polarizations smaller.
 Then the birefringence effect will increase the amplitude ratio during the propagation.
@@
 We can see this effect in other events as well, but not as strongly as in GW170818.
+
+\mi{THOMAS: did you have a comment to make about this case being similar to GW150914?
+Also, were there any other notable differences between the birefringence posterior and the LVK one?
+e.g., in other parameters like the masses or spins?}
 
@@
 
+\subsubsection{GW200129}
+\label{sec:GW200129}
+Past research suggests a potential glitch in GW200129 Virgo data. \citep{GW200129_glitch}
+We perform three sets of \ac{PE} to check if the glitch would affect the population posterior of $\kappa$.
+We pick only two out of three detector data in each set to perform \ac{PE}.
+(i.e.
+LIGO-Hanford and LIGO-Livingston for the first set, LIGO-Hanford and Virgo for the second set, and LIGO-Livingston and Virgo for the third set.)
+The two sets of \ac{PE} results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
+Further investigation is needed to understand the glitch.
+% Thus, we did not include this event in our collective analysis.
+
+Figure~\ref{fig:corner_GW200129} shows the posteriors of the parameters of GW200129 are strongly affected by the Virgo data.
+
+\begin{figure}
+    \script{corner_GW200129.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
+    \caption{
+        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW200129.
+        Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with and without Virgo data.
+        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
+        This plot shows that Virgo data may contain a glitch, which affects the \ac{PE} result.
+    }
+    \label{fig:corner_GW200129}
+\end{figure}
+
 % Case: GW150914 (example of broken degeneracy)
-\subsection{Result on GW150914}
+\subsubsection{Result on GW150914}
 In this study, we included frequency dependence of birefringence, which would affect the posterior of $\kappa$ obtained from \ac{PE}.
@@
 % Case: GW190521 (massive BBH)
-\subsection{Result on GW190521}
+\subsubsection{Result on GW190521}
 Even with the frequency-dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
@@
 % Case: GW200129 (glitch in Virgo data)
-\subsection{Contamination of GW200129}
-\label{sec:GW200129}
-Past research suggests a potential glitch in GW200129 Virgo data. \citep{GW200129_glitch}
-We perform three sets of \ac{PE} to check if the glitch would affect the population posterior of $\kappa$.
-We pick only two out of three detector data in each set to perform \ac{PE}.
-(i.e.
-LIGO-Hanford and LIGO-Livingston for the first set, LIGO-Hanford and Virgo for the second set, and LIGO-Livingston and Virgo for the third set.)
-The two sets of \ac{PE} results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
-Further investigation is needed to understand the glitch.
-% Thus, we did not include this event in our collective analysis.
-
-Figure~\ref{fig:corner_GW200129} shows the posteriors of the parameters of GW200129 are strongly affected by the Virgo data.
-
-\begin{figure}
-    \script{corner_GW200129.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
-    \caption{
-        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW200129.
-        Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with and without Virgo data.
-        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
-        This plot shows that Virgo data may contain a glitch, which affects the \ac{PE} result.
-    }
-    \label{fig:corner_GW200129}
-\end{figure}
-
 \section{Discussion}

### 2023-02-28 src/scripts/posterior_kappa.py

@@ -43,7 +43,7 @@ ll = ll - np.max(ll)
 likelihood = np.exp(ll)
 likelihood = likelihood / np.trapz(likelihood, x=kappa)
 
-plt.axvline(0, color=sns.color_palette()[4])
+plt.axvline(0, color='k', lw=2, ls='--')
 sns.kdeplot(np.array(samples_reweighted).reshape(-1), label="generic", color=sns.color_palette()[3])
 plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5), color=sns.color_palette()[3], linestyle=':')
 plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95), color=sns.color_palette()[3], linestyle=':')

### 2023-02-28 src/tex/ms.tex
GW200129 exposition

@@
 \subsubsection{GW170818}
-GW170818 gave the largest tension with $\kappa=0$.
+GW170818 gave the largest tension with $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
 Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$ for
@@
 \label{sec:GW200129}
-Past research suggests a potential glitch in GW200129 Virgo data. \citep{GW200129_glitch}
-We perform three sets of \ac{PE} to check if the glitch would affect the population posterior of $\kappa$.
-We pick only two out of three detector data in each set to perform \ac{PE}.
-(i.e.
-LIGO-Hanford and LIGO-Livingston for the first set, LIGO-Hanford and Virgo for the second set, and LIGO-Livingston and Virgo for the third set.)
-The two sets of \ac{PE} results with Virgo data return a $\kappa$ posterior that strongly suggests a negative $\kappa$.
-Further investigation is needed to understand the glitch.
-% Thus, we did not include this event in our collective analysis.
 
-Figure~\ref{fig:corner_GW200129} shows the posteriors of the parameters of GW200129 are strongly affected by the Virgo data.
+GW200129 is the event with the second largest $|\mu_i/\sigma_i|$ in Fig.~\ref{fig:violin_kappa}.
+However, data for this event were affected by a non-Gaussian noise disturbance (glitch) in the Virgo instrument, which was subtracted in the publicly-available data used for parameter estimation \cite{Davis:2022ird}.
+Since previous work suggests the degree of glitch subtraction affects the inference for this event \citep{GW200129_glitch}, we consider whether the apparent preference for $\kappa < 0$ could also be tied to the instrumental artifact.
+
+To this end, we perform three additional \ac{PE} runs for GW200129, considering only two detectors at a time: LIGO Hanford and Virgo (HL), LIGO Livingston and Virgo (LV), and LIGO Hanford and LIGO Livingston (HL).
+If the preference for $\kappa < 0$ is tied to the glitch in Virgo, we expect it to disappear in the HL run, which excludes Virgo data.
+
+This is indeed the case, as we show in Fig.~\ref{fig:corner_GW200129}: all runs including Virgo lean towards $\kappa < 0$ (solid curves in color), whereas the LIGO-only run is fully consistent with $\kappa = 0$ (dashed black).
+While this is not conclusive proof that the glitch itself is driving the result, it does indicate that the Virgo data play a key role in the inference of $\kappa$.
+Since more work would be needed to understand the effect of the glitch, this motivates us to consider the effect of excluding this event from the collective analyses above.
 
@@
     \caption{
-        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW200129.
-        Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with and without Virgo data.
-        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
-        This plot shows that Virgo data may contain a glitch, which affects the \ac{PE} result.
+        GW200129 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos{\iota}$, including different sets of detectors in the analysis per the legend.
+        The main run with all three detectors (HLV, filled blue), shows a preference for $\kappa < 0$, as in Fig.~\ref{fig:violin_kappa}; this preference is more pronounced for two-detector runs that include Virgo (HV and LV, orange and green); however, it disappears if we remove Virgo (HL, dashed black).
+        The 2D contours correspond to the $90\%$ credible level.
     }

### 2023-02-28 src/scripts/corner_GW190521.py

@@ -5,7 +5,7 @@ import numpy as np
 import seaborn as sns
 import paths
 
-sns.set_theme(palette='colorblind', font_scale=1.2)
+sns.set_theme(palette='colorblind', font_scale=1.5)
 
 plt.rcParams.update({
     "text.usetex": True,
@@ -36,8 +36,8 @@ g = sns.pairplot(result,
                  diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, levels=[(1.-0.90),(1.-0.3935)]))
 
 g.axes[2,0].set_xlabel(r"$\kappa$")
-g.axes[1,0].set_ylabel(r"$d_L$")
-g.axes[2,1].set_xlabel(r"$d_L$")
+g.axes[1,0].set_ylabel(r"$d_L$ (Mpc)")
+g.axes[2,1].set_xlabel(r"$d_L$ (Mpc)")
 g.axes[2,0].set_ylabel(r"$\cos\iota$")
 g.axes[2,2].set_xlabel(r"$\cos\iota$")
 g.fig.legends[0].set_bbox_to_anchor((0.65,0.8))

### 2023-02-28 src/scripts/corner_GW200129.py

@@ -5,7 +5,7 @@ import seaborn as sns
 from bilby.gw.result import CBCResult
 import paths
 
-sns.set_theme(palette='colorblind', font_scale=1.2)
+sns.set_theme(palette='colorblind', font_scale=1.5)
 
 plt.rcParams.update({
     "text.usetex": True,
@@ -65,8 +65,8 @@ g.map_lower(kdeplot, mask=m, linewidths=2, linestyles='--', common_norm=False, l
 g.map_diag(kdeplot, mask=m, linewidth=2, linestyle='--', common_norm=False, fill=False, zorder=100)
 
 g.axes[2,0].set_xlabel(r"$\kappa$")
-g.axes[1,0].set_ylabel(r"$d_L$")
-g.axes[2,1].set_xlabel(r"$d_L$")
+g.axes[1,0].set_ylabel(r"$d_L$ (Mpc)")
+g.axes[2,1].set_xlabel(r"$d_L$ (Mpc)")
 g.axes[2,0].set_ylabel(r"$\cos\iota$")
 g.axes[2,2].set_xlabel(r"$\cos\iota$")
 #g.fig.legends[0].set_bbox_to_anchor((0.65,0.8))

### 2023-03-01 src/tex/ms.tex
section rearrangement
Add content of GW200129

@@
 
-The two events that show the largest tension with $\kappa = 0$ is GW170818, for which $|\mu_i / \sigma_i| = \variable{output/GW170818_constraint.txt}$, and GW200129\_065458 (henceforth GW200129), for which $|\mu_i / \sigma_i| = \dots$.
-\mi{THOMAS: please add the number for GW200129. Also, we don't need that many significant figures, one decimal will do.}
+The two events that show the largest tension with $\kappa = 0$ is GW170818, for which $|\mu_i / \sigma_i| =$ \variable{output/GW170818_constraint.txt}, and GW200129\_065458 (henceforth GW200129), for which $|\mu_i / \sigma_i| =$ \variable{output/GW200129_constraint.txt}.
 However, as we discuss further below, we have reason to think that the GW200129 preference for $\kappa < 0$ might be driven by noise anomalies in the Virgo detector; for that reason we consider the effect of excluding it from the joint result below.
@@
         The dashed line marks the absence of birefringence ($\kappa=0$).
-        \mi{THOMAS: please add a second set of curves \emph{without} GW200129; use the same colors you have now but plot them with a dashed trace}
     }
@@
 
-% Case study
-
-% Case: GW170818
+% Case studies
 \subsection{Notable events}
@@
 
+% Case: GW170818
 \subsubsection{GW170818}
@@
 
+% Case: GW200129 (glitch in Virgo data)
 \subsubsection{GW200129}
@@
     \label{fig:corner_GW200129}
+\end{figure}
+
+% Case: GW190521 (massive BBH)
+\subsubsection{GW190521}
+Even with the frequency-dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
+Consider GW190521, the most massive binary black hole merger in the events we included.
+The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in Fig.~\ref{fig:corner_GW190521}.
+The reason could be this GW event's frequency range is too narrow to break the degeneracy.
+The effect of birefringence at different frequencies within the range is similar, so the degeneracy cannot be broken.
+Therefore, the narrow frequency range can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
+Note that $\kappa=0$ is still within the $90\%$ confidence interval, with a non-negligible probability.
+The bimodal $\kappa$ distribution only means that the degeneracy between $\kappa$ and $\iota$ cannot be broken in this case.
+
+\begin{figure}[h]
+    \script{corner_GW190521.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
+    \caption{
+        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
+        Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with the frequency-dependent birefringence.
+        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
+        This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
+    }
+    \label{fig:corner_GW190521}
 \end{figure}
@@
 
-% Case: GW190521 (massive BBH)
-\subsubsection{Result on GW190521}
-Even with the frequency-dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
-Consider GW190521, the most massive binary black hole merger in the events we included.
-The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in Fig.~\ref{fig:corner_GW190521}.
-The reason could be this GW event's frequency range is too narrow to break the degeneracy.
-The effect of birefringence at different frequencies within the range is similar, so the degeneracy cannot be broken.
-Therefore, the narrow frequency range can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
-Note that $\kappa=0$ is still within the $90\%$ confidence interval, with a non-negligible probability.
-The bimodal $\kappa$ distribution only means that the degeneracy between $\kappa$ and $\iota$ cannot be broken in this case.
-
-\begin{figure}[h]
-    \script{corner_GW190521.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
-    \caption{
-        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
-        Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with the frequency-dependent birefringence.
-        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
-        This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
-    }
-    \label{fig:corner_GW190521}
-\end{figure}
-
-% Case: GW200129 (glitch in Virgo data)
 \section{Discussion}

### 2023-03-01 src/scripts/GW200129_constraint.py

@@ -1,0 +1,9 @@
+import pandas as pd
+import paths
+
+result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+
+mean_over_std_GW200129 = result_DataFrame[result_DataFrame.event == "GW200129_065458"]['kappa'].mean() / result_DataFrame[result_DataFrame.event == "GW200129_065458"]['kappa'].std()
+
+with open(paths.output/"GW200129_constraint.txt", "w") as f:
+    f.write(rf"${mean_over_std_GW200129:.1f}$")

### 2023-03-01 src/scripts/posterior_kappa.py

@@ -16,6 +16,7 @@ plt.rcParams.update({
 def normal_distribution(x, mu, sigma):
     return (1/(sigma*np.sqrt(2*np.pi)))*np.exp(-0.5*((x-mu)/sigma)**2)
 
+# Result with GW200129
 result_dict = {}
 result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
 for event in result_DataFrame['event'].unique():
@@ -45,11 +46,47 @@ likelihood = likelihood / np.trapz(likelihood, x=kappa)
 
 plt.axvline(0, color='k', lw=2, ls='--')
 sns.kdeplot(np.array(samples_reweighted).reshape(-1), label="generic", color=sns.color_palette()[3])
-plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5), color=sns.color_palette()[3], linestyle=':')
-plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95), color=sns.color_palette()[3], linestyle=':')
+# plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5), color=sns.color_palette()[3], linestyle=':')
+# plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95), color=sns.color_palette()[3], linestyle=':')
 plt.plot(kappa,likelihood, label="restricted", color=sns.color_palette()[0])
-plt.axvline(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[0], linestyle=':')
-plt.axvline(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[0], linestyle=':')
+# plt.axvline(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[0], linestyle=':')
+# plt.axvline(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[0], linestyle=':')
+
+# Result without GW200129
+result_dict = {}
+result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
[…]

### 2023-03-01 src/tex/ms.tex
Update text on GW170818

@@
 Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on three events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, and GW190521.
-When considered in isolation, the first of these is the event with the most significant support for $\kappa \neq 0$; the second shows sign of potential noise anomalies; and the third is representative of a class of events with strongly bimodal $\kappa$ posteriors.
+When considered in isolation, the first of these is the event with the most significant support for $\kappa \neq 0$; the second shows the sign of potential noise anomalies; and the third is representative of a class of events with strongly bimodal $\kappa$ posteriors.
 
@@
 GW170818 gave the largest tension with $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
-Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$ for
- larger distance and a smaller inclination angles than would be allowed in the absence of birefringence.
+Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$ for a larger distance and smaller inclination angles than would be allowed in the absence of birefringence.
 We can make sense of this result because a smaller inclination angle will make the amplitude ratio between the two circular polarizations smaller.
-Then the birefringence effect will increase the amplitude ratio during the propagation.
+On the other hand, the birefringence effect will increase the amplitude ratio during propagation.
 The larger distance is for compensating the combined effect of a nonzero $\kappa$ and a smaller inclination angle.
+This is similar to the degeneracy mentioned in Sec.~\ref{sec:inclination}.
+Even though the degeneracy is broken by the frequency dependence to a large extent, it is still possible for getting an extra region with a nonzero $\kappa$.
+This is because there is still a significant likelihood in the region of $\kappa$ close to zero.
 We can see this effect in other events as well, but not as strongly as in GW170818.
-
-\mi{THOMAS: did you have a comment to make about this case being similar to GW150914?
-Also, were there any other notable differences between the birefringence posterior and the LVK one?
-e.g., in other parameters like the masses or spins?}
+Other parameters such as the spins and polarization angles are notably different from the \ac{GR} case as well.
+They may have some slight effect on the degeneracy, but they are not correlated with $\kappa$ significantly.
 
@@
 GW200129 is the event with the second largest $|\mu_i/\sigma_i|$ in Fig.~\ref{fig:violin_kappa}.
-However, data for this event were affected by a non-Gaussian noise disturbance (glitch) in the Virgo instrument, which was subtracted in the publicly-available data used for parameter estimation \cite{Davis:2022ird}.
+However, data for this event were affected by a non-Gaussian noise disturbance (glitch) in the Virgo instrument, which was subtracted from the publicly-available data used for parameter estimation \cite{Davis:2022ird}.
 Since previous work suggests the degree of glitch subtraction affects the inference for this event \citep{GW200129_glitch}, we consider whether the apparent preference for $\kappa < 0$ could also be tied to the instrumental artifact.

### 2023-03-01 src/tex/ms.tex
plot tweaks and GW170818

@@
 \subsection{Hierarchical modeling}
+\label{sec:results:hier}
 
@@
 Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on three events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, and GW190521.
-When considered in isolation, the first of these is the event with the most significant support for $\kappa \neq 0$; the second shows the sign of potential noise anomalies; and the third is representative of a class of events with strongly bimodal $\kappa$ posteriors.
+When considered in isolation, the first of these is the event with the most significant support for nonzero $\kappa$; the second shows the sign of potential noise anomalies; and the third is representative of a class of events with strongly bimodal $\kappa$ posteriors.
 
@@
 \subsubsection{GW170818}
-GW170818 gave the largest tension with $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
-Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$ for a larger distance and smaller inclination angles than would be allowed in the absence of birefringence.
-We can make sense of this result because a smaller inclination angle will make the amplitude ratio between the two circular polarizations smaller.
-On the other hand, the birefringence effect will increase the amplitude ratio during propagation.
-The larger distance is for compensating the combined effect of a nonzero $\kappa$ and a smaller inclination angle.
+GW170818 produced the posterior in most tension with $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
+Figure \ref{fig:corner_GW170818} shows that this happens because birefringence opens up a region of parameter space with $\kappa>0$ for larger distances and smaller inclination angles than would be allowed in the GR case.
+We can make sense of this because an edge-on, non-precessing source will produce linearly-polarized waves, so a smaller inclination will lead the two circular polarizations to have similar amplitudes.
+On the other hand, having $\kappa > 0$ will enhance the left-handed modes during propagation.
+The two effects can be balanced to yield the correct polarization ratio at the detector, as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
 This is similar to the degeneracy mentioned in Sec.~\ref{sec:inclination}.
-Even though the degeneracy is broken by the frequency dependence to a large extent, it is still possible for getting an extra region with a nonzero $\kappa$.
-This is because there is still a significant likelihood in the region of $\kappa$ close to zero.
-We can see this effect in other events as well, but not as strongly as in GW170818.
-Other parameters such as the spins and polarization angles are notably different from the \ac{GR} case as well.
-They may have some slight effect on the degeneracy, but they are not correlated with $\kappa$ significantly.
+% Even though the degeneracy is broken by the frequency dependence to a large extent, it is still possible for getting an extra region with a nonzero $\kappa$.
+% This is because there is still a significant likelihood in the region of $\kappa$ close to zero.
+% We can see this effect in other events as well, but not as strongly as in GW170818.
+
+It is not clear what specific feature of the GW170818 data leads to this posterior structure.
+However, the reason may be related to the fact that this event provides an uncommonly definite measurement of the polarization angle and spin orientations \cite{Varma:2021csh}.
+Indeed, the inferred polarization angle and spins in the birefringent case are are notably different from the \ac{GR} case (Fig. XXX).
+If our inference on $\kappa$ is indeed tied to those parameters, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
+In any case, as we argued in Sec.~\ref{sec:results:hier}, there is no evidence at the moment that this event is inconsistent with the either collection of measurements or with $\kappa = 0$ (Fig.~\ref{fig:reweighted_kappa}).
 
@@
     \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
-    \caption{Corner plot of GW170818.}
+    \caption{GW170818 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).}
     \label{fig:corner_GW170818}

### 2023-03-01 src/tex/ms.tex
Some comments from Will.

--- 2023-03-01 src/tex/ms.tex
@@
 \newcommand*{\mi}[1]{\textsf{\color{magenta} [\textbf{MAX:} #1]}}
+\newcommand*{\wf}[1]{\textsf{\color{cyan} [\textbf{WILL:} #1]}}
 
@@
 
-We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
-We take the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
-We apply a distance prior which is uniform in comoving volume and source frame time instead of a power law distribution used in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby},
+that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline
+waveform. We take the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3,
+GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with
+\textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform. We
+apply a distance prior which is uniform in comoving volume and source frame time
+instead of a power law distribution used in \citet{GWTC-2.1, GWTC-3,
+GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform
+distribution between $-1$ and $1$.  \wf{Actually, the LIGO references
+\emph{also} used the ``cosmological'' prior that is uniform merger rate density
+in the source frame; maybe just drop the ``instead of\ldots''?}
 
@@
 
-Since birefringence is a property of spacetime and we are targeting a propagation effect not intrinsic to any source, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
-\mi{caveats?}
-Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply obtained from the product of the individual likelihoods, $p(d_i \mid \kappa) \propto p(\kappa \mid d_i)/p(\kappa)$, such that
+Since birefringence is a property of spacetime and we are targeting a
+propagation effect not intrinsic to any source, we should expect $\kappa$ to
+take the same value for all signals, whether it vanishes or not. \mi{caveats?}
+\wf{\emph{If} birefringence were a property of spacetime, this would be true;
+but in (some? most? all?) ``parity violating'' theories, the parity violation is
+due to an extra \emph{field} that is parity-odd and couples to gravity.  In this
+case whether the birefringence is the same along the different lines of sight to
+the different events depends on \emph{whatever that field is doing between us
+and the events}.  The \emph{simplest} assumption is that the field is ``the
+same'' or shared from event-to-event, but it's by no means required that it be
+so.  Hence the analysis of the next section.} Under this assumption, the
+posterior on $\kappa$ inferred collectively from all events with a uniform prior
+is simply obtained from the product of the individual likelihoods, $p(d_i \mid
+\kappa) \propto p(\kappa \mid d_i)/p(\kappa)$, such that
 \begin{equation}
@@
 This could signal further physics than is implied by Eq.~\eqref{eq:waveform_modification} \mi{which physics?
-@Will, any ideas?} or, more prosaically, that there are outliers in our measurements due to mismodeling, e.g., in the waveform approximant or the noise of the detector.
+@Will, any ideas?} \wf{see comment above} or, more prosaically, that there are outliers in our measurements due to mismodeling, e.g., in the waveform approximant or the noise of the detector.
 

### 2023-03-01 src/tex/ms.tex
Add GW170818 corner to appendix
More comments from Will.

@@
 The hierarchical measurement leads to an expectation that $\kappa_i = $\variable{output/restricted_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies $\kappa = $\variable{output/generic_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
+\wf{It is very interesting to me that the analysis with $\mu$ and $\sigma$ provides a \emph{tighter} constraint on $\kappa$ than the one with the shared-$\kappa$ value.
+@Max: maybe we should polish off that note with Katerina and Cole about the benefts / costs of doing the $\mu$-$\sigma$ analysis?
+Do either of you understand how this comes about?
+(I can make up some b.s.\ about the bimodal posteriors yielding a better model fit for the $\mu$-$\sigma$ analysis than the fixed-$\kappa$ analysis, but I'm not sure it will hold water.)}
 

### 2023-03-01 src/tex/ms.tex
Will comments on 190521.

@@
 The bimodal $\kappa$ distribution only means that the degeneracy between $\kappa$ and $\iota$ cannot be broken in this case.
+\wf{You should still explain \emph{why} the model seems to prefer $\left| \kappa \right| > 0$.
+I \emph{think} the reason here is the distance prior---there is a lot more ``phase space'' at large distances that can be accessed by allowing $\kappa$ to be non-zero and choosing a corresponding inclination so that the amplification effects of birefringence are compensated for by the larger distance compared to the pure-GR run.
+Does this make sense to you?
+If so, can you write it up in this paragraph?}
 
@@
 
+\section{GW170818}
+
+\begin{figure}
+    \script{corner_GW170818_appendix.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW170818_appendix.pdf}
+    \caption{
+        Corner plot of GW170818.
+    }
+    \label{fig:corner_GW170818_appendix}
+\end{figure}
+
 \section{Relation between $\kappa$ and $M_{PV}$}

### 2023-03-01 src/tex/ms.tex
Final comments from Will re: Wang, et al.
Correct the kappa constraint

@@
 We show the result of doing this in Fig.~\ref{fig:posterior_kappa}, where we also compare to the posterior on $\kappa$ obtained by assuming a shared value across events (same result shown as a blue band in Fig.~\ref{fig:violin_kappa}).
-The hierarchical measurement leads to an expectation that $\kappa_i = $\variable{output/restricted_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies $\kappa = $\variable{output/generic_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
+The hierarchical measurement leads to an expectation that $\kappa_i = $\variable{output/generic_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies $\kappa = $\variable{output/restricted_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
 \wf{It is very interesting to me that the analysis with $\mu$ and $\sigma$ provides a \emph{tighter} constraint on $\kappa$ than the one with the shared-$\kappa$ value.
@@
 They formulated the birefringence effect as corrections to the GW waveform on a linear basis.
-We convert their constraint to the same units as ours with the derivation in Appendix~\ref{sec:M_PV_derivation}.
+We convert their constraint to the same units as ours with the derivation in Appendix~\ref{sec:M_PV_derivation}.  \wf{Can you give the answer here (as well as in Appendix B)?}
 % Discussion on the difference between our work and theirs

### 2023-03-01 src/tex/ms.tex
Update figure captions

@@
         Individual-event $\kappa$ posteriors (distributions), and joint measurement (blue band, 90\% CI; blue line, median).
-%        Each violin represents a different event.
-%        The color of the violins represents the quotient of the median and standard deviation of the posterior.
-%        The blue horizontal solid line represents the median of the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval of the restricted posterior of $\kappa$.
-%        The pink horizontal solid line represents $\kappa=0$.
-%        GW200129 is shown with light blue edges, as it is excluded from the calculation of collective analysis.
     }
@@
         The generic and restricted posterior of $\kappa$.
-        The blue solid line shows the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
-        The orange solid line shows the generic posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
-        The dashed line marks the absence of birefringence ($\kappa=0$).
+        The blue solid and dashed lines show the restricted posterior of $\kappa$ with and without GW200129, respectively.
+        The orange solid and dashed lines show the generic posterior of $\kappa$ with and without GW200129, respectively.
+        The black dashed line marks the absence of birefringence ($\kappa=0$).
     }
@@
     \caption{
-        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
-        Colors in the plot are the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} and the \ac{PE} result done by us with the frequency-dependent birefringence.
-        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
-        This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
+        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+        % The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
+        % Colors in the plot represent the \ac{PE} result with \ac{GR} and the frequency-dependent birefringence, respectively.
+        % The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
+        % This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
     }
@@
     \caption{
-        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-        Colors in the plot represent the \ac{PE} result with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}, the \ac{PE} results done by us with both frequency-independent and dependent birefringence respectively.
-        The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
-        Note that there is no posterior of $\kappa$ for the \ac{PE} result from LVK, as the LVK \ac{PE} is based on GR, which does not suggest GW amplitude birefringence.
-        This plot shows that frequency-independent birefringence creates a degeneracy between $\kappa$ and $\iota$, while frequency-dependent birefringence can break it.
+        GW150914 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue) and frequency-independent (orange) birefringence analysis, compared to the GR result (green).
+        % The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
+        % Colors in the plot represent the \ac{PE} result with \ac{GR}, frequency-independent and dependent birefringence, respectively.
+        % The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
+        % Note that there is no posterior of $\kappa$ for the \ac{PE} result with \ac{GR}, as the result is based on GR, which does not suggest GW amplitude birefringence.
+        % This plot shows that frequency-independent birefringence creates a degeneracy between $\kappa$ and $\iota$, while frequency-dependent birefringence can break it.
     }

### 2023-03-01 src/tex/ms.tex
Formatting

@@
 \begin{table}
-    \caption{Events that best constrain $\kappa$, sorted by posterior standard deviation $\sigma_i$.}
+    \caption{Events that best constrain $\kappa$, sorted by posterior standard deviation $\sigma_i$. CL is the credible level of $\kappa = 0$.}
     \begin{ruledtabular}
@@
 \begin{table}
-    \caption{Events that show bimodality in the $\kappa$ posterior tend to have high total masses.}
+    \caption{Events that show bimodality in the $\kappa$ posterior tend to have high total masses. CL is the credible level of $\kappa = 0$.}
     \begin{ruledtabular}
@@
 
-\begin{figure}
+\begin{figure}[h]
     \script{corner_GW170818_appendix.py}

### 2023-03-01 src/scripts/bimodal_events_mass.py

@@ -6,7 +6,7 @@ import paths
 result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
 
 with open(paths.output/"bimodal_events_mass.txt", "w") as f:
-    f.write(r"\begin{tabular}{lrrr}Event & Total Mass ($M_{\odot}$) & $\kappa$ & CL($\kappa=0$)\\ \hline ")
+    f.write(r"\begin{tabular}{lrrr}Event & Total Mass ($M_{\odot}$) & $\kappa$ & CL\\ \hline ")
     for event in ["GW170104", "GW190413_134308", "GW190521", "GW190805_211137", "GW191105_143521"]:
         name = event.replace("_", r"\_")
 
@@ -25,7 +25,7 @@ with open(paths.output/"bimodal_events_mass.txt", "w") as f:
         kde = np.array(kde)
         credible_level = len(kde[kde > kernel(0)[0]])/len(kde)
 
-        f.write(rf"{name} & ${mass_median:.1f}^{{+{mass_95:.1f}}}_{{-{mass_5:.1f}}}$ & ${kappa_median:.3f}^{{+{kappe_95:.3f}}}_{{-{kappe_5:.3f}}}$ & ${credible_level:.2f}$")
+        f.write(rf"{name} & ${mass_median:.1f}^{{+{mass_95:.1f}}}_{{-{mass_5:.1f}}}$ & ${kappa_median:.3f}^{{+{kappe_95:.3f}}}_{{-{kappe_5:.3f}}}$ & ${credible_level:.3f}$")
         if event != "GW191105_143521":
             f.write(r"\\")
     f.write(r" \end{tabular}")

### 2023-03-01 src/scripts/corner_GW170818_appendix.py

@@ -23,7 +23,7 @@ result_GR['cos_iota'] = np.cos([float(value) for value in result_GR['iota']])
 result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
 result_bilby = result_bilby[result_bilby.event == "GW170818"]
 result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), r"birefringence (frequency dependent)")
+result_bilby['with'] = np.full(len(result_bilby), r"BR")
 result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
 
 result = pd.concat([result_bilby,result_GR], ignore_index=True)
@@ -56,6 +56,6 @@ g.axes[10,9].set_xlabel("$\\phi_{JL}$")
 g.axes[10,0].set_ylabel("$\\phi_{\\rm{ref}}$")
 g.axes[10,10].set_xlabel("$\\phi_{\\rm{ref}}$")
 
-g.fig.legends[0].set_bbox_to_anchor((0.87,0.5))
+g.fig.legends[0].set_bbox_to_anchor((0.89,0.5))
 
 g.savefig(fname=paths.figures/"corner_GW170818_appendix.pdf", bbox_inches="tight", dpi=300)

### 2023-03-02 src/tex/ms.tex
Update ms.tex
merged in most of Will's comments

@@
 \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform. We
-apply a distance prior which is uniform in comoving volume and source frame time
-instead of a power law distribution used in \citet{GWTC-2.1, GWTC-3,
-GWTC-2.1_dataset, GWTC-3_dataset}, and set the prior on $\kappa$ to a uniform
-distribution between $-1$ and $1$.  \wf{Actually, the LIGO references
-\emph{also} used the ``cosmological'' prior that is uniform merger rate density
-in the source frame; maybe just drop the ``instead of\ldots''?}
+apply a distance prior which is uniform in comoving volume and source frame time, and set the prior on $\kappa$ to a uniform
+distribution between $-1$ and $1$.
 
@@
 
-Since birefringence is a property of spacetime and we are targeting a
-propagation effect not intrinsic to any source, we should expect $\kappa$ to
-take the same value for all signals, whether it vanishes or not. \mi{caveats?}
-\wf{\emph{If} birefringence were a property of spacetime, this would be true;
-but in (some? most? all?) ``parity violating'' theories, the parity violation is
-due to an extra \emph{field} that is parity-odd and couples to gravity.  In this
-case whether the birefringence is the same along the different lines of sight to
-the different events depends on \emph{whatever that field is doing between us
-and the events}.  The \emph{simplest} assumption is that the field is ``the
-same'' or shared from event-to-event, but it's by no means required that it be
-so.  Hence the analysis of the next section.} Under this assumption, the
-posterior on $\kappa$ inferred collectively from all events with a uniform prior
+In the most simplified scenarios, birefringence is a property of spacetime that is not intrinsic to any source or region in space. 
+If so, we should expect $\kappa$ to
+take the same value for all signals, whether it vanishes or not.
+Under this assumption, the
+posterior on $\kappa$ inferred collectively from all events 
 is simply obtained from the product of the individual likelihoods, $p(d_i \mid
@@
 % Bayesian Hierarchical Modeling
-In order to further characterize our set of measurements, and identify potential outliers, we carry out an additional collective analysis that allows $\kappa$ to take different values for each event \cite{Zimmerman:2019wzo,Isi:2022cii}.
+Under many frameworks, birefringence is brokered by an extra parity-odd
+field that couples to gravity. In this case, whether the birefringence
+is the same along the different lines of sight to the different events depends
+on cosmic history and the local evolution of the field. The \emph{simplest}
+assumption is that the field manifests equally for all events, as assumed in the previous subsection, but this is by
+no means required.
+This motivates a collective analysis that does not assume $\kappa$ is shared across events \cite{Zimmerman:2019wzo,Isi:2022cii}.
+Such an analysis has the additional advantage of helping us further characterize our set of measurements, and identify potential outliers.
+
 To do this, we apply hierarchical Bayesian inference \cite{Loredo:2004nn} to model the distribution of $\kappa$'s consistent with the observed data:
@@
 But this analysis also has the power to reveal unexpected physics or systematics in our measurements: if $\sigma$ is confidently found to be nonzero, this would imply that our set of measurements is statistically unlikely to originate from a unique $\kappa$ value.
-This could signal further physics than is implied by Eq.~\eqref{eq:waveform_modification} \mi{which physics?
-@Will, any ideas?} \wf{see comment above} or, more prosaically, that there are outliers in our measurements due to mismodeling, e.g., in the waveform approximant or the noise of the detector.
+This could signal richer physics than is implied by Eq.~\eqref{eq:waveform_modification} or, more prosaically, that there are outliers in our measurements due to mismodeling, e.g., in the waveform approximant or the noise of the detector.
 
@@
 
-\begin{table}
+\begin{table}[h]
     \caption{Events that best constrain $\kappa$, sorted by posterior standard deviation $\sigma_i$. CL is the credible level of $\kappa = 0$.}
@@
 
-\begin{table}
+\begin{table}[h]
     \caption{Events that show bimodality in the $\kappa$ posterior tend to have high total masses. CL is the credible level of $\kappa = 0$.}
@@
 The hierarchical measurement leads to an expectation that $\kappa_i = $\variable{output/generic_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies $\kappa = $\variable{output/restricted_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
-\wf{It is very interesting to me that the analysis with $\mu$ and $\sigma$ provides a \emph{tighter} constraint on $\kappa$ than the one with the shared-$\kappa$ value.
-@Max: maybe we should polish off that note with Katerina and Cole about the benefts / costs of doing the $\mu$-$\sigma$ analysis?
-Do either of you understand how this comes about?
-(I can make up some b.s.\ about the bimodal posteriors yielding a better model fit for the $\mu$-$\sigma$ analysis than the fixed-$\kappa$ analysis, but I'm not sure it will hold water.)}
 
@@
 However, the reason may be related to the fact that this event provides an uncommonly definite measurement of the polarization angle and spin orientations \cite{Varma:2021csh}.
-Indeed, the inferred polarization angle and spins in the birefringent case are are notably different from the \ac{GR} case (Fig. XXX).
+Indeed, the inferred polarization angle and spins in the birefringent case are notably different from the \ac{GR} case (Fig.~\ref{fig:corner_GW170818_appendix}).
 If our inference on $\kappa$ is indeed tied to those parameters, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.

### 2023-03-02 src/scripts/corner_GW191105.py
Add corner plot for GW191105

@@ -1,0 +1,33 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+from bilby.gw.result import CBCResult
+import paths
+
+sns.set_theme(palette='colorblind', font_scale=1.5)
+
+plt.rcParams.update({
+    "text.usetex": True,
+    "font.family": "serif",
+    "font.serif": ["Computer Modern Roman"],
+})
+
+result = CBCResult.from_json(filename=paths.data/"GW191105_143521_birefringence.json.gz").posterior
+result['cos_iota'] = np.cos([float(value) for value in result['iota']])
+
+g = sns.pairplot(result,
+                    vars=['kappa', 'luminosity_distance', 'cos_iota', 'chi_eff', 'chi_p'],
+                    corner=True, kind='hist', diag_kws=dict(common_norm=False), plot_kws=dict(common_norm=False, bins=100, rasterized=True))
+
+g.axes[4,0].set_xlabel(r"$\kappa$")
+g.axes[1,0].set_ylabel(r"$d_L$ (Mpc)")
+g.axes[4,1].set_xlabel(r"$d_L$ (Mpc)")
+g.axes[2,0].set_ylabel(r"$\cos\iota$")
+g.axes[4,2].set_xlabel(r"$\cos\iota$")
+g.axes[3,0].set_ylabel(r"$\chi_{eff}$")
+g.axes[4,3].set_xlabel(r"$\chi_{eff}$")
+g.axes[4,0].set_ylabel(r"$\chi_p$")
+g.axes[4,4].set_xlabel(r"$\chi_p$")
+plt.subplots_adjust(wspace=0.05, hspace=0.05)
+
+g.savefig(fname=paths.figures/"corner_GW191105.pdf", bbox_inches="tight", dpi=300)

### 2023-03-02 src/tex/ms.tex
Add corner plot for GW191105
new text for spin degeneracies

@@
 
-\subsection{Relation to inclination}
+\subsection{Inclination and other degeneracies}
 \label{sec:inclination}
@@
 \end{equation}
-For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination can mask the effect of birefringence, which will affect the time domain waveform nontrivially (Fig.~\ref{fig:birefringence}).
+For an isolated Fourier mode of definite frequency $f$, the effect of birefringence will be degenerate with a change in inclination; however, if multiple modes come into play, then no change in inclination alone can mask the effect of birefringence, which will affect the time domain waveform nontrivially (Fig.~\ref{fig:birefringence}).
 
@@
 As we will see in Sec.~\ref{sec:Results}, this is the case for heavy \acp{BBH}, which are only in the sensitive band of the detectors for a few cycles.
+
+Besides source inclination, the effect of birefringence can be (partially) degenerate with other parameters---most notably, precession, as we find in Sec.~\ref{sec:Results}.
+Indeed, under the right conditions, the frequency-dependent amplification/dampening of polarizations caused by $\kappa$ can be effectively absorbed by changes in intrinsic parameters, like the spins;
+in that case, the degeneracy with inclination may appear even for lighter \acp{BBH} (although usually to a lesser degree than heavy binaries).
+We chart some of these degeneracies as part of the results we present in Sec.~\ref{sec:Results}.
+% In particular, we find that, under certain conditions, the frequency-dependent birefringence of Eq.~\eqref{eq:waveform_modification} can mimic the time-dependent amplitude modulation expected from a precessing system.
 
@@
 
-\subsection{Result on GWTC-3}
+\subsection{GWTC-3 result}
 
@@
 as represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate).
-Posteriors are colored by the distance of the respective posterior mean from the origin in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.%
+Posteriors are colored by the distance to the origin of the respective mean in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.%
 \footnote{We use the $i$ notation for individual events here to disambiguate from the population mean $\mu$ and standard deviation $\sigma$ of Eq.~\eqref{eq:posterior_of_mu_sigma}.}
@@
 Figure \ref{fig:violin_kappa} makes it clear that not all \acp{BBH} in GWTC-3 are equally informative about birefringence.
-When considered individually, the events that best constrain $\kappa$ are listed in Table~\ref{tab:best_events_kappa}.
+When considered individually, the events that best constrain $\kappa$ are listed in Table~\ref{tab:best_events_kappa}, in order of increasing standard deviation $\sigma_i$.
+That table also shows the credible level (CL) at which the posterior supports $\kappa = 0$, whereby $\mathrm{CL} = 0$ ($\mathrm{CL} = 1$) means the posterior supports that value with high (low) probability.
 
-The two events that show the largest tension with $\kappa = 0$ is GW170818, for which $|\mu_i / \sigma_i| =$ \variable{output/GW170818_constraint.txt}, and GW200129\_065458 (henceforth GW200129), for which $|\mu_i / \sigma_i| =$ \variable{output/GW200129_constraint.txt}.
-However, as we discuss further below, we have reason to think that the GW200129 preference for $\kappa < 0$ might be driven by noise anomalies in the Virgo detector; for that reason we consider the effect of excluding it from the joint result below.
+Judging by $\mu_i/\sigma_i$, the two events that show the largest tension with $\kappa = 0$ are GW170818, for which $\mu_i / \sigma_i =$ \variable{output/GW170818_constraint.txt}, and GW200129\_065458 (henceforth GW200129), for which $\mu_i / \sigma_i =$ \variable{output/GW200129_constraint.txt}.
+However, as we discuss in Sec.~\ref{sec:GW200129}, we have reason to think that the preference for $\kappa < 0$ in GW200129 might be driven by noise anomalies in the Virgo detector; for that reason, in the next section we consider the effect of excluding this event from the joint result (we find its impact to be minimal).
 
-\begin{table}[h]
+\begin{table}
     \caption{Events that best constrain $\kappa$, sorted by posterior standard deviation $\sigma_i$. CL is the credible level of $\kappa = 0$.}
@@
 Finally, a set of events stands out in Fig.~\ref{fig:violin_kappa} due to evident bimodality in the $\kappa$ posterior.
-That is the case for the events listed in Table~\ref{tab:bimodal_events_mass}.
-We understand this effect to be a consequence of the narrow sets of frequencies that are visible for these heavy events, as we mentioned in Sec.~\ref{sec:inclination} and further demonstrate below.
+To a varying degree, that is the case for those events listed in Table~\ref{tab:bimodal_events_mass}, which tend to have quite high total masses (Table~\ref{tab:bimodal_events_mass} show total mass as measured in the standard  \ac{GR} analysis).
+\mi{THOMAS: is this correct? do these numbers come from the GR runs? they should}
+For these bimodal posteriors, $\mu_i/\sigma_i$ is not a good proxy for agreement with \ac{GR}; instead, we can rely on $\mathrm{CL}(\kappa = 0)$.
+By this measure, the bimodal events are some of the least consistent with $\kappa = 0$, GW190521 in particular.
+(However, note that configurations with $\kappa = 0$ for this event are still within the higher dimensional 90\%-credible region, when other parameters are considered and not just the one-dimensional marginal; we discuss this in Sec.~\ref{sec:GW190521}.)
 
-\begin{table}[h]
+As we mentioned in Sec.~\ref{sec:inclination}, we understand the bimodality in $\kappa$ to be primarily a consequence of the narrow set of frequencies that are visible for these heavy events, which makes the effect of birefringence (partially) degenerate with a change in inclination.
+Other parameter degeneracies also come into play, especially for the lighter events GW191105\_143521 and GW170104.
+We discuss this in detail in a dedicated section below (Sec.~\ref{sec:notable}).
+
+\begin{table}
     \caption{Events that show bimodality in the $\kappa$ posterior tend to have high total masses. CL is the credible level of $\kappa = 0$.}
@@
 \end{table}
-
-\subsection{Hierarchical modeling}
-\label{sec:results:hier}
-
-Figure \ref{fig:violin_kappa} shows certain variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
-This is not unexpected: assuming independent Gaussian noise instantiations for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$.
-
-To further assess the statistical properties of the set of posteriors in Fig.~\ref{fig:violin_kappa}, we apply the hierarchical analysis described in Sec.~\ref{sec:method:hier}.
-By characterizing the population mean and standard deviation over events, this also allows us to obtain a collective measurement that does not assume all events share the same value of $\kappa$.
 
@@
 \end{figure}
+
+\subsection{Hierarchical modeling}
+\label{sec:results:hier}
+
+Figure \ref{fig:violin_kappa} shows certain degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
+This is not unexpected: assuming independent Gaussian noise instantiations for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$.
+
+To further assess the statistical properties of the set of posteriors in Fig.~\ref{fig:violin_kappa}, we apply the hierarchical analysis described in Sec.~\ref{sec:method:hier}.
+By characterizing the population mean and standard deviation over events, this also allows us to obtain a collective measurement that does not assume all events share the same value of $\kappa$.
 
@@
 \subsection{Notable events}
+\label{sec:notable}
 
@@
 \subsubsection{GW170818}
+\label{sec:GW170818}
+
 GW170818 produced the posterior in most tension with $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
@@
 \subsubsection{GW190521}
+\label{sec:GW190521}
+
 Even with the frequency-dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
@@
 % Case: GW150914 (example of broken degeneracy)
-\subsubsection{Result on GW150914}
+\subsubsection{GW150914}
 In this study, we included frequency dependence of birefringence, which would affect the posterior of $\kappa$ obtained from \ac{PE}.
@@
 
+\section{placeholder for GW191105}
+
+\begin{figure}[h]
+    \script{corner_GW191105.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
+    \caption{
+        Corner plot of GW191105.
+    }
+    \label{fig:corner_GW191105}
+\end{figure}
+
 \section{GW170818}

### 2023-03-02 src/tex/ms.tex
started work on GW170818 precession

@@
 As we mentioned in Sec.~\ref{sec:inclination}, we understand the bimodality in $\kappa$ to be primarily a consequence of the narrow set of frequencies that are visible for these heavy events, which makes the effect of birefringence (partially) degenerate with a change in inclination.
-Other parameter degeneracies also come into play, especially for the lighter events GW191105\_143521 and GW170104.
+Other parameter degeneracies also come into play, especially for the lighter events GW191105\_143521 (henceforth GW191105) and GW170104.
 We discuss this in detail in a dedicated section below (Sec.~\ref{sec:notable}).
@@
 
-Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on three events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, and GW190521.
-When considered in isolation, the first of these is the event with the most significant support for nonzero $\kappa$; the second shows the sign of potential noise anomalies; and the third is representative of a class of events with strongly bimodal $\kappa$ posteriors.
+Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, GW190521 and GW191105\_143521.
+
+When considered in isolation, the first of these is the event with the most significant support for nonzero $\kappa$; the second shows the sign of potential noise anomalies; the third is representative of a class of heavy events with strongly bimodal $\kappa$ posteriors; and the fourth of a small class of lighter events with slightly bimodal posteriors.
 
@@
 Figure \ref{fig:corner_GW170818} shows that this happens because birefringence opens up a region of parameter space with $\kappa>0$ for larger distances and smaller inclination angles than would be allowed in the GR case.
-We can make sense of this because an edge-on, non-precessing source will produce linearly-polarized waves, so a smaller inclination will lead the two circular polarizations to have similar amplitudes.
+We can make sense of this because an edge-on, nonprecessing source will produce linearly-polarized waves, so a smaller inclination will lead the two circular polarizations to have similar amplitudes.
 On the other hand, having $\kappa > 0$ will enhance the left-handed modes during propagation.
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
     \script{corner_GW170818_appendix.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW170818_appendix.pdf}
+    \includegraphics[width=\textwidth]{figures/corner_GW170818_appendix.pdf}
     \caption{
@@
     \label{fig:corner_GW170818_appendix}
+\end{figure*}
+
+\begin{figure}
+    \script{hist_GW170818.py}
+    \includegraphics[width=\columnwidth]{figures/hist_GW170818.pdf}
+    \caption{
+        Corner plot of GW170818.
+    }
+    \label{fig:hist_GW170818}
 \end{figure}

### 2023-03-03 src/tex/ms.tex
table rearrangement
Correct the total mass calculation

@@
 
+\begin{table}
+    \caption{Events with bimodality in the $\kappa$ posterior tend to have high total masses. CL is the credible level of $\kappa = 0$.}
+    \begin{ruledtabular}
+        \variable{output/bimodal_events_mass.txt}
+    \end{ruledtabular}
+    \label{tab:bimodal_events_mass}
+\end{table}
+
 Finally, a set of events stands out in Fig.~\ref{fig:violin_kappa} due to evident bimodality in the $\kappa$ posterior.
 To a varying degree, that is the case for those events listed in Table~\ref{tab:bimodal_events_mass}, which tend to have quite high total masses (Table~\ref{tab:bimodal_events_mass} show total mass as measured in the standard  \ac{GR} analysis).
-\mi{THOMAS: is this correct? do these numbers come from the GR runs? they should}
 For these bimodal posteriors, $\mu_i/\sigma_i$ is not a good proxy for agreement with \ac{GR}; instead, we can rely on $\mathrm{CL}(\kappa = 0)$.
@@
 We discuss this in detail in a dedicated section below (Sec.~\ref{sec:notable}).
-
-\begin{table}
-    \caption{Events that show bimodality in the $\kappa$ posterior tend to have high total masses. CL is the credible level of $\kappa = 0$.}
-    \begin{ruledtabular}
-        \variable{output/bimodal_events_mass.txt}
-    \end{ruledtabular}
-    \label{tab:bimodal_events_mass}
-\end{table}
 

### 2023-03-03 src/scripts/bimodal_events_mass.py

@@ -1,6 +1,7 @@
 import pandas as pd
 import numpy as np
 import scipy
+from bilby.gw.result import CBCResult
 import paths
 
 def sign(x):
@@ -16,7 +17,9 @@ with open(paths.output/"bimodal_events_mass.txt", "w") as f:
     for event in ["GW170104", "GW190413_134308", "GW190521", "GW190805_211137", "GW191105_143521"]:
         name = event.replace("_", r"\_")
 
-        mass = result_DataFrame.loc[result_DataFrame["event"] == event, "mass_1"] + result_DataFrame.loc[result_DataFrame["event"] == event, "mass_2"]
+        file_name = str(event)+"_GR.json.gz"
+        result_DataFrame_GR = CBCResult.from_json(filename=paths.data/file_name).posterior
+        mass = result_DataFrame_GR["mass_1"] + result_DataFrame_GR["mass_2"]
         mass_median = mass.median()
         mass_95 = mass.quantile(0.95) - mass_median
         mass_5 = mass_median - mass.quantile(0.05)

### 2023-03-03 src/tex/ms.tex
Formatting

@@
 In the most simplified scenarios, birefringence is a property of spacetime that is not intrinsic to any source or region in space. 
-If so, we should expect $\kappa$ to
-take the same value for all signals, whether it vanishes or not.
-Under this assumption, the
-posterior on $\kappa$ inferred collectively from all events 
-is simply obtained from the product of the individual likelihoods, $p(d_i \mid
-\kappa) \propto p(\kappa \mid d_i)/p(\kappa)$, such that
+If so, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
+Under this assumption, the posterior on $\kappa$ inferred collectively from all events is simply obtained from the product of the individual likelihoods, $p(d_i \mid\kappa) \propto p(\kappa \mid d_i)/p(\kappa)$, such that
 \begin{equation}
@@
 % Bayesian Hierarchical Modeling
-Under many frameworks, birefringence is brokered by an extra parity-odd
-field that couples to gravity. In this case, whether the birefringence
-is the same along the different lines of sight to the different events depends
-on cosmic history and the local evolution of the field. The \emph{simplest}
-assumption is that the field manifests equally for all events, as assumed in the previous subsection, but this is by
-no means required.
+Under many frameworks, birefringence is brokered by an extra parity-odd field that couples to gravity.
+In this case, whether the birefringence is the same along the different lines of sight to the different events depends on cosmic history and the local evolution of the field.
+The \emph{simplest} assumption is that the field manifests equally for all events, as assumed in the previous subsection, but this is by no means required.
 This motivates a collective analysis that does not assume $\kappa$ is shared across events \cite{Zimmerman:2019wzo,Isi:2022cii}.
@@
 % Violin plot
-Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis
-as represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate).
-Posteriors are colored by the distance to the origin of the respective mean in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.%
-\footnote{We use the $i$ notation for individual events here to disambiguate from the population mean $\mu$ and standard deviation $\sigma$ of Eq.~\eqref{eq:posterior_of_mu_sigma}.}
-The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line);
-this joint measurement is fully consistent with $\kappa = 0$ (dashed line), meaning we find no evidence for birefringence.
+Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis as represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate).
+Posteriors are colored by the distance to the origin of the respective mean in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.\footnote{We use the $i$ notation for individual events here to disambiguate from the population mean $\mu$ and standard deviation $\sigma$ of Eq.~\eqref{eq:posterior_of_mu_sigma}.}
+The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line); this joint measurement is fully consistent with $\kappa = 0$ (dashed line), meaning we find no evidence for birefringence.
 
@@
 
-Figure \ref{fig:violin_kappa} shows certain degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
+Figure \ref{fig:violin_kappa} shows a certain degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
 This is not unexpected: assuming independent Gaussian noise instantiations for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$.
@@
     \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
-    \caption{Individual-event $\kappa_i$ distributions of Fig.~\ref{fig:violin_kappa} reweighted in light of the population-level inference of Fig.~\ref{fig:corner_Gaussian}, marginalizing over $\mu$ and $\sigma$.
-Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i$; we label the events most in tension with $\kappa_i = 0$ by that same measure.}
+    \caption{
+        Individual-event $\kappa_i$ distributions of Fig.~\ref{fig:violin_kappa} reweighted in light of the population-level inference of Fig.~\ref{fig:corner_Gaussian}, marginalizing over $\mu$ and $\sigma$.
+        Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i|$; we label the events most in tension with $\kappa_i = 0$ by that same measure.
+    }
     \label{fig:reweighted_kappa}

### 2023-03-03 src/scripts/hist_GW170818.py

@@ -37,6 +37,7 @@ g = sns.pairplot(result,
             diag_kws=dict(common_norm=False),
             plot_kws=dict(common_norm=False, levels=[(1.-0.90),(1.-0.3935)]))
 
-g.fig.legends[0].set_bbox_to_anchor((0.89,0.5))
+g.fig.legends[0].set_bbox_to_anchor((0.55,0.8))
+g.legend.set_title(None)
 
 g.savefig(fname=paths.figures/"hist_GW170818.pdf", bbox_inches="tight", dpi=300)

### 2023-03-03 src/tex/ms.tex
Add acronym

@@
 \acro{GWOSC}{the Gravitational Wave Open Science Center}
+\acro{SNR}{signal-to-noise ratio}
 \end{acronym}
@@
 
-We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby},
-that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline
-waveform. We take the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3,
-GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with
-\textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform. We
-apply a distance prior which is uniform in comoving volume and source frame time, and set the prior on $\kappa$ to a uniform
-distribution between $-1$ and $1$.
+We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
+We take the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
+We apply a distance prior which is uniform in comoving volume and source frame time, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
 
@@
 % More observations with higher SNR
-Another future work is to apply the same method to more GW events and events with a higher signal-to-noise ratio (SNR).
-LVK will release more events with higher SNR in the future.
-Using data with higher SNR allows us to obtain more precise \ac{PE} results and constrain the birefringence effect more precisely.
+Another future work is to apply the same method to more \ac{GW} events and events with a higher \ac{SNR}.
+\ac{LVK} will release more events with higher \ac{SNR} in the future.
+Using data with higher \ac{SNR} allows us to obtain more precise \ac{PE} results and constrain the birefringence effect more precisely.
 And using data from more events will allow us to calculate a more constrained population posterior of $\kappa$.

### 2023-03-03 src/tex/ms.tex
GW170818 exposition (WIP)

@@
 This assumption results in a full degeneracy between $\kappa$ and $\iota$:
-to reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
+to reconstruct the amplitude ratio from the interferometer data, a value of $\iota$ representing a more face-off inspiral can pair with a positive value of $\kappa$, or a value of $\iota$ representing a more face-on inspiral with a negative value of $\kappa$.
 That fact can be used to constrain frequency-independent birefringence by searching for features in the distribution of inferred inclinations \cite{Okounkova_2022}.
@@
 Besides source inclination, the effect of birefringence can be (partially) degenerate with other parameters---most notably, precession, as we find in Sec.~\ref{sec:Results}.
-Indeed, under the right conditions, the frequency-dependent amplification/dampening of polarizations caused by $\kappa$ can be effectively absorbed by changes in intrinsic parameters, like the spins;
+Indeed, under the right conditions, the frequency-dependent amplification or dampening of polarizations caused by $\kappa$ can be effectively absorbed by changes in intrinsic parameters, like the spins;
 in that case, the degeneracy with inclination may appear even for lighter \acp{BBH} (although usually to a lesser degree than heavy binaries).
@@
 Figure \ref{fig:violin_kappa} shows a certain degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
-This is not unexpected: assuming independent Gaussian noise instantiations for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$.
+This is not unexpected: assuming independent Gaussian noise instantiations for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$ due to noise alone.
 
@@
 
-Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, GW190521 and GW191105\_143521.
-
-When considered in isolation, the first of these is the event with the most significant support for nonzero $\kappa$; the second shows the sign of potential noise anomalies; the third is representative of a class of heavy events with strongly bimodal $\kappa$ posteriors; and the fourth of a small class of lighter events with slightly bimodal posteriors.
+Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, GW190521 and GW191105.
+%
+When considered in isolation, the first of these is the unimodel event with the most significant support for nonzero $\kappa$; the second shows signs of potential noise anomalies; the third is representative of a class of heavy events with strongly bimodal $\kappa$ posteriors; and the fourth of a smaller class of lighter events with slightly bimodal posteriors.
+Through these examples, we elucidate the interaction between $\kappa$ and the source luminosity distance, inclination and spin precession.
 
@@
 
-GW170818 produced the posterior in most tension with $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
+GW170818 produced the posterior most displaced from $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
 Figure \ref{fig:corner_GW170818} shows that this happens because birefringence opens up a region of parameter space with $\kappa>0$ for larger distances and smaller inclination angles than would be allowed in the GR case.
-We can make sense of this because an edge-on, nonprecessing source will produce linearly-polarized waves, so a smaller inclination will lead the two circular polarizations to have similar amplitudes.
-On the other hand, having $\kappa > 0$ will enhance the left-handed modes during propagation.
-The two effects can be balanced to yield the correct polarization ratio at the detector, as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
+We can make sense of this because an edge-on, nonprecessing source will produce linearly-polarized waves, meaning that a smaller inclination will lead the two circular polarizations to have similar amplitudes.
+On the other hand, having $\kappa > 0$ will enhance the left-handed modes during propagation, per Eq.~\eqref{eq:waveform_modification}.
+The two effects can be balanced to yield the correct polarization ratio observed at the detector (clearly predominantly left-handed, per the preference for $\cos\iota \approx -1$), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
 This is similar to the degeneracy mentioned in Sec.~\ref{sec:inclination}.
@@
 
-It is not clear what specific feature of the GW170818 data leads to this posterior structure.
-However, the reason may be related to the fact that this event provides an uncommonly definite measurement of the polarization angle and spin orientations \cite{Varma:2021csh}.
-Indeed, the inferred polarization angle and spins in the birefringent case are notably different from the \ac{GR} case (Fig.~\ref{fig:corner_GW170818_appendix}).
+It is difficult to unequivocally identify a specific feature of the GW170818 data that leads to this posterior structure.
+However, it appears to be related to this event's support for precession, in conjuction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
+The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:hist_GW170818}): allowing for $\kappa \neq 0$, leads us to infer a reduced support for precession.
+We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the time-dependent signal enhancement or dampening due to birefringence can mimic the amplitude modulation in a precession cycle.
+Therefore, under these circumstances, similar morphologies can be obtained by setting $\chi_p > 0$ or $\kappa > 0$, as long as the distance and inclination can also be adjusted accordingly.
+
+The fact that the birefringent analysis favors a high $\kappa$ for rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
+The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and of specific allowed spin angles.
+These control the allowed phasing of the precession cycle, and therefore the specific morphology that $\kappa$ may attempt to match.
+\mi{WIP text}
+The inferred polarization angle in the birefringent case is notably different from the \ac{GR} case (Fig.~\ref{fig:corner_GW170818_appendix}).
+
 If our inference on $\kappa$ is indeed tied to those parameters, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
@@
     \label{fig:corner_GW170818}
+\end{figure}
+
+\begin{figure}
+    \script{hist_GW170818.py}
+    \includegraphics[width=\columnwidth]{figures/hist_GW170818.pdf}
+    \caption{
+        GW170818 posterior on the precessing spin $\chi_p$, as obtained by allowing birefringence (blue) and not (orange). \mi{TODO: make this an inset to Fig.~\ref{fig:corner_GW170818}.}
+    }
+    \label{fig:hist_GW170818}
 \end{figure}
@@
 
-\begin{figure}
-    \script{hist_GW170818.py}
-    \includegraphics[width=\columnwidth]{figures/hist_GW170818.pdf}
-    \caption{
-        Corner plot of GW170818.
-    }
-    \label{fig:hist_GW170818}
-\end{figure}
-
 \section{Relation between $\kappa$ and $M_{PV}$}

### 2023-03-03 src/tex/ms.tex
GW170818 text edits

@@
 Figure \ref{fig:corner_GW170818} shows that this happens because birefringence opens up a region of parameter space with $\kappa>0$ for larger distances and smaller inclination angles than would be allowed in the GR case.
-We can make sense of this because an edge-on, nonprecessing source will produce linearly-polarized waves, meaning that a smaller inclination will lead the two circular polarizations to have similar amplitudes.
-On the other hand, having $\kappa > 0$ will enhance the left-handed modes during propagation, per Eq.~\eqref{eq:waveform_modification}.
-The two effects can be balanced to yield the correct polarization ratio observed at the detector (clearly predominantly left-handed, per the preference for $\cos\iota \approx -1$), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
+We can make sense of this by noting that an edge-on, nonprecessing source produces linearly-polarized waves, meaning that a smaller inclination leads the two circular polarizations to have similar amplitudes.
+On the other hand, having $\kappa > 0$ enhances the left-handed modes during propagation, per Eq.~\eqref{eq:waveform_modification}.
+The two effects can be balanced to yield the correct polarization ratio observed at the detector (predominantly left-handed, per the preference for $\cos\iota \approx -1$), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
 This is similar to the degeneracy mentioned in Sec.~\ref{sec:inclination}.
@@
 It is difficult to unequivocally identify a specific feature of the GW170818 data that leads to this posterior structure.
-However, it appears to be related to this event's support for precession, in conjuction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
+However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
 The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:hist_GW170818}): allowing for $\kappa \neq 0$, leads us to infer a reduced support for precession.
@@
 The fact that the birefringent analysis favors a high $\kappa$ for rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
-The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and of specific allowed spin angles.
-These control the allowed phasing of the precession cycle, and therefore the specific morphology that $\kappa$ may attempt to match.
-\mi{WIP text}
-The inferred polarization angle in the birefringent case is notably different from the \ac{GR} case (Fig.~\ref{fig:corner_GW170818_appendix}).
+The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and of the phasing of the precession cycle.
+The latter manifests as a precise constraint on the spin orientation and phase angles in the \ac{GR} analysis \cite{Varma:2021csh}.
+The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
+Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR}.
+We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix XX.
 
-If our inference on $\kappa$ is indeed tied to those parameters, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
+If our inference on $\kappa$ is tied to precession and the spin orientations, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
 In any case, as we argued in Sec.~\ref{sec:results:hier}, there is no evidence at the moment that this event is inconsistent with the either collection of measurements or with $\kappa = 0$ (Fig.~\ref{fig:reweighted_kappa}).
@@
     \caption{
-        GW170818 posterior on the precessing spin $\chi_p$, as obtained by allowing birefringence (blue) and not (orange). \mi{TODO: make this an inset to Fig.~\ref{fig:corner_GW170818}.}
+        GW170818 posterior on the precessing spin $\chi_p$, as obtained by allowing birefringence (blue) and not (orange).
+\mi{TODO (for Max): maybe make this an inset to Fig.~\ref{fig:corner_GW170818}.}
     }

### 2023-03-03 src/tex/ms.tex
chip definition

@@
 in that case, the degeneracy with inclination may appear even for lighter \acp{BBH} (although usually to a lesser degree than heavy binaries).
-We chart some of these degeneracies as part of the results we present in Sec.~\ref{sec:Results}.
+Specifically, the interplay between birefringence and precession will manifest in the posterior of the precessing spin parameter $\chi_p$, defined as \cite{Schmidt:2014iyl}
+\begin{equation}
+\chi_p \equiv \max\left\{ ||\vec{\chi}_{1\perp}||,\, k ||\vec{\chi}_{2\perp}||\right\} ,
+\end{equation}
+where $\vec{\chi}_{i\perp}$ are the projections of the dimensionless spin vectors onto the orbital plane at a reference time, and $k \equiv q\left(4 q +3\right) / \left(4 + 3q\right)$ for a mass ratio $q \equiv m_2/m_1 \leq 1$.
+We chart this and other degeneracies as part of the results we present in Sec.~\ref{sec:Results}.
+
 % In particular, we find that, under certain conditions, the frequency-dependent birefringence of Eq.~\eqref{eq:waveform_modification} can mimic the time-dependent amplitude modulation expected from a precessing system.

### 2023-03-03 src/tex/ms.tex
minor GW170818 tweaks

@@
 However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
-The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:hist_GW170818}): allowing for $\kappa \neq 0$, leads us to infer a reduced support for precession.
-We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the time-dependent signal enhancement or dampening due to birefringence can mimic the amplitude modulation in a precession cycle.
+The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:hist_GW170818}): allowing for $\kappa \neq 0$ leads to reduced support for precession.
+We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the frequency-dependent signal enhancement or dampening due to birefringence can mimic the time-dependent amplitude modulation in a precession cycle.
 Therefore, under these circumstances, similar morphologies can be obtained by setting $\chi_p > 0$ or $\kappa > 0$, as long as the distance and inclination can also be adjusted accordingly.
 
-The fact that the birefringent analysis favors a high $\kappa$ for rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
-The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and of the phasing of the precession cycle.
+The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
+The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and the specific phasing of the precession cycle.
 The latter manifests as a precise constraint on the spin orientation and phase angles in the \ac{GR} analysis \cite{Varma:2021csh}.
 The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
-Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR}.
+Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR} one.
 We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix XX.

### 2023-03-03 src/scripts/posterior_kappa.py
resized a plot

@@ -44,6 +44,8 @@ ll = ll - np.max(ll)
 likelihood = np.exp(ll)
 likelihood = likelihood / np.trapz(likelihood, x=kappa)
 
+fs = (6.807804068078041, 4.207454302822033)
+fig, ax = plt.subplots(1, figsize=fs)
 plt.axvline(0, color='k', lw=2, ls='--')
 sns.kdeplot(np.array(samples_reweighted).reshape(-1), label="generic", color=sns.color_palette()[3])
 # plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5), color=sns.color_palette()[3], linestyle=':')

### 2023-03-03 src/tex/ms.tex
resized a plot
some 190521 text

@@
 \usepackage[nolist,nohyperlinks]{acronym}
+\usepackage{bookmark}
 
@@
 
-% Case: GW200129 (glitch in Virgo data)
-\subsubsection{GW200129}
-\label{sec:GW200129}
-
-GW200129 is the event with the second largest $|\mu_i/\sigma_i|$ in Fig.~\ref{fig:violin_kappa}.
-However, data for this event were affected by a non-Gaussian noise disturbance (glitch) in the Virgo instrument, which was subtracted from the publicly-available data used for parameter estimation \cite{Davis:2022ird}.
-Since previous work suggests the degree of glitch subtraction affects the inference for this event \citep{GW200129_glitch}, we consider whether the apparent preference for $\kappa < 0$ could also be tied to the instrumental artifact.
-
-To this end, we perform three additional \ac{PE} runs for GW200129, considering only two detectors at a time: LIGO Hanford and Virgo (HL), LIGO Livingston and Virgo (LV), and LIGO Hanford and LIGO Livingston (HL).
-If the preference for $\kappa < 0$ is tied to the glitch in Virgo, we expect it to disappear in the HL run, which excludes Virgo data.
-
-This is indeed the case, as we show in Fig.~\ref{fig:corner_GW200129}: all runs including Virgo lean towards $\kappa < 0$ (solid curves in color), whereas the LIGO-only run is fully consistent with $\kappa = 0$ (dashed black).
-While this is not conclusive proof that the glitch itself is driving the result, it does indicate that the Virgo data play a key role in the inference of $\kappa$.
-Since more work would be needed to understand the effect of the glitch, this motivates us to consider the effect of excluding this event from the collective analyses above.
-
-\begin{figure}
-    \script{corner_GW200129.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
-    \caption{
-        GW200129 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos{\iota}$, including different sets of detectors in the analysis per the legend.
-        The main run with all three detectors (HLV, filled blue), shows a preference for $\kappa < 0$, as in Fig.~\ref{fig:violin_kappa}; this preference is more pronounced for two-detector runs that include Virgo (HV and LV, orange and green); however, it disappears if we remove Virgo (HL, dashed black).
-        The 2D contours correspond to the $90\%$ credible level.
-    }
-    \label{fig:corner_GW200129}
-\end{figure}
-
 % Case: GW190521 (massive BBH)
@@
 
-Even with the frequency-dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
-Consider GW190521, the most massive binary black hole merger in the events we included.
+GW190521 is the most extreme representative of a class of events with bimodal $\kappa$ posteriors (Table \ref{tab:bimodal_events_mass}).
+As we anticipated in Sec.~\ref{sec:inclination}, this is driven by not having sufficient frequencies in band to disambiguate between $\kappa$ and $\cos\iota$.
+As the most massive \ac{BBH} in our set, GW190521 is also the one with the fewest cycles in band, and therefore the most susceptible to this degeneracy.
+
 The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in Fig.~\ref{fig:corner_GW190521}.
-The reason could be this GW event's frequency range is too narrow to break the degeneracy.
 The effect of birefringence at different frequencies within the range is similar, so the degeneracy cannot be broken.
@@
     \label{fig:corner_GW150914}
+\end{figure}
+
+% Case: GW200129 (glitch in Virgo data)
+\subsubsection{GW200129}
+\label{sec:GW200129}
+
+GW200129 is the event with the second largest $|\mu_i/\sigma_i|$ in Fig.~\ref{fig:violin_kappa}.
+However, data for this event were affected by a non-Gaussian noise disturbance (glitch) in the Virgo instrument, which was subtracted from the publicly-available data used for parameter estimation \cite{Davis:2022ird}.
+Since previous work suggests the degree of glitch subtraction affects the inference for this event \citep{GW200129_glitch}, we consider whether the apparent preference for $\kappa < 0$ could also be tied to the instrumental artifact.
+
+To this end, we perform three additional \ac{PE} runs for GW200129, considering only two detectors at a time: LIGO Hanford and Virgo (HL), LIGO Livingston and Virgo (LV), and LIGO Hanford and LIGO Livingston (HL).
+If the preference for $\kappa < 0$ is tied to the glitch in Virgo, we expect it to disappear in the HL run, which excludes Virgo data.
+
+This is indeed the case, as we show in Fig.~\ref{fig:corner_GW200129}: all runs including Virgo lean towards $\kappa < 0$ (solid curves in color), whereas the LIGO-only run is fully consistent with $\kappa = 0$ (dashed black).
+While this is not conclusive proof that the glitch itself is driving the result, it does indicate that the Virgo data play a key role in the inference of $\kappa$.
+Since more work would be needed to understand the effect of the glitch, this motivates us to consider the effect of excluding this event from the collective analyses above.
+
+\begin{figure}
+    \script{corner_GW200129.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
+    \caption{
+        GW200129 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos{\iota}$, including different sets of detectors in the analysis per the legend.
+        The main run with all three detectors (HLV, filled blue), shows a preference for $\kappa < 0$, as in Fig.~\ref{fig:violin_kappa}; this preference is more pronounced for two-detector runs that include Virgo (HV and LV, orange and green); however, it disappears if we remove Virgo (HL, dashed black).
+        The 2D contours correspond to the $90\%$ credible level.
+    }
+    \label{fig:corner_GW200129}
 \end{figure}

### 2023-03-06 src/tex/ms.tex
Add GW190521 freq independent plot

@@
 Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR} one.
-We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix XX.
+We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix~\ref{sec:corner_GW170818_appendix}.
 
@@
     \caption{
-        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue) and frequency-independent (orange) birefringence analysis, compared to the GR result (green).
         % The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
@@
 
-\section{GW170818}
+\section{Full corner plot of GW170818}
+\label{sec:corner_GW170818_appendix}
 

### 2023-03-06 src/scripts/corner_GW190521.py

@@ -25,10 +25,15 @@ result_GR['cos_iota'] = np.cos([float(value) for value in result_GR['iota']])
 result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
 result_bilby = result_bilby[result_bilby.event == "GW190521"]
 result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), "BR")
+result_bilby['with'] = np.full(len(result_bilby), "BR (frequency dependent)")
 result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
 
-result = pd.concat([result_bilby,result_GR], ignore_index=True)
+result_extra = CBCResult.from_json(filename=paths.data/"GW190521_birefringence_freq_independent.json.gz").posterior
+result_extra = result_extra.sample(n=nsamples)
+result_extra['with'] = np.full(len(result_extra), "BR (frequency independent)")
+result_extra['cos_iota'] = np.cos([float(value) for value in result_extra['iota']])
+
+result = pd.concat([result_bilby, result_extra, result_GR], ignore_index=True)
 
 def kdeplot2d(x, y, **kws):
     kws.pop('label', None)

### 2023-03-06 src/scripts/corner_GW200129.py

@@ -90,11 +90,11 @@ g.map_diag(kdeplot1d, mask=m, linewidth=2, linestyle='--', fill=False, zorder=10
 # g.axes[2,1].set_ylim(-1., 1.)
 
 for i in range(len(vars)):
-    g.axes[i,i].set_xlim(result[vars[i]].min(), result[vars[i]].max())
+    g.axes[i,i].set_xlim(result_all[vars[i]].min(), result_all[vars[i]].max())
     # g.axes[i,i].set_ylim(0)
     for j in range(i):
-        g.axes[i,j].set_xlim(result[vars[j]].min(), result[vars[j]].max())
-        g.axes[i,j].set_ylim(result[vars[i]].min(), result[vars[i]].max())
+        g.axes[i,j].set_xlim(result_all[vars[j]].min(), result_all[vars[j]].max())
+        g.axes[i,j].set_ylim(result_all[vars[i]].min(), result_all[vars[i]].max())
 
 g.axes[2,0].set_xlabel(r"$\kappa$")
 g.axes[1,0].set_ylabel(r"$d_L$ (Mpc)")

### 2023-03-10 src/tex/ms.tex
precession wip

@@
 \usepackage{bookmark}
+
+\newcommand{\infd}{\mathrm{d}}
 
@@
 For instance, unequal propagation of \ac{GW} polarization eigenstates would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm,Alexander:2009tp,Sopuerta:2009iy}.
-Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC-3 \citep{GWTC-3}.
+Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC\nobreakdash-3 \citep{GWTC-3}.
 
@@
 Nevertheless, there exist systems for which the degeneracy cannot be perfectly broken because not enough frequencies are available in the data.
-As we will see in Sec.~\ref{sec:Results}, this is the case for heavy \acp{BBH}, which are only in the sensitive band of the detectors for a few cycles.
+This may be the case for quasimonochromatic sources, like nonaxisymmetric pulsars or very light binaries, which are well approximated by a single Fourier mode.
 
-Besides source inclination, the effect of birefringence can be (partially) degenerate with other parameters---most notably, precession, as we find in Sec.~\ref{sec:Results}.
-Indeed, under the right conditions, the frequency-dependent amplification or dampening of polarizations caused by $\kappa$ can be effectively absorbed by changes in intrinsic parameters, like the spins;
-in that case, the degeneracy with inclination may appear even for lighter \acp{BBH} (although usually to a lesser degree than heavy binaries).
-Specifically, the interplay between birefringence and precession will manifest in the posterior of the precessing spin parameter $\chi_p$, defined as \cite{Schmidt:2014iyl}
+As we will find in Sec.~\ref{sec:Results}, the effect of birefringence can be (partially) degenerate with other parameters besides source inclination.
+Indeed, the frequency-dependent amplification or dampening of polarizations caused by $\kappa$ can sometimes be absorbed by changes in intrinsic parameters, like the spins, with concurrent adjustments to the source inclination and distance.
+We will study the interplay between birefringence and precession by focusing on the posterior of the precessing spin parameter $\chi_p$, defined as \cite{Schmidt:2014iyl}
 \begin{equation}
@@
 where $\vec{\chi}_{i\perp}$ are the projections of the dimensionless spin vectors onto the orbital plane at a reference time, and $k \equiv q\left(4 q +3\right) / \left(4 + 3q\right)$ for a mass ratio $q \equiv m_2/m_1 \leq 1$.
-We chart this and other degeneracies as part of the results we present in Sec.~\ref{sec:Results}.
+We chart this and other degeneracies as part of the results presented in Sec.~\ref{sec:Results}.
 
@@
 We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
-We take the \ac{PE} configuration as in \citet{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
-We apply a distance prior which is uniform in comoving volume and source frame time, and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
+We take the \ac{PE} configuration in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
+We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time, and set the prior on $\kappa$ to be uniform between $-1$ and $1$.
 
-For GW190521, we increase the maximum distance of the distance prior to $1.5\times$ the original value, as the birefringence effect enables the probability of a larger distance.
-For GW190720, we decrease the analysis segment to 8 s from 16 s, in order to accommodate missing data near the edges of the 16 s segment in Virgo.
+For GW190521, we increase the maximum distance of the distance prior to $1.5\times$ the original value in \cite{GWTC-2.1_dataset}, as the birefringence effect results in posterior support at larger distances.
+For GW190720, we decrease the analysis segment from 16 s to 8 s, in order to accommodate missing data near the edges of the 16 s segment in Virgo.
 
@@
 Under many frameworks, birefringence is brokered by an extra parity-odd field that couples to gravity.
-In this case, whether the birefringence is the same along the different lines of sight to the different events depends on cosmic history and the local evolution of the field.
-The \emph{simplest} assumption is that the field manifests equally for all events, as assumed in the previous subsection, but this is by no means required.
-This motivates a collective analysis that does not assume $\kappa$ is shared across events \cite{Zimmerman:2019wzo,Isi:2022cii}.
+In that case, the effective strength of birefringence may vary along the different lines of sight to each event, depending on cosmic history and the local evolution of the field.
+The \emph{simplest} assumption is that the field manifests equally for all events, as assumed in the previous subsection, but this is not strictly required.
+These considerations motivate a collective analysis that does not assume $\kappa$ is shared across events \cite{Zimmerman:2019wzo,Isi:2022cii}.
 Such an analysis has the additional advantage of helping us further characterize our set of measurements, and identify potential outliers.
@@
 \begin{equation}
-    p(\mu,\sigma \mid \{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i\mid d_i)}{p(\kappa_i)}p(\kappa_i\mid\mu,\sigma)\,d\kappa_i\,,
+    p(\mu,\sigma \mid \{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i\mid d_i)}{p(\kappa_i)}p(\kappa_i\mid\mu,\sigma)\,\infd\kappa_i,
     \label{eq:posterior_of_mu_sigma}
 \end{equation}
-where $p(\kappa)$ is the prior initially applied to $\kappa$ during \ac{PE}, which in our case is a uniform distribution $\mathcal{U}[-1,1]$.
-Furhter choosing the hyperpriors on $\mu$ and $\sigma$ to be uniform, Eq.~\eqref{eq:posterior_of_mu_sigma} simplifies to
+where $p(\kappa)$ is the prior initially applied to $\kappa$ during \ac{PE}, which in our case is a uniform distribution, $\mathcal{U}[-1,1]$.
+Further choosing the hyperpriors on $\mu$ and $\sigma$ to be uniform, Eq.~\eqref{eq:posterior_of_mu_sigma} simplifies to
 \begin{equation}
-    p(\mu,\sigma\mid\{d\})\propto\prod_{i}\int p(\kappa_i\mid d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
+    p(\mu,\sigma\mid\{d\})\propto\prod_{i}\int p(\kappa_i\mid d_i)p(\kappa_i\mid\mu,\sigma)\,\infd\kappa_i\,
 \end{equation}
+where $p(\kappa_i\mid\mu,\sigma) \propto \exp(-|\kappa_i - \mu|^2/2\sigma^2)$ is the usual Gaussian likelihood.
 
-We also calculate the expected population distribution of $\kappa$ marginalized over $\mu$ and $\sigma$.
+Beyond measuring the population mean and variance, we also calculate the expected population distribution of $\kappa$ marginalized over $\mu$ and $\sigma$.
 This is given by
 \begin{equation}
-            p(\kappa\mid \{d\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma\mid \{d\})\,d\mu\,d\sigma\, , 
+            p(\kappa\mid \{d\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma\mid \{d\})\,\infd\mu\,\infd\sigma\, , 
     \label{eq:generic_posterior}
 \end{equation}
-and represents our overall expectation for $\kappa$, given the observed set of individual measurements.
+and represents our overall expectation for $\kappa$, given the observed set of individual measurements and, crucially, assuming the underlying distribution is Gaussian.%
+\footnote{The measurement of the population mean and variance is valid irrespective of whether the underlying $\kappa_i$ distribution is truly Gaussian; the specific shape of the population-marginalized distribution of Eq.~\eqref{eq:generic_posterior} is not.}
 To sample the posterior distribution of $\mu$ and $\sigma$, we use the sampling package \textsc{flowMC} \citep{flowMC}.
@@
 We first show the \ac{PE} results of all events and present the results of Bayesian hierarchical modeling.
-Then, we will discuss some special events individually.
+Then, we discuss some special events individually, and outline the degeneracies that arise between birefringence and orbital precession.
 
@@
 % Violin plot
-Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis as represented by the posterior distribution on $\kappa$ (abscissa) obtained for each event (ordinate).
+Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis as represented by the posterior distribution on $\kappa$ (abscissa) obtained individually for each event (ordinate).
 Posteriors are colored by the distance to the origin of the respective mean in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.\footnote{We use the $i$ notation for individual events here to disambiguate from the population mean $\mu$ and standard deviation $\sigma$ of Eq.~\eqref{eq:posterior_of_mu_sigma}.}
@@
 
-As we mentioned in Sec.~\ref{sec:inclination}, we understand the bimodality in $\kappa$ to be primarily a consequence of the narrow set of frequencies that are visible for these heavy events, which makes the effect of birefringence (partially) degenerate with a change in inclination.
+As we anticipated in Sec.~\ref{sec:inclination}, we understand the bimodality in $\kappa$ to be linked to precessing morphologies.
 Other parameter degeneracies also come into play, especially for the lighter events GW191105\_143521 (henceforth GW191105) and GW170104.
@@
 This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Callister:T2100301}.
-Figure \ref{fig:corner_Gaussian} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its preference for $\kappa > 0$.
+Figure \ref{fig:corner_Gaussian} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its higher support for $\kappa > 0$.
 This feature appears to offset a few other events which tend to favor $\kappa < 0$.
@@
 %
-When considered in isolation, the first of these is the unimodel event with the most significant support for nonzero $\kappa$; the second shows signs of potential noise anomalies; the third is representative of a class of heavy events with strongly bimodal $\kappa$ posteriors; and the fourth of a smaller class of lighter events with slightly bimodal posteriors.
+When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$; the second shows signs of potential noise anomalies; the third is representative of a class of heavy events with strongly bimodal $\kappa$ posteriors; and the fourth of a smaller class of lighter events with slightly bimodal posteriors.
 Through these examples, we elucidate the interaction between $\kappa$ and the source luminosity distance, inclination and spin precession.
@@
 On the other hand, having $\kappa > 0$ enhances the left-handed modes during propagation, per Eq.~\eqref{eq:waveform_modification}.
-The two effects can be balanced to yield the correct polarization ratio observed at the detector (predominantly left-handed, per the preference for $\cos\iota \approx -1$), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
+The two effects can be balanced to yield the correct polarization ratio observed at the detector (predominantly left-handed, per the preference for $\cos\iota \approx -1$ in the \ac{GR} analysis), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
 This is similar to the degeneracy mentioned in Sec.~\ref{sec:inclination}.
@@
 We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix~\ref{sec:corner_GW170818_appendix}.
-
-If our inference on $\kappa$ is tied to precession and the spin orientations, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
-In any case, as we argued in Sec.~\ref{sec:results:hier}, there is no evidence at the moment that this event is inconsistent with the either collection of measurements or with $\kappa = 0$ (Fig.~\ref{fig:reweighted_kappa}).
 
@@
 \subsection{Future work}
+
+\mi{WIP}
+If our inference on $\kappa$ is tied to precession and the spin orientations, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
+In any case, as we argued in Sec.~\ref{sec:results:hier}, there is no evidence at the moment that this event is inconsistent with the either collection of measurements or with $\kappa = 0$ (Fig.~\ref{fig:reweighted_kappa}).
+
 % BNS

### 2023-03-10 src/tex/ms.tex
190521 wip

@@
 GW190521 is the most extreme representative of a class of events with bimodal $\kappa$ posteriors (Table \ref{tab:bimodal_events_mass}).
-As we anticipated in Sec.~\ref{sec:inclination}, this is driven by not having sufficient frequencies in band to disambiguate between $\kappa$ and $\cos\iota$.
-As the most massive \ac{BBH} in our set, GW190521 is also the one with the fewest cycles in band, and therefore the most susceptible to this degeneracy.
+This appears to arise out of a confluence of factors related to this event's high mass and support for precession: this being the most massive event in our set, the signal is only in-band for a very short time before merger, meaning that spin effects can more easily be approximated by frequency-dependent birefringent.
+In particular, a precession cycle can more easily by approximated by a monotonic amplitude modulation.
+This effect gives rise to a similar degeneracy between $\kappa$ and $\chi_p$ as outlined for GW170818, wherein the prior volume favored configurations with $|\kappa| >0$.
+However, in the case of GW190521, the polarization measurement at the detectors is not accurate enough to fully disambiguate between right- and left-handed states (as implied by the bimodal $\cos\iota$ posterior obtained assuming \ac{GR} in Fig.~\ref{fig:corner_GW190521}).
+As a result, the $\kappa$ posterior is bimodal for GW190521.
 

### 2023-03-10 src/tex/ms.tex
190521 edits

@@
 However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
-The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:hist_GW170818}): allowing for $\kappa \neq 0$ leads to reduced support for precession.
+The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:corner_GW170818}): allowing for $\kappa \neq 0$ leads to reduced support for precession.
 We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the frequency-dependent signal enhancement or dampening due to birefringence can mimic the time-dependent amplitude modulation in a precession cycle.
@@
     \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
-    \caption{GW170818 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).}
+    \caption{
+        GW170818 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+        The marginalized posterior on $\chi_p$ is shown in the top right panel.
+    }
     \label{fig:corner_GW170818}
-\end{figure}
-
-\begin{figure}
-    \script{hist_GW170818.py}
-    \includegraphics[width=\columnwidth]{figures/hist_GW170818.pdf}
-    \caption{
-        GW170818 posterior on the precessing spin $\chi_p$, as obtained by allowing birefringence (blue) and not (orange).
-\mi{TODO (for Max): maybe make this an inset to Fig.~\ref{fig:corner_GW170818}.}
-    }
-    \label{fig:hist_GW170818}
 \end{figure}
@@
 GW190521 is the most extreme representative of a class of events with bimodal $\kappa$ posteriors (Table \ref{tab:bimodal_events_mass}).
-This appears to arise out of a confluence of factors related to this event's high mass and support for precession: this being the most massive event in our set, the signal is only in-band for a very short time before merger, meaning that spin effects can more easily be approximated by frequency-dependent birefringent.
-In particular, a precession cycle can more easily by approximated by a monotonic amplitude modulation.
+This appears to arise out of a confluence of factors related to this event's high mass and support for precession: GW190521 being the most massive event in our set, its signal is only in-band for a very short time before merger, meaning that frequency-dependent birefringence can more easily resemble spin effects.
+In particular, a precession cycle can be approximated by a monotonic amplitude modulation for a short time.
 This effect gives rise to a similar degeneracy between $\kappa$ and $\chi_p$ as outlined for GW170818, wherein the prior volume favored configurations with $|\kappa| >0$.
 However, in the case of GW190521, the polarization measurement at the detectors is not accurate enough to fully disambiguate between right- and left-handed states (as implied by the bimodal $\cos\iota$ posterior obtained assuming \ac{GR} in Fig.~\ref{fig:corner_GW190521}).
-As a result, the $\kappa$ posterior is bimodal for GW190521.
-
-The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in Fig.~\ref{fig:corner_GW190521}.
-The effect of birefringence at different frequencies within the range is similar, so the degeneracy cannot be broken.
-Therefore, the narrow frequency range can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
-Note that $\kappa=0$ is still within the $90\%$ confidence interval, with a non-negligible probability.
-The bimodal $\kappa$ distribution only means that the degeneracy between $\kappa$ and $\iota$ cannot be broken in this case.
-\wf{You should still explain \emph{why} the model seems to prefer $\left| \kappa \right| > 0$.
-I \emph{think} the reason here is the distance prior---there is a lot more ``phase space'' at large distances that can be accessed by allowing $\kappa$ to be non-zero and choosing a corresponding inclination so that the amplification effects of birefringence are compensated for by the larger distance compared to the pure-GR run.
-Does this make sense to you?
-If so, can you write it up in this paragraph?}
+As a result, the GW190521 posterior does not favor $\kappa > 0$ exclusively, but supports both $\kappa > 0$ and $\kappa < 0$, leading to a bimodal posterior.
 
@@
     \caption{
-        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue) and frequency-independent (orange) birefringence analysis, compared to the GR result (green).
-        % The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
-        % Colors in the plot represent the \ac{PE} result with \ac{GR} and the frequency-dependent birefringence, respectively.
-        % The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
-        % This plot shows that frequency-dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
+        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue), compared to the GR result (green).
+        The marginalized posterior on $\chi_p$ is shown in the top right panel.
     }
@@
     \caption{
-        GW150914 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue) and frequency-independent (orange) birefringence analysis, compared to the GR result (green).
-        % The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
-        % Colors in the plot represent the \ac{PE} result with \ac{GR}, frequency-independent and dependent birefringence, respectively.
-        % The 2D contours correspond to the $39.35\%$ ($1\sigma$) and $90\%$ credible levels.
-        % Note that there is no posterior of $\kappa$ for the \ac{PE} result with \ac{GR}, as the result is based on GR, which does not suggest GW amplitude birefringence.
-        % This plot shows that frequency-independent birefringence creates a degeneracy between $\kappa$ and $\iota$, while frequency-dependent birefringence can break it.
+        GW150914 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue), compared to the GR result (green).
+        The marginalized posterior on $\chi_p$ is shown in the top right panel.
     }
@@
         GW200129 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos{\iota}$, including different sets of detectors in the analysis per the legend.
+        The marginalized posterior on $\chi_p$ is shown in the top right panel.
         The main run with all three detectors (HLV, filled blue), shows a preference for $\kappa < 0$, as in Fig.~\ref{fig:violin_kappa}; this preference is more pronounced for two-detector runs that include Virgo (HV and LV, orange and green); however, it disappears if we remove Virgo (HL, dashed black).

### 2023-03-10 src/scripts/corner_GW150914.py
Add chi_p plots

@@ -22,13 +22,13 @@ result_GR['kappa'] = np.full(len(result_GR), None)
 result_GR['with'] = np.full(len(result_GR), "GR")
 result_GR['cos_iota'] = np.cos([float(value) for value in result_GR['iota']])
 
-result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_bilby = result_bilby[result_bilby.event == "GW150914"]
-result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), "BR (frequency dependent)")
-result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+result_BR = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_BR = result_BR[result_BR.event == "GW150914"]
+result_BR = result_BR.sample(n=nsamples)
+result_BR['with'] = np.full(len(result_BR), "BR")
+result_BR['cos_iota'] = np.cos(result_BR['iota'])
 
-result = pd.concat([result_bilby,result_GR], ignore_index=True)
+result = pd.concat([result_BR,result_GR], ignore_index=True)
 
 def kdeplot2d(x, y, **kws):
     kws.pop('label', None)
@@ -57,6 +57,7 @@ g.map_diag(kdeplot1d)
 
 for i in range(len(vars)):
     g.axes[i,i].set_xlim(result[vars[i]].min(), result[vars[i]].max())
+    g.axes[i,i].set_ylim(0)
     for j in range(i):
         g.axes[i,j].set_xlim(result[vars[j]].min(), result[vars[j]].max())
         g.axes[i,j].set_ylim(result[vars[i]].min(), result[vars[i]].max())
@@ -72,5 +73,26 @@ for k, c in zip(result['with'].unique(), sns.color_palette()):
 g.axes[0,0].legend(loc='center left', bbox_to_anchor=((1.1, 0.5)), frameon=False)
 
 plt.subplots_adjust(wspace=0.05, hspace=0.05)
+
+ax = g.fig.add_axes([g.axes[2,2].get_position().x0, g.axes[0,0].get_position().y0, g.axes[0,0].get_position().width, g.axes[0,0].get_position().height])
+
[…]

### 2023-03-10 src/scripts/corner_GW170818.py

@@ -22,13 +22,13 @@ result_GR['kappa'] = np.full(len(result_GR), None)
 result_GR['with'] = np.full(len(result_GR), "GR")
 result_GR['cos_iota'] = np.cos([float(value) for value in result_GR['iota']])
 
-result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_bilby = result_bilby[result_bilby.event == "GW170818"]
-result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), "BR")
-result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+result_BR = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_BR = result_BR[result_BR.event == "GW170818"]
+result_BR = result_BR.sample(n=nsamples)
+result_BR['with'] = np.full(len(result_BR), "BR")
+result_BR['cos_iota'] = np.cos(result_BR['iota'])
 
-result = pd.concat([result_bilby,result_GR], ignore_index=True)
+result = pd.concat([result_BR,result_GR], ignore_index=True)
 
 def kdeplot2d(x, y, **kws):
     kws.pop('label', None)
@@ -41,8 +41,8 @@ def kdeplot1d(x, **kws):
         kws.pop(key, None)
     df = pd.DataFrame({'x': x, 'y': Bounded_1d_kde(x, xlow=min(x), xhigh=max(x), **kws)(x)})
     df = df.sort_values(['x'])
-    plt.fill_between(df['x'], df['y'], np.zeros(len(x)), alpha=0.2)
     plt.plot(df['x'], df['y'])
+    plt.fill_between(df['x'], df['y'], np.zeros(len(x)), alpha=0.2)
 
 vars = ['kappa', 'luminosity_distance', 'cos_iota']
 g = sns.PairGrid(data=result,
@@ -55,23 +55,9 @@ g = sns.PairGrid(data=result,
 g.map_lower(kdeplot2d, levels=[0.90,0.3935])
 g.map_diag(kdeplot1d)
 
-# g.axes[0,0].set_xlim(-1., 1.)
-# g.axes[0,0].set_ylim(0)
-# g.axes[1,1].set_xlim(0, 3000)
[…]

### 2023-03-10 src/scripts/corner_GW190521.py

@@ -22,13 +22,13 @@ result_GR['kappa'] = np.full(len(result_GR), None)
 result_GR['with'] = np.full(len(result_GR), "GR")
 result_GR['cos_iota'] = np.cos([float(value) for value in result_GR['iota']])
 
-result_bilby = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_bilby = result_bilby[result_bilby.event == "GW190521"]
-result_bilby = result_bilby.sample(n=nsamples)
-result_bilby['with'] = np.full(len(result_bilby), "BR (frequency dependent)")
-result_bilby['cos_iota'] = np.cos(result_bilby['iota'])
+result_BR = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
+result_BR = result_BR[result_BR.event == "GW190521"]
+result_BR = result_BR.sample(n=nsamples)
+result_BR['with'] = np.full(len(result_BR), "BR")
+result_BR['cos_iota'] = np.cos(result_BR['iota'])
 
-result = pd.concat([result_bilby, result_GR], ignore_index=True)
+result = pd.concat([result_BR, result_GR], ignore_index=True)
 
 def kdeplot2d(x, y, **kws):
     kws.pop('label', None)
@@ -57,6 +57,7 @@ g.map_diag(kdeplot1d)
 
 for i in range(len(vars)):
     g.axes[i,i].set_xlim(result[vars[i]].min(), result[vars[i]].max())
+    g.axes[i,i].set_ylim(0)
     for j in range(i):
         g.axes[i,j].set_xlim(result[vars[j]].min(), result[vars[j]].max())
         g.axes[i,j].set_ylim(result[vars[i]].min(), result[vars[i]].max())
@@ -72,5 +73,26 @@ for k, c in zip(result['with'].unique(), sns.color_palette()):
 g.axes[0,0].legend(loc='center left', bbox_to_anchor=((1.1, 0.5)), frameon=False)
 
 plt.subplots_adjust(wspace=0.05, hspace=0.05)
+
+ax = g.fig.add_axes([g.axes[2,2].get_position().x0, g.axes[0,0].get_position().y0, g.axes[0,0].get_position().width, g.axes[0,0].get_position().height])
+
[…]

### 2023-03-10 src/tex/ms.tex
empty GW191105 section

--- 2023-03-10 src/tex/ms.tex
@@
 
-% Case: GW150914 (example of broken degeneracy)
-\subsubsection{GW150914}
-In this study, we included frequency dependence of birefringence, which would affect the posterior of $\kappa$ obtained from \ac{PE}.
-Consider GW150914, the first GW detected by LIGO, as an example.
-In Fig.~\ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
-With the frequency-independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
-This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
-% explain the kappa gap
+% % Case: GW150914 (example of broken degeneracy)
+% \subsubsection{GW150914}
+% In this study, we included frequency dependence of birefringence, which would affect the posterior of $\kappa$ obtained from \ac{PE}.
+% Consider GW150914, the first GW detected by LIGO, as an example.
+% In Fig.~\ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
+% With the frequency-independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
+% This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
+% % explain the kappa gap
+% 
+% On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
+% This is because the frequency dependence broke the degeneracy, as the effect of birefringence will differ from the effect of changing $\iota$.
+% In this case, the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent.
+% Therefore, the \ac{PE} result with frequency dependence is consistent with GR.
+% 
+% \begin{figure}
+%     \script{corner_GW150914.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
+%     \caption{
+%         GW150914 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue), compared to the GR result (green).
+%         The marginalized posterior on $\chi_p$ is shown in the top right panel.
+%     }
+%     \label{fig:corner_GW150914}
+% \end{figure}
+\subsubsection{GW191105}
+\label{sec:GW191105}
 
-On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
-This is because the frequency dependence broke the degeneracy, as the effect of birefringence will differ from the effect of changing $\iota$.
-In this case, the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent.
-Therefore, the \ac{PE} result with frequency dependence is consistent with GR.
-
-\begin{figure}
-    \script{corner_GW150914.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
-    \caption{
-        GW150914 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue), compared to the GR result (green).
-        The marginalized posterior on $\chi_p$ is shown in the top right panel.
-    }
-    \label{fig:corner_GW150914}
-\end{figure}
+The situation is similar for GW191105.
 

### 2023-03-10 src/scripts/corner_GW150914.py
linewidth setting

@@ -30,9 +30,11 @@ result_BR['cos_iota'] = np.cos(result_BR['iota'])
 
 result = pd.concat([result_BR,result_GR], ignore_index=True)
 
+lw = 1
+
 def kdeplot2d(x, y, **kws):
     kws.pop('label', None)
-    kdeplot_2d_clevels(xs=x, ys=y, auto_bound=True, **kws)
+    kdeplot_2d_clevels(xs=x, ys=y, auto_bound=True, linewidths=lw, **kws)
 
 def kdeplot1d(x, **kws):
     if np.all(x.isna()):
@@ -42,7 +44,7 @@ def kdeplot1d(x, **kws):
     df = pd.DataFrame({'x': x, 'y': Bounded_1d_kde(x, xlow=min(x), xhigh=max(x), **kws)(x)})
     df = df.sort_values(['x'])
     plt.fill_between(df['x'], df['y'], np.zeros(len(x)), alpha=0.2)
-    plt.plot(df['x'], df['y'])
+    plt.plot(df['x'], df['y'], lw=lw)
 
 vars = ['kappa', 'luminosity_distance', 'cos_iota']
 g = sns.PairGrid(data=result,
@@ -83,9 +85,9 @@ chi_p_GR = result_GR['chi_p']
 bounded_kde_BR = Bounded_1d_kde(chi_p_BR, xlow=0, xhigh=1)(chi_p_BR)
 bounded_kde_GR = Bounded_1d_kde(chi_p_GR, xlow=0, xhigh=1)(chi_p_GR)
 
-ax.plot(chi_p_BR, bounded_kde_BR, color=sns.color_palette()[0])
+ax.plot(chi_p_BR, bounded_kde_BR, color=sns.color_palette()[0], lw=lw)
 ax.fill_between(chi_p_BR, bounded_kde_BR, np.zeros(len(chi_p_BR)), alpha=0.2, color=sns.color_palette()[0])
-ax.plot(chi_p_GR, bounded_kde_GR, color=sns.color_palette()[1])
+ax.plot(chi_p_GR, bounded_kde_GR, color=sns.color_palette()[1], lw=lw)
 ax.fill_between(chi_p_GR, bounded_kde_GR, np.zeros(len(chi_p_GR)), alpha=0.2, color=sns.color_palette()[1])
 
 ax.set_xlabel(r"$\chi_p$")

### 2023-03-10 src/scripts/corner_GW190521.py

@@ -30,9 +30,11 @@ result_BR['cos_iota'] = np.cos(result_BR['iota'])
 
 result = pd.concat([result_BR, result_GR], ignore_index=True)
 
+lw = 1
+
 def kdeplot2d(x, y, **kws):
     kws.pop('label', None)
-    kdeplot_2d_clevels(xs=x, ys=y, auto_bound=True, **kws)
+    kdeplot_2d_clevels(xs=x, ys=y, auto_bound=True, linewidths=lw, **kws)
 
 def kdeplot1d(x, **kws):
     if np.all(x.isna()):
@@ -42,7 +44,7 @@ def kdeplot1d(x, **kws):
     df = pd.DataFrame({'x': x, 'y': Bounded_1d_kde(x, xlow=min(x), xhigh=max(x), **kws)(x)})
     df = df.sort_values(['x'])
     plt.fill_between(df['x'], df['y'], np.zeros(len(x)), alpha=0.2)
-    plt.plot(df['x'], df['y'])
+    plt.plot(df['x'], df['y'], lw=lw)
 
 vars = ['kappa', 'luminosity_distance', 'cos_iota']
 g = sns.PairGrid(data=result,
@@ -69,7 +71,7 @@ g.axes[2,0].set_ylabel(r"$\cos\iota$")
 g.axes[2,2].set_xlabel(r"$\cos\iota$")
 
 for k, c in zip(result['with'].unique(), sns.color_palette()):
-    g.axes[0,0].plot([], c=c, lw=2, label=k)
+    g.axes[0,0].plot([], c=c, lw=lw, label=k)
 g.axes[0,0].legend(loc='center left', bbox_to_anchor=((1.1, 0.5)), frameon=False)
 
 plt.subplots_adjust(wspace=0.05, hspace=0.05)
@@ -83,9 +85,9 @@ chi_p_GR = result_GR['chi_p']
 bounded_kde_BR = Bounded_1d_kde(chi_p_BR, xlow=0, xhigh=1)(chi_p_BR)
 bounded_kde_GR = Bounded_1d_kde(chi_p_GR, xlow=0, xhigh=1)(chi_p_GR)
 
-ax.plot(chi_p_BR, bounded_kde_BR, color=sns.color_palette()[0])
+ax.plot(chi_p_BR, bounded_kde_BR, color=sns.color_palette()[0], lw=lw)
[…]

### 2023-03-10 src/tex/ms.tex
corner plot updates

--- 2023-03-10 src/tex/ms.tex
@@
 However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
-The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:corner_GW170818}): allowing for $\kappa \neq 0$ leads to reduced support for precession.
+The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:corner_GW170818}, top right): allowing for $\kappa \neq 0$ leads to reduced support for precession.
 We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the frequency-dependent signal enhancement or dampening due to birefringence can mimic the time-dependent amplitude modulation in a precession cycle.
@@
 
-The situation is similar for GW191105.
+The situation is similar for GW191105 (Fig.~\ref{fig:corner_GW191105}).
+
+\begin{figure}[h]
+    \script{corner_GW191105.py}
+    \includegraphics[width=\columnwidth]{figures/hist_GW191105.pdf}
+    \caption{
+        Corner plot of GW191105.
+    }
+    \label{fig:corner_GW191105}
+\end{figure}
 
@@
 
-\section{placeholder for GW191105}
+% \section{placeholder for GW191105}
 
-\begin{figure}[h]
-    \script{corner_GW191105.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
-    \caption{
-        Corner plot of GW191105.
-    }
-    \label{fig:corner_GW191105}
-\end{figure}
+% \begin{figure}[h]
+%     \script{corner_GW191105.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
+%     \caption{
+%         Corner plot of GW191105.
+%     }
+%     \label{fig:corner_GW191105}
+% \end{figure}
 

### 2023-03-10 src/scripts/corner_GW190521.py

@@ -78,23 +78,23 @@ plt.subplots_adjust(wspace=0.05, hspace=0.05)
 
 ax = g.fig.add_axes([g.axes[2,2].get_position().x0, g.axes[0,0].get_position().y0, g.axes[0,0].get_position().width, g.axes[0,0].get_position().height])
 
-result_BR = result_BR.sort_values(['chi_p'])
-result_GR = result_GR.sort_values(['chi_p'])
-chi_p_BR = result_BR['chi_p']
-chi_p_GR = result_GR['chi_p']
-bounded_kde_BR = Bounded_1d_kde(chi_p_BR, xlow=0, xhigh=1)(chi_p_BR)
-bounded_kde_GR = Bounded_1d_kde(chi_p_GR, xlow=0, xhigh=1)(chi_p_GR)
+result_BR = result_BR.sort_values(['chi_eff'])
+result_GR = result_GR.sort_values(['chi_eff'])
+chi_eff_BR = result_BR['chi_eff']
+chi_eff_GR = result_GR['chi_eff']
+bounded_kde_BR = Bounded_1d_kde(chi_eff_BR, xlow=-1, xhigh=1)(chi_eff_BR)
+bounded_kde_GR = Bounded_1d_kde(chi_eff_GR, xlow=-1, xhigh=1)(chi_eff_GR)
 
-ax.plot(chi_p_BR, bounded_kde_BR, color=sns.color_palette()[0], lw=lw)
-ax.fill_between(chi_p_BR, bounded_kde_BR, np.zeros(len(chi_p_BR)), alpha=0.2, color=sns.color_palette()[0])
-ax.plot(chi_p_GR, bounded_kde_GR, color=sns.color_palette()[1], lw=lw)
-ax.fill_between(chi_p_GR, bounded_kde_GR, np.zeros(len(chi_p_GR)), alpha=0.2, color=sns.color_palette()[1])
+ax.plot(chi_eff_BR, bounded_kde_BR, color=sns.color_palette()[0])
+ax.fill_between(chi_eff_BR, bounded_kde_BR, np.zeros(len(chi_eff_BR)), alpha=0.2, color=sns.color_palette()[0])
+ax.plot(chi_eff_GR, bounded_kde_GR, color=sns.color_palette()[1])
[…]

### 2023-03-11 src/tex/ms.tex
Update figures

@@
 Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, GW190521 and GW191105.
-%
 When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$; the second shows signs of potential noise anomalies; the third is representative of a class of heavy events with strongly bimodal $\kappa$ posteriors; and the fourth of a smaller class of lighter events with slightly bimodal posteriors.
@@
 
-\begin{figure}[h]
+\begin{figure}
     \script{corner_GW190521.py}
@@
     \caption{
-        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue), compared to the GR result (green).
+        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
         The marginalized posterior on $\chi_p$ is shown in the top right panel.
@@
 % \end{figure}
+
+% Case: GW191105
 \subsubsection{GW191105}
@@
 
-\begin{figure}[h]
+\begin{figure}
     \script{corner_GW191105.py}
-    \includegraphics[width=\columnwidth]{figures/hist_GW191105.pdf}
+    \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
     \caption{
-        Corner plot of GW191105.
+        GW191105 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+        The posterior on $\chi_{eff}$ and $\kappa$ from our birefringence analysis is shown in the top right panel.
     }
@@
 
-% \section{placeholder for GW191105}
-
-% \begin{figure}[h]
-%     \script{corner_GW191105.py}
-%     \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
-%     \caption{
-%         Corner plot of GW191105.
-%     }
-%     \label{fig:corner_GW191105}
-% \end{figure}
-
 \section{Full corner plot of GW170818}

### 2023-03-11 src/scripts/corner_GW191105.py

@@ -14,7 +14,7 @@ plt.rcParams.update({
     "font.serif": ["Computer Modern Roman"],
 })
 
-nsamples = 4000
+nsamples = 5000
 
 result_GR = CBCResult.from_json(filename=paths.data/"GW191105_143521_GR.json.gz").posterior
 result_GR = result_GR.sample(n=nsamples)
@@ -86,4 +86,4 @@ ax.set_xlabel(r"$\kappa$")
 ax.set_ylim(-0.25, 0.25)
 ax.set_xlim(-0.3, 0.3)
 
-g.savefig(fname=paths.figures/"hist_GW191105.pdf", bbox_inches="tight", dpi=300)
+g.savefig(fname=paths.figures/"corner_GW191105.pdf", bbox_inches="tight", dpi=300)

### 2023-03-30 src/tex/ms.tex
text edits up to GW170817

@@
 This model is a better approximation of the birefringence effect expected from theory than the frequency-independent model used in \citet{Okounkova_2022}.
-Compared to other studies, we perform \ac{PE} on more events, including events new to the third \ac{GW} transient catalog \citep{GWTC-3}, GWTC-3.
+Compared to other studies, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
 We consider 71 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$, as listed in Table I of \citet{GWTC-3_population}.
+We discuss single-event results in detail, and identify degeneracies between birefringence and spin effects, in addition to the already known correlations with source orientation and distance.
 We use the results from individual events to place a collective population constraint on the strength of \ac{GW} amplitude birefringence from GWTC-3.
@@
 In \ac{GR}, \acp{GW} are comprised of two independent polarization modes, usually represented in the linear basis of plus ($+$) and cross ($\times$) states.
-In the Fourier domain, these can be combined into left-handed (L) and right-handed (R) circular states \cite{Isi:2022mbx},
+In the Fourier domain, these can be combined into left-handed (L) and right-handed (R) circular states (see, e.g., \cite{Isi:2022mbx}),
 \begin{equation}
@@
 % Waveform modification
-Yet, parity odd extensions of \ac{GR} may make distinctions between the two circular polarizations, which appear in both the generation and propagation of \acp{GW}.
-The latter can manifest in changes to the relative amplitude and phase of the polarizations that accrue as the wave propagates, giving us hope to detect initially small effects that compound over long propagation distances.
+Yet, parity odd extensions of \ac{GR} may make distinctions between the two circular polarizations, potentially appearing in both the generation and propagation of \acp{GW}.
+The latter can manifest in changes to the relative amplitude and phase of the polarizations that accrue as the wave propagates, giving us hope of detecting initially small effects that compound over long propagation distances.
 
 In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
-To first order in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
+To first order, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
 \begin{equation}
@@
 The overall magnitude of this effect for a given distance and frequency $f$ is set by a dimensionless opacity parameter, $\kappa$, which encodes the intrinsic strength of the birefringence.
-The emitted waveform for a given source (i.e., the waveform observed in the near zone, very close to the source) will generally differ from the analogous waveform predicted by \ac{GR}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small, it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
+The emitted waveform for a given source (i.e., the waveform observed in the near zone, very close to the source) will generally differ from the analogous waveform predicted by \ac{GR} \cite{Alexander:2009tp,Okounkova:2019zjf}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small (e.g., \cite{Okounkova:2022grv}), it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
 
 Although the intrinsic modification is small, the effect targeted by Eq.~\eqref{eq:waveform_modification} accumulates as the \ac{GW} propagates.
-During propagation, the effect of birefringence will be built up with the number of cycles which depends on the distance traveled and the frequency of the \acp{GW}.
+During propagation, the effect of birefringence will be built up with the number of cycles, which is itself a function of the distance traveled and the frequency of the \acp{GW}.
 According to Eq.~\eqref{eq:waveform_modification}, a positive $\kappa$ means the left-handed polarization is enhanced over the right-handed polarization, while a negative $\kappa$ means the opposite;
@@
 Equation \eqref{eq:waveform_modification} can be derived as the first order effect in an expansion away from \ac{GR} under multiple frameworks.
-In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity.
+In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity \cite{Alexander:2009tp}.
 Since it originates from a truncated series expansion, Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
@@
 
+\begin{figure}
+    \script{birefringence.py}
+    \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
+    \caption{
+        \emph{Illustration of amplitude birefringence.} The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
+        However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
+        In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
+        For this example, the black holes do not spin and have equal masses $m_1 = m_2 = 10\, M_\odot$, and we have chosen a luminosity distance $d_L = 400\, {\rm Mpc}$ and $\kappa = 0.6$.
+        }
+    \label{fig:birefringence}
+\end{figure}
+
 \subsection{Inclination and other degeneracies}
@@
 % Effect on inclination PE results
-Under certain conditions, the effect of birefringence can be degenerate with a change in the orientation of the source with respect to the line of sight.
-Concretely, for a nonprecessing compact binary inspiral in \ac{GR}, the observed amplitude ratio of the left-handed and right-handed polarizations is only a function of the inclination $\iota$, the angle between the orbital angular momentum of the source and the line of sight.
+Under certain conditions, the effect of birefringence can be degenerate with a change in the orientation of the source with respect to the line of sight \cite{Alexander:2009tp}.
+Concretely, for a nonprecessing compact binary inspiral in \ac{GR}, the observed amplitude ratio of the left-handed and right-handed polarizations is only a function of the inclination $\iota$, the angle between the orbital angular momentum of the source and the line of sight \cite{Blanchet:2013haa}.
 For  the dominant $\ell = |m| = 2$ angular mode of the radiation, the relation between the amplitude ratio and the inclination is
@@
 
-\begin{figure}
-    \script{birefringence.py}
-    \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
-    \caption{
-        \emph{Illustration of amplitude birefringence.} The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
-        However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
-        In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
-        For this example, the black holes do not spin and have equal masses $m_1 = m_2 = 10\, M_\odot$, and we have chosen a luminosity distance $d_L = 400\, {\rm Mpc}$ and $\kappa = 0.6$.
-        }
-    \label{fig:birefringence}
-\end{figure}
-
 \citet{Okounkova_2022} took the effect of birefringence to be independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
@@
 Indeed, the frequency-dependent amplification or dampening of polarizations caused by $\kappa$ can sometimes be absorbed by changes in intrinsic parameters, like the spins, with concurrent adjustments to the source inclination and distance.
+As a measure of the \ac{BH} spin magnitudes along the orbital angular momentum, we will use the effective spin parameter \cite{Damour:2001tu,Ajith:2009bn,Santamaria:2010yb}
+% \begin{equation}
+% \chi_{\rm eff} \equiv \frac{1}{M} \left(m_1 \left|\vec{\chi}_{1\parallel}\right| + m_2\left|\vec{\chi}_{2\parallel}\right|\right)
+% \end{equation}
+\begin{equation}
+\chi_{\rm eff} \equiv \frac{1}{1+q} \left(\left|\vec{\chi}_{1\parallel}\right| + q\left|\vec{\chi}_{2\parallel}\right|\right)
+\end{equation}
+where $|\vec{\chi}_{i\parallel}|$ are the norms of the projections of the dimensionless spin vectors along the orbital angular momentum, and the  mass ratio is $q \equiv m_2/m_1 \leq 1$.
 We will study the interplay between birefringence and precession by focusing on the posterior of the precessing spin parameter $\chi_p$, defined as \cite{Schmidt:2014iyl}
 \begin{equation}
-\chi_p \equiv \max\left\{ ||\vec{\chi}_{1\perp}||,\, k ||\vec{\chi}_{2\perp}||\right\} ,
+\chi_p \equiv \max\left\{ |\vec{\chi}_{1\perp}|,\, k |\vec{\chi}_{2\perp}|\right\} ,
 \end{equation}
-where $\vec{\chi}_{i\perp}$ are the projections of the dimensionless spin vectors onto the orbital plane at a reference time, and $k \equiv q\left(4 q +3\right) / \left(4 + 3q\right)$ for a mass ratio $q \equiv m_2/m_1 \leq 1$.
-We chart this and other degeneracies as part of the results presented in Sec.~\ref{sec:Results}.
+where $|\vec{\chi}_{i\perp}|$ are the norms of the projections of the dimensionless spin vectors onto the orbital plane at a reference time, and $k \equiv q\left(4 q +3\right) / \left(4 + 3q\right)$.
+We chart this and other approximate degeneracies as part of the results presented in Sec.~\ref{sec:Results}.
 
@@
 
-We carry out \ac{PE} using a custom version of \textsc{Bilby} \citep{Bilby}, that applies Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
+We estimate source parameters using a custom version of the \textsc{Bilby} software \citep{Bilby}, modified from the baseline version to apply Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
 We take the \ac{PE} configuration in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
@@
 
-For GW190521, we increase the maximum distance of the distance prior to $1.5\times$ the original value in \cite{GWTC-2.1_dataset}, as the birefringence effect results in posterior support at larger distances.
+For GW190521, we increase the maximum distance allowed by the prior to $1.5\times$ the original value in \cite{GWTC-2.1_dataset}, as the birefringence effect results in posterior support at larger distances.
 For GW190720, we decrease the analysis segment from 16 s to 8 s, in order to accommodate missing data near the edges of the 16 s segment in Virgo.
+Otherwise, the configuration is as in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset}.
 
@@
 In the most simplified scenarios, birefringence is a property of spacetime that is not intrinsic to any source or region in space. 
-If so, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
+Consequently, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
 Under this assumption, the posterior on $\kappa$ inferred collectively from all events is simply obtained from the product of the individual likelihoods, $p(d_i \mid\kappa) \propto p(\kappa \mid d_i)/p(\kappa)$, such that
@@
 To do this, we apply hierarchical Bayesian inference \cite{Loredo:2004nn} to model the distribution of $\kappa$'s consistent with the observed data:
-we posit that there is a specific value of the parameter, $\kappa_i$, associated with each event, drawn from some unknown distribution of underlying values; from the imperfect measurements of $\kappa_i$ for each event, we may reconstruct the underlying distribution.
+we posit that, rather than a unique global value of $\kappa$, there is a specific value of the parameter, $\kappa_i$, associated with each event, and that this is drawn from some unknown distribution of true underlying values; from the imperfect measurements of $\kappa_i$ for each event, we may reconstruct the underlying distribution.
 If we are interested in constraining the first two moments of the distribution, it is convenient to parametrize the $\kappa_i$'s as drawn from a Gaussian with unknown mean $\mu$ and variance $\sigma^2$, i.e., $\kappa_i \sim \mathcal{N}(\mu, \sigma^2)$ \cite{Isi:2019asy}, and measure those hyperparameters from the collection of observed data.
 
-If \ac{GR} is correct and there is no birefringence, we should find the observed $\kappa$ distribution to be consistent with a delta function at the origin ($\kappa_i = 0$ for all $i$, or $\mu=\sigma=0$); on the other hand, if spacetime is birefringent, we expect to find a delta function at some nonzero value ($\kappa_i = \kappa \neq 0$, or $\mu = \kappa$ and $\sigma=0$).
+If \ac{GR} is correct and there is no birefringence, we should find the observed $\kappa$ distribution to be consistent with a delta function at the origin ($\kappa_i = 0$ for all $i$, or $\mu=\sigma=0$); on the other hand, if spacetime is globally birefringent, we expect to find a delta function at some nonzero value ($\kappa_i = \kappa \neq 0$, or $\mu = \kappa$ and $\sigma=0$).
 But this analysis also has the power to reveal unexpected physics or systematics in our measurements: if $\sigma$ is confidently found to be nonzero, this would imply that our set of measurements is statistically unlikely to originate from a unique $\kappa$ value.
@@
 
-Starting from the posterior on $\kappa$ from each event, the posterior on the hyperparameters $\mu$ and $\sigma$ can be calculated by
+Starting from the posterior on $\kappa$ from each $i$\textsuperscript{th} event, $p(\kappa_i\mid d_i)$, the posterior on the hyperparameters $\mu$ and $\sigma$ can be calculated by
 \begin{equation}
@@
 \begin{equation}
-    p(\mu,\sigma\mid\{d\})\propto\prod_{i}\int p(\kappa_i\mid d_i)p(\kappa_i\mid\mu,\sigma)\,\infd\kappa_i\,
+    p(\mu,\sigma\mid\{d\})\propto\prod_{i}\int p(\kappa_i\mid d_i)\, p(\kappa_i\mid\mu,\sigma)\,\infd\kappa_i\,
 \end{equation}
@@
 Beyond measuring the population mean and variance, we also calculate the expected population distribution of $\kappa$ marginalized over $\mu$ and $\sigma$.
-This is given by
+This is defined by
 \begin{equation}
-            p(\kappa\mid \{d\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma\mid \{d\})\,\infd\mu\,\infd\sigma\, , 
+p(\kappa_i \mid \{d\})=\int p(\kappa_i \mid \mu,\sigma)p(\mu,\sigma\mid \{d\})\,\infd\mu\,\infd\sigma\, , 
     \label{eq:generic_posterior}
 \end{equation}
-and represents our overall expectation for $\kappa$, given the observed set of individual measurements and, crucially, assuming the underlying distribution is Gaussian.%
-\footnote{The measurement of the population mean and variance is valid irrespective of whether the underlying $\kappa_i$ distribution is truly Gaussian; the specific shape of the population-marginalized distribution of Eq.~\eqref{eq:generic_posterior} is not.}
+and represents our overall expectation for the true values of $\kappa$, given the observed set of individual measurements and, crucially, assuming the underlying distribution is Gaussian.
+Although the measurement of the population mean and variance applies irrespective of whether the underlying $\kappa_i$ distribution is truly Gaussian, the specific shape of the population-marginalized distribution of Eq.~\eqref{eq:generic_posterior} is not; therefore, Eq.~\eqref{eq:generic_posterior} should not be interpreted as giving a generic inference on the $\kappa$ distribution---more expressive models than a Gaussian would be better suited to that purpose.
 To sample the posterior distribution of $\mu$ and $\sigma$, we use the sampling package \textsc{flowMC} \citep{flowMC}.
@@
 In this section, we present the results of our study.
-We first show the \ac{PE} results of all events and present the results of Bayesian hierarchical modeling.
-Then, we discuss some special events individually, and outline the degeneracies that arise between birefringence and orbital precession.
+We first show the $\kappa$ measurements from of all events in our set, as well as the resulting global measurement of $\kappa$ that represents our primary constraint on birefringence (Sec.~\ref{sec:results:gwtc}).
+We then assess the collection of measurements in more detail through a hierarchical analysis (Sec.~\ref{sec:results:hier}).
+Finally, we discuss some special events individually, and outline the degeneracies that arise between birefringence and orbital precession (Sec.~\ref{sec:results:notable}).
 
@@
 \subsection{GWTC-3 result}
+\label{sec:results:gwtc}
 
@@
 Figure \ref{fig:violin_kappa} shows the main result of our GWTC-3 analysis as represented by the posterior distribution on $\kappa$ (abscissa) obtained individually for each event (ordinate).
-Posteriors are colored by the distance to the origin of the respective mean in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.\footnote{We use the $i$ notation for individual events here to disambiguate from the population mean $\mu$ and standard deviation $\sigma$ of Eq.~\eqref{eq:posterior_of_mu_sigma}.}
+Posteriors are colored by the respective posterior mean distance to the origin in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.\footnote{We use the $i$ notation for individual events here to disambiguate from the population mean $\mu$ and standard deviation $\sigma$ of Eq.~\eqref{eq:posterior_of_mu_sigma}.}
 The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line); this joint measurement is fully consistent with $\kappa = 0$ (dashed line), meaning we find no evidence for birefringence.
+\mi{THOMAS: can you quote the CL of 0 obtained from the joint measurement?}
 
@@
 Judging by $\mu_i/\sigma_i$, the two events that show the largest tension with $\kappa = 0$ are GW170818, for which $\mu_i / \sigma_i =$ \variable{output/GW170818_constraint.txt}, and GW200129\_065458 (henceforth GW200129), for which $\mu_i / \sigma_i =$ \variable{output/GW200129_constraint.txt}.
-However, as we discuss in Sec.~\ref{sec:GW200129}, we have reason to think that the preference for $\kappa < 0$ in GW200129 might be driven by noise anomalies in the Virgo detector; for that reason, in the next section we consider the effect of excluding this event from the joint result (we find its impact to be minimal).
+However, as we discuss in Sec.~\ref{sec:GW200129}, we have reason to think that the preference for $\kappa < 0$ in GW200129 might be driven by noise anomalies in the Virgo detector; with that in mind, in the next section we consider the effect of excluding this event from the joint result (we find its impact to be minimal).
 
@@
 \begin{table}
-    \caption{Events with bimodality in the $\kappa$ posterior tend to have high total masses. CL is the credible level of $\kappa = 0$.}
+    \caption{Events with bimodality in the $\kappa$ posterior, the \ac{GR} measurement of their detector-frame total mass ($M$), their $\chi_p$ measured without (GR) and with (BR) brefringence, and the credible level of $\kappa = 0$ (CL).}
     \begin{ruledtabular}
@@
 Finally, a set of events stands out in Fig.~\ref{fig:violin_kappa} due to evident bimodality in the $\kappa$ posterior.
-To a varying degree, that is the case for those events listed in Table~\ref{tab:bimodal_events_mass}, which tend to have quite high total masses (Table~\ref{tab:bimodal_events_mass} show total mass as measured in the standard  \ac{GR} analysis).
+To a varying degree, that is the case for those events listed in Table~\ref{tab:bimodal_events_mass}, which tend to have quite high total masses in the detector frame (Table~\ref{tab:bimodal_events_mass} shows total mass as measured in the standard \ac{GR} analysis).
 For these bimodal posteriors, $\mu_i/\sigma_i$ is not a good proxy for agreement with \ac{GR}; instead, we can rely on $\mathrm{CL}(\kappa = 0)$.
 By this measure, the bimodal events are some of the least consistent with $\kappa = 0$, GW190521 in particular.
-(However, note that configurations with $\kappa = 0$ for this event are still within the higher dimensional 90\%-credible region, when other parameters are considered and not just the one-dimensional marginal; we discuss this in Sec.~\ref{sec:GW190521}.)
+% (Nevertheless, configurations with $\kappa = 0$ for this event are still within the higher dimensional 90\%-credible region, when other parameters are considered and not just the one-dimensional marginal; we discuss this in Sec.~\ref{sec:GW190521}.)
 
-As we anticipated in Sec.~\ref{sec:inclination}, we understand the bimodality in $\kappa$ to be linked to precessing morphologies.
+As we anticipated in Sec.~\ref{sec:inclination}, we understand the bimodality in $\kappa$ to be linked to spin effects, and often to precessing morphologies in particular.
 Other parameter degeneracies also come into play, especially for the lighter events GW191105\_143521 (henceforth GW191105) and GW170104.
-We discuss this in detail in a dedicated section below (Sec.~\ref{sec:notable}).
+We discuss this further in a dedicated section below (Sec.~\ref{sec:results:notable}).
 
@@
 
-Figure \ref{fig:violin_kappa} shows a certain degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818.
+Figure \ref{fig:violin_kappa} shows a certain degree of variance in the distribution of $\kappa$ posteriors for different events, including some apparent outliers like GW170818 or GW190521.
 This is not unexpected: assuming independent Gaussian noise instantiations for each event, we might expect up to ${\sim}3$ out of the 71 posteriors (i.e., ${\sim}5\%$) to deviate away from the true $\kappa$ value by over ${\sim}2\sigma$ due to noise alone.
@@
 Both versions of the posterior support the lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals of the result including all events, we constrain $\mu =$ \variable{output/mu_median.txt} for 90\%-credible symmetric intervals around the median, and $\sigma <$ \variable{output/sigma_median.txt} for the 90\%-credible one-sided upper limit.
-However, even though $\sigma = 0$ is well supported, the $\sigma$ posterior peaks away from the origin, indicating some preference for a nonzero variance.
+However, even though $\sigma = 0$ is well supported, the $\sigma$ posterior peaks visibly away from the origin, indicating some preference for a nonzero variance.
 This could be a sign of the presence of outliers in our sample.
 
-As a check for outliers, we reconsider the set of measurements in Fig.~\ref{fig:violin_kappa} in light of the hierarchical result for $\mu$ and $\sigma$ in Fig.~\ref{fig:corner_Gaussian}, including GW200129 (blue curve).
+As a visual check for outliers, we reconsider the set of measurements in Fig.~\ref{fig:violin_kappa} in light of the hierarchical result for $\mu$ and $\sigma$ in Fig.~\ref{fig:corner_Gaussian}, including GW200129 (blue curve).
 This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Callister:T2100301}.
@@
 This feature appears to offset a few other events which tend to favor $\kappa < 0$.
-The interaction between these distributions leads to a hyperposterior that is fully consistent with $\mu = 0$ while indicating some support for $\sigma > 0$ (Fig.~\ref{fig:corner_Gaussian}).
+The interaction between these distributions leads to a hyperposterior that is fully consistent with $\mu = 0$ while offering some support for $\sigma > 0$ (Fig.~\ref{fig:corner_Gaussian}).
 Future observations will determine whether there is truly evidence for a nonvanishing variance in this population.
@@
 % Constraint on $\kappa$
-Finally, the hierarchical result in Fig.~\ref{fig:corner_Gaussian} can be translated into an overall expectation for $\kappa_i$ via Eq.~\eqref{eq:generic_posterior}.
+Finally, the hierarchical result in Fig.~\ref{fig:corner_Gaussian} can be translated into an overall expectation for $\kappa_i$ under the assumption of a Gaussian distribution via Eq.~\eqref{eq:generic_posterior}.
 We show the result of doing this in Fig.~\ref{fig:posterior_kappa}, where we also compare to the posterior on $\kappa$ obtained by assuming a shared value across events (same result shown as a blue band in Fig.~\ref{fig:violin_kappa}).
@@
 \subsection{Notable events}
-\label{sec:notable}
+\label{sec:results:notable}
 
@@
 \label{sec:GW170818}
+
+\begin{figure}
+    \script{corner_GW170818.py}
+    \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
+    \caption{
+        GW170818 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+        The top right panel shows the marginalized posterior on $\chi_p$: allowing for birefringence reduces the preference for precession.
+        (See Fig.~\ref{fig:corner_GW170818_appendix} for a full corner plot.)
+    }
+    \label{fig:corner_GW170818}
+\end{figure}
 
@@
 The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:corner_GW170818}, top right): allowing for $\kappa \neq 0$ leads to reduced support for precession.
-We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the frequency-dependent signal enhancement or dampening due to birefringence can mimic the time-dependent amplitude modulation in a precession cycle.
+We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the frequency-dependent signal enhancement or dampening due to birefringence can mimic the time-dependent amplitude modulation produced by a precession cycle.
 Therefore, under these circumstances, similar morphologies can be obtained by setting $\chi_p > 0$ or $\kappa > 0$, as long as the distance and inclination can also be adjusted accordingly.
 
-The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
+The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ is likely a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
 The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and the specific phasing of the precession cycle.
@@
 We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix~\ref{sec:corner_GW170818_appendix}.
-
-\begin{figure}
-    \script{corner_GW170818.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
-    \caption{
-        GW170818 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
-        The marginalized posterior on $\chi_p$ is shown in the top right panel.
-    }
-    \label{fig:corner_GW170818}
-\end{figure}
 
@@
 
-\section{Full corner plot of GW170818}
+\section{Extended corner plot for GW170818}
 \label{sec:corner_GW170818_appendix}
+
+Here we present the measurement on all parameters we consider relevant for GW170818, of which Fig.~\ref{fig:corner_GW170818} in the main text represents a subset.
+There are several differences between the birefringent analysis (BR; blue) and the baseline \ac{GR} result (orange).
+The fact that the former has a reduced support for precession relative to the latter, as shown in Fig.~\ref{fig:corner_GW170818}, can here be seen in the fact that the BR analysis favors weaker spin magnitudes, $\chi_{1/2}$.
 

### 2023-03-31 src/tex/ms.tex
add credible level

@@
 Posteriors are colored by the respective posterior mean distance to the origin in units of standard deviation, i.e., $|\mu_i/\sigma_i|$ for each event $i$.\footnote{We use the $i$ notation for individual events here to disambiguate from the population mean $\mu$ and standard deviation $\sigma$ of Eq.~\eqref{eq:posterior_of_mu_sigma}.}
-The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line); this joint measurement is fully consistent with $\kappa = 0$ (dashed line), meaning we find no evidence for birefringence.
-\mi{THOMAS: can you quote the CL of 0 obtained from the joint measurement?}
+The collective measurement obtained by assuming a shared $\kappa$ across events, Eq.~\eqref{eq:restricted_posterior}, is represented by its 90\%-credible symmetric interval (blue band) around the median (blue line); this joint measurement is fully consistent with $\kappa = 0$ (dashed line) with the credible level of \variable{output/CL_kappa_0.txt}, meaning we find no evidence for birefringence.
 
@@
 We show the result of doing this in Fig.~\ref{fig:posterior_kappa}, where we also compare to the posterior on $\kappa$ obtained by assuming a shared value across events (same result shown as a blue band in Fig.~\ref{fig:violin_kappa}).
-The hierarchical measurement leads to an expectation that $\kappa_i = $\variable{output/generic_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies $\kappa = $\variable{output/restricted_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
+The hierarchical measurement leads to an expectation that $\kappa_i =$ \variable{output/generic_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies $\kappa =$ \variable{output/restricted_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
 

### 2023-03-31 src/scripts/best_events_kappa.py

@@ -31,8 +31,7 @@ with open(paths.output/"best_events_kappa.txt", "w") as f:
         result_dict[event]['kappe_5'] = result_dict[event]['kappa_median'] - np.quantile(result_dict[event]['kappa'], 0.05)
 
         kernel = scipy.stats.gaussian_kde(result_dict[event]['kappa'])
-        kde = [kernel(kappa) for kappa in result_dict[event]['kappa']]
-        kde = np.array(kde)
+        kde = np.array([kernel(kappa) for kappa in result_dict[event]['kappa']])
         result_dict[event]['credible_level'] = len(kde[kde > kernel(0)[0]])/len(kde)
 
         f.write(rf"{result_dict[event]['event']} & ${sign(result_dict[event]['kappa_median'])}{result_dict[event]['kappa_median']:.3f}^{{+{result_dict[event]['kappe_95']:.3f}}}_{{-{result_dict[event]['kappe_5']:.3f}}}$ & ${result_dict[event]['kappa_std']:.3f}$ & ${result_dict[event]['credible_level']:.3f}$")

### 2023-03-31 src/scripts/kappa_constraint.py

@@ -35,8 +35,13 @@ likelihood = likelihood / np.trapz(likelihood, x=kappa)
 restricted_kappa_median = np.interp(0.5,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)
 generic_kappa_median = np.median(np.array(samples_reweighted).reshape(-1))
 
+credible_level = len(likelihood[likelihood > np.interp(0, kappa, likelihood)])/len(likelihood)
+
 with open(paths.output/"restricted_kappa_median.txt", "w") as f:
     f.write(f"${restricted_kappa_median:.3f}^{{+{(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)-restricted_kappa_median):.3f}}}_ {{{(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)-restricted_kappa_median):.3f}}}$")
 
 with open(paths.output/"generic_kappa_median.txt", "w") as f:
     f.write(f"${generic_kappa_median:.3f}^{{+{(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95)-generic_kappa_median):.3f}}}_{{{(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5)-generic_kappa_median):.3f}}}$")
+
+with open(paths.output/"CL_kappa_0.txt", "w") as f:
+    f.write(f"${credible_level:.3f}$")

### 2023-04-03 src/tex/ms.tex
appendix text WIP

@@
 The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and the specific phasing of the precession cycle.
-The latter manifests as a precise constraint on the spin orientation and phase angles in the \ac{GR} analysis \cite{Varma:2021csh}.
+The latter manifests as a precise constraint on the spin orientation and phase angles in the \ac{GR} analysis \cite{Varma:2021csh} (see also Appendix \ref{sec:corner_GW170818_appendix}).
 The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
@@
 Here we present the measurement on all parameters we consider relevant for GW170818, of which Fig.~\ref{fig:corner_GW170818} in the main text represents a subset.
+The regular \ac{GR} analysis (GR; orange) and the birefringent analysis (BR; blue) both show notable features.
+
+The GR analysis stands out for its relatively confident identification of the spin angles, $\theta_{1/2}$ and $\phi_{JL}$, as well as the phase and polarization angles, $\phi_{\rm ref}$ and $\psi$.
+Of the former, the two $\theta_{1/2}$ parameters encode the tilts of the component \acp{BH} with respect to the orbital angular momentum, $\vec{L}$, while $\phi_{JL}$ is the angle between the projections of $\vec{L}$ and the total angular momentum, $\vec{J}$, along the orbital plane;
+of the latter, $\phi_{\rm ref}$ is an overall reference phase, and $\psi$ encodes the orientation of the binary within the plane of the sky \cite{Isi:2022mbx}.
+All these parameters are anchored to a reference point in the evolution of the inspiral, which in this case corresponds to the time at which the dominant mode of the observed \ac{GW} signal reaches 20 Hz at the detector (spin angles may be better identified by using a more physical reference point \cite{Varma:2021csh}).
+It is unusual for these angles to be well constrained, which suggests that this event displays a particular phase and polarization signature.
+
+The joint posterior on the spin magnitudes, $\chi_1$ and $\chi_2$, indicates that at least one of the component \acp{BH} must have been highly spinning according to this \ac{GR} waveform, as evidenced by the lack of support for $\chi_1=\chi_2 = 0$ (this is not apparent from the marginals because the \acp{BH} were equal in mass \cite{Biscoveanu:2020are}).
+Furthermore, the fact that the tilts are favored to be close to $\theta_{1/2} \approx \pi/2$ implies that the spins must lie along the orbital plane.
+This is a restatement of the support for high $\chi_p$ in Fig.~\ref{fig:corner_GW170818}.
+
 There are several differences between the birefringent analysis (BR; blue) and the baseline \ac{GR} result (orange).
@@
     \caption{
-        Corner plot of GW170818.
+        Extended corner plot for GW170818: a supplement to Fig.~\ref{fig:corner_GW170818} discussed in Appendix~\ref{sec:corner_GW170818_appendix}.
     }

### 2023-04-03 src/tex/ms.tex
minor wording

@@
 Here we present the measurement on all parameters we consider relevant for GW170818, of which Fig.~\ref{fig:corner_GW170818} in the main text represents a subset.
-The regular \ac{GR} analysis (GR; orange) and the birefringent analysis (BR; blue) both show notable features.
+Figure \ref{fig:corner_GW170818_appendix} shows the the regular \ac{GR} analysis (GR; orange) and the birefringent analysis (BR; blue), both of which show notable features.
 
 The GR analysis stands out for its relatively confident identification of the spin angles, $\theta_{1/2}$ and $\phi_{JL}$, as well as the phase and polarization angles, $\phi_{\rm ref}$ and $\psi$.
-Of the former, the two $\theta_{1/2}$ parameters encode the tilts of the component \acp{BH} with respect to the orbital angular momentum, $\vec{L}$, while $\phi_{JL}$ is the angle between the projections of $\vec{L}$ and the total angular momentum, $\vec{J}$, along the orbital plane;
+Of the former, the two $\theta_{1/2}$ angles encode the tilts of the component \acp{BH} with respect to the orbital angular momentum, $\vec{L}$, while $\phi_{JL}$ is the angle between the projections of $\vec{L}$ and the total angular momentum, $\vec{J}$, along the orbital plane;
 of the latter, $\phi_{\rm ref}$ is an overall reference phase, and $\psi$ encodes the orientation of the binary within the plane of the sky \cite{Isi:2022mbx}.
 All these parameters are anchored to a reference point in the evolution of the inspiral, which in this case corresponds to the time at which the dominant mode of the observed \ac{GW} signal reaches 20 Hz at the detector (spin angles may be better identified by using a more physical reference point \cite{Varma:2021csh}).
-It is unusual for these angles to be well constrained, which suggests that this event displays a particular phase and polarization signature.
+It is unusual for these angles to be well constrained, which suggests that data for this event display a particular phase and polarization signature.
 

### 2023-04-04 src/tex/ms.tex
GW170818 appendix done

--- 2023-03-31 src/tex/ms.tex
@@
 
-Here we present the measurement on all parameters we consider relevant for GW170818, of which Fig.~\ref{fig:corner_GW170818} in the main text represents a subset.
-There are several differences between the birefringent analysis (BR; blue) and the baseline \ac{GR} result (orange).
-The fact that the former has a reduced support for precession relative to the latter, as shown in Fig.~\ref{fig:corner_GW170818}, can here be seen in the fact that the BR analysis favors weaker spin magnitudes, $\chi_{1/2}$.
+Here we provide further details on the GW170818 measurement. 
+Figure \ref{fig:corner_GW170818_appendix} displays the posterior on all parameters we consider relevant for this event, of which Fig.~\ref{fig:corner_GW170818} in the main text represents a subset.
+As in that figure, Fig.~\ref{fig:corner_GW170818_appendix} shows the regular \textsc{IMRPhenomXPHM} \ac{GR} analysis (GR; orange) and the birefringent analysis (BR; blue), both of which show notable features.
+
+The GR analysis stands out for its relatively confident identification of the spin angles, $\theta_{1/2}$, $\Delta \phi$ and $\phi_{JL}$, as well as the phase and polarization angles, $\phi_{\rm ref}$ and $\psi$.
+Of the former, the two $\theta_{1/2}$ angles encode the tilts of the component \acp{BH} with respect to the orbital angular momentum, $\vec{L}$, $\Delta \phi$ is the angle between the spin projections onto the orbital plane, while $\phi_{JL}$ is a similar angle separating the projections of $\vec{L}$ and the total angular momentum, $\vec{J}$;
+of the latter, $\phi_{\rm ref}$ is an overall reference phase, and $\psi$ encodes the orientation of the binary within the plane of the sky \cite{Isi:2022mbx}.
+All these parameters are anchored to a reference point in the inspiral, which in this case corresponds to the time at which the dominant multipole of the observed \ac{GW} signal reaches 20 Hz at the detector (spin angles may be better identified by using a more physical reference point \cite{Varma:2021csh}).
+It is unusual for all these angles to be well constrained, which suggests that data for this event display a particular phase and polarization signature.
+
+The joint posterior on the spin magnitudes, $\chi_1$ and $\chi_2$, disfavors the origin $\chi_1 = \chi_2 = 0$ with high credibility, indicating that at least one of the component \acp{BH} must have been highly spinning (this is not apparent from the marginals because the \acp{BH} were equal in mass \cite{Biscoveanu:2020are}).
+Furthermore, the fact that the tilts are favored to be close to $\theta_{1/2} \approx \pi/2$ implies that the spins must lie along the orbital plane;
+this is a restatement of the support for high $\chi_p$ in Fig.~\ref{fig:corner_GW170818}.
+
+On the other hand, the birefringent analysis does not strongly favor high spins; this is evidenced by the joint posterior on $\chi_1$ and $\chi_2$, which now favors $\chi_1 = \chi_2 = 0$ and assigns relatively low probability densities to higher values.
+Although the spin tilts are similarly constrained to $\theta_{1/2} \approx \pi/2$, the low spin magnitudes would imply weak if any precession, which is reflected in the support for lower values of $\chi_p$ in Fig.~\ref{fig:corner_GW170818} of the main text.
+As discussed in Sec.~\ref{sec:GW170818}, this can be explained by the introduction of $\kappa$, which can result in morphologies that are approximately degenerate with precession over a small number of waveform cycles (cf.~Fig.~\ref{fig:birefringence}).
+See the main text for further discussion.
 

### 2023-04-04 src/tex/ms.tex
Add appendix plots

@@
 
+\section{Extended corner plot for GW190521}
+\label{sec:corner_GW190521_appendix}
+
+\begin{figure*}[h]
+    \script{corner_GW190521_appendix.py}
+    \includegraphics[width=\textwidth]{figures/corner_GW190521_appendix.pdf}
+    \caption{
+        Extended corner plot for GW190521: a supplement to Fig.~\ref{fig:corner_GW190521} discussed in Appendix~\ref{sec:corner_GW190521_appendix}.
+    }
+    \label{fig:corner_GW190521_appendix}
+\end{figure*}
+
+\section{Extended corner plot for GW191105}
+\label{sec:corner_GW191105_appendix}
+
+\begin{figure*}[h]
+    \script{corner_GW191105_appendix.py}
+    \includegraphics[width=\textwidth]{figures/corner_GW191105_appendix.pdf}
+    \caption{
+        Extended corner plot for GW191105: a supplement to Fig.~\ref{fig:corner_GW191105} discussed in Appendix~\ref{sec:corner_GW191105_appendix}.
+    }
+    \label{fig:corner_GW191105_appendix}
+\end{figure*}
+
+\section{Extended corner plot for GW200129}
+\label{sec:corner_GW200129_appendix}
+
+\begin{figure*}[h]
+    \script{corner_GW200129_appendix.py}
+    \includegraphics[width=\textwidth]{figures/corner_GW200129_appendix.pdf}
+    \caption{
+        Extended corner plot for GW200129: a supplement to Fig.~\ref{fig:corner_GW200129} discussed in Appendix~\ref{sec:corner_GW200129_appendix}.
+    }
+    \label{fig:corner_GW200129_appendix}
+\end{figure*}
+
 \section{Relation between $\kappa$ and $M_{PV}$}

### 2023-04-04 src/scripts/corner_GW190521_appendix.py

@@ -16,14 +16,14 @@ plt.rcParams.update({
 
 nsamples = 5000
 
-result_GR = CBCResult.from_json(filename=paths.data/"GW170818_GR.json.gz").posterior
+result_GR = CBCResult.from_json(filename=paths.data/"GW190521_GR.json.gz").posterior
 result_GR = result_GR.sample(n=nsamples)
 result_GR['kappa'] = np.full(len(result_GR), None)
 result_GR['with'] = np.full(len(result_GR), "GR")
 result_GR['cos_iota'] = np.cos([float(value) for value in result_GR['iota']])
 
 result_BR = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_BR = result_BR[result_BR.event == "GW170818"]
+result_BR = result_BR[result_BR.event == "GW190521"]
 result_BR = result_BR.sample(n=nsamples)
 result_BR['with'] = np.full(len(result_BR), r"BR")
 result_BR['cos_iota'] = np.cos(result_BR['iota'])
@@ -94,4 +94,4 @@ g.axes[5,5].legend(loc='center left', bbox_to_anchor=((1.1,0.5)), frameon=False)
 
 plt.subplots_adjust(wspace=0.05, hspace=0.05)
 
-g.savefig(fname=paths.figures/"corner_GW170818_appendix.pdf", bbox_inches="tight", dpi=300)
+g.savefig(fname=paths.figures/"corner_GW190521_appendix.pdf", bbox_inches="tight", dpi=300)

### 2023-04-04 src/scripts/corner_GW200129_appendix.py

@@ -16,14 +16,14 @@ plt.rcParams.update({
 
 nsamples = 5000
 
-result_GR = CBCResult.from_json(filename=paths.data/"GW170818_GR.json.gz").posterior
+result_GR = CBCResult.from_json(filename=paths.data/"GW200129_065458_GR.json.gz").posterior
 result_GR = result_GR.sample(n=nsamples)
 result_GR['kappa'] = np.full(len(result_GR), None)
 result_GR['with'] = np.full(len(result_GR), "GR")
 result_GR['cos_iota'] = np.cos([float(value) for value in result_GR['iota']])
 
 result_BR = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
-result_BR = result_BR[result_BR.event == "GW170818"]
+result_BR = result_BR[result_BR.event == "GW200129_065458"]
 result_BR = result_BR.sample(n=nsamples)
 result_BR['with'] = np.full(len(result_BR), r"BR")
 result_BR['cos_iota'] = np.cos(result_BR['iota'])
@@ -53,6 +53,7 @@ g = sns.PairGrid(data=result,
                  corner=True, hue='with', 
                  diag_sharey=False,
                  layout_pad=0.,
+                 #hue_kws={'cmap': ['Blues', 'Oranges']},
                  hue_kws={'colors': [[c0[3], c0[0]], [c0[7], c0[5]]]}
                 )
 
@@ -94,4 +95,4 @@ g.axes[5,5].legend(loc='center left', bbox_to_anchor=((1.1,0.5)), frameon=False)
 
 plt.subplots_adjust(wspace=0.05, hspace=0.05)
 
-g.savefig(fname=paths.figures/"corner_GW170818_appendix.pdf", bbox_inches="tight", dpi=300)
+g.savefig(fname=paths.figures/"corner_GW200129_appendix.pdf", bbox_inches="tight", dpi=300)

### 2023-04-04 src/tex/ms.tex
gw170818 wording

@@
 \begin{equation}
-\chi_{\rm eff} \equiv \frac{1}{1+q} \left(\left|\vec{\chi}_{1\parallel}\right| + q\left|\vec{\chi}_{2\parallel}\right|\right)
+\chi_{\rm eff} \equiv \frac{1}{1+q} \left(\chi_{1\parallel} + q\chi_{2\parallel}\right)
 \end{equation}
-where $|\vec{\chi}_{i\parallel}|$ are the norms of the projections of the dimensionless spin vectors along the orbital angular momentum, and the  mass ratio is $q \equiv m_2/m_1 \leq 1$.
+where $\chi_{i\parallel}$ are the norms of the projections of the dimensionless spin vectors along the orbital angular momentum, and the  mass ratio is $q \equiv m_2/m_1 \leq 1$.
 We will study the interplay between birefringence and precession by focusing on the posterior of the precessing spin parameter $\chi_p$, defined as \cite{Schmidt:2014iyl}
 \begin{equation}
-\chi_p \equiv \max\left\{ |\vec{\chi}_{1\perp}|,\, k |\vec{\chi}_{2\perp}|\right\} ,
+\chi_p \equiv \max\left\{ \chi_{1\perp},\, k \chi_{2\perp}\right\} ,
 \end{equation}
-where $|\vec{\chi}_{i\perp}|$ are the norms of the projections of the dimensionless spin vectors onto the orbital plane at a reference time, and $k \equiv q\left(4 q +3\right) / \left(4 + 3q\right)$.
+where $\chi_{i\perp}$ are the norms of the projections of the dimensionless spin vectors onto the orbital plane at a reference time, and $k \equiv q\left(4 q +3\right) / \left(4 + 3q\right)$.
 We chart this and other approximate degeneracies as part of the results presented in Sec.~\ref{sec:Results}.
@@
 On the other hand, having $\kappa > 0$ enhances the left-handed modes during propagation, per Eq.~\eqref{eq:waveform_modification}.
-The two effects can be balanced to yield the correct polarization ratio observed at the detector (predominantly left-handed, per the preference for $\cos\iota \approx -1$ in the \ac{GR} analysis), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
+The two effects can be balanced to match the polarization ratio observed at the detector (predominantly left-handed, per the preference for $\cos\iota \approx -1$ in the \ac{GR} analysis), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
 This is similar to the degeneracy mentioned in Sec.~\ref{sec:inclination}.
@@
 It is difficult to unequivocally identify a specific feature of the GW170818 data that leads to this posterior structure.
-However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
+However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh} (see Appendix \ref{sec:corner_GW170818_appendix}).
 The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:corner_GW170818}, top right): allowing for $\kappa \neq 0$ leads to reduced support for precession.
@@
 
-The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ is likely a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
+The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ can be explained a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
 The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and the specific phasing of the precession cycle.
@@
 The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
-Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR} one.
+Indeed, the inference on the polarization angle in the birefringent case is notably narrower than in \ac{GR}.
 We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix~\ref{sec:corner_GW170818_appendix}.

### 2023-04-04 src/tex/ms.tex
GW910521 text

@@
 GW190521 is the most extreme representative of a class of events with bimodal $\kappa$ posteriors (Table \ref{tab:bimodal_events_mass}).
-This appears to arise out of a confluence of factors related to this event's high mass and support for precession: GW190521 being the most massive event in our set, its signal is only in-band for a very short time before merger, meaning that frequency-dependent birefringence can more easily resemble spin effects.
-In particular, a precession cycle can be approximated by a monotonic amplitude modulation for a short time.
-This effect gives rise to a similar degeneracy between $\kappa$ and $\chi_p$ as outlined for GW170818, wherein the prior volume favored configurations with $|\kappa| >0$.
-However, in the case of GW190521, the polarization measurement at the detectors is not accurate enough to fully disambiguate between right- and left-handed states (as implied by the bimodal $\cos\iota$ posterior obtained assuming \ac{GR} in Fig.~\ref{fig:corner_GW190521}).
-As a result, the GW190521 posterior does not favor $\kappa > 0$ exclusively, but supports both $\kappa > 0$ and $\kappa < 0$, leading to a bimodal posterior.
+Figure \ref{fig:corner_GW190521} shows that the two peaks of the $\kappa$ distribution arise from respective modes in the inclination (lower left corner): one solution corresponds to an intrinsically face-on source ($\cos\iota \approx 1$) with negative $\kappa$, and the other to a face-off source ($\cos\iota \approx -1$) with positive $\kappa$.
+In the former, an initially right-handed signal is birefringently enhanced by $\kappa < 0$ as it propagates towards the detector; in the latter, an initially left-handed signal is enhanced by $\kappa > 0$.
+The two outcomes (a right-handed or left-handed signal at the detector) match the two possibilities allowed also in the reference \ac{GR} run, which itself yields a bimodal $\cos\iota$ posterior.
+In both cases, the overall amplification of the signal introduced by a nonzero $\kappa$ is balanced by a large luminosity distance ($d_L \gtrsim 5 \, {\rm Gpc}$), so as to match the observed signal-to-noise ratio.
+
+Introducing birefringence affects our inference of the spins in the GW190521 system.
+We illustrate this in Fig.~\ref{fig:corner_GW190521} through the posterior on the effective spin $\chi_{\rm eff}$, which is favored to take values closer to zero in the birefringent analysis;
+this can also be seen clearly in the joint posterior for the individual spin magnitudes, provided in Appendix \ref{sec:corner_GW190521_appendix}.
+This suggests that at least part of the data features that lead the \ac{GR} analysis to infer near extremal spins can be alternatively explained as a consequence of birefringence.
+The fact that this is possible is not fully surprising given that the signal from this heavy system is only observed for a very small number of cycles \cite{LIGOScientific:2020iuh,LIGOScientific:2020ufj}.
+
+All in all, the situation for GW190521 is not unlike that of GW170818 discussed above (Sec.~\ref{sec:GW170818}): features in the data that would normally be interpreted as traces of high component spins are instead (at least partially) attributed to birefringent propagation, driven by a prior preference for high distances.
+The degeneracies in the case of GW190521 are likely exacerbated by the heavier mass of the system, which reduces the observed number of cycles.
+Crucially, GW190521 also lacks a definite measurement of the polarization state at the detector: without certainty about the ratio of right- versus left-handed components in the observed signal, it is not possible to disambiguate between $\kappa > 0$ versus $\kappa < 0$; hence the $\kappa$ distribution remains bimodal for GW190521, unlike for GW170918.
+The poor determination of the polarization state of the detected signal is evidenced by the bimodal $\cos\iota$ posterior inferred assuming \ac{GR} in Fig.~\ref{fig:corner_GW190521}, and is a consequence of several factors including sky location.
 

### 2023-04-04 src/tex/ms.tex
notable events done

@@
 
-Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW200129, GW190521 and GW191105.
-When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$; the second shows signs of potential noise anomalies; the third is representative of a class of heavy events with strongly bimodal $\kappa$ posteriors; and the fourth of a smaller class of lighter events with slightly bimodal posteriors.
-Through these examples, we elucidate the interaction between $\kappa$ and the source luminosity distance, inclination and spin precession.
+Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW190521 and GW200129.
+When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$; the third is representative of a class of events with bimodal $\kappa$ posteriors; and the third shows signs of potential noise anomalies.
+Through these examples, we elucidate the interaction between $\kappa$ and the source luminosity distance, inclination and spin parameters.
 
@@
 
-% Case: GW191105
-\subsubsection{GW191105}
-\label{sec:GW191105}
-
-The situation is similar for GW191105 (Fig.~\ref{fig:corner_GW191105}).
-
-\begin{figure}
-    \script{corner_GW191105.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
-    \caption{
-        GW191105 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
-        The posterior on $\chi_{eff}$ and $\kappa$ from our birefringence analysis is shown in the top right panel.
-    }
-    \label{fig:corner_GW191105}
-\end{figure}
+% % Case: GW191105
+% \subsubsection{GW191105}
+% \label{sec:GW191105}
+% 
+% The situation is similar for GW191105 (Fig.~\ref{fig:corner_GW191105}).
+% 
+% \begin{figure}
+%     \script{corner_GW191105.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
+%     \caption{
+%         GW191105 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+%         The posterior on $\chi_{eff}$ and $\kappa$ from our birefringence analysis is shown in the top right panel.
+%     }
+%     \label{fig:corner_GW191105}
+% \end{figure}
 
@@
 
-\section{Extended corner plot for GW170818}
+\section{Extended results for notable events}
+
+\subsection{GW170818}
 \label{sec:corner_GW170818_appendix}
@@
         Extended corner plot for GW170818: a supplement to Fig.~\ref{fig:corner_GW170818} discussed in Appendix~\ref{sec:corner_GW170818_appendix}.
+The shaded regions contain 90\% and 39.35\% ($1\sigma$) of the probability mass.
+        The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.
     }
@@
 
-\section{Extended corner plot for GW190521}
+\subsection{GW190521}
 \label{sec:corner_GW190521_appendix}
+
+As in the previous section, here we display additional parameters from the GW190521 measurement in Fig.~\ref{fig:corner_GW190521_appendix}, expanding upon Fig.~\ref{fig:corner_GW190521} in the main text.
+The reference \ac{GR} analysis, obtained with the \textsc{IMRPhenomXPHM} waveform, favors near extremal spins, in particular for the heavier component, and rules out nonspinning objects (i.e., $\chi_1 = \chi_2 = 0$) with $\geq 90\%$ credibility.
+This can be seen most clearly in the joint posterior for $\chi_1$ and $\chi_2$, rather than in the respective 1D marginals because the \acp{BH} in this system were inferred to have equal masses \cite{Biscoveanu:2020are}.
+
+Similar to GW170818, the \ac{GR} analysis for GW190521 also provides an informative measurement of the polarization angle $\psi$ and reference phase $\phi_{\rm ref}$.
+However, the fact that $\psi$ is constrained does not alone imply that the degeneracy between right- and left-handed states is broken (for a nonprecessing source, $\psi$ is related to the difference in phase between the circular polarization modes \cite{Isi:2022mbx}).
+Additionally, the posterior for tilt angle of the primary \ac{BH}, $\theta_1$, shows support for an antialigned spin ($\cos\theta_1 \approx -1$).
+For this events, spin and phase angles are referred to 10 Hz at the detector.
+
+Unlike the \ac{GR} analysis, the birefringent analysis does not favor extremal spins and is, in fact, consistent with nonspinning objects within 90\% credibility (lighter blue region encloses $\chi_1 = \chi_2 = 0$ in the respective panel of Fig.~\ref{fig:corner_GW190521_appendix}).
+In Fig.~\ref{fig:corner_GW190521}, this same fact manifested as a $\chi_{\rm eff}$ posterior with higher support for the origin.
+As we discussed in Sec.~\ref{sec:GW190521}, this suggests that some of the spin effects in these data are being absorbed by a nonzero $\kappa$.
+Other than the spins, the birefringent analysis differs most notably from the \ac{GR} one in the inference of the luminosity distance, which is favored to be much greater.
+See Sec.~\ref{sec:GW190521} in the main text for related exposition.
 
@@
         Extended corner plot for GW190521: a supplement to Fig.~\ref{fig:corner_GW190521} discussed in Appendix~\ref{sec:corner_GW190521_appendix}.
+The shaded regions contain 90\% and 39.35\% ($1\sigma$) of
+ the probability mass.
+        The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.
     }
@@
 
-\section{Extended corner plot for GW191105}
-\label{sec:corner_GW191105_appendix}
+% \subsection{Extended corner plot for GW191105}
+% \label{sec:corner_GW191105_appendix}
+% 
+% \begin{figure*}[h]
+%     \script{corner_GW191105_appendix.py}
+%     \includegraphics[width=\textwidth]{figures/corner_GW191105_appendix.pdf}
+%     \caption{
+%         Extended corner plot for GW191105: a supplement to Fig.~\ref{fig:corner_GW191105} discussed in Appendix~\ref{sec:corner_GW191105_appendix}.
+%     }
+%     \label{fig:corner_GW191105_appendix}
+% \end{figure*}
 
-\begin{figure*}[h]
-    \script{corner_GW191105_appendix.py}
-    \includegraphics[width=\textwidth]{figures/corner_GW191105_appendix.pdf}
-    \caption{
-        Extended corner plot for GW191105: a supplement to Fig.~\ref{fig:corner_GW191105} discussed in Appendix~\ref{sec:corner_GW191105_appendix}.
-    }
-    \label{fig:corner_GW191105_appendix}
-\end{figure*}
-
-\section{Extended corner plot for GW200129}
-\label{sec:corner_GW200129_appendix}
-
-\begin{figure*}[h]
-    \script{corner_GW200129_appendix.py}
-    \includegraphics[width=\textwidth]{figures/corner_GW200129_appendix.pdf}
-    \caption{
-        Extended corner plot for GW200129: a supplement to Fig.~\ref{fig:corner_GW200129} discussed in Appendix~\ref{sec:corner_GW200129_appendix}.
-    }
-    \label{fig:corner_GW200129_appendix}
-\end{figure*}
+% \section{Extended corner plot for GW200129}
+% \label{sec:corner_GW200129_appendix}
+% 
+% \begin{figure*}[h]
+%     \script{corner_GW200129_appendix.py}
+%     \includegraphics[width=\textwidth]{figures/corner_GW200129_appendix.pdf}
+%     \caption{
+%         Extended corner plot for GW200129: a supplement to Fig.~\ref{fig:corner_GW200129} discussed in Appendix~\ref{sec:corner_GW200129_appendix}.
+%     }
+%     \label{fig:corner_GW200129_appendix}
+% \end{figure*}
 

### 2023-04-04 src/scripts/corner_GW170818_appendix.py

@@ -6,6 +6,8 @@ from bilby.gw.result import CBCResult
 from kde_contour import Bounded_1d_kde, kdeplot_2d_clevels
 import paths
 
+rng = np.random.default_rng(12345)
+
 sns.set_theme(palette='colorblind', font_scale=2.0)
 
 plt.rcParams.update({
@@ -17,16 +19,20 @@ plt.rcParams.update({
 nsamples = 5000
 
 result_GR = CBCResult.from_json(filename=paths.data/"GW170818_GR.json.gz").posterior
-result_GR = result_GR.sample(n=nsamples)
+result_GR = result_GR.sample(n=nsamples, random_state=rng)
 result_GR['kappa'] = np.full(len(result_GR), None)
 result_GR['with'] = np.full(len(result_GR), "GR")
 result_GR['cos_iota'] = np.cos([float(value) for value in result_GR['iota']])
+result_GR['cos_tilt_1'] = np.cos([float(value) for value in result_GR['tilt_1']])
+result_GR['cos_tilt_2'] = np.cos([float(value) for value in result_GR['tilt_2']])
 
 result_BR = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
 result_BR = result_BR[result_BR.event == "GW170818"]
-result_BR = result_BR.sample(n=nsamples)
+result_BR = result_BR.sample(n=nsamples, random_state=rng)
 result_BR['with'] = np.full(len(result_BR), r"BR")
 result_BR['cos_iota'] = np.cos(result_BR['iota'])
+result_BR['cos_tilt_1'] = np.cos([float(value) for value in result_BR['tilt_1']])
+result_BR['cos_tilt_2'] = np.cos([float(value) for value in result_BR['tilt_2']])
 
 result = pd.concat([result_BR,result_GR], ignore_index=True)
 
@@ -46,7 +52,7 @@ def kdeplot1d(x, **kws):
     plt.fill_between(df['x'], df['y'], np.zeros(len(x)), alpha=0.2)
     plt.plot(df['x'], df['y'], lw=lw)
 
-vars = ['kappa','luminosity_distance','cos_iota','psi','a_1','a_2','tilt_1','tilt_2','phi_12','phi_jl','phase']
[…]

### 2023-04-04 src/scripts/corner_Gaussian.py

@@ -76,12 +76,17 @@ axs[1,1].set_xlim(0)
 axs[1,1].set_ylim(0)
 axs[1,1].set_yticks([])
 
+c0 = sns.color_palette('tab20c')
+colors = [[c0[2], c0[0]], [c0[7], c0[5]]]
+
 # axs[1,0].axvline(mu_median, color=sns.color_palette()[0])
 # axs[1,0].axhline(sigma_median, color=sns.color_palette()[0])
-kdeplot_2d_clevels(xs=df['mu'], ys=df['sigma'], ylow=0, ax=axs[1,0], color=sns.color_palette()[0], levels=[0.90,0.3935], linewidths=lw)
+kdeplot_2d_clevels(xs=df['mu'], ys=df['sigma'], ylow=0, ax=axs[1,0], colors=colors[0], levels=[0.90,0.3935], fill=True, alpha=0.7)
+kdeplot_2d_clevels(xs=df['mu'], ys=df['sigma'], ylow=0, ax=axs[1,0], color=c0[0], levels=[0.90,0.3935], fill=False, linewidhts=[0.5,], alpha=0.5)
+
 #axs[1,0].axvline(mu_median_without_GW200129, color=sns.color_palette()[3])
 #axs[1,0].axhline(sigma_median_without_GW200129, color=sns.color_palette()[3])
-kdeplot_2d_clevels(xs=df_without_GW200129['mu'], ys=df_without_GW200129['sigma'], ylow=0, ax=axs[1,0], color=sns.color_palette()[1], levels=[0.90,0.3935], linewidths=lw)
+kdeplot_2d_clevels(xs=df_without_GW200129['mu'], ys=df_without_GW200129['sigma'], ylow=0, ax=axs[1,0], colors=colors[1], levels=[0.90,0.3935], fill=True, alpha=0.7)
 
 axs[1,0].set_xlabel("$\\mu$")
 axs[1,0].set_ylabel("$\\sigma$")

### 2023-04-05 src/tex/ms.tex
Edit table 2 columns

@@
 \begin{table}
-    \caption{Events with bimodality in the $\kappa$ posterior, the \ac{GR} measurement of their detector-frame total mass ($M$), their $\chi_p$ measured without (GR) and with (BR) brefringence, and the credible level of $\kappa = 0$ (CL).}
+    \caption{Events with bimodality in the $\kappa$ posterior, the \ac{GR} measurement of their detector-frame total mass ($M$), their $\chi_p$ and $\chi_{\rm eff}$ measured without birefringence, and the credible level of $\kappa = 0$ (CL) with birefringence.}
     \begin{ruledtabular}
@@
 Additionally, the posterior for tilt angle of the primary \ac{BH}, $\theta_1$, shows support for an antialigned spin ($\cos\theta_1 \approx -1$).
-For this events, spin and phase angles are referred to 10 Hz at the detector.
+For these events, spin and phase angles are referred to 10 Hz at the detector.
 

### 2023-04-05 src/scripts/bimodal_events_mass.py

@@ -4,10 +4,16 @@ import scipy
 from bilby.gw.result import CBCResult
 import paths
 
+def sign(x):
+    if x > 0:
+        return "+"
+    elif x <= 0:
+        return ""
+
 result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
 
 with open(paths.output/"bimodal_events_mass.txt", "w") as f:
-    f.write(r"\begin{tabular}{lcccc}Event & $M$ ($M_{\odot}$) & $\chi_p$ (GR) & $\chi_p$ (BR) & CL\\ \hline ")
+    f.write(r"\begin{tabular}{lcccc}Event & $M$ ($M_{\odot}$) & $\chi_p$ & $\chi_{\rm eff}$ & CL\\ \hline ")
     for event in ["GW170104", "GW190413_134308", "GW190521", "GW190805_211137", "GW191105_143521"]:
         name = event.replace("_", r"\_")
 
@@ -23,19 +29,18 @@ with open(paths.output/"bimodal_events_mass.txt", "w") as f:
         chi_p_GR_95 = chi_p_GR.quantile(0.95) - chi_p_GR_median
         chi_p_GR_5 = chi_p_GR_median - chi_p_GR.quantile(0.05)
 
-        chi_p_BR = result_DataFrame[result_DataFrame.event == event]['chi_p']
-        chi_p_BR_median  = chi_p_BR.median()
-        chi_p_BR_95 = chi_p_BR.quantile(0.95) - chi_p_BR_median
-        chi_p_BR_5 = chi_p_BR_median - chi_p_BR.quantile(0.05)
+        chi_eff_GR = result_DataFrame[result_DataFrame.event == event]['chi_eff']
+        chi_eff_GR_median  = chi_eff_GR.median()
+        chi_eff_GR_95 = chi_eff_GR.quantile(0.95) - chi_eff_GR_median
+        chi_eff_GR_5 = chi_eff_GR_median - chi_eff_GR.quantile(0.05)
 
         kappa = result_DataFrame[result_DataFrame.event == event]['kappa'].values
 
         kernel = scipy.stats.gaussian_kde(kappa)
-        kde = [kernel(kappa) for kappa in kappa]
-        kde = np.array(kde)
[…]

### 2023-04-05 src/tex/ms.tex
conclusion WIP

@@
 \usepackage{bookmark}
+%\usepackage[all]{hypcap}
 
@@
 Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW190521 and GW200129.
-When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$; the third is representative of a class of events with bimodal $\kappa$ posteriors; and the third shows signs of potential noise anomalies.
-Through these examples, we elucidate the interaction between $\kappa$ and the source luminosity distance, inclination and spin parameters.
+When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$, the third is the most extreme representative of a class of events with bimodal $\kappa$ posteriors, and the third shows signs of potential noise anomalies.
+Through these examples, we elucidate some of the interactions between $\kappa$ and the source luminosity distance, inclination and spin parameters.
 
@@
 It is difficult to unequivocally identify a specific feature of the GW170818 data that leads to this posterior structure.
-However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh} (see Appendix \ref{sec:corner_GW170818_appendix}).
+However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization, phase and spin angles \cite{Varma:2021csh} (see Appendix \ref{sec:corner_GW170818_appendix}).
 The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:corner_GW170818}, top right): allowing for $\kappa \neq 0$ leads to reduced support for precession.
@@
 The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
-Indeed, the inference on the polarization angle in the birefringent case is notably narrower than in \ac{GR}.
+%Indeed, the inference on the polarization angle in the birefringent case is narrower than in \ac{GR}, suggesting a degree of fine tuning.
 We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix~\ref{sec:corner_GW170818_appendix}.
@@
 We illustrate this in Fig.~\ref{fig:corner_GW190521} through the posterior on the effective spin $\chi_{\rm eff}$, which is favored to take values closer to zero in the birefringent analysis;
-this can also be seen clearly in the joint posterior for the individual spin magnitudes, provided in Appendix \ref{sec:corner_GW190521_appendix}.
+this can also be seen, perhaps more clearly, in the joint posterior for the individual spin magnitudes, provided in Appendix \ref{sec:corner_GW190521_appendix}.
 This suggests that at least part of the data features that lead the \ac{GR} analysis to infer near extremal spins can be alternatively explained as a consequence of birefringence.
-The fact that this is possible is not fully surprising given that the signal from this heavy system is only observed for a very small number of cycles \cite{LIGOScientific:2020iuh,LIGOScientific:2020ufj}.
+The fact that this is possible is not fully surprising given that the signal from this heavy system is only observed for a very small number of cycles \cite{LIGOScientific:2020iuh,LIGOScientific:2020ufj} (cf.~Fig.~\ref{fig:birefringence}).
 
-All in all, the situation for GW190521 is not unlike that of GW170818 discussed above (Sec.~\ref{sec:GW170818}): features in the data that would normally be interpreted as traces of high component spins are instead (at least partially) attributed to birefringent propagation, driven by a prior preference for high distances.
+All in all, the situation for GW190521 is not unlike that for GW170818 discussed above (Sec.~\ref{sec:GW170818}): structures in the data that would normally be interpreted as traces of high component spins are instead (at least partially) attributed to birefringent propagation, driven by a prior preference for high distances.
 The degeneracies in the case of GW190521 are likely exacerbated by the heavier mass of the system, which reduces the observed number of cycles.
-Crucially, GW190521 also lacks a definite measurement of the polarization state at the detector: without certainty about the ratio of right- versus left-handed components in the observed signal, it is not possible to disambiguate between $\kappa > 0$ versus $\kappa < 0$; hence the $\kappa$ distribution remains bimodal for GW190521, unlike for GW170918.
-The poor determination of the polarization state of the detected signal is evidenced by the bimodal $\cos\iota$ posterior inferred assuming \ac{GR} in Fig.~\ref{fig:corner_GW190521}, and is a consequence of several factors including sky location.
+Crucially, GW190521 also lacks a definite measurement of the polarization state at the detector: without certainty about the ratio of right- versus left-handed components in the observed signal, it is not possible to disambiguate between $\kappa > 0$ versus $\kappa < 0$; hence the $\kappa$ distribution remains bimodal for GW190521, unlike for GW170818.
+The poor determination of the polarization state of the detected signal is evidenced by the bimodal $\cos\iota$ posterior inferred assuming \ac{GR} in Fig.~\ref{fig:corner_GW190521}, and is attributable to several factors including sky location.
 
@@
 While this is not conclusive proof that the glitch itself is driving the result, it does indicate that the Virgo data play a key role in the inference of $\kappa$.
-Since more work would be needed to understand the effect of the glitch, this motivates us to consider the effect of excluding this event from the collective analyses above.
+Since more work would be needed to understand the effect of the glitch, this motivated us to consider the effect of excluding this event from the collective analyses above (Sec.~\ref{sec:results:hier}).
+In any case, the outcome of that test showed the role of GW200129 to not be significant (e.g., Fig.~\ref{fig:corner_Gaussian}).
 
@@
 \subsection{Comparison with previous studies}
+
 \citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
@@
 
-% Future work
-\subsection{Future work}
+\subsection{Parameter degeneracies}
+
+In examining our results, we identified a number of interactions between birefringence and the inference of source properties like the component spin parameters, in addition to partial degeneracies with inclination and distance already discussed in the literature.
+We found that the frequency-dependent modification to the waveform induced by birefringence can effectively mimic spin effects, including precession, especially for heavier systems which are observed over short durations; we expect something similar to occur with mass ratio.
+
+As a consequence, a system that is inferred to be highly spinning in \ac{GR} may instead be inferred to be low-spinning but highly birefringent.
+These degeneracies can interact with the distribution of prior volume to yield a preference for nonzero $\kappa$ in some events.
+The resulting posterior can be bimodal or unimodal depending on the specific phase and polarization states observed at the detector.
+
+Conversely, we expect that a preference for birefringence can also appear for signals that do not show significant spin effects as long as the spin or other source parameters can be tuned to counteract the frequency-dependent dephasing induced by a nonzero $\kappa$.
+In those situations, $\kappa$ may become at least partially degenerate with inclination, also leading to a bimodal posterior.
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+We have reanalyzed all \acp{BBH} in GWTC-3 with $\mathrm{FAR} < 1/\mathrm{yr}$ to constrain amplitude birefringence in the propagation of \acp{GW}.
+To this end, we implemented a model in which right or left handed polarizations are amplified or suppressed over distance as a function of frequency, with an overall strength parametrized by a birefringent ``opacity'' parameter, $\kappa$, following Eq.~\eqref{eq:waveform_modification}.
+This parameterization is consistent with parity-odd theories like Chern-Simons gravity, and can be used to constrain them where applicable.
+
+We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events.
+As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources.
+The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
+Future catalogs, with more events, will shed light on this feature.
 
@@
 Therefore, we may need to wait for future \ac{PE} methods and tools to further our work.
-
-% More observations with higher SNR
-Another future work is to apply the same method to more \ac{GW} events and events with a higher \ac{SNR}.
-\ac{LVK} will release more events with higher \ac{SNR} in the future.
-Using data with higher \ac{SNR} allows us to obtain more precise \ac{PE} results and constrain the birefringence effect more precisely.
-And using data from more events will allow us to calculate a more constrained population posterior of $\kappa$.
 

### 2023-04-05 src/tex/ms.tex
Constraint by Wang et. al.

@@
 They formulated the birefringence effect as corrections to the GW waveform on a linear basis.
-We convert their constraint to the same units as ours with the derivation in Appendix~\ref{sec:M_PV_derivation}.  \wf{Can you give the answer here (as well as in Appendix B)?}
-% Discussion on the difference between our work and theirs
+We convert their constraint to the same units as ours with the derivation in Appendix~\ref{sec:M_PV_derivation}.
+They obtained a constraint equivalent to $|\kappa| \lesssim 0.30$.
 
@@
 \end{equation}
-Using $H_0d_C=cz$, where $H_0$ is the Hubble constant and $c$ is the speed of light, we can rewrite the equation as
+Using $H_0d_C=cz$, where $H_0$ is the Hubble constant and $c$ is the speed of light, and adding back the factor of Planck constant $h$, we can rewrite the equation as
 \begin{equation}
-    \kappa=\frac{\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{PV}^{-1}\,.
+    \kappa=\frac{h\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{PV}^{-1}\,.
 \end{equation}
+They gave the constraint on $M_{PV}$ as $M_{PV}>10^{-22}\,\mathrm{GeV}$.
+This constraint only considered the amplitude birefringence modification.
+Equivalently, the constraint on $\kappa$ is $\kappa \lesssim 0.30$.
 

### 2023-04-05 src/tex/ms.tex
Add absolute sign
mass ratio plot

@@
 
-In examining our results, we identified a number of interactions between birefringence and the inference of source properties like the component spin parameters, in addition to partial degeneracies with inclination and distance already discussed in the literature.
-We found that the frequency-dependent modification to the waveform induced by birefringence can effectively mimic spin effects, including precession, especially for heavier systems which are observed over short durations; we expect something similar to occur with mass ratio.
+\begin{figure}
+    \script{mass_ratio.py}
+    \includegraphics[width=\columnwidth]{figures/mass_ratio.pdf}
+    \caption{
+    Impact of birefringence on mass ratio inference.
+Posterior on $q \equiv m_2/m_1$ for GW170104 (top) and GW191105 (bottom), as obtained from both the birefringent analysis (orange) and the regular \ac{GR} analysis (blue).
+    }
+    \label{fig:mass_ratio}
+\end{figure}
+
+In examining our results in Sec.~\ref{sec:results:notable}, we identified a number of interactions between birefringence and the inference of source properties like the component spin parameters---in addition to partial degeneracies with inclination and distance already discussed in the literature.
+We found that the frequency-dependent modification to the waveform induced by birefringence can effectively mimic spin effects, including precession, especially for heavier systems which are observed over short durations.
 
@@
 In those situations, $\kappa$ may become at least partially degenerate with inclination, also leading to a bimodal posterior.
+Either way, a preference for nonzero $\kappa$ will only manifest if the data offer enough flexibility to unlock a region of parameter space favored by prior volume, e.g., in letting the distance grow.
+
+Since mass ratio and spins can have similar effects, we expect some interactions could take place between $q$ and $\kappa$.
+This seems to indeed be the case for some events, like GW170104 or GW191105, as we demonstrate in Fig.~\ref{fig:mass_ratio}.
+This degeneracy is likely more limited since mass ratio alone cannot produce significant amplitude modulations.
 
@@
 
-We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events.
-As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources.
+We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
+As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).
 The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
@@
 
+From the set of results from individual events, we highlighted two, GW170818 and GW190521, whose posterior manifests interactions between birefringence and our inference of source parameters (Sec.~\ref{sec:results:notable}).
+In particular, we identified the relevance of spins their (partial) degeneracy with $\kappa$, in addition to expected correlations with source inclination and distance.
+
 \mi{WIP}
 If our inference on $\kappa$ is tied to precession and the spin orientations, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
-In any case, as we argued in Sec.~\ref{sec:results:hier}, there is no evidence at the moment that this event is inconsistent with the either collection of measurements or with $\kappa = 0$ (Fig.~\ref{fig:reweighted_kappa}).
 
@@
 \begin{equation}
-    \kappa=\frac{h\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{PV}^{-1}\,.
+    |\kappa|=\frac{h\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{PV}^{-1}\,.
 \end{equation}
@@
 This constraint only considered the amplitude birefringence modification.
-Equivalently, the constraint on $\kappa$ is $\kappa \lesssim 0.30$.
+Equivalently, the constraint on $\kappa$ is $|\kappa| \lesssim 0.30$.
 

### 2023-04-06 src/tex/ms.tex
Add calculated constraints for comparison

@@
 
-Both versions of the posterior support the lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals of the result including all events, we constrain $\mu =$ \variable{output/mu_median.txt} for 90\%-credible symmetric intervals around the median, and $\sigma <$ \variable{output/sigma_median.txt} for the 90\%-credible one-sided upper limit.
+Both versions of the posterior support the lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals of the result including all events, we constrain $\mu =$ \variable{output/mu_median.txt}for 90\%-credible symmetric intervals around the median, and $\sigma <$ \variable{output/sigma_median.txt}for the 90\%-credible one-sided upper limit.
 However, even though $\sigma = 0$ is well supported, the $\sigma$ posterior peaks visibly away from the origin, indicating some preference for a nonzero variance.
@@
 We convert the constraint they gave to the same units as ours to make a comparison.
-They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
-We obtained a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
-Our result is an order of magnitude improvement in constraining $\kappa$.
-The main reason is that we have more events from GWTC-3 than GWTC-2.
+They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at 1 standard deviation.
+We obtained a tighter constraint on $|\kappa|$ with $|\kappa| <$ \variable{output/restricted_kappa_std.txt}at 1 standard deviation.
 
@@
 We convert their constraint to the same units as ours with the derivation in Appendix~\ref{sec:M_PV_derivation}.
-They obtained a constraint equivalent to $|\kappa| \lesssim 0.30$.
+They obtained a constraint equivalent to $|\kappa| \lesssim 0.30$, given by the 90\% upper limit.
+We obtained a tighter 90\% upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_90.txt}.
+
+Our result is an order of magnitude improvement in constraining $\kappa$.
+The main reason is that we included more events from GWTC-3 compared to \citet{Okounkova_2022} and \citet{Wang_2021}.
 
@@
 
-We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
+We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained $\kappa =$ \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
 As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).
@@
         Extended corner plot for GW190521: a supplement to Fig.~\ref{fig:corner_GW190521} discussed in Appendix~\ref{sec:corner_GW190521_appendix}.
-The shaded regions contain 90\% and 39.35\% ($1\sigma$) of
- the probability mass.
+        The shaded regions contain 90\% and 39.35\% ($1\sigma$) of the probability mass.
         The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.

### 2023-04-06 src/scripts/kappa_constraint.py

@@ -18,13 +18,16 @@ df_Gaussian_samples_all = pd.DataFrame()
 df_Gaussian_samples_all['mu'] = samples_all[:,0]
 df_Gaussian_samples_all['sigma'] = samples_all[:,1]
 
+# generic constraint
 samples_reweighted = []
 for _ in range(1000):
     Gaussian = np.full(shape=(len(result_DataFrame), 2), fill_value=df_Gaussian_samples_all.sample(1))
     weight = normal_distribution(result_DataFrame['kappa'], Gaussian[:,0], Gaussian[:,1])
     weight /= weight.sum()
     samples_reweighted.append(np.random.choice(a=np.array(result_DataFrame['kappa']), size=1, p=weight))
+samples_reweighted = np.array(samples_reweighted).reshape(-1)
 
+# restricted constraint
 kernels = [scipy.stats.gaussian_kde(result_dict[event]['kappa']) for event in events]
 kappa = np.linspace(-.1, .1, 1000)
 ll = [np.sum([np.log(ker(k)) for ker in kernels]) for k in kappa]
@@ -32,16 +35,38 @@ ll = ll - np.max(ll)
 likelihood = np.exp(ll)
 likelihood = likelihood / np.trapz(likelihood, x=kappa)
 
-restricted_kappa_median = np.interp(0.5,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa)
-generic_kappa_median = np.median(np.array(samples_reweighted).reshape(-1))
+# output results
+restricted_cdf = np.array([np.trapz(likelihood[0:i],kappa[0:i]) for i in range(len(kappa))])
+restricted_kappa_median = np.interp(0.5,restricted_cdf,kappa)
+restricted_kappa_95 = np.interp(0.95,restricted_cdf,kappa)
+restricted_kappa_5 = np.interp(0.05,restricted_cdf,kappa)
+
+generic_kappa_median = np.median(samples_reweighted)
+generic_kappa_95 = np.percentile(a=samples_reweighted,q=95)
+generic_kappa_5 = np.percentile(a=samples_reweighted,q=5)
 
 credible_level = len(likelihood[likelihood > np.interp(0, kappa, likelihood)])/len(likelihood)
 
+restricted_kappa_mean = np.trapz(kappa*likelihood,kappa)
[…]

### 2023-04-06 src/tex/ms.tex
mass ratio discussion

@@
     \caption{
-    Impact of birefringence on mass ratio inference.
-Posterior on $q \equiv m_2/m_1$ for GW170104 (top) and GW191105 (bottom), as obtained from both the birefringent analysis (orange) and the regular \ac{GR} analysis (blue).
+    Impact of birefringence on mass ratio and precession inference.
+Posterior on mass ratio $q \equiv m_2/m_1$ (left) and precessing spin $\chi_p$ (right) for GW170104 (top) and GW191105 (bottom), as obtained from both the birefringent analysis (orange) and the regular \ac{GR} analysis (blue).
     }
@@
 
-In examining our results in Sec.~\ref{sec:results:notable}, we identified a number of interactions between birefringence and the inference of source properties like the component spin parameters---in addition to partial degeneracies with inclination and distance already discussed in the literature.
-We found that the frequency-dependent modification to the waveform induced by birefringence can effectively mimic spin effects, including precession, especially for heavier systems which are observed over short durations.
+In examining our results in Sec.~\ref{sec:results:notable}, we identified interactions between birefringence and the inference of source properties like the component spin parameters---in addition to expected partial degeneracies with inclination and distance (Sec.~\ref{sec:inclination}).
+In particular, we found that the frequency-dependent modification to the waveform induced by birefringence can effectively mimic spin effects, including precession, especially for heavier systems which are observed over short durations.
 
-As a consequence, a system that is inferred to be highly spinning in \ac{GR} may instead be inferred to be low-spinning but highly birefringent.
-These degeneracies can interact with the distribution of prior volume to yield a preference for nonzero $\kappa$ in some events.
-The resulting posterior can be bimodal or unimodal depending on the specific phase and polarization states observed at the detector.
+As a consequence, a system that would be inferred to be highly spinning in \ac{GR} may instead be inferred to be low-spinning but highly birefringent if nonzero values of $\kappa$ are allowed.
+These morphological degeneracies can interact with the distribution of prior probability mass to yield a preference for nonzero $\kappa$ in some events: in absence of a strong constraint from the data, it can be more favorable to place a source at far distances with a large $|\kappa|$, than at close distances with a fine tuned spin configuration.
+The resulting posterior can be unimodal or bimodal depending on the specific phase and polarization states observed at the detector.
+The events GW170818 (Sec.~\ref{sec:GW170818}) and GW190521 (Sec.~\ref{sec:GW190521}) appear to fall in these two categories respectively.
 
-Conversely, we expect that a preference for birefringence can also appear for signals that do not show significant spin effects as long as the spin or other source parameters can be tuned to counteract the frequency-dependent dephasing induced by a nonzero $\kappa$.
+Conversely, we expect that support for birefringence can also appear for signals that do not necessarily show significant spin effects in the \ac{GR} analysis, as long as the spins or other source parameters can be tuned to \emph{counteract} the frequency-dependent dephasing that would be otherwise induced by a nonzero $\kappa$.
 In those situations, $\kappa$ may become at least partially degenerate with inclination, also leading to a bimodal posterior.
-Either way, a preference for nonzero $\kappa$ will only manifest if the data offer enough flexibility to unlock a region of parameter space favored by prior volume, e.g., in letting the distance grow.
+This is because the intrinsic source parameters may, to an extent, ``absorb'' the frequency dependence of the birefringent effect, approximately reinstating the degeneracy expected for a single Fourier mode (Sec.~\ref{sec:inclination}).
+%Either way, a preference for nonzero $\kappa$ will only manifest if the data offer enough flexibility to unlock a region of parameter space favored by prior volume, e.g., in letting the distance grow.
 
-Since mass ratio and spins can have similar effects, we expect some interactions could take place between $q$ and $\kappa$.
+Besides spins, other parameters can interact with $\kappa$ either directly or through their own couplings with the spins.
+Since the mass ratio and spins can affect the phasing of a \ac{GW} inspiral at similar order \cite{Blanchet:2013haa}, we might expect some degree of interaction between $q$ and $\kappa$.
 This seems to indeed be the case for some events, like GW170104 or GW191105, as we demonstrate in Fig.~\ref{fig:mass_ratio}.
+In the former, introducing birefringence increases the preference for asymmetric masses ($q < 1$), while it decreases the inferred level of precession; in the latter, birefringence increases the preference for symmetric masses ($q = 1$) and does not affect the inference on precession.
+In neither case is $\chi_{\rm eff}$ significantly impacted by birefringence.
 This degeneracy is likely more limited since mass ratio alone cannot produce significant amplitude modulations.

### 2023-04-06 src/scripts/kappa_constraint.py
Add calculation of Wang's constraint

@@ -41,7 +41,7 @@ restricted_kappa_median = np.interp(0.5,restricted_cdf,kappa)
 restricted_kappa_95 = np.interp(0.95,restricted_cdf,kappa)
 restricted_kappa_5 = np.interp(0.05,restricted_cdf,kappa)
 with open(paths.output/"restricted_kappa_median.txt", "w") as f:
-    f.write(f"${restricted_kappa_median:.3f}^{{+{(restricted_kappa_95-restricted_kappa_median):.3f}}}_ {{{(restricted_kappa_5-restricted_kappa_median):.3f}}}$")
+    f.write(f"${restricted_kappa_median:.3f}^{{+{(restricted_kappa_95-restricted_kappa_median):.3f}}}_{{{(restricted_kappa_5-restricted_kappa_median):.3f}}}$")
 
 generic_kappa_median = np.median(samples_reweighted)
 generic_kappa_95 = np.percentile(a=samples_reweighted,q=95)

### 2023-04-06 src/scripts/kappa_constraint_Wang.py

@@ -1,0 +1,13 @@
+import numpy as np
+import paths
+
+h = 4.135667696e-15 # eV s
+c = 299792.458 # km/s
+H_0 = 68.3 # km/s/Mpc
+
+M_PV_lower_limit = 1e-22*1e9 # eV
+
+kappa_Wang = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(M_PV_lower_limit)
+
+with open(paths.output/"kappa_Wang.txt", "w") as f:
+    f.write(f"${kappa_Wang:.2f}$")

### 2023-04-06 src/tex/ms.tex
Add calculation of Wang's constraint
Temporary text for comparison

@@
 
-\citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
-We convert the constraint they gave to the same units as ours to make a comparison.
-They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at 1 standard deviation.
-We obtained a tighter constraint on $|\kappa|$ with $|\kappa| <$ \variable{output/restricted_kappa_std.txt}at 1 standard deviation.
+\citet{Okounkova_2022} gave a constraint on frequency-independent \ac{GW} amplitude birefringence by performing statistical analysis on GWTC-2.
+We take $f=100\, \mathrm{Hz}$ as the reference frequency to make the comparison.
+They were able to constrain $\tilde{\kappa}$ at $d_C=d_C(z=0.3)$ to be $|\tilde{\kappa}| \lesssim 0.74$ at 1 $\sigma$, where
+\begin{equation}
+    \tilde{\kappa}d_C = \left[{\kappa}\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\, \mathrm{Hz}}\right]_{f=100\, \mathrm{Hz}}\,.
+\end{equation}
+Note that $\tilde{\kappa}$ is in units of $1/\mathrm{Gpc}$, while $\kappa$ is dimensionless.
+Equivalently, their constraint is $|\kappa| \lesssim 0.74$ at 1 $\sigma$.
+We obtained a tighter constraint on $|\kappa|$ with $|\kappa| <$ \variable{output/restricted_absolute_kappa_68.txt}at $68.27\%$ upper limit.
 
@@
 We convert their constraint to the same units as ours with the derivation in Appendix~\ref{sec:M_PV_derivation}.
-They obtained a constraint equivalent to $|\kappa| \lesssim 0.30$, given by the 90\% upper limit.
+They obtained a constraint equivalent to $|\kappa| <$ \variable{output/kappa_Wang.txt}, given by the 90\% upper limit.
 We obtained a tighter 90\% upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_90.txt}.
@@
 This constraint only considered the amplitude birefringence modification.
-Equivalently, the constraint on $\kappa$ is $|\kappa| \lesssim 0.30$.
+Equivalently, the constraint on $\kappa$ is $|\kappa| <$ \variable{output/kappa_Wang.txt}.
 

### 2023-04-06 src/scripts/kappa_constraint.py

@@ -53,15 +53,12 @@ credible_level = len(likelihood[likelihood > np.interp(0, kappa, likelihood)])/l
 with open(paths.output/"CL_kappa_0.txt", "w") as f:
     f.write(f"${credible_level:.3f}$")
 
-restricted_kappa_mean = np.trapz(kappa*likelihood,kappa)
-restricted_kappa_var = np.trapz(np.square(kappa-restricted_kappa_mean)*likelihood,kappa)
-restricted_kappa_std = np.sqrt(restricted_kappa_var)
-with open(paths.output/"restricted_kappa_std.txt", "w") as f:
-    f.write(f"${restricted_kappa_std:.2f}$")
-
 absolute_kappa = kappa[len(kappa)//2:]
 likelihood_absolute_kappa = likelihood[len(kappa)//2:]+likelihood[0:len(kappa)//2][::-1] # for len(kappa) even
 restricted_cdf_absolute_kappa = np.array([np.trapz(likelihood_absolute_kappa[0:i],absolute_kappa[0:i]) for i in range(len(absolute_kappa))])
+restricted_absolute_kappa_68 = np.interp(0.6827,restricted_cdf_absolute_kappa,absolute_kappa)
+with open(paths.output/"restricted_absolute_kappa_68.txt", "w") as f:
+    f.write(f"${restricted_absolute_kappa_68:.2f}$")
 restricted_absolute_kappa_90 = np.interp(0.9,restricted_cdf_absolute_kappa,absolute_kappa)
 with open(paths.output/"restricted_absolute_kappa_90.txt", "w") as f:
     f.write(f"${restricted_absolute_kappa_90:.2f}$")

### 2023-04-06 src/tex/ms.tex
Update ms.tex
q discussion refinement

@@
 \begin{table}
-    \caption{Events with bimodality in the $\kappa$ posterior, the \ac{GR} measurement of their detector-frame total mass ($M$), their $\chi_p$ and $\chi_{\rm eff}$ measured without birefringence, and the credible level of $\kappa = 0$ (CL) with birefringence.}
+    \caption{Events with bimodality in the $\kappa$ posterior, the \ac{GR} measurement of their detector-frame total mass ($M$), precessing spin $\chi_p$ and effective spin $\chi_{\rm eff}$, as well as the credible level of $\kappa = 0$ (CL) from the birefringence analysis.}
     \begin{ruledtabular}
@@
 Equivalently, their constraint is $|\kappa| \lesssim 0.74$ at 1 $\sigma$.
-We obtained a tighter constraint on $|\kappa|$ with $|\kappa| <$ \variable{output/restricted_absolute_kappa_68.txt}at $68.27\%$ upper limit.
+We obtained a tighter $68.27\%$ upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_68.txt}.
 
@@
     \caption{
-    Impact of birefringence on mass ratio and precession inference.
     Posterior on mass ratio $q \equiv m_2/m_1$ (left) and precessing spin $\chi_p$ (right) for GW170104 (top) and GW191105 (bottom), as obtained from both the birefringent analysis (orange) and the regular \ac{GR} analysis (blue).
@@
 As a consequence, a system that would be inferred to be highly spinning in \ac{GR} may instead be inferred to be low-spinning but highly birefringent if nonzero values of $\kappa$ are allowed.
-These morphological degeneracies can interact with the distribution of prior probability mass to yield a preference for nonzero $\kappa$ in some events: in absence of a strong constraint from the data, it can be more favorable to place a source at far distances with a large $|\kappa|$, than at close distances with a fine tuned spin configuration.
+These morphological degeneracies can interact with the distribution of prior probability mass to yield a preference for nonzero $\kappa$ in some events: in absence of a strong constraint from the data, it can be more favorable to place a source at far distances with large $|\kappa|$, than at close distances with a fine tuned spin configuration and low $|\kappa|$.
 The resulting posterior can be unimodal or bimodal depending on the specific phase and polarization states observed at the detector.
-The events GW170818 (Sec.~\ref{sec:GW170818}) and GW190521 (Sec.~\ref{sec:GW190521}) appear to fall in these two categories respectively.
+The events GW170818 (Sec.~\ref{sec:GW170818}) and GW190521 (Sec.~\ref{sec:GW190521}) appear to fall into these two categories respectively.
 
@@
 Besides spins, other parameters can interact with $\kappa$ either directly or through their own couplings with the spins.
-Since the mass ratio and spins can affect the phasing of a \ac{GW} inspiral at similar order \cite{Blanchet:2013haa}, we might expect some degree of interaction between $q$ and $\kappa$.
+Since the mass ratio and can affect the phasing of a \ac{GW} inspiral at similar order as the spins \cite{Blanchet:2013haa}, we might expect some degree of interaction between $q$ and $\kappa$.
 This seems to indeed be the case for some events, like GW170104 or GW191105, as we demonstrate in Fig.~\ref{fig:mass_ratio}.
-In the former, introducing birefringence increases the preference for asymmetric masses ($q < 1$), while it decreases the inferred level of precession; in the latter, birefringence increases the preference for symmetric masses ($q = 1$) and does not affect the inference on precession.
+In the former, introducing birefringence increases the preference for asymmetric masses ($q < 1$) while decreasing the inferred level of precession; in the latter, birefringence increases the preference for symmetric masses ($q = 1$) but does not affect the inference on precession.
+The $\kappa$ posteriors are bimodal for both events, but do not rule out $\kappa = 0$ (Fig.~\ref{fig:violin_kappa}).
 In neither case is $\chi_{\rm eff}$ significantly impacted by birefringence.
-This degeneracy is likely more limited since mass ratio alone cannot produce significant amplitude modulations.
+The coupling between $q$ and $\kappa$ is likely limited by the fact that changes in the mass ratio alone cannot produce significant amplitude modulations.
 

### 2023-04-06 src/scripts/corner_GW200129.py

@@ -54,6 +54,7 @@ def kdeplot1d(x, mask=None, **kws):
     df = df.sort_values(['x'])
     if kws.pop('fill', False):
         plt.fill_between(df['x'], df['y'], np.zeros(len(x)), alpha=0.2, **kws)
+    lw = kws.pop('lw', 1)
     plt.plot(df['x'], df['y'], lw=lw, **kws)
 
 p = list(sns.color_palette())[:3] + ['0.35']
@@ -73,8 +74,8 @@ g.map_lower(kdeplot2d, mask=m, levels=[0.90], linewidths=lw)
 g.map_diag(kdeplot1d, mask=m, fill=False)
 
 m = result_all['with'] == "HL"
-g.map_lower(kdeplot2d, mask=m, linestyles='--', levels=[0.90], zorder=100, linewidths=lw)
-g.map_diag(kdeplot1d, mask=m, linestyle='--', fill=False, zorder=100)
+g.map_lower(kdeplot2d, mask=m, linestyles='--', levels=[0.90], zorder=100, linewidths=2*lw)
+g.map_diag(kdeplot1d, mask=m, linestyle='--', fill=False, zorder=100, lw=2*lw)
 
 for i in range(len(vars)):
     g.axes[i,i].set_xlim(result_all[vars[i]].min(), result_all[vars[i]].max())
@@ -114,7 +115,7 @@ ax.plot(chi_p, bounded_kde, c=p[0], lw=lw)
 ax.fill_between(chi_p, bounded_kde, np.zeros(len(chi_p)), alpha=0.2, color=p[0])
 ax.plot(chi_p_without_H1, bounded_kde_without_H1, c=p[1], lw=lw)
 ax.plot(chi_p_without_L1, bounded_kde_without_L1, c=p[2], lw=lw)
-ax.plot(chi_p_without_V1, bounded_kde_without_V1, c=p[3], lw=lw, ls='--')
+ax.plot(chi_p_without_V1, bounded_kde_without_V1, c=p[3], lw=lw*2, ls='--')
 
 ax.set_xlabel(r"$\chi_p$")
 ax.set_ylabel("")

### 2023-04-06 src/tex/ms.tex
edits to comparison with Masha's paper

@@
 
-\citet{Okounkova_2022} gave a constraint on frequency-independent \ac{GW} amplitude birefringence by performing statistical analysis on GWTC-2.
-We take $f=100\, \mathrm{Hz}$ as the reference frequency to make the comparison.
-They were able to constrain $\tilde{\kappa}$ at $d_C=d_C(z=0.3)$ to be $|\tilde{\kappa}| \lesssim 0.74$ at 1 $\sigma$, where
+Our primary constraint on birefringence is given by the GWTC-3 combined constraint on the birefringent opacity, $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility, obtained in Sec.~\ref{sec:results:gwtc} (blue band in Fig.~\ref{fig:violin_kappa}).
+This can be compared to previous measurements in the literature.
+
+\citet{Okounkova_2022} produced a constraint on frequency-independent \ac{GW} amplitude birefringence from the distribution of measured inclinations of \acp{BBH} in GWTC-2.
+As opposed to our Eq.~\eqref{eq:waveform_modification}, this reference parameterized the effect of birefringence as
+\begin{equation}
+    h_{L/R}^{\mathrm{br}}(f) =
+    h_{L/R}^{\mathrm{GR}}(f) \times
+    \exp\left(\pm \tilde{\kappa}\, d_C\right)\,,
+\end{equation}
+where $\tilde{\kappa}$ is a frequency-independent opacity (simply denoted ``$\kappa$'' in Ref.~\cite{Okounkova_2022}).
+Assuming a fiducial redshift of $z=0.3$ for all events in GWTC-2, \citet{Okounkova_2022} obtained a constraint of $|\tilde{\kappa}| \lesssim 0.74\, {\rm Gpc}^{-1}$ at 68\% credibility.
+Taking a reference frequency of $f=100\, \mathrm{Hz}$, which is representative of LIGO-Virgo events, we may compare that constraint to ours by noting that
 \begin{equation}
@@
 \end{equation}
-Note that $\tilde{\kappa}$ is in units of $1/\mathrm{Gpc}$, while $\kappa$ is dimensionless.
-Equivalently, their constraint is $|\kappa| \lesssim 0.74$ at 1 $\sigma$.
-We obtained a tighter $68.27\%$ upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_68.txt}.
+per Eq.~\eqref{eq:waveform_modification}, and hence $\tilde{\kappa} = \kappa / {\rm Gpc}$ at $f=100\, {\rm Hz}$.
+In terms of this fiducial quantity, our constraint becomes $|\tilde{\kappa}| <$ \variable{output/restricted_absolute_kappa_68.txt} ${\rm Gpc}^{-1}$ at 68\% credibility.
 

### 2023-04-07 src/tex/ms.tex
moved wang et al appendix into main text

@@
 
-Our primary constraint on birefringence is given by the GWTC-3 combined constraint on the birefringent opacity, $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility, obtained in Sec.~\ref{sec:results:gwtc} (blue band in Fig.~\ref{fig:violin_kappa}).
-This can be compared to previous measurements in the literature.
+Our primary constraint on birefringence is given by the combined constraint on the birefringent opacity from GWTC-3, $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility, obtained in Sec.~\ref{sec:results:gwtc} (blue band in Fig.~\ref{fig:violin_kappa}).
+We can compare this to previous measurements of amplitude birefringence in the literature.
 
+\subsubsection{Okounkova et al.}
 \citet{Okounkova_2022} produced a constraint on frequency-independent \ac{GW} amplitude birefringence from the distribution of measured inclinations of \acp{BBH} in GWTC-2.
@@
 \begin{equation}
+\label{eq:freqindep}
     h_{L/R}^{\mathrm{br}}(f) =
@@
 Assuming a fiducial redshift of $z=0.3$ for all events in GWTC-2, \citet{Okounkova_2022} obtained a constraint of $|\tilde{\kappa}| \lesssim 0.74\, {\rm Gpc}^{-1}$ at 68\% credibility.
-Taking a reference frequency of $f=100\, \mathrm{Hz}$, which is representative of LIGO-Virgo events, we may compare that constraint to ours by noting that
+Taking a reference frequency of $f=100\, \mathrm{Hz}$, which is representative of LIGO-Virgo events, and adopting the same fiducial distance $d_C(z=0.3)$, we may compare that constraint to ours by noting that
 \begin{equation}
-    \tilde{\kappa}d_C = \left[{\kappa}\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\, \mathrm{Hz}}\right]_{f=100\, \mathrm{Hz}}\,.
+    \left[ \tilde{\kappa} = \frac{\kappa}{1\, \mathrm{Gpc}} \right]_{\raisebox{1ex}{$\begin{subarray}{l} z=0.3 \\ f=100\,{\rm Hz} \end{subarray}$}}
 \end{equation}
-per Eq.~\eqref{eq:waveform_modification}, and hence $\tilde{\kappa} = \kappa / {\rm Gpc}$ at $f=100\, {\rm Hz}$.
-In terms of this fiducial quantity, our constraint becomes $|\tilde{\kappa}| <$ \variable{output/restricted_absolute_kappa_68.txt} ${\rm Gpc}^{-1}$ at 68\% credibility.
+per Eq.~\eqref{eq:waveform_modification}.
+In terms of this fiducial quantity, our constraint from Sec.~\ref{sec:results:gwtc} can be expressed as $|\tilde{\kappa}| <$ \variable{output/restricted_absolute_kappa_68.txt} ${\rm Gpc}^{-1}$ at 68\% credibility.
+This is a factor of ${\sim}25\times$ \mi{Thomas, can you make this number be dynamically generated?
+it should be the ratio of Masha's constraint to ours, rounded to the closest integer} more stringent than Ref.~\cite{Okounkova_2022}, besides arising from a less simplified model.
 
-\citet{Wang_2021} performed \ac{PE} on GWTC-1 events with a frequency-dependent birefringence model.
-They formulated the birefringence effect as corrections to the GW waveform on a linear basis.
-We convert their constraint to the same units as ours with the derivation in Appendix~\ref{sec:M_PV_derivation}.
-They obtained a constraint equivalent to $|\kappa| <$ \variable{output/kappa_Wang.txt}, given by the 90\% upper limit.
+\subsubsection{Wang et al.}
+
+\citet{Wang_2021} performed \ac{PE} on GWTC-1 events with a frequency-dependent model of birefringence resembling ours.
+In contrast to our Eq.~\eqref{eq:waveform_modification}, this reference parameterized the birefringent waveform in terms of some amplitude and phase modifications to the linear polarizations ($\delta h$ and $\delta \Psi$ respectively), such that
+\begin{equation}
+    h_{+/\times}^{\rm BR}(f) = h_{+/\times}^{\rm GR}(f)\mp h_{\times/+}^{\rm GR}(f)(i\delta h-\delta\Psi)\,,
+\end{equation}
+with the plus (minus) sign for the cross (plus) polarization.
+In terms of circular polarizations, this is equivalent to 
+\begin{align}
+h^{\rm BR}_{L/R}(f) &= h^{\rm GR}_{L/R}(f) \left(1 \mp \delta h \mp i \delta \Psi \right) \nonumber\\
+&\approx h^{\rm GR}_{L/R}(f) \exp(\mp \delta h \mp i \delta \Psi) \, ,
+\end{align}
+assuming a small $\delta $ and $\delta \Psi \ll 1$.
+To consider only amplitude birefringence, as in this work, we must compare to the result in Ref.~\citep{Wang_2021} that set $\delta \Psi = 0$ and parameterized $\delta h = \pi f z / M_{\rm PV}$, where $M_{\rm PV}$ is the energy scale of the birefringent (parity-violating) correction in units of frequency.%
+\footnote{Concretely, \citet{Wang_2021} write $\delta h = - A_\nu \pi f$ with $A_\nu = M_{\rm PV}^{-1} \left[\alpha_\nu (z=0) - \alpha_\nu(z) \left(1+z\right)\right]$, for $\alpha_\nu$ some function of redshift encoding the evolution of a birefringence-mediating field; to produce their constraint, however, they further set $\alpha_\nu(z) = 1$, yielding $A_\nu = -z / M_{\rm PV}$ and hence $\delta h = \pi f z / M_{\rm PV}$.}
+Comparing to our Eq.~\eqref{eq:waveform_modification} and approximating $z \approx d_C H_0/c$ with the Hubble constant $H_0$, this means
+\begin{equation}
+    |\kappa| = \frac{\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{\rm PV}^{-1}\,.
+\end{equation}
+\citet{Wang_2021} quote a constraint of $M_{\rm PV} > 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into $|\kappa| <$ \variable{output/kappa_Wang.txt}.
 We obtained a tighter 90\% upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_90.txt}.
 
-Our result is an order of magnitude improvement in constraining $\kappa$.
-The main reason is that we included more events from GWTC-3 compared to \citet{Okounkova_2022} and \citet{Wang_2021}.
+\subsection{Parameter degeneracies}
 
-\subsection{Parameter degeneracies}
+In examining our results in Sec.~\ref{sec:results:notable}, we identified interactions between birefringence and the inference of source properties like the component spin parameters---in addition to expected partial degeneracies with inclination and distance (Sec.~\ref{sec:inclination}).
+In particular, we found that the frequency-dependent modification to the waveform induced by birefringence can effectively mimic spin effects, including precession, especially for heavier systems which are observed over short durations.
 
@@
 \end{figure}
-
-In examining our results in Sec.~\ref{sec:results:notable}, we identified interactions between birefringence and the inference of source properties like the component spin parameters---in addition to expected partial degeneracies with inclination and distance (Sec.~\ref{sec:inclination}).
-In particular, we found that the frequency-dependent modification to the waveform induced by birefringence can effectively mimic spin effects, including precession, especially for heavier systems which are observed over short durations.
 
@@
 
-% \subsection{Extended corner plot for GW191105}
-% \label{sec:corner_GW191105_appendix}
-% 
-% \begin{figure*}[h]
-%     \script{corner_GW191105_appendix.py}
-%     \includegraphics[width=\textwidth]{figures/corner_GW191105_appendix.pdf}
-%     \caption{
-%         Extended corner plot for GW191105: a supplement to Fig.~\ref{fig:corner_GW191105} discussed in Appendix~\ref{sec:corner_GW191105_appendix}.
-%     }
-%     \label{fig:corner_GW191105_appendix}
-% \end{figure*}
-
-% \section{Extended corner plot for GW200129}
-% \label{sec:corner_GW200129_appendix}
-% 
-% \begin{figure*}[h]
-%     \script{corner_GW200129_appendix.py}
-%     \includegraphics[width=\textwidth]{figures/corner_GW200129_appendix.pdf}
-%     \caption{
-%         Extended corner plot for GW200129: a supplement to Fig.~\ref{fig:corner_GW200129} discussed in Appendix~\ref{sec:corner_GW200129_appendix}.
-%     }
-%     \label{fig:corner_GW200129_appendix}
-% \end{figure*}
-
-\section{Relation between $\kappa$ and $M_{PV}$}
-\label{sec:M_PV_derivation}
-
-\citet{Wang_2021} formulated the birefringence effect as
-\begin{equation}
-    h_{+/\times}^{PV}(f) = h_{+/\times}^{GR}(f)\mp h_{\times/+}^{GR}(f)(i\delta h-\delta\Psi)\,,
-\end{equation}
-where $h_{+/\times}^{PV}(f)$ is the modified \ac{GW} waveform in the linear basis, $h_{+/\times}^{GR}(f)$ is the \ac{GW} waveform in the linear basis as \ac{GR} predicts, $\delta h$ is the amplitude modification, and $\delta\Psi$ is the phase modification.
-We can rewrite the equation as
-\begin{equation}
-\begin{split}
-    h_{L/R}^{PV}(f)&=h_{L/R}^{GR}(f)(1\mp\delta h\mp i\delta\Psi)\\
-    &\approx h_{L/R}^{GR}(f)\exp\left(\mp\delta h\mp i\delta\Psi\right)
-\end{split}
-\,,
-\end{equation}
-where $h_{L/R}^{PV}(f)$ is the modified \ac{GW} waveform in the circular basis, $h_{L/R}^{GR}(f)$ is the \ac{GW} waveform in the circular basis as \ac{GR} predicts.
-This equation applies to any birefringence model that allows the assumption that the modification to \ac{GR} is small when observed very near the source.
-To make a comparison with our work, we can choose $\delta\Psi=0$, which is a case that \citet{Wang_2021} considered.
-In their work, they parametrized $\delta h=-A_\nu\pi f$, while
-\begin{equation}
-    A_\nu=M_{PV}^{-1}(\alpha_\nu(z=0)-\alpha_\nu(z)(1+z))\,,
-\end{equation}
-where $A_\nu$ is the parity-violating parameter for amplitude birefringence, $M_{PV}$ is the characteristic energy scale, and $\alpha_\nu$ is a arbitrary functions determined by the birefringence model.
-They then chose $\alpha_\nu=1$.
-Therefore, we can rewrite the equation as
-\begin{equation}
-    h_{L/R}^{PV}(f)= h_{L/R}^{GR}(f)\exp\left(\mp\frac{\pi zf}{M_{PV}}\right)\,.
-\end{equation}
-This equation takes the same form as Eq.~\eqref{eq:waveform_modification}.
-We can then find the relation between $M_{PV}$ and $\kappa$ by
-\begin{equation}
-    \left|\mp\frac{\pi zf}{M_{PV}}\right|=\left|\pm\kappa\frac{d_C}{1\,\mathrm{Gpc}}\frac{f}{100\,\mathrm{Hz}}\right|\,.
-\end{equation}
-Using $H_0d_C=cz$, where $H_0$ is the Hubble constant and $c$ is the speed of light, and adding back the factor of Planck constant $h$, we can rewrite the equation as
-\begin{equation}
-    |\kappa|=\frac{h\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{PV}^{-1}\,.
-\end{equation}
-They gave the constraint on $M_{PV}$ as $M_{PV}>10^{-22}\,\mathrm{GeV}$.
-This constraint only considered the amplitude birefringence modification.
-Equivalently, the constraint on $\kappa$ is $|\kappa| <$ \variable{output/kappa_Wang.txt}.
-
 \bibliography{bib}

### 2023-04-07 src/tex/ms.tex
conclusion

@@
 \begin{equation}
-    \left[ \tilde{\kappa} = \frac{\kappa}{1\, \mathrm{Gpc}} \right]_{\raisebox{1ex}{$\begin{subarray}{l} z=0.3 \\ f=100\,{\rm Hz} \end{subarray}$}}
+    \left[ \kappa = \tilde{\kappa}\left({1\, \mathrm{Gpc}}\right) \right]_{\raisebox{1ex}{$\begin{subarray}{l} z=0.3 \\ f=100\,{\rm Hz} \end{subarray}$}}
 \end{equation}
 per Eq.~\eqref{eq:waveform_modification}.
-In terms of this fiducial quantity, our constraint from Sec.~\ref{sec:results:gwtc} can be expressed as $|\tilde{\kappa}| <$ \variable{output/restricted_absolute_kappa_68.txt} ${\rm Gpc}^{-1}$ at 68\% credibility.
-This is a factor of ${\sim}25\times$ \mi{Thomas, can you make this number be dynamically generated?
-it should be the ratio of Masha's constraint to ours, rounded to the closest integer} more stringent than Ref.~\cite{Okounkova_2022}, besides arising from a less simplified model.
+In terms of this fiducial quantity, the constraint given by \citet{Okounkova_2022} can be expressed by $|{\kappa}| \lesssim 0.74$  at 68\% credibility.
+On the other hand, our constraint from Sec.~\ref{sec:results:gwtc} is $|\kappa| <$ \variable{output/restricted_absolute_kappa_68.txt} at 68\% credibility.
+This is a factor of \variable{output/improvement_Okounkova.txt} more stringent than Ref.~\cite{Okounkova_2022}, besides arising from a less simplified model.
 
@@
 \footnote{Concretely, \citet{Wang_2021} write $\delta h = - A_\nu \pi f$ with $A_\nu = M_{\rm PV}^{-1} \left[\alpha_\nu (z=0) - \alpha_\nu(z) \left(1+z\right)\right]$, for $\alpha_\nu$ some function of redshift encoding the evolution of a birefringence-mediating field; to produce their constraint, however, they further set $\alpha_\nu(z) = 1$, yielding $A_\nu = -z / M_{\rm PV}$ and hence $\delta h = \pi f z / M_{\rm PV}$.}
-Comparing to our Eq.~\eqref{eq:waveform_modification} and approximating $z \approx d_C H_0/c$ with the Hubble constant $H_0$, this means
+Comparing to our Eq.~\eqref{eq:waveform_modification} and approximating $z \approx d_C H_0/c$ via the Hubble constant, $H_0$, this means
 \begin{equation}
-    |\kappa| = \frac{\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{\rm PV}^{-1}\,.
+    |\kappa| \simeq \frac{\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{\rm PV}^{-1}\,.
 \end{equation}
-\citet{Wang_2021} quote a constraint of $M_{\rm PV} > 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into $|\kappa| <$ \variable{output/kappa_Wang.txt}.
+\citet{Wang_2021} quote a constraint of $M_{\rm PV} > 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into $|\kappa| \lesssim$ \variable{output/kappa_Wang.txt}.
 We obtained a tighter 90\% upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_90.txt}.
 
+\subsubsection{Comparison summary}
+
+We summarize the constraint on birefringence in Table~\ref{tab:comparison_summary}.
+The constraint we obtain is the most stringent, followed by \citet{Okounkova_2022} and \citet{Wang_2021}.
+
+\begin{table}
+    \caption{
+        Comparison of our constraints to \citet{Okounkova_2022} and \citet{Wang_2021}.
+        Note that the constraint from \citet{Okounkova_2022} is given at 68\% credibility, while the other two are given at 90\% credibility.
+        The constraint on $M_{PV}$ is a lower limit, while the constraint on $\tilde{\kappa}$ and $\kappa$ is an upper limit.
+    }
+    \begin{ruledtabular}
+        \variable{output/comparison_summary.txt}
+    \end{ruledtabular}
+    \label{tab:comparison_summary}
+\end{table}
+
 \subsection{Parameter degeneracies}
+\label{sec:degeneracies}
 
@@
 Besides spins, other parameters can interact with $\kappa$ either directly or through their own couplings with the spins.
-Since the mass ratio and can affect the phasing of a \ac{GW} inspiral at similar order as the spins \cite{Blanchet:2013haa}, we might expect some degree of interaction between $q$ and $\kappa$.
+Since the mass ratio can affect the phasing of a \ac{GW} inspiral at similar order as the spins \cite{Blanchet:2013haa}, we might expect some degree of interaction between $q$ and $\kappa$.
 This seems to indeed be the case for some events, like GW170104 or GW191105, as we demonstrate in Fig.~\ref{fig:mass_ratio}.
@@
 The coupling between $q$ and $\kappa$ is likely limited by the fact that changes in the mass ratio alone cannot produce significant amplitude modulations.
+
+Since our inference on $\kappa$ can be tied to precession and the spin orientations, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
 
@@
 The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
-Future catalogs, with more events, will shed light on this feature.
+Future catalogs with more events will shed light on this feature.
 
 From the set of results from individual events, we highlighted two, GW170818 and GW190521, whose posterior manifests interactions between birefringence and our inference of source parameters (Sec.~\ref{sec:results:notable}).
-In particular, we identified the relevance of spins their (partial) degeneracy with $\kappa$, in addition to expected correlations with source inclination and distance.
+In particular, we identified the relevance of spins and their (partial) degeneracy with $\kappa$, in addition to expected correlations with source inclination and distance.
+The mass ratio can also play a role, by coupling with the spins or $\kappa$ directly (Sec.~\ref{sec:degeneracies}).
 
-\mi{WIP}
-If our inference on $\kappa$ is tied to precession and the spin orientations, we might be susceptible to systematics in the modeling of spin angles in \textsc{IMRPhenomXPHM} and might thus benefit from further analysis with other waveforms like \textsc{NRSur7dq4} \cite{Varma:2018mmi}.
-
-% BNS
-Future work is to perform \ac{PE} on binary neutron star mergers, such as GW170817.
-The frequency range of the signal is much wider compared to the binary black hole mergers.
-Thus, the difference in the effect of birefringence at different frequencies within the range can be more significant.
-The \ac{PE} on binary neutron star mergers can allow us to further constrain the birefringence effect and the beyond-GR theories that predict it.
-However, performing \ac{PE} on binary neutron star mergers requires much more computational resources.
-Therefore, we may need to wait for future \ac{PE} methods and tools to further our work.
+This study was restricted to \acp{BBH} because of the expectation that they should dominate the birefringence constraint thanks to their larger redshifts.
+However, lower mass systems involving neutron stars may also be informative thanks to the wide band of frequencies spanned by their signals, in spite of their closer distances.
+Analysis of future catalogs will benefit from the inclusion of those events, as well as a much larger number of \ac{BBH} sources at increasingly greater distances.
+Future measurements will also be enriched by the addition of KAGRA \cite{KAGRA} and LIGO India to the \ac{GW} detector network, which will allow us to better disambiguate between polarization states and hence between models of birefringent propagation.
 

### 2023-04-07 src/tex/ms.tex
Add comparison table

--- 2023-04-07 src/tex/ms.tex
@@
 We obtained a tighter 90\% upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_90.txt}.
+
+\subsubsection{Comparison summary}
+
+We summarize the constraint on birefringence in Table~\ref{tab:comparison_summary}.
+The constraint we obtain is the most stringent, followed by \citet{Okounkova_2022} and \citet{Wang_2021}.
+
+\begin{table}
+    \caption{
+        Comparison of our constraints to \citet{Okounkova_2022} and \citet{Wang_2021}.
+        Note that the constraint from \citet{Okounkova_2022} is given at 68\% credibility, while the other two are given at 90\% credibility.
+        The constraint on $M_{PV}$ is a lower limit, while the constraint on $\tilde{\kappa}$ and $\kappa$ is an upper limit.
+    }
+    \begin{ruledtabular}
+        \variable{output/comparison_summary.txt}
+    \end{ruledtabular}
+    \label{tab:comparison_summary}
+\end{table}
 
@@
 From the set of results from individual events, we highlighted two, GW170818 and GW190521, whose posterior manifests interactions between birefringence and our inference of source parameters (Sec.~\ref{sec:results:notable}).
-In particular, we identified the relevance of spins their (partial) degeneracy with $\kappa$, in addition to expected correlations with source inclination and distance.
+In particular, we identified the relevance of spins and their (partial) degeneracy with $\kappa$, in addition to expected correlations with source inclination and distance.
 

### 2023-04-07 src/scripts/corner_GW190521.py

@@ -49,10 +49,10 @@ def kdeplot1d(x, **kws):
     plt.plot(df['x'], df['y'], lw=lw)
     plt.xlim(df['x'].min(), df['x'].max())
     current_ymax = plt.ylim()[1]
-    if current_ymax > df['y'].max()*1.1:
+    if current_ymax > df['y'].max()*1.05:
         plt.ylim(0,current_ymax)
     else:
-        plt.ylim(0,df['y'].max()*1.1)
+        plt.ylim(0,df['y'].max()*1.05)
 
 vars = ['kappa', 'luminosity_distance', 'cos_iota']
 g = sns.PairGrid(data=result,

### 2023-04-07 src/scripts/corner_GW200129.py

@@ -58,10 +58,10 @@ def kdeplot1d(x, mask=None, **kws):
     plt.plot(df['x'], df['y'], lw=lw, **kws)
     plt.xlim(df['x'].min(), df['x'].max())
     current_ymax = plt.ylim()[1]
-    if current_ymax > df['y'].max()*1.1:
+    if current_ymax > df['y'].max()*1.05:
         plt.ylim(0,current_ymax)
     else:
-        plt.ylim(0,df['y'].max()*1.1)
+        plt.ylim(0,df['y'].max()*1.05)
 
 p = list(sns.color_palette())[:3] + ['0.35']
 vars = ['kappa', 'luminosity_distance', 'cos_iota']

### 2023-04-07 src/scripts/kappa_constraint_Wang.py

@@ -1,13 +1,0 @@
-import numpy as np
-import paths
-
-h = 4.135667696e-15 # eV s
-c = 299792.458 # km/s
-H_0 = 68.3 # km/s/Mpc
-
-M_PV_lower_limit = 1e-22*1e9 # eV
-
-kappa_Wang = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(M_PV_lower_limit)
-
-with open(paths.output/"kappa_Wang.txt", "w") as f:
-    f.write(f"${kappa_Wang:.2f}$")

### 2023-04-07 src/scripts/kappa_constraint_restricted.py

@@ -3,6 +3,11 @@ import pandas as pd
 import scipy
 import paths
 
+# constants
+h = 4.135667696e-15 # eV s
+c = 299792.458 # km/s
+H_0 = 68.3 # km/s/Mpc
+
 result_dict = {}
 result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.feather")
 events = result_DataFrame['event'].unique()
@@ -35,6 +40,36 @@ restricted_cdf_absolute_kappa = np.array([np.trapz(likelihood_absolute_kappa[0:i
 restricted_absolute_kappa_68 = np.interp(0.68,restricted_cdf_absolute_kappa,absolute_kappa)
 with open(paths.output/"restricted_absolute_kappa_68.txt", "w") as f:
     f.write(f"${restricted_absolute_kappa_68:.2f}$")
+
+improvement_Okounkova = 0.74/restricted_absolute_kappa_68
+with open(paths.output/"improvement_Okounkova.txt", "w") as f:
+    f.write(rf"${{\sim}}{int(improvement_Okounkova)}\times$")
+
 restricted_absolute_kappa_90 = np.interp(0.9,restricted_cdf_absolute_kappa,absolute_kappa)
 with open(paths.output/"restricted_absolute_kappa_90.txt", "w") as f:
     f.write(f"${restricted_absolute_kappa_90:.2f}$")
+
+# Wang's constraint in kappa
+M_PV_Wang = 1e-22 # GeV
+kappa_Wang = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(M_PV_Wang*1e9)
+with open(paths.output/"kappa_Wang.txt", "w") as f:
+    f.write(f"${kappa_Wang:.2f}$")
+
+# Constraint comparison
+tilde_kappa_Okounkova = 0.74 # Gpc^-1
+kappa_this_work = restricted_absolute_kappa_90
+
+tilde_kappa_Wang = kappa_Wang
+
+kappa_Okounkova = tilde_kappa_Okounkova
+M_PV_Okounkova = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(kappa_Okounkova)/1e9
+
+tilde_kappa_this_work = kappa_this_work
[…]

### 2023-04-07 src/tex/ms.tex
abstract and intro

--- 2023-04-07 src/tex/ms.tex
@@
     Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
-    From 71 binary-black-hole signals, we obtain the most precise constraints on amplitude birefringence yet, an order of magnitude more stringent than previous results.
+    From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent opacity of $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility, equivalent to an energy scale of $M_{\rm PV} \lesssim XX\, {\rm GeV}$ \mi{Thomas please provide the equivalent $M_{\rm PV}$ constraint from $|\kappa|$)}.
 \end{abstract}
@@
 In Sec.~\ref{sec:Background}, we briefly review the background of \ac{GW} amplitude birefringence.
-In Sec.~\ref{sec:Method}, we describe the modification we made to the waveform model, mention the configuration we used in the \ac{PE} and show the method we used to obtain the population constraint on \ac{GW} amplitude birefringence.
-In Sec.~\ref{sec:Results}, we present the population constraint on \ac{GW} amplitude birefringence we obtained and show the results of individual \ac{PE}.
-In Sec.~\ref{sec:Discussion}, we discuss the limitation of this study and provide suggestions for future studies.
+In Sec.~\ref{sec:Method}, we describe the modification we apply to the baseline \ac{GR} waveform, summarize our \ac{PE} configuration and outline our statistical methods for combining information from multiple events.
+In Sec.~\ref{sec:Results}, we present our constraint on \ac{GW} amplitude birefringence, discussing individual events and the catalog collectively.
+In Sec.~\ref{sec:Discussion}, we discuss the implications of our result, comparing to previous constraints in the literature and outlining correlation structures that appear in our measurements.
+Finally, we conclude in Sec.~\ref{sec:Discussion} with a summary.
 
@@
 Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
-\mi{specify what the parameter for the expansion is, add citations, and check with a theorist}
 

### 2023-04-07 src/tex/ms.tex
Modify variables

@@
     Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
-    From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent opacity of $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility, equivalent to an energy scale of $M_{\rm PV} \lesssim XX\, {\rm GeV}$ \mi{Thomas please provide the equivalent $M_{\rm PV}$ constraint from $|\kappa|$)}.
+    From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent opacity of \variable{output/restricted_kappa_median.txt}with 90\% credibility, equivalent to an energy scale of \variable{output/M_PV_constraint.txt}.
 \end{abstract}
@@
 
-Judging by $\mu_i/\sigma_i$, the two events that show the largest tension with $\kappa = 0$ are GW170818, for which $\mu_i / \sigma_i =$ \variable{output/GW170818_constraint.txt}, and GW200129\_065458 (henceforth GW200129), for which $\mu_i / \sigma_i =$ \variable{output/GW200129_constraint.txt}.
+Judging by $\mu_i/\sigma_i$, the two events that show the largest tension with $\kappa = 0$ are GW170818, for which \variable{output/GW170818_constraint.txt}, and GW200129\_065458 (henceforth GW200129), for which \variable{output/GW200129_constraint.txt}.
 However, as we discuss in Sec.~\ref{sec:GW200129}, we have reason to think that the preference for $\kappa < 0$ in GW200129 might be driven by noise anomalies in the Virgo detector; with that in mind, in the next section we consider the effect of excluding this event from the joint result (we find its impact to be minimal).
@@
 
-Both versions of the posterior support the lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals of the result including all events, we constrain $\mu =$ \variable{output/mu_median.txt}for 90\%-credible symmetric intervals around the median, and $\sigma <$ \variable{output/sigma_median.txt}for the 90\%-credible one-sided upper limit.
+Both versions of the posterior support the lack of birefringence ($\mu = \sigma = 0$) within 90\% credibility; from the marginals of the result including all events, we constrain \variable{output/mu_median.txt}for 90\%-credible symmetric intervals around the median, and \variable{output/sigma_median.txt}for the 90\%-credible one-sided upper limit.
 However, even though $\sigma = 0$ is well supported, the $\sigma$ posterior peaks visibly away from the origin, indicating some preference for a nonzero variance.
@@
 We show the result of doing this in Fig.~\ref{fig:posterior_kappa}, where we also compare to the posterior on $\kappa$ obtained by assuming a shared value across events (same result shown as a blue band in Fig.~\ref{fig:violin_kappa}).
-The hierarchical measurement leads to an expectation that $\kappa_i =$ \variable{output/generic_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies $\kappa =$ \variable{output/restricted_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
+The hierarchical measurement leads to an expectation that \variable{output/generic_kappa_median.txt}, whereas the shared-$\kappa$ measurement implies \variable{output/restricted_kappa_median.txt}, both symmetric 90\%-credible intervals around the median.
 
@@
 
-Our primary constraint on birefringence is given by the combined constraint on the birefringent opacity from GWTC-3, $\kappa =$ \variable{output/restricted_kappa_median.txt} with 90\% credibility, obtained in Sec.~\ref{sec:results:gwtc} (blue band in Fig.~\ref{fig:violin_kappa}).
+Our primary constraint on birefringence is given by the combined constraint on the birefringent opacity from GWTC-3, \variable{output/restricted_kappa_median.txt}with 90\% credibility, obtained in Sec.~\ref{sec:results:gwtc} (blue band in Fig.~\ref{fig:violin_kappa}).
 We can compare this to previous measurements of amplitude birefringence in the literature.
@@
 In terms of this fiducial quantity, the constraint given by \citet{Okounkova_2022} can be expressed by $|{\kappa}| \lesssim 0.74$  at 68\% credibility.
-On the other hand, our constraint from Sec.~\ref{sec:results:gwtc} is $|\kappa| <$ \variable{output/restricted_absolute_kappa_68.txt} at 68\% credibility.
-This is a factor of \variable{output/improvement_Okounkova.txt} more stringent than Ref.~\cite{Okounkova_2022}, besides arising from a less simplified model.
+On the other hand, our constraint from Sec.~\ref{sec:results:gwtc} is \variable{output/restricted_absolute_kappa_68.txt}at 68\% credibility.
+This is a factor of \variable{output/improvement_Okounkova.txt}more stringent than Ref.~\cite{Okounkova_2022}, besides arising from a less simplified model.
 
@@
 \end{equation}
-\citet{Wang_2021} quote a constraint of $M_{\rm PV} > 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into $|\kappa| \lesssim$ \variable{output/kappa_Wang.txt}.
-We obtained a tighter 90\% upper limit of $|\kappa| <$ \variable{output/restricted_absolute_kappa_90.txt}.
+\citet{Wang_2021} quote a constraint of $M_{\rm PV} > 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Wang.txt}.
+We obtained a tighter 90\% upper limit of \variable{output/restricted_absolute_kappa_90.txt}.
 
@@
 
-We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained $\kappa =$ \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
+We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
 As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).

### 2023-04-07 src/scripts/GW170818_constraint.py

@@ -6,4 +6,4 @@ result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.f
 mean_over_std_GW170818 = result_DataFrame[result_DataFrame.event == "GW170818"]['kappa'].mean() / result_DataFrame[result_DataFrame.event == "GW170818"]['kappa'].std()
 
 with open(paths.output/"GW170818_constraint.txt", "w") as f:
-    f.write(rf"${mean_over_std_GW170818:.1f}$")
+    f.write(rf"$\mu_i / \sigma_i = {mean_over_std_GW170818:.1f}$")

### 2023-04-07 src/scripts/GW200129_constraint.py

@@ -6,4 +6,4 @@ result_DataFrame = pd.read_feather(paths.data/"samples_posterior_birefringence.f
 mean_over_std_GW200129 = result_DataFrame[result_DataFrame.event == "GW200129_065458"]['kappa'].mean() / result_DataFrame[result_DataFrame.event == "GW200129_065458"]['kappa'].std()
 
 with open(paths.output/"GW200129_constraint.txt", "w") as f:
-    f.write(rf"${mean_over_std_GW200129:.1f}$")
+    f.write(rf"$\mu_i / \sigma_i = {mean_over_std_GW200129:.1f}$")

### 2023-04-07 src/scripts/kappa_constraint_generic.py

@@ -28,4 +28,4 @@ generic_kappa_median = np.median(samples_reweighted)
 generic_kappa_95 = np.percentile(a=samples_reweighted,q=95)
 generic_kappa_5 = np.percentile(a=samples_reweighted,q=5)
 with open(paths.output/"generic_kappa_median.txt", "w") as f:
-    f.write(f"${generic_kappa_median:.3f}^{{+{(generic_kappa_95-generic_kappa_median):.3f}}}_{{{(generic_kappa_5-generic_kappa_median):.3f}}}$")
+    f.write(f"$\kappa_i = {generic_kappa_median:.3f}^{{+{(generic_kappa_95-generic_kappa_median):.3f}}}_{{{(generic_kappa_5-generic_kappa_median):.3f}}}$")

### 2023-04-07 src/scripts/kappa_constraint_restricted.py

@@ -28,7 +28,7 @@ restricted_kappa_median = np.interp(0.5,restricted_cdf,kappa)
 restricted_kappa_95 = np.interp(0.95,restricted_cdf,kappa)
 restricted_kappa_5 = np.interp(0.05,restricted_cdf,kappa)
 with open(paths.output/"restricted_kappa_median.txt", "w") as f:
-    f.write(f"${restricted_kappa_median:.3f}^{{+{(restricted_kappa_95-restricted_kappa_median):.3f}}}_{{{(restricted_kappa_5-restricted_kappa_median):.3f}}}$")
+    f.write(f"$\kappa = {restricted_kappa_median:.3f}^{{+{(restricted_kappa_95-restricted_kappa_median):.3f}}}_{{{(restricted_kappa_5-restricted_kappa_median):.3f}}}$")
 
 credible_level = len(likelihood[likelihood > np.interp(0, kappa, likelihood)])/len(likelihood)
 with open(paths.output/"CL_kappa_0.txt", "w") as f:
@@ -39,7 +39,7 @@ likelihood_absolute_kappa = likelihood[len(kappa)//2:]+likelihood[0:len(kappa)//
 restricted_cdf_absolute_kappa = np.array([np.trapz(likelihood_absolute_kappa[0:i],absolute_kappa[0:i]) for i in range(len(absolute_kappa))])
 restricted_absolute_kappa_68 = np.interp(0.68,restricted_cdf_absolute_kappa,absolute_kappa)
 with open(paths.output/"restricted_absolute_kappa_68.txt", "w") as f:
-    f.write(f"${restricted_absolute_kappa_68:.2f}$")
+    f.write(f"$|\kappa| < {restricted_absolute_kappa_68:.2f}$")
 
 improvement_Okounkova = 0.74/restricted_absolute_kappa_68
 with open(paths.output/"improvement_Okounkova.txt", "w") as f:
@@ -47,28 +47,30 @@ with open(paths.output/"improvement_Okounkova.txt", "w") as f:
 
 restricted_absolute_kappa_90 = np.interp(0.9,restricted_cdf_absolute_kappa,absolute_kappa)
[…]

### 2023-04-07 src/tex/ms.tex
tiny edits

@@
 We can compare this to previous measurements of amplitude birefringence in the literature.
+We summarize these constraints in Table~\ref{tab:comparison_summary}.
+
+\begin{table}
+    \caption{
+        Comparison of our constraints to \citet{Okounkova_2022} and \citet{Wang_2021}.
+        The constraint from \citet{Okounkova_2022} is given at 68\% credibility, while the other two are given at 90\% credibility.
+        The constraint on $M_{PV}$ is a lower limit, while the constraint on $\tilde{\kappa}$ and $\kappa$ is an upper limit.
+    }
+    \begin{ruledtabular}
+        \variable{output/comparison_summary.txt}
+    \end{ruledtabular}
+    \label{tab:comparison_summary}
+\end{table}
 
@@
 \begin{equation}
-    \left[ \kappa = \tilde{\kappa}\left({1\, \mathrm{Gpc}}\right) \right]_{\raisebox{1ex}{$\begin{subarray}{l} z=0.3 \\ f=100\,{\rm Hz} \end{subarray}$}}
+    \left[ \tilde{\kappa} = \frac{\kappa}{1\, \mathrm{Gpc}} \right]_{\raisebox{1ex}{$\begin{subarray}{l} z=0.3 \\ f=100\,{\rm Hz} \end{subarray}$}}
 \end{equation}
@@
 We obtained a tighter 90\% upper limit of \variable{output/restricted_absolute_kappa_90.txt}.
-
-\subsubsection{Comparison summary}
-
-We summarize the constraint on birefringence in Table~\ref{tab:comparison_summary}.
-The constraint we obtain is the most stringent, followed by \citet{Okounkova_2022} and \citet{Wang_2021}.
-
-\begin{table}
-    \caption{
-        Comparison of our constraints to \citet{Okounkova_2022} and \citet{Wang_2021}.
-        Note that the constraint from \citet{Okounkova_2022} is given at 68\% credibility, while the other two are given at 90\% credibility.
-        The constraint on $M_{PV}$ is a lower limit, while the constraint on $\tilde{\kappa}$ and $\kappa$ is an upper limit.
-    }
-    \begin{ruledtabular}
-        \variable{output/comparison_summary.txt}
-    \end{ruledtabular}
-    \label{tab:comparison_summary}
-\end{table}
 
@@
 We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
+This measurement is signficantly more stringent than past constraints (Table \ref{tab:comparison_summary}).
 As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).
@@
 
+All our data are available in \citet{dataset}.
+
 \begin{acknowledgments}
@@
 The Flatiron Institute provided the computational resources used in this work.
-
 This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gwosc.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
@@
 KAGRA is supported by Ministry of Education, Culture, Sports, Science and Technology (MEXT), Japan Society for the Promotion of Science (JSPS) in Japan; National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea; Academia Sinica (AS) and National Science and Technology Council (NSTC) in Taiwan.
+This paper was compiled using \textsc{showyourwork} \cite{Luger2021} to facilitate reproducibility.
 \end{acknowledgments}

### 2023-04-07 src/tex/ms.tex
acks

@@
 \begin{acknowledgments}
+We thank Macarena Lagos and Nicol\'as Yunes for helpful discussions.
 M.~I., K.~W.~K.~W.~, and W.~M.~F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute.

### 2023-04-07 src/tex/ms.tex
zhao citation

@@
 \citet{Wang_2021} performed \ac{PE} on GWTC-1 events with a frequency-dependent model of birefringence resembling ours.
-In contrast to our Eq.~\eqref{eq:waveform_modification}, this reference parameterized the birefringent waveform in terms of some amplitude and phase modifications to the linear polarizations ($\delta h$ and $\delta \Psi$ respectively), such that
+Following \cite{Zhao:2019xmm}, that reference parameterized the birefringent waveform in terms of some amplitude and phase modifications to the linear polarizations ($\delta h$ and $\delta \Psi$ respectively), such that
 \begin{equation}

### 2023-04-07 src/tex/ms.tex
Modify Table 3

@@
         Comparison of our constraints to \citet{Okounkova_2022} and \citet{Wang_2021}.
-        The constraint from \citet{Okounkova_2022} is given at 68\% credibility, while the other two are given at 90\% credibility.
-        The constraint on $M_{PV}$ is a lower limit, while the constraint on $\tilde{\kappa}$ and $\kappa$ is an upper limit.
+        CL is the percentage of the one-sided credible limit of each work.
     }

### 2023-04-07 src/scripts/kappa_constraint_restricted.py

@@ -56,22 +56,18 @@ with open(paths.output/"kappa_Wang.txt", "w") as f:
     f.write(f"$|\kappa| \lesssim {kappa_Wang:.2f}$")
 
 # Constraint comparison
-tilde_kappa_Wang = kappa_Wang
-
-tilde_kappa_Okounkova = 0.74 # Gpc^-1
-kappa_Okounkova = tilde_kappa_Okounkova
+kappa_Okounkova = 0.74
 M_PV_Okounkova = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(kappa_Okounkova)/1e9
 
 kappa_this_work = restricted_absolute_kappa_90
-tilde_kappa_this_work = kappa_this_work
 M_PV_this_work = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(kappa_this_work)/1e9
 
 with open(paths.output/"M_PV_constraint.txt", "w") as f:
-    f.write(rf"$M_{{\rm PV}} \gtrsim {M_PV_this_work/1e-21:.2f} \times 10^{{-21}}\, {{\rm GeV}}$")
+    f.write(rf"$M_{{\rm PV}} \gtrsim {M_PV_this_work/1e-21:.1f} \times 10^{{-21}}\, {{\rm GeV}}$")
 
 with open(paths.output/"comparison_summary.txt", "w") as f:
-    f.write(r"\begin{tabular}{lccc} & $M_{\rm PV}$ ($10^{-21}\, {\rm GeV}$) & $\tilde{\kappa}$ (${\rm Gpc}^{-1}$) & $\kappa$ \\ \hline ")
-    f.write(rf"\citet{{Wang_2021}} & ${M_PV_Wang/1e-21:.2f}$ & ${tilde_kappa_Wang:.2f}$ & ${kappa_Wang:.2f}$ \\ ")
-    f.write(rf"\citet{{Okounkova_2022}} & ${M_PV_Okounkova/1e-21:.2f}$ & ${tilde_kappa_Okounkova:.2f}$ & ${kappa_Okounkova:.2f}$ \\ ")
-    f.write(rf"This work & ${M_PV_this_work/1e-21:.2f}$ & ${tilde_kappa_this_work:.2f}$ & ${kappa_this_work:.2f}$ ")
[…]

### 2023-04-07 src/scripts/posterior_kappa.py
fill Fig 5

@@ -5,6 +5,8 @@ import matplotlib.pyplot as plt
 import scipy
 import paths
 
+rng = np.random.default_rng(12345)
+
 sns.set_theme(palette='colorblind', font_scale=1.5)
 
 plt.rcParams.update({
@@ -31,10 +33,10 @@ df_Gaussian_samples_all['sigma'] = samples_all[:,1]
 
 samples_reweighted = []
 for _ in range(1000):
-    Gaussian = np.full(shape=(len(result_DataFrame), 2), fill_value=df_Gaussian_samples_all.sample(1))
+    Gaussian = np.full(shape=(len(result_DataFrame), 2), fill_value=df_Gaussian_samples_all.sample(1, random_state=rng))
     weight = normal_distribution(result_DataFrame['kappa'], Gaussian[:,0], Gaussian[:,1])
     weight /= weight.sum()
-    samples_reweighted.append(np.random.choice(a=np.array(result_DataFrame['kappa']), size=1, p=weight))
+    samples_reweighted.append(rng.choice(a=np.array(result_DataFrame['kappa']), size=1, p=weight))
 
 events = result_DataFrame['event'].unique()
 kernels = [scipy.stats.gaussian_kde(result_dict[event]['kappa']) for event in events]
@@ -51,6 +53,9 @@ sns.kdeplot(np.array(samples_reweighted).reshape(-1), label="generic", color=sns
 # plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=5), color=sns.color_palette()[3], linestyle=':')
 # plt.axvline(np.percentile(a=np.array(samples_reweighted).reshape(-1),q=95), color=sns.color_palette()[3], linestyle=':')
 plt.plot(kappa,likelihood, label="restricted", color=sns.color_palette()[0])
+yl = plt.gca().get_ylim()
+plt.fill_between(kappa, likelihood, color=sns.color_palette()[0], alpha=0.2)
+plt.ylim(yl)
 # plt.axvline(np.interp(0.05,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[0], linestyle=':')
 # plt.axvline(np.interp(0.95,[np.trapz(likelihood[0:i],kappa[0:i]) for i in range(1000)],kappa), color=sns.color_palette()[0], linestyle=':')
 

### 2023-04-07 src/tex/ms.tex
fill Fig 5
fig 5

@@
     \caption{
-        The generic and restricted posterior of $\kappa$.
-        The blue solid and dashed lines show the restricted posterior of $\kappa$ with and without GW200129, respectively.
-        The orange solid and dashed lines show the generic posterior of $\kappa$ with and without GW200129, respectively.
+        Restricted and generic posteriors on $\kappa$, which respectively do and do not assume that $\kappa$ is shared by all events (color).
+        Solid and dashed traces indicate whether the analysis included GW200129 or not.
         The black dashed line marks the absence of birefringence ($\kappa=0$).
+        The shaded distribution is the primary result in this work (blue band in Fig.~\ref{fig:violin_kappa}).
     }

### 2023-04-13 src/tex/ms.tex
typo fixed
Put proper units on Kappa, call it attenuation (coefficient).

@@
     Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
-    From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent opacity of \variable{output/restricted_kappa_median.txt}with 90\% credibility, equivalent to an energy scale of \variable{output/M_PV_constraint.txt}.
+    From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent attenuation of \variable{output/restricted_kappa_median.txt} at $100 \, \mathrm{Hz}$ with 90\% credibility, equivalent to a parity-violation energy scale of \variable{output/M_PV_constraint.txt}.
 \end{abstract}
@@
     h_{L/R}^{\mathrm{GR}}(f) \times
-    \exp\left(\pm\kappa\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\,\mathrm{Hz}}\right)\,,
+    \exp\left(\pm\kappa d_C \frac{f}{100\,\mathrm{Hz}}\right)\,,
     \label{eq:waveform_modification}
@@
 where the emitted waveform $h_{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_{L/R}^{\mathrm{br}}$.
-The overall magnitude of this effect for a given distance and frequency $f$ is set by a dimensionless opacity parameter, $\kappa$, which encodes the intrinsic strength of the birefringence.
+The overall magnitude of this effect for a given frequency $f$ is set by an attenuation coefficient, $\kappa$, which encodes the intrinsic strength of the birefringence.
 The emitted waveform for a given source (i.e., the waveform observed in the near zone, very close to the source) will generally differ from the analogous waveform predicted by \ac{GR} \cite{Alexander:2009tp,Okounkova:2019zjf}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small (e.g., \cite{Okounkova:2022grv}), it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
@@
 \begin{equation}
-\left|\kappa\right| \left(d_C/1\,\mathrm{Gpc}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, .
+\left|\kappa\right| d_C \left(f/100\, \mathrm{Hz}\right) < 1\, .
 \end{equation}
@@
     \frac{h_{L}^\mathrm{br}}{h_{R}^\mathrm{br}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2
-    \exp\left({2\kappa\frac{d_C}{1\, \mathrm{Gpc}}\frac{f}{100\, \mathrm{Hz}}}\right)\, .
+    \exp\left(2\kappa d_C \frac{f}{100\, \mathrm{Hz}}\right)\, .
     \label{eq:modified_amplitude_ratio}
@@
 
-Our primary constraint on birefringence is given by the combined constraint on the birefringent opacity from GWTC-3, \variable{output/restricted_kappa_median.txt}with 90\% credibility, obtained in Sec.~\ref{sec:results:gwtc} (blue band in Fig.~\ref{fig:violin_kappa}).
+Our primary constraint on birefringence is given by the combined constraint on the birefringent attenuation from GWTC-3, \variable{output/restricted_kappa_median.txt}with 90\% credibility, obtained in Sec.~\ref{sec:results:gwtc} (blue band in Fig.~\ref{fig:violin_kappa}).
 We can compare this to previous measurements of amplitude birefringence in the literature.
@@
         Comparison of our constraints to \citet{Okounkova_2022} and \citet{Wang_2021}.
-        CL is the percentage of the one-sided credible limit of each work.
+        CL is the credible level.
     }
@@
 \end{equation}
-where $\tilde{\kappa}$ is a frequency-independent opacity (simply denoted ``$\kappa$'' in Ref.~\cite{Okounkova_2022}).
+where $\tilde{\kappa}$ is a frequency-independent attenuation (simply denoted ``$\kappa$'' in Ref.~\cite{Okounkova_2022}).
 Assuming a fiducial redshift of $z=0.3$ for all events in GWTC-2, \citet{Okounkova_2022} obtained a constraint of $|\tilde{\kappa}| \lesssim 0.74\, {\rm Gpc}^{-1}$ at 68\% credibility.
-Taking a reference frequency of $f=100\, \mathrm{Hz}$, which is representative of LIGO-Virgo events, and adopting the same fiducial distance $d_C(z=0.3)$, we may compare that constraint to ours by noting that
-\begin{equation}
-    \left[ \tilde{\kappa} = \frac{\kappa}{1\, \mathrm{Gpc}} \right]_{\raisebox{1ex}{$\begin{subarray}{l} z=0.3 \\ f=100\,{\rm Hz} \end{subarray}$}}
-\end{equation}
-per Eq.~\eqref{eq:waveform_modification}.
-In terms of this fiducial quantity, the constraint given by \citet{Okounkova_2022} can be expressed by $|{\kappa}| \lesssim 0.74$  at 68\% credibility.
-On the other hand, our constraint from Sec.~\ref{sec:results:gwtc} is \variable{output/restricted_absolute_kappa_68.txt}at 68\% credibility.
+On the other hand, our constraint from Sec.~\ref{sec:results:gwtc} is \variable{output/restricted_absolute_kappa_68.txt} at $100 \, \mathrm{Hz}$ at 68\% credibility.
 This is a factor of \variable{output/improvement_Okounkova.txt}more stringent than Ref.~\cite{Okounkova_2022}, besides arising from a less simplified model.
@@
 \end{align}
-assuming a small $\delta $ and $\delta \Psi \ll 1$.
+assuming a small $\delta h$ and $\delta \Psi$ in the last line.
 To consider only amplitude birefringence, as in this work, we must compare to the result in Ref.~\citep{Wang_2021} that set $\delta \Psi = 0$ and parameterized $\delta h = \pi f z / M_{\rm PV}$, where $M_{\rm PV}$ is the energy scale of the birefringent (parity-violating) correction in units of frequency.%
@@
 \begin{equation}
-    |\kappa| \simeq \frac{\pi H_0}{c}\left(1\,\mathrm{Gpc}\right)\left(100\,\mathrm{Hz}\right)M_{\rm PV}^{-1}\,.
+    |\kappa| \simeq \frac{\pi H_0}{c}\left(100\,\mathrm{Hz}\right)M_{\rm PV}^{-1}\,.
 \end{equation}
@@
 We have reanalyzed all \acp{BBH} in GWTC-3 with $\mathrm{FAR} < 1/\mathrm{yr}$ to constrain amplitude birefringence in the propagation of \acp{GW}.
-To this end, we implemented a model in which right or left handed polarizations are amplified or suppressed over distance as a function of frequency, with an overall strength parametrized by a birefringent ``opacity'' parameter, $\kappa$, following Eq.~\eqref{eq:waveform_modification}.
+To this end, we implemented a model in which right or left handed polarizations are amplified or suppressed over distance as a function of frequency, with an overall strength parametrized by a birefringent ``attenuation'' parameter, $\kappa$, following Eq.~\eqref{eq:waveform_modification}.
 This parameterization is consistent with parity-odd theories like Chern-Simons gravity, and can be used to constrain them where applicable.

### 2023-04-13 src/scripts/Gaussian_result.py

@@ -10,7 +10,7 @@ mu_median = np.median(df['mu'])
 sigma_median = np.median(df['sigma'])
 
 with open(paths.output/"mu_median.txt", "w") as f:
-    f.write(rf"$\mu = {mu_median:.3f}^{{+{(np.percentile(a=df['mu'],q=95)-mu_median):.3f}}}_{{{(np.percentile(a=df['mu'],q=5)-mu_median):.3f}}}$")
+    f.write(rf"$\mu = {mu_median:.3f}^{{+{(np.percentile(a=df['mu'],q=95)-mu_median):.3f}}}_{{{(np.percentile(a=df['mu'],q=5)-mu_median):.3f}}} \, \mathrm{{Gpc}}^{{-1}}$")
 
 with open(paths.output/"sigma_median.txt", "w") as f:
-    f.write(rf"$\sigma < {np.percentile(a=df['sigma'],q=90):.3f}$")
+    f.write(rf"$\sigma < {np.percentile(a=df['sigma'],q=90):.3f} \, \mathrm{{Gpc}}^{{-1}}$")

### 2023-04-13 src/scripts/best_events_kappa.py

@@ -22,7 +22,7 @@ for event in events:
 result_dict = {k: v for k, v in sorted(result_dict.items(), key=lambda item: item[1]['kappa_std'])}
 
 with open(paths.output/"best_events_kappa.txt", "w") as f:
-    f.write(r"\begin{tabular}{lccc}Event & $\kappa$ & $\sigma_i$ & CL\\ \hline ")
+    f.write(r"\begin{tabular}{lccc}Event & $\kappa$ ($\mathrm{Gpc}^{-1}$) & $\sigma_i$ ($\mathrm{Gpc}^{-1}$) & CL\\ \hline ")
     for i, event in enumerate(result_dict):
         result_dict[event]['event'] = event.replace("_", r"\_")
         

### 2023-04-13 src/scripts/corner_GW190521_appendix.py

@@ -76,7 +76,7 @@ for i in range(len(vars)):
         g.axes[i,j].set_xlim(result[vars[j]].min(), result[vars[j]].max())
         g.axes[i,j].set_ylim(result[vars[i]].min(), result[vars[i]].max())
 
-g.axes[10,0].set_xlabel("$\kappa$")
+g.axes[10,0].set_xlabel("$\kappa$ ($\mathrm{Gpc}$)")
 g.axes[1,0].set_ylabel("$d_L$ (Mpc)")
 g.axes[10,1].set_xlabel("$d_L$ (Mpc)")
 g.axes[2,0].set_ylabel("$\\cos\\iota$")

### 2023-04-13 src/scripts/corner_GW200129.py

@@ -88,7 +88,7 @@ for i in range(len(vars)):
         g.axes[i,j].set_xlim(result_all[vars[j]].min(), result_all[vars[j]].max())
         g.axes[i,j].set_ylim(result_all[vars[i]].min(), result_all[vars[i]].max())
 
-g.axes[2,0].set_xlabel(r"$\kappa$")
+g.axes[2,0].set_xlabel(r"$\kappa$ ($\mathrm{Gpc}^{-1}$)")
 g.axes[1,0].set_ylabel(r"$d_L$ (Mpc)")
 g.axes[2,1].set_xlabel(r"$d_L$ (Mpc)")
 g.axes[2,0].set_ylabel(r"$\cos\iota$")

### 2023-04-13 src/tex/ms.tex
Clean up translation of our limit into M_PV.

@@
 assuming a small $\delta h$ and $\delta \Psi$ in the last line.
-To consider only amplitude birefringence, as in this work, we must compare to the result in Ref.~\citep{Wang_2021} that set $\delta \Psi = 0$ and parameterized $\delta h = \pi f z / M_{\rm PV}$, where $M_{\rm PV}$ is the energy scale of the birefringent (parity-violating) correction in units of frequency.%
+To consider only amplitude birefringence, as in this work, we must compare to the result in Ref.~\citep{Wang_2021} that set $\delta \Psi = 0$ and parameterized $\delta h = \pi f z / M_{\rm PV}$, where $M_{\rm PV}$ is the energy scale of the birefringent (parity-violating) correction.%
 \footnote{Concretely, \citet{Wang_2021} write $\delta h = - A_\nu \pi f$ with $A_\nu = M_{\rm PV}^{-1} \left[\alpha_\nu (z=0) - \alpha_\nu(z) \left(1+z\right)\right]$, for $\alpha_\nu$ some function of redshift encoding the evolution of a birefringence-mediating field; to produce their constraint, however, they further set $\alpha_\nu(z) = 1$, yielding $A_\nu = -z / M_{\rm PV}$ and hence $\delta h = \pi f z / M_{\rm PV}$.}
@@
 \begin{equation}
-    |\kappa| \simeq \frac{\pi H_0}{c}\left(100\,\mathrm{Hz}\right)M_{\rm PV}^{-1}\,.
+    |\kappa| \simeq \frac{\pi H_0}{c} \frac{h \left( 100 \, \mathrm{Hz} \right)}{M_{\rm PV}} \simeq \frac{3.0 \times 10^{-22} \, \mathrm{GeV}}{M_\mathrm{PV}} \, \mathrm{Gpc}^{-1}
 \end{equation}
 \citet{Wang_2021} quote a constraint of $M_{\rm PV} > 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Wang.txt}.
-We obtained a tighter 90\% upper limit of \variable{output/restricted_absolute_kappa_90.txt}.
+We obtained a tighter 90\% upper limit of \variable{output/restricted_absolute_kappa_90.txt}, or \variable{output/M_PV_constraint.txt}.
 

### 2023-04-13 src/tex/ms.tex
Fix hierarchical modeling references.

@@
 By implementing Eq.~\eqref{eq:waveform_modification}, which is a first-order approximation of the birefringence model, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
-Nevertheless, there exist systems for which the degeneracy cannot be perfectly broken because not enough frequencies are available in the data.
+Nevertheless, there exist systems for which the degeneracy cannot be broken in practice because not enough frequencies are available in the data.
 This may be the case for quasimonochromatic sources, like nonaxisymmetric pulsars or very light binaries, which are well approximated by a single Fourier mode.
@@
 
-To do this, we apply hierarchical Bayesian inference \cite{Loredo:2004nn} to model the distribution of $\kappa$'s consistent with the observed data:
+To do this, we apply hierarchical Bayesian inference \cite{Mandel2010,Hogg2010} to model the distribution of $\kappa$'s consistent with the observed data:
 we posit that, rather than a unique global value of $\kappa$, there is a specific value of the parameter, $\kappa_i$, associated with each event, and that this is drawn from some unknown distribution of true underlying values; from the imperfect measurements of $\kappa_i$ for each event, we may reconstruct the underlying distribution.

### 2023-04-13 src/tex/ms.tex
Clean up figure labels.

@@
 This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Callister:T2100301}.
-Figure \ref{fig:corner_Gaussian} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its higher support for $\kappa > 0$.
+The result in Figure \ref{fig:reweighted_kappa} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its higher support for $\kappa > 0$.
 This feature appears to offset a few other events which tend to favor $\kappa < 0$.
@@
 
-Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:posterior_kappa}: GW170818, GW190521 and GW200129.
+Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:violin_kappa}: GW170818, GW190521 and GW200129.
 When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$, the third is the most extreme representative of a class of events with bimodal $\kappa$ posteriors, and the third shows signs of potential noise anomalies.

### 2023-04-13 src/tex/ms.tex
Minor text change

@@
 Compared to other studies, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
-We consider 71 binary black hole merger events with a \ac{FAR} $\leq1\mathrm{yr^{-1}}$, as listed in Table I of \citet{GWTC-3_population}.
+We consider 71 binary black hole merger events with a \ac{FAR} $\leq1/\mathrm{yr}$, as listed in Table I of \citet{GWTC-3_population}.
 We discuss single-event results in detail, and identify degeneracies between birefringence and spin effects, in addition to the already known correlations with source orientation and distance.
@@
 We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
-This measurement is signficantly more stringent than past constraints (Table \ref{tab:comparison_summary}).
+This measurement is significantly more stringent than past constraints (Table \ref{tab:comparison_summary}).
 As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).

### 2023-04-13 src/tex/ms.tex
Somehow my python bugs out with tex in the figure. Disabling them

@@
 
-\begin{figure}
-    \script{birefringence.py}
-    \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
-    \caption{
-        \emph{Illustration of amplitude birefringence.} The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
-        However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
-        In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
-        For this example, the black holes do not spin and have equal masses $m_1 = m_2 = 10\, M_\odot$, and we have chosen a luminosity distance $d_L = 400\, {\rm Mpc}$ and $\kappa = 0.6$.
-        }
-    \label{fig:birefringence}
-\end{figure}
+% \begin{figure}
+%     \script{birefringence.py}
+%     \includegraphics[width=\columnwidth]{figures/birefringence.pdf}
+%     \caption{
+%         \emph{Illustration of amplitude birefringence.} The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
+%         However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
+%         In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
+%         For this example, the black holes do not spin and have equal masses $m_1 = m_2 = 10\, M_\odot$, and we have chosen a luminosity distance $d_L = 400\, {\rm Mpc}$ and $\kappa = 0.6$.
+%         }
+%     \label{fig:birefringence}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \script{violin_kappa.py}
-    \includegraphics[width=\columnwidth]{figures/violin_kappa.pdf}
-    \caption{
-        Individual-event $\kappa$ posteriors (distributions), and joint measurement (blue band, 90\% CI; blue line, median).
-    }
-    \label{fig:violin_kappa}
-\end{figure}
+% \begin{figure}
+%     \script{violin_kappa.py}
+%     \includegraphics[width=\columnwidth]{figures/violin_kappa.pdf}
+%     \caption{
+%         Individual-event $\kappa$ posteriors (distributions), and joint measurement (blue band, 90\% CI; blue line, median).
+%     }
+%     \label{fig:violin_kappa}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \script{corner_Gaussian.py}
-    \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
-    \caption{
-        The posterior of the $\kappa$ population hyperparameters $\mu$ and $\sigma$, including (blue) and excluding (orange) GW200129 from the collection of events.
-        The 2D contours correspond to the $39.35\%$ and $90\%$ credible levels.
-        The plot shows that the population constraint on $\kappa$ is consistent with no birefringence ($\mu=\sigma=0$) at the 90\% credible level.
-    }
-    \label{fig:corner_Gaussian}
-\end{figure}
+% \begin{figure}
+%     \script{corner_Gaussian.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
+%     \caption{
+%         The posterior of the $\kappa$ population hyperparameters $\mu$ and $\sigma$, including (blue) and excluding (orange) GW200129 from the collection of events.
+%         The 2D contours correspond to the $39.35\%$ and $90\%$ credible levels.
+%         The plot shows that the population constraint on $\kappa$ is consistent with no birefringence ($\mu=\sigma=0$) at the 90\% credible level.
+%     }
+%     \label{fig:corner_Gaussian}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \script{reweighed_kappa.py}
-    \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
-    \caption{
-        Individual-event $\kappa_i$ distributions of Fig.~\ref{fig:violin_kappa} reweighted in light of the population-level inference of Fig.~\ref{fig:corner_Gaussian}, marginalizing over $\mu$ and $\sigma$.
-        Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i|$; we label the events most in tension with $\kappa_i = 0$ by that same measure.
-    }
-    \label{fig:reweighted_kappa}
-\end{figure}
+% \begin{figure}
+%     \script{reweighed_kappa.py}
+%     \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
+%     \caption{
+%         Individual-event $\kappa_i$ distributions of Fig.~\ref{fig:violin_kappa} reweighted in light of the population-level inference of Fig.~\ref{fig:corner_Gaussian}, marginalizing over $\mu$ and $\sigma$.
+%         Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i|$; we label the events most in tension with $\kappa_i = 0$ by that same measure.
+%     }
+%     \label{fig:reweighted_kappa}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \script{posterior_kappa.py}
-    \includegraphics[width=\columnwidth]{figures/posterior_kappa.pdf}
-    \caption{
-        Restricted and generic posteriors on $\kappa$, which respectively do and do not assume that $\kappa$ is shared by all events (color).
-        Solid and dashed traces indicate whether the analysis included GW200129 or not.
-        The black dashed line marks the absence of birefringence ($\kappa=0$).
-        The shaded distribution is the primary result in this work (blue band in Fig.~\ref{fig:violin_kappa}).
-    }
-    \label{fig:posterior_kappa}
-\end{figure}
+% \begin{figure}
+%     \script{posterior_kappa.py}
+%     \includegraphics[width=\columnwidth]{figures/posterior_kappa.pdf}
+%     \caption{
+%         Restricted and generic posteriors on $\kappa$, which respectively do and do not assume that $\kappa$ is shared by all events (color).
+%         Solid and dashed traces indicate whether the analysis included GW200129 or not.
+%         The black dashed line marks the absence of birefringence ($\kappa=0$).
+%         The shaded distribution is the primary result in this work (blue band in Fig.~\ref{fig:violin_kappa}).
+%     }
+%     \label{fig:posterior_kappa}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \script{corner_GW170818.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
-    \caption{
-        GW170818 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
-        The top right panel shows the marginalized posterior on $\chi_p$: allowing for birefringence reduces the preference for precession.
-        (See Fig.~\ref{fig:corner_GW170818_appendix} for a full corner plot.)
-    }
-    \label{fig:corner_GW170818}
-\end{figure}
+% \begin{figure}
+%     \script{corner_GW170818.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_GW170818.pdf}
+%     \caption{
+%         GW170818 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+%         The top right panel shows the marginalized posterior on $\chi_p$: allowing for birefringence reduces the preference for precession.
+%         (See Fig.~\ref{fig:corner_GW170818_appendix} for a full corner plot.)
+%     }
+%     \label{fig:corner_GW170818}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \script{corner_GW190521.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
-    \caption{
-        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
-        The marginalized posterior on $\chi_p$ is shown in the top right panel.
-    }
-    \label{fig:corner_GW190521}
-\end{figure}
-
-% % Case: GW150914 (example of broken degeneracy)
-% \subsubsection{GW150914}
-% In this study, we included frequency dependence of birefringence, which would affect the posterior of $\kappa$ obtained from \ac{PE}.
-% Consider GW150914, the first GW detected by LIGO, as an example.
-% In Fig.~\ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
-% With the frequency-independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
-% This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
-% % explain the kappa gap
-% 
-% On the other hand, with the frequency-dependent birefringence model, the posterior looks similar to the GR posterior.
-% This is because the frequency dependence broke the degeneracy, as the effect of birefringence will differ from the effect of changing $\iota$.
-% In this case, the most probable value of $\kappa$ is close to $0$, which means the birefringence is weak or absent.
-% Therefore, the \ac{PE} result with frequency dependence is consistent with GR.
-% 
 % \begin{figure}
-%     \script{corner_GW150914.py}
-%     \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
+%     \script{corner_GW190521.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
 %     \caption{
-%         GW150914 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our frequency-dependent (blue), compared to the GR result (green).
+%         GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
 %         The marginalized posterior on $\chi_p$ is shown in the top right panel.
 %     }
-%     \label{fig:corner_GW150914}
-% \end{figure}
-
-% % Case: GW191105
-% \subsubsection{GW191105}
-% \label{sec:GW191105}
-% 
-% The situation is similar for GW191105 (Fig.~\ref{fig:corner_GW191105}).
-% 
-% \begin{figure}
-%     \script{corner_GW191105.py}
-%     \includegraphics[width=\columnwidth]{figures/corner_GW191105.pdf}
-%     \caption{
-%         GW191105 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
-%         The posterior on $\chi_{eff}$ and $\kappa$ from our birefringence analysis is shown in the top right panel.
-%     }
-%     \label{fig:corner_GW191105}
+%     \label{fig:corner_GW190521}
 % \end{figure}
@@
 
-\begin{figure}
-    \script{corner_GW200129.py}
-    \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
-    \caption{
-        GW200129 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos{\iota}$, including different sets of detectors in the analysis per the legend.
-        The marginalized posterior on $\chi_p$ is shown in the top right panel.
-        The main run with all three detectors (HLV, filled blue), shows a preference for $\kappa < 0$, as in Fig.~\ref{fig:violin_kappa}; this preference is more pronounced for two-detector runs that include Virgo (HV and LV, orange and green); however, it disappears if we remove Virgo (HL, dashed black).
-        The 2D contours correspond to the $90\%$ credible level.
-    }
-    \label{fig:corner_GW200129}
-\end{figure}
+% \begin{figure}
+%     \script{corner_GW200129.py}
+%     \includegraphics[width=\columnwidth]{figures/corner_GW200129.pdf}
+%     \caption{
+%         GW200129 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos{\iota}$, including different sets of detectors in the analysis per the legend.
+%         The marginalized posterior on $\chi_p$ is shown in the top right panel.
+%         The main run with all three detectors (HLV, filled blue), shows a preference for $\kappa < 0$, as in Fig.~\ref{fig:violin_kappa}; this preference is more pronounced for two-detector runs that include Virgo (HV and LV, orange and green); however, it disappears if we remove Virgo (HL, dashed black).
+%         The 2D contours correspond to the $90\%$ credible level.
+%     }
+%     \label{fig:corner_GW200129}
+% \end{figure}
 
@@
 
-\begin{figure}
-    \script{mass_ratio.py}
-    \includegraphics[width=\columnwidth]{figures/mass_ratio.pdf}
-    \caption{
-    Posterior on mass ratio $q \equiv m_2/m_1$ (left) and precessing spin $\chi_p$ (right) for GW170104 (top) and GW191105 (bottom), as obtained from both the birefringent analysis (orange) and the regular \ac{GR} analysis (blue).
-    }
-    \label{fig:mass_ratio}
-\end{figure}
+% \begin{figure}
+%     \script{mass_ratio.py}
+%     \includegraphics[width=\columnwidth]{figures/mass_ratio.pdf}
+%     \caption{
+%     Posterior on mass ratio $q \equiv m_2/m_1$ (left) and precessing spin $\chi_p$ (right) for GW170104 (top) and GW191105 (bottom), as obtained from both the birefringent analysis (orange) and the regular \ac{GR} analysis (blue).
+%     }
+%     \label{fig:mass_ratio}
+% \end{figure}
 
@@
 
-\begin{figure*}[h]
-    \script{corner_GW170818_appendix.py}
-    \includegraphics[width=\textwidth]{figures/corner_GW170818_appendix.pdf}
-    \caption{
-        Extended corner plot for GW170818: a supplement to Fig.~\ref{fig:corner_GW170818} discussed in Appendix~\ref{sec:corner_GW170818_appendix}.
-        The shaded regions contain 90\% and 39.35\% ($1\sigma$) of the probability mass.
-        The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.
-    }
-    \label{fig:corner_GW170818_appendix}
-\end{figure*}
+% \begin{figure*}[h]
+%     \script{corner_GW170818_appendix.py}
+%     \includegraphics[width=\textwidth]{figures/corner_GW170818_appendix.pdf}
+%     \caption{
+%         Extended corner plot for GW170818: a supplement to Fig.~\ref{fig:corner_GW170818} discussed in Appendix~\ref{sec:corner_GW170818_appendix}.
+%         The shaded regions contain 90\% and 39.35\% ($1\sigma$) of the probability mass.
+%         The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.
+%     }
+%     \label{fig:corner_GW170818_appendix}
+% \end{figure*}
 
@@
 
-\begin{figure*}[h]
-    \script{corner_GW190521_appendix.py}
-    \includegraphics[width=\textwidth]{figures/corner_GW190521_appendix.pdf}
-    \caption{
-        Extended corner plot for GW190521: a supplement to Fig.~\ref{fig:corner_GW190521} discussed in Appendix~\ref{sec:corner_GW190521_appendix}.
-        The shaded regions contain 90\% and 39.35\% ($1\sigma$) of the probability mass.
-        The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.
-    }
-    \label{fig:corner_GW190521_appendix}
-\end{figure*}
+% \begin{figure*}[h]
+%     \script{corner_GW190521_appendix.py}
+%     \includegraphics[width=\textwidth]{figures/corner_GW190521_appendix.pdf}
+%     \caption{
+%         Extended corner plot for GW190521: a supplement to Fig.~\ref{fig:corner_GW190521} discussed in Appendix~\ref{sec:corner_GW190521_appendix}.
+%         The shaded regions contain 90\% and 39.35\% ($1\sigma$) of the probability mass.
+%         The prior is uniform in all shown quantities except $d_L$, whose prior corresponds to a distribution uniform in comoving volume.
+%     }
+%     \label{fig:corner_GW190521_appendix}
+% \end{figure*}
 

### 2023-04-14 src/tex/ms.tex
Explicitly mention that h is the Planck constant.

--- 2023-04-13 src/tex/ms.tex
@@
 \begin{equation}
-    |\kappa| \simeq \frac{\pi H_0}{c} \frac{h \left( 100 \, \mathrm{Hz} \right)}{M_{\rm PV}} \simeq \frac{3.0 \times 10^{-22} \, \mathrm{GeV}}{M_\mathrm{PV}} \, \mathrm{Gpc}^{-1}
+    |\kappa| \simeq \frac{\pi H_0}{c} \frac{h \left( 100 \, \mathrm{Hz} \right)}{M_{\rm PV}} \simeq \frac{3.0 \times 10^{-22} \, \mathrm{GeV}}{M_\mathrm{PV}} \, \mathrm{Gpc}^{-1},
 \end{equation}
+where $h$ is Planck's constant.
 \citet{Wang_2021} quote a constraint of $M_{\rm PV} > 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Wang.txt}.

### 2023-04-14 src/tex/ms.tex
Added footnote pointing out that these are not, actually, "low redshift."

@@
 In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity \cite{Alexander:2009tp}.
-Since it originates from a truncated series expansion, Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
+Since it originates from a truncated series expansion%
+\footnote{Often the birefringent effect is written as an expansion in redshift, $z$ \cite[e.g.][]{Zhao:2019xmm}, rather than distance.
+Here we use $d_C$ to emphasize that the effect accumulates \emph{per cycle}, or with propagation distance.
+Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$, are already observing propagation over a large fraction of the universe and constrain this effect meaningfully.}, %
+Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
 \begin{equation}

### 2023-04-15 src/tex/ms.tex
minor text change

@@
 Here we use $d_C$ to emphasize that the effect accumulates \emph{per cycle}, or with propagation distance.
-Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$, are already observing propagation over a large fraction of the universe and constrain this effect meaningfully.}, %
+Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$ (Hubble distance), are already observing propagation over a large fraction of the universe and constrain this effect meaningfully.}, %
 Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 

### 2023-04-17 src/tex/ms.tex
Minor typo fixes, additional reference for re-weighting posteriors.

@@
 Nevertheless, there exist systems for which the degeneracy cannot be broken in practice because not enough frequencies are available in the data.
-This may be the case for quasimonochromatic sources, like nonaxisymmetric pulsars or very light binaries, which are well approximated by a single Fourier mode.
+This may also be the case for quasimonochromatic sources, like nonaxisymmetric pulsars or very light binaries, which are well approximated by a single Fourier mode.
 
@@
 We take the \ac{PE} configuration in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
-We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time, and set the prior on $\kappa$ to be uniform between $-1$ and $1$.
+We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time, and set the prior on $\kappa$ to be uniform between $-1 \, \mathrm{Gpc}^{-1}$ and $1 \, \mathrm{Gpc}^{-1}$.
 
@@
 As a visual check for outliers, we reconsider the set of measurements in Fig.~\ref{fig:violin_kappa} in light of the hierarchical result for $\mu$ and $\sigma$ in Fig.~\ref{fig:corner_Gaussian}, including GW200129 (blue curve).
-This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Callister:T2100301}.
+This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Miller2020,Callister:T2100301}.
 The result in Figure \ref{fig:reweighted_kappa} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its higher support for $\kappa > 0$.

### 2023-05-04 src/tex/ms.tex
Correct number of notable events

@@
 
-Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on four events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:violin_kappa}: GW170818, GW190521 and GW200129.
+Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on three events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:violin_kappa}: GW170818, GW190521 and GW200129.
 When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$, the third is the most extreme representative of a class of events with bimodal $\kappa$ posteriors, and the third shows signs of potential noise anomalies.

### 2023-05-08 src/tex/ms.tex

@@
 Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on three events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:violin_kappa}: GW170818, GW190521 and GW200129.
-When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$, the third is the most extreme representative of a class of events with bimodal $\kappa$ posteriors, and the third shows signs of potential noise anomalies.
+When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$, the second is the most extreme representative of a class of events with bimodal $\kappa$ posteriors, and the third shows signs of potential noise anomalies.
 Through these examples, we elucidate some of the interactions between $\kappa$ and the source luminosity distance, inclination and spin parameters.

### 2023-05-09 src/tex/ms.tex
new references

@@
 \citet{Okounkova_2022} considered the distribution of observed inclinations of the \ac{GW} events in the second \ac{GW} transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.
+While this manuscript was finalized, \citet{Zhu:2023wci} reported an analysis of GWTC-3.
 
@@
 
+The parameterization of \citet{Wang_2021} is the same as that in the more recent work by \citet{Zhu:2023wci}.
+That work, which appeared while this manuscript was finalized, reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}$ based on GWTC-3.
+
 \subsection{Parameter degeneracies}
@@
 The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
-Future catalogs with more events will shed light on this feature.
+Future catalogs with more events will shed light on this feature, and enable richer models with explicit correlations between birefringent attenuation and source parameters like sky location \citep{Goyal:2023uvm,Ezquiaga:2021ler}.
 

### 2023-05-09 src/scripts/kappa_constraint_restricted.py
Add Zhu's constraint to table

@@ -62,6 +62,9 @@ M_PV_Okounkova = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(kappa_Okounkova)/1e9
 kappa_this_work = restricted_absolute_kappa_90
 M_PV_this_work = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(kappa_this_work)/1e9
 
+M_PV_Zhu = 4.1e-22 # GeV
+kappa_Zhu = (h*np.pi*H_0/c)*(1e3)*(100)*np.reciprocal(M_PV_Zhu*1e9)
+
 with open(paths.output/"M_PV_constraint.txt", "w") as f:
     f.write(rf"$M_{{\rm PV}} \gtrsim {M_PV_this_work/1e-21:.1f} \times 10^{{-21}}\, {{\rm GeV}}$")
 
@@ -69,5 +72,6 @@ with open(paths.output/"comparison_summary.txt", "w") as f:
     f.write(r"\begin{tabular}{lccc} & $M_{\rm PV}$ ($10^{-21}\, {\rm GeV}$) & $|\kappa|$ ($\mathrm{Gpc}^{-1}$) & CL (\%) \\ \hline ")
     f.write(rf"\citet{{Wang_2021}} & $> {M_PV_Wang/1e-21:.2f}$ & $< {kappa_Wang:.2f}$ & 90 \\ ")
     f.write(rf"\citet{{Okounkova_2022}} & $> {M_PV_Okounkova/1e-21:.2f}$ & $< {kappa_Okounkova:.2f}$ & 68 \\ ")
+    f.write(rf"\citet{{Zhu:2023wci}} & $> {M_PV_Zhu/1e-21:.2f}$ & $< {kappa_Zhu:.2f}$ & 90 \\ ")
     f.write(rf"Ng \emph{{et al.}} (this work) & $> {M_PV_this_work/1e-21:.2f}$ & $< {kappa_this_work:.2f}$ & 90")
     f.write(r" \end{tabular}")

### 2023-05-10 src/tex/ms.tex
Add Zhu's constraint to table
Minor wording
no floatfix for arxiv

@@
-\documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,floatfix,nofootinbib]{revtex4-2}
+\documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,nofootinbib]{revtex4-2}
 
@@
 
+\begin{figure}
+    \script{reweighed_kappa.py}
+    \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
+    \caption{
+        Individual-event $\kappa_i$ distributions of Fig.~\ref{fig:violin_kappa} reweighted in light of the population-level inference of Fig.~\ref{fig:corner_Gaussian}, marginalizing over $\mu$ and $\sigma$.
+        Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i|$; we label the events most in tension with $\kappa_i = 0$ by that same measure.
+    }
+    \label{fig:reweighted_kappa}
+\end{figure}
+
 We summarize the result of this exercise in Fig.~\ref{fig:corner_Gaussian}, which shows the posterior on the population mean $\mu$ and standard deviation $\sigma$ inferred from the collection of observations in Fig.~\ref{fig:violin_kappa}.
@@
 Future observations will determine whether there is truly evidence for a nonvanishing variance in this population.
-
-\begin{figure}
-    \script{reweighed_kappa.py}
-    \includegraphics[width=\columnwidth]{figures/reweighed_kappa.pdf}
-    \caption{
-        Individual-event $\kappa_i$ distributions of Fig.~\ref{fig:violin_kappa} reweighted in light of the population-level inference of Fig.~\ref{fig:corner_Gaussian}, marginalizing over $\mu$ and $\sigma$.
-        Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i|$; we label the events most in tension with $\kappa_i = 0$ by that same measure.
-    }
-    \label{fig:reweighted_kappa}
-\end{figure}
 
@@
     \caption{
-        Comparison of our constraints to \citet{Okounkova_2022} and \citet{Wang_2021}.
+        Comparison of our constraints with previous studies.
         CL is the credible level.

### 2023-05-11 src/tex/ms.tex
salvo citation

@@
 \citet{Yamada_2020} and \citet{Wang_2021} both performed \ac{PE} on the events in the first \ac{GW} transient catalog \citep{GWTC-1}, GWTC-1, using birefringent templates.
-\citet{Okounkova_2022} considered the distribution of observed inclinations of the \ac{GW} events in the second \ac{GW} transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.
+\citet{Okounkova_2022} considered the distribution of observed inclinations of the \ac{GW} events in the second \ac{GW} transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence; \citet{Vitale:2022pmu} carried out a related analysis.
 While this manuscript was finalized, \citet{Zhu:2023wci} reported an analysis of GWTC-3.
@@
 
-\citet{Okounkova_2022} took the effect of birefringence to be independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
+\citet{Okounkova_2022} took the effect of birefringence to be independent of the frequency, which is a simplified approximation of the birefringence model in Chern-Simons gravity.
 This assumption results in a full degeneracy between $\kappa$ and $\iota$:
@@
 
-By implementing Eq.~\eqref{eq:waveform_modification}, which is a first-order approximation of the birefringence model, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
+By implementing Eq.~\eqref{eq:waveform_modification}, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
 Nevertheless, there exist systems for which the degeneracy cannot be broken in practice because not enough frequencies are available in the data.

### 2023-05-12 src/tex/ms.tex
Lue et al ref

@@
 Among those, measurements of the basic properties of \acp{GW}, like their speed and polarization, can directly probe the fundamental symmetries of the underlying theory of gravity \citep{Will:2018bme}.
-For instance, unequal propagation of \ac{GW} polarization eigenstates would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Jackiw:2003pm,Alexander:2009tp,Sopuerta:2009iy}.
+For instance, unequal propagation of \ac{GW} polarization eigenstates would reveal that spacetime is birefringent, a smoking gun for parity-odd theories like Chern-Simons gravity \citep{Lue:1998mq,Jackiw:2003pm,Alexander:2009tp,Sopuerta:2009iy}.
 Here we constrain the magnitude of possible amplitude birefringence using \ac{BBH} signals from the latest \ac{LVK} catalog, GWTC\nobreakdash-3 \citep{GWTC-3}.

### 2023-05-13 src/tex/ms.tex
Added 'graph about Okounkova l0 constraint.

--- 2023-05-11 src/tex/ms.tex
@@
 
+The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale,\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into a FRW cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.}
+\begin{equation}
+    l_0 = \frac{c d_H \kappa}{3 \pi f} = 1400 \, \mathrm{km} \left( \frac{\kappa}{1 \, \mathrm{Gpc}^{-1}} \right) \left( \frac{100 \, \mathrm{Hz}}{f} \right);
+\end{equation}
+our 68\% credible bound on $\kappa$ implies $l_0 \lesssim 40 \, \mathrm{km}$ at $100 \, \mathrm{Hz}$.
+This compares favorably to the constraint in \citet{Okounkova_2022}, which had $l_0 \lesssim 1000 \, \mathrm{km}$.
+
 \subsubsection{Wang et al.}

### 2023-07-06 src/tex/ms.tex
Minor changes for B.3 & B.5

@@
 \begin{equation}
-\left|\kappa\right| d_C \left(f/100\, \mathrm{Hz}\right) < 1\, .
+\left|\kappa\right| d_C \left(f/100\, \mathrm{Hz}\right) < 1\,,
 \end{equation}
+note that $d_C$ is in units of Gpc.
 Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
@@
 In this section, we present the results of our study.
-We first show the $\kappa$ measurements from of all events in our set, as well as the resulting global measurement of $\kappa$ that represents our primary constraint on birefringence (Sec.~\ref{sec:results:gwtc}).
+We first show the $\kappa$ measurements of all events in our set, as well as the resulting global measurement of $\kappa$ that represents our primary constraint on birefringence (Sec.~\ref{sec:results:gwtc}).
 We then assess the collection of measurements in more detail through a hierarchical analysis (Sec.~\ref{sec:results:hier}).

### 2023-07-07 src/tex/ms.tex
Changes for A.1 & A.2

@@
 This model is a better approximation of the birefringence effect expected from theory than the frequency-independent model used in \citet{Okounkova_2022}.
-Compared to other studies, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
+Including the frequecy dependence allows us to break the degeneracy between birefringence and source inclination, which will be discussed in Sec.~\ref{sec:inclination}.
+Compared to \citet{Yamada_2020,Wang_2021} and \citet{Okounkova_2022}, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
 We consider 71 binary black hole merger events with a \ac{FAR} $\leq1/\mathrm{yr}$, as listed in Table I of \citet{GWTC-3_population}.

### 2023-07-07 src/tex/ms.tex
Changes for A.2,4,5,7

@@
 Including the frequecy dependence allows us to break the degeneracy between birefringence and source inclination, which will be discussed in Sec.~\ref{sec:inclination}.
-Compared to \citet{Yamada_2020,Wang_2021} and \citet{Okounkova_2022}, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
+Compared to Ref.~\citep{Yamada_2020,Wang_2021,Okounkova_2022}, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
 We consider 71 binary black hole merger events with a \ac{FAR} $\leq1/\mathrm{yr}$, as listed in Table I of \citet{GWTC-3_population}.
@@
         Restricted and generic posteriors on $\kappa$, which respectively do and do not assume that $\kappa$ is shared by all events (color).
-        Solid and dashed traces indicate whether the analysis included GW200129 or not.
+        Solid (dashed) traces indicate the analysis including (excluding) GW200129.
         The black dashed line marks the absence of birefringence ($\kappa=0$).
@@
 
+Encouraged by Chern-Simons gravity, this study only focused on amplitude birefringence.
+Other parity-violating gravity theories also predict velocity birefringence, which dominates over amplitude birefringence in terms of amplitude and phase corrections to the waveform when both birefringence are present \cite{Zhao:2019xmm}.
+Velocity birefringence was tested with GWTC-3 in \citet{Wang:2021gqm} and \citet{Haegel:2022ymk}.
+
 All our data are available in \citet{dataset}.

### 2023-07-29 src/tex/ms.tex
Include unit in Eqn 3

@@
 \begin{equation}
-\left|\kappa\right| d_C \left(f/100\, \mathrm{Hz}\right) < 1\,,
+\left|\kappa\right| \left(d_c/1\, \mathrm{Gpc}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, \mathrm{Gpc}^{-1}\,.
 \end{equation}
-note that $d_C$ is in units of Gpc.
 Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.

### 2023-07-31 src/tex/ms.tex
Changes for A.4

@@
         Individual-event $\kappa_i$ distributions of Fig.~\ref{fig:violin_kappa} reweighted in light of the population-level inference of Fig.~\ref{fig:corner_Gaussian}, marginalizing over $\mu$ and $\sigma$.
-        Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i|$; we label the events most in tension with $\kappa_i = 0$ by that same measure.
+        Curves are colored by the magnitude of the supported deviation, as represented by the respective posterior $|\mu_i / \sigma_i|$.
     }

### 2023-07-31 src/scripts/reweighed_kappa.py

@@ -35,6 +35,7 @@ for i, (z,line) in enumerate(zip(zscores, g.get_lines()[::-1])):
     line.set_alpha(norm_alpha(z))
     if zscores.index[i]=='GW170818':
       line.set_label(zscores.index[i].replace('_', r'\_'))
+      line.set_linestyle('--')
     else:
       line.set_label(None)
 g.legend(fontsize=14)

### 2023-08-08 src/tex/ms.tex
Change for A.2
Changes for A.6

@@
 In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
+Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening.
+Similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci}.
 To first order, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
@@
 \subsubsection{Wang et al.}
+\label{sec:comparison_Wang}
 
@@
 
-The parameterization of \citet{Wang_2021} is the same as that in the more recent work by \citet{Zhu:2023wci}.
-That work, which appeared while this manuscript was finalized, reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}$ based on GWTC-3.
+\subsubsection{Zhu et al.}
+A more recent work by \citet{Zhu:2023wci}, which appeared while this manuscript was finalized, performed \ac{PE} on GWTC-3 events with a frequency-dependent model of amplitude birefringence.
+Their model is similar to the one used in \citet{Wang_2021}, which allows us to compare our results, as we did in Sec.~\ref{sec:comparison_Wang}.
+\citet{Zhu:2023wci} reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Zhu.txt}.
+This result constrain the birefringence effect much weaker than our result.
+This is an unexpected result, as the analysis in \citet{Zhu:2023wci} is similar to ours\footnote{Note that the prior on the birefringence parameter and luminosity distance are different between the analysis by \citet{Zhu:2023wci}and ours, but it is unlikely to be the cause of the difference in the constraint.} and they included more events than we did.
+However, we could not compare the analysis in \citet{Zhu:2023wci} with ours in detail, as the analysis is not publicly available, and the list of events included is not provided in the paper.
 

### 2023-08-08 src/tex/ms.tex
Change for B.6
minor text change

@@
 That table also shows the credible level (CL) at which the posterior supports $\kappa = 0$, whereby $\mathrm{CL} = 0$ ($\mathrm{CL} = 1$) means the posterior supports that value with high (low) probability.
+This quantity can be understood as the relative height of the probability density function at a given value, and does not assume any symmetric interval or modality.
 
@@
 
-The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale,\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into a FRW cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.}
+The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into a FRW cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.},
 \begin{equation}

### 2023-08-20 src/tex/ms.tex
Changes for B.7
Minor change for B.7
started change tracking (WIP)

@@
 \usepackage{bookmark}
+\usepackage{changes}
 %\usepackage[all]{hypcap}
@@
 This model is a better approximation of the birefringence effect expected from theory than the frequency-independent model used in \citet{Okounkova_2022}.
-Including the frequecy dependence allows us to break the degeneracy between birefringence and source inclination, which will be discussed in Sec.~\ref{sec:inclination}.
-Compared to Ref.~\citep{Yamada_2020,Wang_2021,Okounkova_2022}, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
+\added{Including the frequency dependence allows us to break the degeneracy between birefringence and source inclination, which we discuss below in Sec.~\ref{sec:inclination}.}
+Compared to \replaced{Refs.~\citep{Yamada_2020,Wang_2021,Okounkova_2022}}{other studies}, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
 We consider 71 binary black hole merger events with a \ac{FAR} $\leq1/\mathrm{yr}$, as listed in Table I of \citet{GWTC-3_population}.
@@
 In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
-Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening.
-Similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci}.
+Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening;
+\added{similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci}.}
 To first order, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
@@
     \left|\kappa\right| \left(d_c/1\, \mathrm{Gpc}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, \mathrm{Gpc}^{-1}\,.
+    \label{eq:small_exponent}
 \end{equation}
@@
         Restricted and generic posteriors on $\kappa$, which respectively do and do not assume that $\kappa$ is shared by all events (color).
-        Solid (dashed) traces indicate the analysis including (excluding) GW200129.
+        \replaced{%
+        Solid (dashed) traces indicate the analysis included (excluded) GW200129.}
+        {Solid and dashed traces indicate whether the analysis included GW200129 or not.}
         The black dashed line marks the absence of birefringence ($\kappa=0$).
@@
         GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
-        The marginalized posterior on $\chi_p$ is shown in the top right panel.
+        The marginalized posterior on $\chi_p$ is shown in the top right panel. (See Fig.~\ref{fig:corner_GW190521_appendix} for a full corner plot.)
+        Note that the condition in Eq.~\eqref{eq:small_exponent} is not violated as the distance posterior shown here is the posterior on the luminosity distance $d_L$, not on the comoving distance $d_C$.
     }
@@
 
-\subsubsection{Zhu et al.}
+\deleted{The parameterization of \cite{Wang_2021} is the same as that in the more recent work by \cite{Zhu:2023wci}.
+That work, which appeared while this manuscript was finalized, reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}$ based on GWTC-3.}
+
+\subsubsection{\added{Zhu et al.}}
+
 A more recent work by \citet{Zhu:2023wci}, which appeared while this manuscript was finalized, performed \ac{PE} on GWTC-3 events with a frequency-dependent model of amplitude birefringence.

### 2023-08-28 src/tex/ms.tex
Add plot for A.6

@@
 We obtained a tighter 90\% upper limit of \variable{output/restricted_absolute_kappa_90.txt}, or \variable{output/M_PV_constraint.txt}.
+Note that our constraints from each event are stronger compared to the constraint from \citet{Wang_2021}, details of which are given in Sec.~\ref{sec:comparison_Zhu}.
 
@@
 \subsubsection{\added{Zhu et al.}}
+\label{sec:comparison_Zhu}
 
 A more recent work by \citet{Zhu:2023wci}, which appeared while this manuscript was finalized, performed \ac{PE} on GWTC-3 events with a frequency-dependent model of amplitude birefringence.
-Their model is similar to the one used in \citet{Wang_2021}, which allows us to compare our results, as we did in Sec.~\ref{sec:comparison_Wang}.
+Their model is the same as used in \citet{Wang_2021}, which allows us to compare our results, as we did in Sec.~\ref{sec:comparison_Wang}.
 \citet{Zhu:2023wci} reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Zhu.txt}.
-This result constrain the birefringence effect much weaker than our result.
+This result constrains the birefringence effect much weaker than our result.
 This is an unexpected result, as the analysis in \citet{Zhu:2023wci} is similar to ours\footnote{Note that the prior on the birefringence parameter and luminosity distance are different between the analysis by \citet{Zhu:2023wci}and ours, but it is unlikely to be the cause of the difference in the constraint.} and they included more events than we did.
-However, we could not compare the analysis in \citet{Zhu:2023wci} with ours in detail, as the analysis is not publicly available, and the list of events included is not provided in the paper.
+
+In \ref{fig:posterior_MPV}, we show the posterior on $M_{\rm PV}^{-1}$ for each event in our analysis.
+Compared to Fig.~I in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different between our analysis and theirs.
+For example, the posterior for GW190727\_060333 peaks at zero in our analysis, but peaks at a nonzero value in \citet{Zhu:2023wci}.
+This shows that the difference in the combined constraint may come from the difference in the individual event analyses.
+However, we could not compare the analysis in \citet{Zhu:2023wci} with ours in detail, as their analysis result is not publicly available.
+
+\begin{figure}
+    \script{posterior_MPV.py}
+    \includegraphics[width=\columnwidth]{figures/posterior_MPV.pdf}
+    \caption{
+        Individual-event $M_\mathrm{PV}^{-1}$ posteriors from our analysis, converted from the $\kappa$ posteriors in Fig.~\ref{fig:violin_kappa}.
+        Chosen events are those labeled in Fig.~I of \citet{Zhu:2023wci}.
+    }
+    \label{fig:posterior_MPV}
+\end{figure}
 

### 2023-08-28 src/tex/ms.tex
minor text change

@@
 In \ref{fig:posterior_MPV}, we show the posterior on $M_{\rm PV}^{-1}$ for each event in our analysis.
-Compared to Fig.~I in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different between our analysis and theirs.
+Compared to Fig.~1 in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different between our analysis and theirs.
 For example, the posterior for GW190727\_060333 peaks at zero in our analysis, but peaks at a nonzero value in \citet{Zhu:2023wci}.
@@
         Individual-event $M_\mathrm{PV}^{-1}$ posteriors from our analysis, converted from the $\kappa$ posteriors in Fig.~\ref{fig:violin_kappa}.
-        Chosen events are those labeled in Fig.~I of \citet{Zhu:2023wci}.
+        Chosen events are those labeled in Fig.~1 of \citet{Zhu:2023wci}.
     }

### 2023-08-28 src/tex/ms.tex
Will's minor edits to Zhu+ discussion.

@@
 \newcommand*{\wf}[1]{\textsf{\color{cyan} [\textbf{WILL:} #1]}}
+
+\definechangesauthor[name=Will, color=cyan]{WF}
 
@@
 In \ref{fig:posterior_MPV}, we show the posterior on $M_{\rm PV}^{-1}$ for each event in our analysis.
-Compared to Fig.~1 in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different between our analysis and theirs.
+Compared to Fig.~1 in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different \added[id=WF]{in both scale and shape} between our analysis and theirs.
 For example, the posterior for GW190727\_060333 peaks at zero in our analysis, but peaks at a nonzero value in \citet{Zhu:2023wci}.

### 2023-08-29 src/tex/ms.tex
edits

--- 2023-08-28 src/tex/ms.tex
@@
 Comparing to our Eq.~\eqref{eq:waveform_modification} and approximating $z \approx d_C H_0/c$ via the Hubble constant, $H_0$, this means
-\begin{equation}
+\begin{equation} \label{eq:kappa-to-Mpv}
     |\kappa| \simeq \frac{\pi H_0}{c} \frac{h \left( 100 \, \mathrm{Hz} \right)}{M_{\rm PV}} \simeq \frac{3.0 \times 10^{-22} \, \mathrm{GeV}}{M_\mathrm{PV}} \, \mathrm{Gpc}^{-1},
@@
 We obtained a tighter 90\% upper limit of \variable{output/restricted_absolute_kappa_90.txt}, or \variable{output/M_PV_constraint.txt}.
-Note that our constraints from each event are stronger compared to the constraint from \citet{Wang_2021}, details of which are given in Sec.~\ref{sec:comparison_Zhu}.
 
@@
 \subsubsection{\added{Zhu et al.}}
-\label{sec:comparison_Zhu}
 
 A more recent work by \citet{Zhu:2023wci}, which appeared while this manuscript was finalized, performed \ac{PE} on GWTC-3 events with a frequency-dependent model of amplitude birefringence.
-Their model is the same as used in \citet{Wang_2021}, which allows us to compare our results, as we did in Sec.~\ref{sec:comparison_Wang}.
-\citet{Zhu:2023wci} reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Zhu.txt}.
-This result constrains the birefringence effect much weaker than our result.
-This is an unexpected result, as the analysis in \citet{Zhu:2023wci} is similar to ours\footnote{Note that the prior on the birefringence parameter and luminosity distance are different between the analysis by \citet{Zhu:2023wci}and ours, but it is unlikely to be the cause of the difference in the constraint.} and they included more events than we did.
+Their model is the same as the one used in \citet{Wang_2021}, which allows us to compare to our results as we did in Sec.~\ref{sec:comparison_Wang}.
+\citet{Zhu:2023wci} reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Zhu.txt} per Eq.~\eqref{eq:kappa-to-Mpv} above.
 
-In \ref{fig:posterior_MPV}, we show the posterior on $M_{\rm PV}^{-1}$ for each event in our analysis.
-Compared to Fig.~1 in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different between our analysis and theirs.
-For example, the posterior for GW190727\_060333 peaks at zero in our analysis, but peaks at a nonzero value in \citet{Zhu:2023wci}.
-This shows that the difference in the combined constraint may come from the difference in the individual event analyses.
-However, we could not compare the analysis in \citet{Zhu:2023wci} with ours in detail, as their analysis result is not publicly available.
-
-\begin{figure}
-    \script{posterior_MPV.py}
-    \includegraphics[width=\columnwidth]{figures/posterior_MPV.pdf}
-    \caption{
-        Individual-event $M_\mathrm{PV}^{-1}$ posteriors from our analysis, converted from the $\kappa$ posteriors in Fig.~\ref{fig:violin_kappa}.
-        Chosen events are those labeled in Fig.~1 of \citet{Zhu:2023wci}.
-    }
-    \label{fig:posterior_MPV}
-\end{figure}
+This constraint is weaker than ours in spite of the similar number of events considered, a discrepancy that might be explained by a number of factors.
+First, there are difference in priors, since \citet{Zhu:2023wci} assume a distribution uniform in $z/M_{\rm PV}$ rather than $\kappa$ or $M_{\rm PV}$ itself.
+Second, as reported in \citep{Zhu:2023wci}, their combined result was computed by multiplying $M_{\rm PV}^{-1}$ posteriors from individual events; however, this is not the correct way of combining $M_{\rm PV}$ measurements from independent observations: instead of multiplying posteriors, one should multiply the \emph{likelihoods}Z to avoid double counting the prior every time a new event is added to the combined constraint.
+Finally, \citet{Zhu:2023wci} report \emph{ad hoc} exclusion of certain events from the combined constraint, without explicitly providing a list of events considered.
+All these factors combined preclude a clear interpretation of the results in \citep{Zhu:2023wci}.
 

### 2023-08-29 src/tex/ms.tex

@@
 \subsubsection{\added{Zhu et al.}}
+\label{sec:comparison_Zhu}
 
@@
 All these factors combined preclude a clear interpretation of the results in \citep{Zhu:2023wci}.
+In \ref{fig:posterior_MPV}, we show the posterior on $M_{\rm PV}^{-1}$ for each event in our analysis.
+Compared to Fig.~1 in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different \added[id=WF]{in both scale and shape} between our analysis and theirs.
+For example, the posterior for GW190727\_060333 peaks at zero in our analysis, but peaks at a nonzero value in \citet{Zhu:2023wci}.
+This shows that the difference in the combined constraint may come from the difference in the individual event analyses.
+However, we could not compare the analysis in \citet{Zhu:2023wci} with ours in detail, as their analysis result is not publicly available.
+
+\begin{figure}
+    \script{posterior_MPV.py}
+    \includegraphics[width=\columnwidth]{figures/posterior_MPV.pdf}
+    \caption{
+        Individual-event $M_\mathrm{PV}^{-1}$ posteriors from our analysis, converted from the $\kappa$ posteriors in Fig.~\ref{fig:violin_kappa}.
+        Chosen events are those labeled in Fig.~1 of \citet{Zhu:2023wci}.
+    }
+    \label{fig:posterior_MPV}
+\end{figure}
 

### 2023-08-29 src/tex/ms.tex
Mark previous changes

@@
 In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
-Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening;
-\added{similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci}.}
+\added{Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening; similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci}.}
 To first order, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
@@
 \begin{equation}
-    \left|\kappa\right| \left(d_c/1\, \mathrm{Gpc}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, \mathrm{Gpc}^{-1}\,.
+    \left|\kappa\right| \left(d_c\added{/1\, \mathrm{Gpc}}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, \added{\mathrm{Gpc}^{-1}}\,.
     \label{eq:small_exponent}
@@
 Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
-
 \begin{figure}
@@
 That table also shows the credible level (CL) at which the posterior supports $\kappa = 0$, whereby $\mathrm{CL} = 0$ ($\mathrm{CL} = 1$) means the posterior supports that value with high (low) probability.
-This quantity can be understood as the relative height of the probability density function at a given value, and does not assume any symmetric interval or modality.
+\added{This quantity can be understood as the relative height of the probability density function at a given value, and does not assume any symmetric interval or modality.}
 
@@
         The marginalized posterior on $\chi_p$ is shown in the top right panel. (See Fig.~\ref{fig:corner_GW190521_appendix} for a full corner plot.)
-        Note that the condition in Eq.~\eqref{eq:small_exponent} is not violated as the distance posterior shown here is the posterior on the luminosity distance $d_L$, not on the comoving distance $d_C$.
+        \added{Note that the condition in Eq.~\eqref{eq:small_exponent} is not violated as the distance posterior shown here is the posterior on the luminosity distance $d_L$, not on the comoving distance $d_C$.}
     }
@@
 Comparing to our Eq.~\eqref{eq:waveform_modification} and approximating $z \approx d_C H_0/c$ via the Hubble constant, $H_0$, this means
-\begin{equation} \label{eq:kappa-to-Mpv}
+\begin{equation}
+    \label{eq:kappa-to-Mpv}
     |\kappa| \simeq \frac{\pi H_0}{c} \frac{h \left( 100 \, \mathrm{Hz} \right)}{M_{\rm PV}} \simeq \frac{3.0 \times 10^{-22} \, \mathrm{GeV}}{M_\mathrm{PV}} \, \mathrm{Gpc}^{-1},
@@
 We obtained a tighter 90\% upper limit of \variable{output/restricted_absolute_kappa_90.txt}, or \variable{output/M_PV_constraint.txt}.
-Note that our constraints from each event are stronger compared to the constraint from \citet{Wang_2021}, details of which are given in Sec.~\ref{sec:comparison_Zhu}.
+\added{Note that our constraints from each event are stronger compared to the constraint from \citet{Wang_2021}, details of which are given in Sec.~\ref{sec:comparison_Zhu}.}
 
@@
 First, there are difference in priors, since \citet{Zhu:2023wci} assume a distribution uniform in $z/M_{\rm PV}$ rather than $\kappa$ or $M_{\rm PV}$ itself.
-Second, as reported in \citep{Zhu:2023wci}, their combined result was computed by multiplying $M_{\rm PV}^{-1}$ posteriors from individual events; however, this is not the correct way of combining $M_{\rm PV}$ measurements from independent observations: instead of multiplying posteriors, one should multiply the \emph{likelihoods}Z to avoid double counting the prior every time a new event is added to the combined constraint.
+Second, as reported in \citep{Zhu:2023wci}, their combined result was computed by multiplying $M_{\rm PV}^{-1}$ posteriors from individual events; however, this is not the correct way of combining $M_{\rm PV}$ measurements from independent observations: instead of multiplying posteriors, one should multiply the \emph{likelihoods} to avoid double counting the prior every time a new event is added to the combined constraint.
 Finally, \citet{Zhu:2023wci} report \emph{ad hoc} exclusion of certain events from the combined constraint, without explicitly providing a list of events considered.
@@
 In \ref{fig:posterior_MPV}, we show the posterior on $M_{\rm PV}^{-1}$ for each event in our analysis.
-Compared to Fig.~1 in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different \added[id=WF]{in both scale and shape} between our analysis and theirs.
+Compared to Fig.~1 in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different in both scale and shape between our analysis and theirs.
 For example, the posterior for GW190727\_060333 peaks at zero in our analysis, but peaks at a nonzero value in \citet{Zhu:2023wci}.
@@
 
+\added{%
 Encouraged by Chern-Simons gravity, this study only focused on amplitude birefringence.
@@
 Velocity birefringence was tested with GWTC-3 in \citet{Wang:2021gqm} and \citet{Haegel:2022ymk}.
+}
 

### 2023-08-31 src/tex/ms.tex
tentative text on discrepancy from Zhu et al

@@
 We obtained a tighter 90\% upper limit of \variable{output/restricted_absolute_kappa_90.txt}, or \variable{output/M_PV_constraint.txt}.
-\added{Note that our constraints from each event are stronger compared to the constraint from \citet{Wang_2021}, details of which are given in Sec.~\ref{sec:comparison_Zhu}.}
 
@@
 
-This constraint is weaker than ours in spite of the similar number of events considered, a discrepancy that might be explained by a number of factors.
-First, there are difference in priors, since \citet{Zhu:2023wci} assume a distribution uniform in $z/M_{\rm PV}$ rather than $\kappa$ or $M_{\rm PV}$ itself.
-Second, as reported in \citep{Zhu:2023wci}, their combined result was computed by multiplying $M_{\rm PV}^{-1}$ posteriors from individual events; however, this is not the correct way of combining $M_{\rm PV}$ measurements from independent observations: instead of multiplying posteriors, one should multiply the \emph{likelihoods} to avoid double counting the prior every time a new event is added to the combined constraint.
-Finally, \citet{Zhu:2023wci} report \emph{ad hoc} exclusion of certain events from the combined constraint, without explicitly providing a list of events considered.
-All these factors combined preclude a clear interpretation of the results in \citep{Zhu:2023wci}.
-In \ref{fig:posterior_MPV}, we show the posterior on $M_{\rm PV}^{-1}$ for each event in our analysis.
-Compared to Fig.~1 in \citet{Zhu:2023wci}, we can see that the posteriors on some events are different in both scale and shape between our analysis and theirs.
+This constraint is weaker than ours in spite of the similar number of events considered, a discrepancy that we trace back to inference at the individual-event level.
+Figure \ref{fig:posterior_MPV} shows the posteriors on $M_\mathrm{PV}^{-1}$ derived from our analysis via Eq.~\eqref{eq:kappa-to-Mpv} for each event highlighted in Fig.~1 of \citet{Zhu:2023wci}.
+Our posteriors differ from those in \citet{Zhu:2023wci} in both overall scale and specific shape.
 For example, the posterior for GW190727\_060333 peaks at zero in our analysis, but peaks at a nonzero value in \citet{Zhu:2023wci}.
-This shows that the difference in the combined constraint may come from the difference in the individual event analyses.
-However, we could not compare the analysis in \citet{Zhu:2023wci} with ours in detail, as their analysis result is not publicly available.
+All distributions are narrower in our analysis by a factor of ${\sim}10\times$.
+% However, we could not compare the analysis in \citet{Zhu:2023wci} with ours in detail, as their analysis result is not publicly available.
+This might be explained by a number of analysis differences.
+The fact that posteriors disagree at the individual-event level suggests that the discrepancy originates in the different choice of parameterization and corresponding priors, in addition to potentially unstated differences in implementation.
+
+% First, there are differences in priors, since \citet{Zhu:2023wci} assume a distribution uniform in $z/M_{\rm PV}$ rather than $\kappa$ or $M_{\rm PV}$ itself.
+% Second, as reported in \citep{Zhu:2023wci}, their combined result was computed by multiplying $M_{\rm PV}^{-1}$ posteriors from individual events; however, this is not the correct way of combining $M_{\rm PV}$ measurements from independent observations: instead of multiplying posteriors, one should multiply the \emph{likelihoods} to avoid double counting the prior every time a new event is added to the combined constraint.
+% Finally, \citet{Zhu:2023wci} report \emph{ad hoc} exclusion of certain events from the combined constraint, without explicitly providing a list of events considered.
+% All these factors combined preclude a clear interpretation of the results in \citep{Zhu:2023wci}.
+% In \ref{fig:posterior_MPV}, we show the posterior on $M_{\rm PV}^{-1}$ for each event in our analysis.
 
@@
     \caption{
-        Individual-event $M_\mathrm{PV}^{-1}$ posteriors from our analysis, converted from the $\kappa$ posteriors in Fig.~\ref{fig:violin_kappa}.
-        Chosen events are those labeled in Fig.~1 of \citet{Zhu:2023wci}.
+        Individual-event $M_\mathrm{PV}^{-1}$ posteriors from our analysis, converted from the $\kappa$ posteriors in Fig.~\ref{fig:violin_kappa} via Eq.~\eqref{eq:kappa-to-Mpv}.
+        Chosen events are those highlighted in Fig.~1 of \citet{Zhu:2023wci}.
     }

### 2023-09-01 src/tex/ms.tex
ref edits

@@
 To constrain birefringence, we reanalyze events from GWTC-3 \citep{GWTC-2.1, GWTC-3} implementing Eq.~\eqref{eq:waveform_modification} to directly obtain a posterior on $\kappa$ from the strain of each event.
-We analyze the 71 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$; to avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
+We analyze the 71 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$\added{, less stringent than the typical \ac{LVK} threshold of $1/1000\,\mathrm{yr}$ \cite{LIGOScientific:2020tif,LIGOScientific:2021sio}.
+The \ac{FAR} values are typically determined by \ac{GR} pipelines, which could down-rank signals beyond \ac{GR} \cite{LIGOScientific:2020tif,Chia:2020psj,tgrsel}; however, since detectability through matched-filtering is most sensitive to the phase, not the amplitude, we expect only a minor decrease in sensitivity to the kind of birefringent signals explored here.}
+To avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
 We procure strain data from \ac{GWOSC} \citep{GWOSC}.
@@
 assuming a small $\delta h$ and $\delta \Psi$ in the last line.
-To consider only amplitude birefringence, as in this work, we must compare to the result in Ref.~\citep{Wang_2021} that set $\delta \Psi = 0$ and parameterized $\delta h = \pi f z / M_{\rm PV}$, where $M_{\rm PV}$ is the energy scale of the birefringent (parity-violating) correction.%
-\footnote{Concretely, \citet{Wang_2021} write $\delta h = - A_\nu \pi f$ with $A_\nu = M_{\rm PV}^{-1} \left[\alpha_\nu (z=0) - \alpha_\nu(z) \left(1+z\right)\right]$, for $\alpha_\nu$ some function of redshift encoding the evolution of a birefringence-mediating field; to produce their constraint, however, they further set $\alpha_\nu(z) = 1$, yielding $A_\nu = -z / M_{\rm PV}$ and hence $\delta h = \pi f z / M_{\rm PV}$.}
-Comparing to our Eq.~\eqref{eq:waveform_modification} and approximating $z \approx d_C H_0/c$ via the Hubble constant, $H_0$, this means
+To consider only amplitude birefringence, as in this work, we must compare to the result in Ref.~\citep{Wang_2021} that set $\delta \Psi = 0$ and parameterized $\delta h = \pi f z h_P / M_{\rm PV}$, where $h_P$ is Planck's constant and $M_{\rm PV}$ is the energy scale of the birefringent (parity-violating) correction.%
+\footnote{Concretely, \citet{Wang_2021} write $\delta h = - A_\nu \pi f$ with $A_\nu = M_{\rm PV}^{-1} \left[\alpha_\nu (z=0) - \alpha_\nu(z) \left(1+z\right)\right]$, for $\alpha_\nu$ some function of redshift encoding the evolution of a birefringence-mediating field; to produce their constraint, however, they further set $\alpha_\nu(z) = 1$, yielding $A_\nu = -z / M_{\rm PV}$ and hence $\delta h = \pi f z h_P / M_{\rm PV}$, multiplying by $h_P$ to obtain the right dimensions.}
+Comparing to our Eq.~\eqref{eq:waveform_modification} and approximating $z \approx d_C H_0/c$ via the Hubble constant, $H_0$, this means that
 \begin{equation}
     \label{eq:kappa-to-Mpv}
-    |\kappa| \simeq \frac{\pi H_0}{c} \frac{h \left( 100 \, \mathrm{Hz} \right)}{M_{\rm PV}} \simeq \frac{3.0 \times 10^{-22} \, \mathrm{GeV}}{M_\mathrm{PV}} \, \mathrm{Gpc}^{-1},
+    |\kappa| \simeq \frac{\pi H_0}{c} \frac{h_P \left( 100 \, \mathrm{Hz} \right)}{M_{\rm PV}} \simeq \frac{3.0 \times 10^{-22} \, \mathrm{GeV}}{M_\mathrm{PV}} \, \mathrm{Gpc}^{-1} .
 \end{equation}
-where $h$ is Planck's constant.
 \citet{Wang_2021} quote a constraint of $M_{\rm PV} > 10^{-22}\, {\rm GeV}$ at 90\% credibility, which translates into \variable{output/kappa_Wang.txt}.

### 2023-09-01 src/tex/ms.tex
final ref pass

@@
 \usepackage{bookmark}
-\usepackage{changes}
+\usepackage[final]{changes}
 %\usepackage[all]{hypcap}
@@
 In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
-\added{Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening; similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci}.}
-To first order, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
+\added{Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening; similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci,Jenks:2023pmk}.}
+To first order \added{in the equations of motion}, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
 \begin{equation}
@@
     h_{L/R}^{\mathrm{GR}}(f) \times
-    \exp\left(\pm\kappa d_C \frac{f}{100\,\mathrm{Hz}}\right)\,,
+    \exp\left(\pm\kappa\, d_C \frac{f}{100\,\mathrm{Hz}}\right)\,,
     \label{eq:waveform_modification}
@@
 where the emitted waveform $h_{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_{L/R}^{\mathrm{br}}$.
-The overall magnitude of this effect for a given frequency $f$ is set by an attenuation coefficient, $\kappa$, which encodes the intrinsic strength of the birefringence.
+The overall magnitude of this effect for a given frequency $f$ is set by an attenuation coefficient, $\kappa$, which encodes the intrinsic strength of the birefringence:
+\added{$\kappa^{-1}$ represents an ``attenuation length'' encoding the typical distance that yields an $e$-folding in the amplification or dampening of the polarizations at a fiducial signal frequency of 100 Hz at the detector.}
 The emitted waveform for a given source (i.e., the waveform observed in the near zone, very close to the source) will generally differ from the analogous waveform predicted by \ac{GR} \cite{Alexander:2009tp,Okounkova:2019zjf}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small (e.g., \cite{Okounkova:2022grv}), it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
@@
 In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity \cite{Alexander:2009tp}.
-Since it originates from a truncated series expansion%
+Since it originates from a truncated series expansion,%
 \footnote{Often the birefringent effect is written as an expansion in redshift, $z$ \cite[e.g.][]{Zhao:2019xmm}, rather than distance.
 Here we use $d_C$ to emphasize that the effect accumulates \emph{per cycle}, or with propagation distance.
-Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$ (Hubble distance), are already observing propagation over a large fraction of the universe and constrain this effect meaningfully.}, %
-Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
+Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$ (Hubble distance), are already observing propagation over a large fraction of the universe and constrain this effect meaningfully.} %
+Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, namely
 \begin{equation}
-    \left|\kappa\right| \left(d_c\added{/1\, \mathrm{Gpc}}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, \added{\mathrm{Gpc}^{-1}}\,.
+    \left(\frac{\left|\kappa\right|}{\added{\mathrm{Gpc}^{-1}}}\right) \left(\frac{d_c}{\added{\mathrm{Gpc}}}\right) \left(\frac{f}{100\, \mathrm{Hz}}\right) < 1\,,
     \label{eq:small_exponent}
 \end{equation}
-Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
+recalling that $\kappa$ has dimensions of inverse length.
+Otherwise, more frequency-dependent terms could enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
 \begin{figure}
@@
 We take the \ac{PE} configuration in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
-We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time, and set the prior on $\kappa$ to be uniform between $-1 \, \mathrm{Gpc}^{-1}$ and $1 \, \mathrm{Gpc}^{-1}$.
+We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time \added{(see, e.g., Eq.~10 in \cite{LIGOScientific:2019zcs})}, and set the prior on $\kappa$ to be uniform between $-1 \, \mathrm{Gpc}^{-1}$ and $1 \, \mathrm{Gpc}^{-1}$.
 
@@
 In this section, we present the results of our study.
-We first show the $\kappa$ measurements of all events in our set, as well as the resulting global measurement of $\kappa$ that represents our primary constraint on birefringence (Sec.~\ref{sec:results:gwtc}).
+We first show the $\kappa$ measurements from \deleted{of} all events in our set, as well as the resulting global measurement of $\kappa$ that represents our primary constraint on birefringence (Sec.~\ref{sec:results:gwtc}).
 We then assess the collection of measurements in more detail through a hierarchical analysis (Sec.~\ref{sec:results:hier}).
@@
 That table also shows the credible level (CL) at which the posterior supports $\kappa = 0$, whereby $\mathrm{CL} = 0$ ($\mathrm{CL} = 1$) means the posterior supports that value with high (low) probability.
-\added{This quantity can be understood as the relative height of the probability density function at a given value, and does not assume any symmetric interval or modality.}
+\added{This quantity represents the relative height of the probability density function at $\kappa = 0$, and is not tied to a symmetric interval, making it particularly useful in assessing bimodal posteriors.}
 
@@
         The marginalized posterior on $\chi_p$ is shown in the top right panel. (See Fig.~\ref{fig:corner_GW190521_appendix} for a full corner plot.)
-        \added{Note that the condition in Eq.~\eqref{eq:small_exponent} is not violated as the distance posterior shown here is the posterior on the luminosity distance $d_L$, not on the comoving distance $d_C$.}
+        \added{Note that, although we show $d_L$ in this plot, our definition of $\kappa$ is associated with $d_C$, meaning that the condition in Eq.~\eqref{eq:small_exponent} is satisfied even for the largest values of $d_L$ supported in this posterior.}
     }
@@
 
+\added{%
+Motivated by Chern-Simons gravity, this study only focused on amplitude birefringence.
+Other parity-violating gravity theories also predict velocity birefringence, which can dominate over amplitude birefringence when both are present \cite{Zhao:2019xmm}.
+Velocity birefringence was tested with GWTC-3 in \citet{Wang:2021gqm} and \citet{Haegel:2022ymk}.
+}
+
 This study was restricted to \acp{BBH} because of the expectation that they should dominate the birefringence constraint thanks to their larger redshifts.
@@
 Future measurements will also be enriched by the addition of KAGRA \cite{KAGRA} and LIGO India to the \ac{GW} detector network, which will allow us to better disambiguate between polarization states and hence between models of birefringent propagation.
-
-\added{%
-Encouraged by Chern-Simons gravity, this study only focused on amplitude birefringence.
-Other parity-violating gravity theories also predict velocity birefringence, which dominates over amplitude birefringence in terms of amplitude and phase corrections to the waveform when both birefringence are present \cite{Zhao:2019xmm}.
-Velocity birefringence was tested with GWTC-3 in \citet{Wang:2021gqm} and \citet{Haegel:2022ymk}.
-}
-
 All our data are available in \citet{dataset}.

### 2023-10-17 src/tex/ms.tex
Update changes from PRD

--- 2023-09-01 src/tex/ms.tex
@@
 \usepackage{changes}
-%\usepackage[all]{hypcap}
 
 \newcommand{\infd}{\mathrm{d}}
-
-\newcommand*{\mi}[1]{\textsf{\color{magenta} [\textbf{MAX:} #1]}}
-\newcommand*{\wf}[1]{\textsf{\color{cyan} [\textbf{WILL:} #1]}}
-
-\definechangesauthor[name=Will, color=cyan]{WF}
 
@@
 \email{misi@flatironinstitute.org}
-\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, United States}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Avenue, New York, New York 10010, USA}
 
@@
 \email{kwong@flatironinstitute.org}
-\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, United States}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Avenue, New York, New York 10010, USA}
 
@@
 \email{wfarr@flatironinstitute.org}
-\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, United States}
-\affiliation{Department of Physics and Astronomy, Stony Brook University, Stony Brook NY 11794, United States}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Avenue, New York, New York 10010, USA}
+\affiliation{Department of Physics and Astronomy, Stony Brook University, Stony Brook, New York 11794, USA}
 
@@
     For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun for parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
-    Here we look for evidence of amplitude birefringence in the latest LIGO-Virgo catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
+    Here we look for evidence of amplitude birefringence in the \replaced{third gravitational-wave transient}{latest Laser Interferometer Gravitational Wave Observatory-Virgo} catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
     From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent attenuation of \variable{output/restricted_kappa_median.txt} at $100 \, \mathrm{Hz}$ with 90\% credibility, equivalent to a parity-violation energy scale of \variable{output/M_PV_constraint.txt}.
@@
 \acro{FAR}{false-alarm rate}
-\acro{GWOSC}{the Gravitational Wave Open Science Center}
 \acro{SNR}{signal-to-noise ratio}
+\acro{2D}{two-dimensional}
 \end{acronym}
@@
 Previous studies have constrained amplitude birefringence by performing different statistical analyses.
-\citet{Yamada_2020} and \citet{Wang_2021} both performed \ac{PE} on the events in the first \ac{GW} transient catalog \citep{GWTC-1}, GWTC-1, using birefringent templates.
+\citet{Yamada_2020}, \citet{Wang_2021} both performed \ac{PE} on the events in the first \ac{GW} transient catalog \citep{GWTC-1}, GWTC-1, using birefringent templates.
 \citet{Okounkova_2022} considered the distribution of observed inclinations of the \ac{GW} events in the second \ac{GW} transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence; \citet{Vitale:2022pmu} carried out a related analysis.
@@
 In this study, we use a frequency-dependent birefringence model to constrain the strength of \ac{GW} amplitude birefringence by performing \ac{PE} on \ac{LVK} binaries.
-This model is a better approximation of the birefringence effect expected from theory than the frequency-independent model used in \citet{Okounkova_2022}.
-\added{Including the frequency dependence allows us to break the degeneracy between birefringence and source inclination, which we discuss below in Sec.~\ref{sec:inclination}.}
-Compared to \replaced{Refs.~\citep{Yamada_2020,Wang_2021,Okounkova_2022}}{other studies}, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
+This model is a better approximation of the birefringence effect expected from theory rather than the frequency-independent model used in \citet{Okounkova_2022}.
+Including the frequency dependence allows us to break the degeneracy between birefringence and source inclination, which we discuss below in Sec.~\ref{sec:inclination}.
+Compared to Refs.~\citep{Yamada_2020,Wang_2021,Okounkova_2022}, we perform \ac{PE} on more events, including events new to GWTC-3 \citep{GWTC-3}, and use a phenomenology-oriented parametrization.
 We consider 71 binary black hole merger events with a \ac{FAR} $\leq1/\mathrm{yr}$, as listed in Table I of \citet{GWTC-3_population}.
@@
 In Sec.~\ref{sec:Discussion}, we discuss the implications of our result, comparing to previous constraints in the literature and outlining correlation structures that appear in our measurements.
-Finally, we conclude in Sec.~\ref{sec:Discussion} with a summary.
+Finally, we conclude in Sec.~\replaced{\ref{sec:Conclusion}}{\ref{sec:Discussion}} with a summary.
 
@@
 In \ac{GR}, \acp{GW} are comprised of two independent polarization modes, usually represented in the linear basis of plus ($+$) and cross ($\times$) states.
-In the Fourier domain, these can be combined into left-handed (L) and right-handed (R) circular states (see, e.g., \cite{Isi:2022mbx}),
+In the Fourier domain, these can be combined into left-handed (\textit{L}) and right-handed (\textit{R}) circular states (see, e.g., \cite{Isi:2022mbx}),
 \begin{equation}
@@
 In particular, \emph{amplitude} birefringence would enhance one polarization mode over the other.
-\added{Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening; similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci}.}
-To first order, in theories like Chern-Simons gravity, the Fourier-domain waveform observed a comoving distance $d_C$ away from the source can be written as
+Following \cite{Alexander:2009tp}, the effect of birefringence can be modeled as a frequency-dependent amplification or dampening; similar derivations can be found in \cite{Zhao:2019xmm, Ezquiaga:2021ler, Zhu:2023wci,Jenks:2023pmk}.
+To first order in the equations of motion, in theories like Chern-Simons gravity, the Fourier-domain waveform observed at a comoving distance $d_C$ away from the source can be written as
 \begin{equation}
@@
 where the emitted waveform $h_{L/R}^{\mathrm{GR}}$ is modified by an exponential birefringent factor to yield the observed waveform $h_{L/R}^{\mathrm{br}}$.
-The overall magnitude of this effect for a given frequency $f$ is set by an attenuation coefficient, $\kappa$, which encodes the intrinsic strength of the birefringence.
+The overall magnitude of this effect for a given frequency $f$ is set by an attenuation coefficient $\kappa$, which encodes the intrinsic strength of the birefringence:
+$\kappa^{-1}$ represents an ``attenuation length'' encoding the typical distance that yields an $e$-folding in the amplification or dampening of the polarizations at a fiducial signal frequency of 100 Hz at the detector.
 The emitted waveform for a given source (i.e., the waveform observed in the near zone, very close to the source) will generally differ from the analogous waveform predicted by \ac{GR} \cite{Alexander:2009tp,Okounkova:2019zjf}; however, since we expect most viable modifications to \ac{GR} to be intrinsically small (e.g., \cite{Okounkova:2022grv}), it is standard to approximate the emitted waveform by the prediction from \ac{GR} (hence the notation ``$h^{\rm GR}$'' above).
@@
 
-Equation \eqref{eq:waveform_modification} can be derived as the first order effect in an expansion away from \ac{GR} under multiple frameworks.
-In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudo-scalar field and its derivative in Chern-Simons gravity \cite{Alexander:2009tp}.
-Since it originates from a truncated series expansion%
-\footnote{Often the birefringent effect is written as an expansion in redshift, $z$ \cite[e.g.][]{Zhao:2019xmm}, rather than distance.
+Equation \eqref{eq:waveform_modification} can be derived as the first-order effect in an expansion away from \ac{GR} under multiple frameworks.
+In general, $\kappa$ will be a function of the theory parameters and the cosmological history, e.g., the value of the pseudoscalar field and its derivative in Chern-Simons gravity \cite{Alexander:2009tp}.
+Since it originates from a truncated series expansion,%
+\footnote{Often the birefringent effect is written as an expansion in redshift $z$ \cite[e.g.][]{Zhao:2019xmm}, rather than distance.
 Here we use $d_C$ to emphasize that the effect accumulates \emph{per cycle}, or with propagation distance.
-Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$ (Hubble distance), are already observing propagation over a large fraction of the universe and constrain this effect meaningfully.}, %
-Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, 
+Thus, current ground-based experiments, which can detect gravitational waves to $z \simeq 1$, $d_C \simeq d_H$ (Hubble distance), are already observing propagation over a large fraction of the Universe and constrain this effect meaningfully.} %
+Eq.~\eqref{eq:waveform_modification} is a good approximation only for small exponents, namely
 \begin{equation}
-    \left|\kappa\right| \left(d_c\added{/1\, \mathrm{Gpc}}\right) \left(f/100\, \mathrm{Hz}\right) < 1\, \added{\mathrm{Gpc}^{-1}}\,.
+    \left(\frac{\left|\kappa\right|}{\mathrm{Gpc}^{-1}}\right) \left(\frac{d_c}{\mathrm{Gpc}}\right) \left(\frac{f}{100\, \mathrm{Hz}}\right) < 1\,,
     \label{eq:small_exponent}
 \end{equation}
-Otherwise, more frequency-dependent terms would enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
+recalling that $\kappa$ has dimensions of inverse length.
+Otherwise, more frequency-dependent terms could enter the exponent of Eq.~\eqref{eq:waveform_modification} in a theory-dependent way.
 \begin{figure}
@@
     \caption{
-        \emph{Illustration of amplitude birefringence.} The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
-        However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
+        Illustration of amplitude birefringence.
+The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
+        However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, then the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
         In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
@@
 \end{equation}
-for all frequencies (see, e.g., Sec.~IIIC in \cite{Isi:2022mbx}).
+for all frequencies (see, e.g., Sec.~III C in \cite{Isi:2022mbx}).
 
@@
 However, the two effects are degenerate only if the frequency dependence of Eq.~\eqref{eq:waveform_modification} is neglected.
-This is easy to see from Eq.~\eqref{eq:waveform_modification}, since the implied polarization ratio for the $\ell = |m| = 2$ mode of a nonprecessing source is
+This is easy to see from Eq.~\eqref{eq:waveform_modification} since the implied polarization ratio for the $\ell = |m| = 2$ mode of a nonprecessing source is
 \begin{equation}
@@
 By implementing Eq.~\eqref{eq:waveform_modification}, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
-Nevertheless, there exist systems for which the degeneracy cannot be broken in practice because not enough frequencies are available in the data.
+Nevertheless, there exist systems for which the degeneracy cannot be broken, in practice, because not enough frequencies are available in the data.
 This may also be the case for quasimonochromatic sources, like nonaxisymmetric pulsars or very light binaries, which are well approximated by a single Fourier mode.
@@
 To constrain birefringence, we reanalyze events from GWTC-3 \citep{GWTC-2.1, GWTC-3} implementing Eq.~\eqref{eq:waveform_modification} to directly obtain a posterior on $\kappa$ from the strain of each event.
-We analyze the 71 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$\added{, less stringent than the typical \ac{LVK} threshold of $1/1000\,\mathrm{yr}$ \cite{LIGOScientific:2020tif,LIGOScientific:2021sio}.
-The \ac{FAR} values are typically determined by \ac{GR} pipelines, which could down-rank signals beyond \ac{GR} \cite{LIGOScientific:2020tif,Chia:2020psj,tgrsel}; however, since detectability through matched-filtering is most sensitive to the phase, not the amplitude, we expect only a minor decrease in sensitivity to the kind of birefringent signals explored here.}
+We analyze the 71 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$, less stringent than the typical \ac{LVK} threshold of $1/1000\,\mathrm{yr}$ \cite{LIGOScientific:2020tif,LIGOScientific:2021sio}.
+The \ac{FAR} values are typically determined by \ac{GR} pipelines, which could down-rank signals beyond \ac{GR} \cite{LIGOScientific:2020tif,Chia:2020psj,tgrsel}; however, since detectability through matched-filtering is most sensitive to the phase, not the amplitude, we expect only a minor decrease in sensitivity to the kind of birefringent signals explored here.
 To avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
-We procure strain data from \ac{GWOSC} \citep{GWOSC}.
+We procure strain data from \add{the} Gravitational Wave Open Science Center \citep{GWOSC}.
 
@@
 We take the \ac{PE} configuration in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
-We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time, and set the prior on $\kappa$ to be uniform between $-1 \, \mathrm{Gpc}^{-1}$ and $1 \, \mathrm{Gpc}^{-1}$.
+We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~10 in \cite{LIGOScientific:2019zcs}), and set the prior on $\kappa$ to be uniform between $-1 \, \mathrm{Gpc}^{-1}$ and $1 \, \mathrm{Gpc}^{-1}$.
 
@@
 \end{equation}
-where $d_i$ is the strain data for the $i$\textsuperscript{th} event, and $p(\kappa)$ is the prior on $\kappa$; since the prior is uniform, in our case Eq.~\eqref{eq:restricted_posterior} reduces to the product of the posteriors, namely $p(\kappa \mid \{d_i\}) \propto \prod_{i}p(\kappa \mid d_i)$.
+where $d_i$ is the strain data for the $i$th event, and $p(\kappa)$ is the prior on $\kappa$; since the prior is uniform, in our case Eq.~\eqref{eq:restricted_posterior} reduces to the product of the posteriors, namely $p(\kappa \mid \{d_i\}) \propto \prod_{i}p(\kappa \mid d_i)$.
 We use Eq.~\eqref{eq:restricted_posterior} to obtain the primary constraint presented in this work.
@@
 we posit that, rather than a unique global value of $\kappa$, there is a specific value of the parameter, $\kappa_i$, associated with each event, and that this is drawn from some unknown distribution of true underlying values; from the imperfect measurements of $\kappa_i$ for each event, we may reconstruct the underlying distribution.
-If we are interested in constraining the first two moments of the distribution, it is convenient to parametrize the $\kappa_i$'s as drawn from a Gaussian with unknown mean $\mu$ and variance $\sigma^2$, i.e., $\kappa_i \sim \mathcal{N}(\mu, \sigma^2)$ \cite{Isi:2019asy}, and measure those hyperparameters from the collection of observed data.
+If we are interested in constraining the first two moments of the distribution, then it is convenient to parametrize the $\kappa_i$'s as drawn from a Gaussian with unknown mean $\mu$ and variance $\sigma^2$, i.e., $\kappa_i \sim \mathcal{N}(\mu, \sigma^2)$ \cite{Isi:2019asy}, and measure those hyperparameters from the collection of observed data.
 
-If \ac{GR} is correct and there is no birefringence, we should find the observed $\kappa$ distribution to be consistent with a delta function at the origin ($\kappa_i = 0$ for all $i$, or $\mu=\sigma=0$); on the other hand, if spacetime is globally birefringent, we expect to find a delta function at some nonzero value ($\kappa_i = \kappa \neq 0$, or $\mu = \kappa$ and $\sigma=0$).
-But this analysis also has the power to reveal unexpected physics or systematics in our measurements: if $\sigma$ is confidently found to be nonzero, this would imply that our set of measurements is statistically unlikely to originate from a unique $\kappa$ value.
+If \ac{GR} is correct and there is no birefringence, then we should find the observed $\kappa$ distribution to be consistent with a delta function at the origin ($\kappa_i = 0$ for all $i$, or $\mu=\sigma=0$); on the other hand, if spacetime is globally birefringent, then we expect to find a delta function at some nonzero value ($\kappa_i = \kappa \neq 0$, or $\mu = \kappa$ and $\sigma=0$).
+But this analysis also has the power to reveal unexpected physics or systematics in our measurements: if $\sigma$ is confidently found to be nonzero, then this would imply that our set of measurements is statistically unlikely to originate from a unique $\kappa$ value.
 This could signal richer physics than is implied by Eq.~\eqref{eq:waveform_modification} or, more prosaically, that there are outliers in our measurements due to mismodeling, e.g., in the waveform approximant or the noise of the detector.
@@
 In this section, we present the results of our study.
-We first show the $\kappa$ measurements of all events in our set, as well as the resulting global measurement of $\kappa$ that represents our primary constraint on birefringence (Sec.~\ref{sec:results:gwtc}).
+We first show the $\kappa$ measurements from all events in our set, as well as the resulting global measurement of $\kappa$ that represents our primary constraint on birefringence (Sec.~\ref{sec:results:gwtc}).
 We then assess the collection of measurements in more detail through a hierarchical analysis (Sec.~\ref{sec:results:hier}).
-Finally, we discuss some special events individually, and outline the degeneracies that arise between birefringence and orbital precession (Sec.~\ref{sec:results:notable}).
+Finally, we discuss some special events, individually, and outline the degeneracies that arise between birefringence and orbital precession (Sec.~\ref{sec:results:notable}).
 
@@
     \caption{
-        Individual-event $\kappa$ posteriors (distributions), and joint measurement (blue band, 90\% CI; blue line, median).
+        Individual-event $\kappa$ posteriors (distributions) and joint measurement (blue band, 90\% CI; blue line, median).
     }
@@
 That table also shows the credible level (CL) at which the posterior supports $\kappa = 0$, whereby $\mathrm{CL} = 0$ ($\mathrm{CL} = 1$) means the posterior supports that value with high (low) probability.
-\added{This quantity can be understood as the relative height of the probability density function at a given value, and does not assume any symmetric interval or modality.}
+This quantity represents the relative height of the probability density function at $\kappa = 0$, and is not tied to a symmetric interval, making it particularly useful in assessing bimodal posteriors.
 
@@
 \begin{table}
-    \caption{Events with bimodality in the $\kappa$ posterior, the \ac{GR} measurement of their detector-frame total mass ($M$), precessing spin $\chi_p$ and effective spin $\chi_{\rm eff}$, as well as the credible level of $\kappa = 0$ (CL) from the birefringence analysis.}
+    \caption{Events with bimodality in the $\kappa$ posterior, the \ac{GR} measurement of their detector-frame total mass ($M$), precessing spin $\chi_p$, and effective spin $\chi_{\rm eff}$, as well as the CL of $\kappa = 0$ from the birefringence analysis.}
     \begin{ruledtabular}
@@
 For these bimodal posteriors, $\mu_i/\sigma_i$ is not a good proxy for agreement with \ac{GR}; instead, we can rely on $\mathrm{CL}(\kappa = 0)$.
-By this measure, the bimodal events are some of the least consistent with $\kappa = 0$, GW190521 in particular.
+By this measure, the bimodal events are some of the least consistent with $\kappa = 0$, GW190521, in particular.
 % (Nevertheless, configurations with $\kappa = 0$ for this event are still within the higher dimensional 90\%-credible region, when other parameters are considered and not just the one-dimensional marginal; we discuss this in Sec.~\ref{sec:GW190521}.)
 
-As we anticipated in Sec.~\ref{sec:inclination}, we understand the bimodality in $\kappa$ to be linked to spin effects, and often to precessing morphologies in particular.
+As we anticipated in Sec.~\ref{sec:inclination}, we understand the bimodality in $\kappa$ to be linked to spin effects and often to precessing morphologies in particular.
 Other parameter degeneracies also come into play, especially for the lighter events GW191105\_143521 (henceforth GW191105) and GW170104.
@@
         The posterior of the $\kappa$ population hyperparameters $\mu$ and $\sigma$, including (blue) and excluding (orange) GW200129 from the collection of events.
-        The 2D contours correspond to the $39.35\%$ and $90\%$ credible levels.
+        The \ac{2D} contours correspond to the $39.35\%$ and $90\%$ credible levels.
         The plot shows that the population constraint on $\kappa$ is consistent with no birefringence ($\mu=\sigma=0$) at the 90\% credible level.
@@
 This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Miller2020,Callister:T2100301}.
-The result in Figure \ref{fig:reweighted_kappa} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its higher support for $\kappa > 0$.
+The result in Fig.~\ref{fig:reweighted_kappa} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its higher support for $\kappa > 0$.
 This feature appears to offset a few other events which tend to favor $\kappa < 0$.
@@
     \caption{
-        Restricted and generic posteriors on $\kappa$, which respectively do and do not assume that $\kappa$ is shared by all events (color).
-        \replaced{%
-        Solid (dashed) traces indicate the analysis included (excluded) GW200129.}
-        {Solid and dashed traces indicate whether the analysis included GW200129 or not.}
+        Restricted and generic posteriors on $\kappa$, which, respectively, do and do not assume that $\kappa$ is shared by all events (color).
+        Solid (dashed) traces indicate the analysis included (excluded) GW200129.
         The black dashed line marks the absence of birefringence ($\kappa=0$).
@@
 
-Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on three events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:violin_kappa}: GW170818, GW190521 and GW200129.
+Having established that the collection of detections is globally consistent with $\kappa=0$, here we focus on three events whose $\kappa$ posteriors stand out in Fig.~\ref{fig:violin_kappa}: GW170818, GW190521, and GW200129.
 When considered in isolation, the first of these is the unimodal event with the most significant support for nonzero $\kappa$, the second is the most extreme representative of a class of events with bimodal $\kappa$ posteriors, and the third shows signs of potential noise anomalies.
-Through these examples, we elucidate some of the interactions between $\kappa$ and the source luminosity distance, inclination and spin parameters.
+Through these examples, we elucidate some of the interactions between $\kappa$ and the source luminosity distance, inclination, and spin parameters.
 
@@
     \caption{
-        GW170818 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+        GW170818 posterior on $\kappa$, luminosity distance $d_L$, and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
         The top right panel shows the marginalized posterior on $\chi_p$: allowing for birefringence reduces the preference for precession.
@@
 Figure \ref{fig:corner_GW170818} shows that this happens because birefringence opens up a region of parameter space with $\kappa>0$ for larger distances and smaller inclination angles than would be allowed in the GR case.
-We can make sense of this by noting that an edge-on, nonprecessing source produces linearly-polarized waves, meaning that a smaller inclination leads the two circular polarizations to have similar amplitudes.
+We can make sense of this by noting that an edge-on nonprecessing source produces linearly-polarized waves, meaning that a smaller inclination leads the two circular polarizations to have similar amplitudes.
 On the other hand, having $\kappa > 0$ enhances the left-handed modes during propagation, per Eq.~\eqref{eq:waveform_modification}.
-The two effects can be balanced to match the polarization ratio observed at the detector (predominantly left-handed, per the preference for $\cos\iota \approx -1$ in the \ac{GR} analysis), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
+The two effects can be balanced to match the polarization ratio observed at the detector (predominantly left handed, per the preference for $\cos\iota \approx -1$ in the \ac{GR} analysis), as long as the distance is also enhanced to yield the right amount of birefringence and overall signal power.
 This is similar to the degeneracy mentioned in Sec.~\ref{sec:inclination}.
@@
 It is difficult to unequivocally identify a specific feature of the GW170818 data that leads to this posterior structure.
-However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization, phase and spin angles \cite{Varma:2021csh} (see Appendix \ref{sec:corner_GW170818_appendix}).
+However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization, phase, and spin angles \cite{Varma:2021csh} (see Appendix \ref{sec:corner_GW170818_appendix}).
 The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:corner_GW170818}, top right): allowing for $\kappa \neq 0$ leads to reduced support for precession.
@@
 
-The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ can be explained a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
+The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ can be explained by a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
 The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and the specific phasing of the precession cycle.
@@
     \caption{
-        GW190521 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
+        GW190521 posterior on $\kappa$, luminosity distance $d_L$, and inclination $\cos\iota$ from our birefringence analysis (blue), compared to the GR result (orange).
         The marginalized posterior on $\chi_p$ is shown in the top right panel. (See Fig.~\ref{fig:corner_GW190521_appendix} for a full corner plot.)
-        \added{Note that the condition in Eq.~\eqref{eq:small_exponent} is not violated as the distance posterior shown here is the posterior on the luminosity distance $d_L$, not on the comoving distance $d_C$.}
+        Note that, although we show $d_L$ in this plot, our definition of $\kappa$ is associated with $d_C$, meaning that the condition in Eq.~\eqref{eq:small_exponent} is satisfied even for the largest values of $d_L$ supported in this posterior.
     }
@@
 
-To this end, we perform three additional \ac{PE} runs for GW200129, considering only two detectors at a time: LIGO Hanford and Virgo (HL), LIGO Livingston and Virgo (LV), and LIGO Hanford and LIGO Livingston (HL).
+To this end, we perform three additional \ac{PE} runs for GW200129, considering only two detectors at a time: LIGO Hanford and Virgo (\replaced{HV}{HVL}), LIGO Livingston and Virgo (LV), and LIGO Hanford and LIGO Livingston (HL).
 If the preference for $\kappa < 0$ is tied to the glitch in Virgo, we expect it to disappear in the HL run, which excludes Virgo data.
@@
     \caption{
-        GW200129 posterior on $\kappa$, luminosity distance $d_L$ and inclination $\cos{\iota}$, including different sets of detectors in the analysis per the legend.
+        GW200129 posterior on $\kappa$, luminosity distance $d_L$, and inclination $\cos{\iota}$, including different sets of detectors in the analysis per the legend.
         The marginalized posterior on $\chi_p$ is shown in the top right panel.
         The main run with all three detectors (HLV, filled blue), shows a preference for $\kappa < 0$, as in Fig.~\ref{fig:violin_kappa}; this preference is more pronounced for two-detector runs that include Virgo (HV and LV, orange and green); however, it disappears if we remove Virgo (HL, dashed black).
-        The 2D contours correspond to the $90\%$ credible level.
+        The \ac{2D} contours correspond to the $90\%$ credible level.
     }
@@
 \citet{Okounkova_2022} produced a constraint on frequency-independent \ac{GW} amplitude birefringence from the distribution of measured inclinations of \acp{BBH} in GWTC-2.
-As opposed to our Eq.~\eqref{eq:waveform_modification}, this reference parameterized the effect of birefringence as
+As opposed to our Eq.~\eqref{eq:waveform_modification}, this reference parametrized the effect of birefringence as
 \begin{equation}
@@
 
-The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into a FRW cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.},
+The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale,\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into an Friedmann-Robertson-Walker cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.}
 \begin{equation}
@@
 \citet{Wang_2021} performed \ac{PE} on GWTC-1 events with a frequency-dependent model of birefringence resembling ours.
-Following \cite{Zhao:2019xmm}, that reference parameterized the birefringent waveform in terms of some amplitude and phase modifications to the linear polarizations ($\delta h$ and $\delta \Psi$ respectively), such that
+Following \cite{Zhao:2019xmm}, that reference parametrized the birefringent waveform in terms of some amplitude and phase modifications to the linear polarizations ($\delta h$ and $\delta \Psi$ respectively), such that
 \begin{equation}
@@
 assuming a small $\delta h$ and $\delta \Psi$ in the last line.
-To consider only amplitude birefringence, as in this work, we must compare to the result in Ref.~\citep{Wang_2021} that set $\delta \Psi = 0$ and parameterized $\delta h = \pi f z h_P / M_{\rm PV}$, where $h_P$ is Planck's constant and $M_{\rm PV}$ is the energy scale of the birefringent (parity-violating) correction.%
+To consider only amplitude birefringence, as in this work, we must compare to the result in Ref.~\citep{Wang_2021} that set $\delta \Psi = 0$ and parametrized $\delta h = \pi f z h_P / M_{\rm PV}$, where $h_P$ is Planck's constant and $M_{\rm PV}$ is the energy scale of the birefringent (parity-violating) correction.%
 \footnote{Concretely, \citet{Wang_2021} write $\delta h = - A_\nu \pi f$ with $A_\nu = M_{\rm PV}^{-1} \left[\alpha_\nu (z=0) - \alpha_\nu(z) \left(1+z\right)\right]$, for $\alpha_\nu$ some function of redshift encoding the evolution of a birefringence-mediating field; to produce their constraint, however, they further set $\alpha_\nu(z) = 1$, yielding $A_\nu = -z / M_{\rm PV}$ and hence $\delta h = \pi f z h_P / M_{\rm PV}$, multiplying by $h_P$ to obtain the right dimensions.}
@@
 
-\deleted{The parameterization of \cite{Wang_2021} is the same as that in the more recent work by \cite{Zhu:2023wci}.
-That work, which appeared while this manuscript was finalized, reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}$ based on GWTC-3.}
-
-\subsubsection{\added{Zhu et al.}}
+\subsubsection{Zhu et al.}
 \label{sec:comparison_Zhu}
@@
 This might be explained by a number of analysis differences.
-The fact that posteriors disagree at the individual-event level suggests that the discrepancy originates in the different choice of parameterization and corresponding priors, in addition to potentially unstated differences in implementation.
+The fact that posteriors disagree at the individual-event level suggests that the discrepancy originates in the different choice of parametrization and corresponding priors, in addition to potentially unstated differences in implementation.
 
@@
 The resulting posterior can be unimodal or bimodal depending on the specific phase and polarization states observed at the detector.
-The events GW170818 (Sec.~\ref{sec:GW170818}) and GW190521 (Sec.~\ref{sec:GW190521}) appear to fall into these two categories respectively.
+The events GW170818 (Sec.~\ref{sec:GW170818}) and GW190521 (Sec.~\ref{sec:GW190521}) appear to fall into these two categories, respectively.
 
@@
 \section{Conclusion}
-\label{sec:conclusion}
+\label{sec:Conclusion}
 
 We have reanalyzed all \acp{BBH} in GWTC-3 with $\mathrm{FAR} < 1/\mathrm{yr}$ to constrain amplitude birefringence in the propagation of \acp{GW}.
-To this end, we implemented a model in which right or left handed polarizations are amplified or suppressed over distance as a function of frequency, with an overall strength parametrized by a birefringent ``attenuation'' parameter, $\kappa$, following Eq.~\eqref{eq:waveform_modification}.
-This parameterization is consistent with parity-odd theories like Chern-Simons gravity, and can be used to constrain them where applicable.
+To this end, we implemented a model in which right or left handed polarizations are amplified or suppressed over distance as a function of frequency, with an overall strength parametrized by a birefringent ``attenuation'' parameter $\kappa$, following Eq.~\eqref{eq:waveform_modification}.
+This parametrization is consistent with parity-odd theories like Chern-Simons gravity and can be used to constrain them where applicable.
 
-We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
+We found no evidence of amplitude birefringence in the GWTC-3 data and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
 This measurement is significantly more stringent than past constraints (Table \ref{tab:comparison_summary}).
@@
 The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
-Future catalogs with more events will shed light on this feature, and enable richer models with explicit correlations between birefringent attenuation and source parameters like sky location \citep{Goyal:2023uvm,Ezquiaga:2021ler}.
+Future catalogs with more events will shed light on this feature and enable richer models with explicit correlations between birefringent attenuation and source parameters like sky location \citep{Goyal:2023uvm,Ezquiaga:2021ler}.
 
@@
 
+Motivated by Chern-Simons gravity, this study only focused on amplitude birefringence.
+Other parity-violating gravity theories also predict velocity birefringence, which can dominate over amplitude birefringence when both are present \cite{Zhao:2019xmm}.
+Velocity birefringence was tested with GWTC-3 in \citet{Wang:2021gqm} and \citet{Haegel:2022ymk}.
+
 This study was restricted to \acp{BBH} because of the expectation that they should dominate the birefringence constraint thanks to their larger redshifts.
@@
 We thank Macarena Lagos and Nicol\'as Yunes for helpful discussions.
-M.~I., K.~W.~K.~W.~, and W.~M.~F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute.
+M.I., K.W.K.W.~, and W.M.F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute.
 The Flatiron Institute provided the computational resources used in this work.
-This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gwosc.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
+This research has made use of data or software obtained from the Gravitational Wave Open Science Center, a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
 LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
@@
 
-\section{Extended results for notable events}
+\section{EXTENDED RESULTS FOR NOTABLE EVENTS}
 
@@
 
-The GR analysis stands out for its relatively confident identification of the spin angles, $\theta_{1/2}$, $\Delta \phi$ and $\phi_{JL}$, as well as the phase and polarization angles, $\phi_{\rm ref}$ and $\psi$.
-Of the former, the two $\theta_{1/2}$ angles encode the tilts of the component \acp{BH} with respect to the orbital angular momentum, $\vec{L}$, $\Delta \phi$ is the angle between the spin projections onto the orbital plane, while $\phi_{JL}$ is a similar angle separating the projections of $\vec{L}$ and the total angular momentum, $\vec{J}$;
+The GR analysis stands out for its relatively confident identification of the spin angles, $\theta_{1/2}$, $\Delta \phi$, and $\phi_{JL}$, as well as the phase and polarization angles, $\phi_{\rm ref}$ and $\psi$.
+Of the former, the two $\theta_{1/2}$ angles encode the tilts of the component \acp{BH} with respect to the orbital angular momentum $\vec{L}$, and $\Delta \phi$ is the angle between the spin projections onto the orbital plane, while $\phi_{JL}$ is a similar angle separating the projections of $\vec{L}$ and the total angular momentum $\vec{J}$;
 of the latter, $\phi_{\rm ref}$ is an overall reference phase, and $\psi$ encodes the orientation of the binary within the plane of the sky \cite{Isi:2022mbx}.
@@
 As in the previous section, here we display additional parameters from the GW190521 measurement in Fig.~\ref{fig:corner_GW190521_appendix}, expanding upon Fig.~\ref{fig:corner_GW190521} in the main text.
-The reference \ac{GR} analysis, obtained with the \textsc{IMRPhenomXPHM} waveform, favors near extremal spins, in particular for the heavier component, and rules out nonspinning objects (i.e., $\chi_1 = \chi_2 = 0$) with $\geq 90\%$ credibility.
+The reference \ac{GR} analysis, obtained with the \textsc{IMRPhenomXPHM} waveform, favors near extremal spins, in particular, for the heavier component, and rules out nonspinning objects (i.e., $\chi_1 = \chi_2 = 0$) with $\geq 90\%$ credibility.
 This can be seen most clearly in the joint posterior for $\chi_1$ and $\chi_2$, rather than in the respective 1D marginals because the \acp{BH} in this system were inferred to have equal masses \cite{Biscoveanu:2020are}.
@@
 However, the fact that $\psi$ is constrained does not alone imply that the degeneracy between right- and left-handed states is broken (for a nonprecessing source, $\psi$ is related to the difference in phase between the circular polarization modes \cite{Isi:2022mbx}).
-Additionally, the posterior for tilt angle of the primary \ac{BH}, $\theta_1$, shows support for an antialigned spin ($\cos\theta_1 \approx -1$).
+Additionally, the posterior for tilt angle of the primary \ac{BH} $\theta_1$ shows support for an antialigned spin ($\cos\theta_1 \approx -1$).
 For these events, spin and phase angles are referred to 10 Hz at the detector.

### 2023-10-17 src/tex/ms.tex
fix bug
Update ms.tex

@@
     For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun for parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
-    Here we look for evidence of amplitude birefringence in the \replaced{third gravitational-wave transient}{latest Laser Interferometer Gravitational Wave Observatory-Virgo} catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
+    Here we look for evidence of amplitude birefringence in the \replaced{third catalog of detections by the Laser Interferometer Gravitational Wave Observatory and Virgo}{latest Laser Interferometer Gravitational Wave Observatory-Virgo catalog (GWTC-3)} through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
     From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent attenuation of \variable{output/restricted_kappa_median.txt} at $100 \, \mathrm{Hz}$ with 90\% credibility, equivalent to a parity-violation energy scale of \variable{output/M_PV_constraint.txt}.
@@
 Finally, we conclude in Sec.~\replaced{\ref{sec:Conclusion}}{\ref{sec:Discussion}} with a summary.
+\added{Additional details on the results for some notable events are presented in \ref{sec:Appendix}.}
 
@@
 To avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
-We procure strain data from \add{the} Gravitational Wave Open Science Center \citep{GWOSC}.
+We procure strain data from Gravitational Wave Open Science Center \citep{GWOSC}.
 
@@
 
-The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale,\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into an Friedmann-Robertson-Walker cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.}
+The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale,\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into a Friedmann-Robertson-Walker cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.}
 \begin{equation}
@@
 \section{EXTENDED RESULTS FOR NOTABLE EVENTS}
+\label{sec:Appendix}
 

### 2023-10-17 src/tex/ms.tex
Update ms.tex

@@
 Finally, we conclude in Sec.~\replaced{\ref{sec:Conclusion}}{\ref{sec:Discussion}} with a summary.
-\added{Additional details on the results for some notable events are presented in \ref{sec:Appendix}.}
+\added{In the Appendix, we provide extended results and discussion for two notable events (GW170818 and GW190521).}
 
@@
 The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
-        However, if spacetime were birefringent following Eq.~\protect\eqref{eq:waveform_modification}, then the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
+        However, if spacetime were birefringent \added{(BR)} following Eq.~\protect\eqref{eq:waveform_modification}, then the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
         In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
@@
 
-\appendix
+\appendix*
 
-\section{EXTENDED RESULTS FOR NOTABLE EVENTS}
+\section{Extended results for notable events}
 \label{sec:Appendix}

### 2023-10-17 src/tex/ms.tex
Update ms.tex
Finalized all changes

@@
 \usepackage{bookmark}
-\usepackage{changes}
 
@@
     For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun for parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
-    Here we look for evidence of amplitude birefringence in the \replaced{third catalog of detections by the Laser Interferometer Gravitational Wave Observatory and Virgo}{latest Laser Interferometer Gravitational Wave Observatory-Virgo catalog (GWTC-3)} through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
+    Here we look for evidence of amplitude birefringence in the third catalog of detections by the Laser Interferometer Gravitational Wave Observatory and Virgo through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
     From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent attenuation of \variable{output/restricted_kappa_median.txt} at $100 \, \mathrm{Hz}$ with 90\% credibility, equivalent to a parity-violation energy scale of \variable{output/M_PV_constraint.txt}.
@@
 In Sec.~\ref{sec:Discussion}, we discuss the implications of our result, comparing to previous constraints in the literature and outlining correlation structures that appear in our measurements.
-Finally, we conclude in Sec.~\replaced{\ref{sec:Conclusion}}{\ref{sec:Discussion}} with a summary.
-\added{In the Appendix, we provide extended results and discussion for two notable events (GW170818 and GW190521).}
+Finally, we conclude in Sec.~\ref{sec:Discussion} with a summary.
+In the Appendix, we provide extended results and discussion for two notable events (GW170818 and GW190521).
 
@@
 The GR waveform for the $\ell=|m|=2$ mode of a nonprecessing BBH seen edge-on ($\cos\iota = 0$) is linearly polarized and thus contains equal amounts of left- and right-handed modes for all frequencies (dotted, top).
-        However, if spacetime were birefringent \added{(BR)} following Eq.~\protect\eqref{eq:waveform_modification}, then the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
+        However, if spacetime were birefringent (BR) following Eq.~\protect\eqref{eq:waveform_modification}, then the waveform observed on Earth would contain different fractions of the two circular modes, with higher frequencies affected more strongly (solid, top).
         In the time domain, this manifests as a time-dependent amplification of the waveform, with a stronger effect at later times when the chirp reaches a higher instantaneous frequency (bottom).
@@
 
-To this end, we perform three additional \ac{PE} runs for GW200129, considering only two detectors at a time: LIGO Hanford and Virgo (\replaced{HV}{HVL}), LIGO Livingston and Virgo (LV), and LIGO Hanford and LIGO Livingston (HL).
+To this end, we perform three additional \ac{PE} runs for GW200129, considering only two detectors at a time: LIGO Hanford and Virgo (HV), LIGO Livingston and Virgo (LV), and LIGO Hanford and LIGO Livingston (HL).
 If the preference for $\kappa < 0$ is tied to the glitch in Virgo, we expect it to disappear in the HL run, which excludes Virgo data.
